% =====================================================================
% Foundations of Practice: A BABOK v3-Aligned Business Analysis Curriculum
%
% Uses the curriculumstyle.sty house-style package. All shared visual
% conventions (colors, fonts, geometry, callouts, headings, title page,
% TOC, PDF metadata, helper macros) are provided by the package.
% This document supplies only:
%   - subject-specific metadata (via \CurriculumSetup)
%   - subject-specific content (modules, tables, capstone, appendices)
%
% Source: BABOK Guide v3 (IIBA, 2015) alignment.
% Target: TeX Live 2022/2023, pdflatex, zero-warning compilation.
% =====================================================================

\documentclass[11pt,letterpaper,oneside]{report}
\usepackage{curriculumstyle}

\CurriculumSetup{
  title={Foundations of Practice},
  subtitle={A BABOK v3-Aligned Business Analysis Curriculum},
  purpose={A modular, outcomes-based, blended program for practicing and aspiring business analysts, with explicit alignment to the IIBA BABOK Guide v3.},
  audience={Practicing and aspiring business analysts at the intermediate level; technically-oriented professionals (engineers, architects, product/project managers, governance and compliance staff); CCBA and CBAP candidates},
  format={11 modules across 14 weeks at ${\sim}$20 hours/week (${\sim}$280 learning hours); accelerated 8-week and self-paced 16-week variants},
  modality={Blended cohort (synchronous instructor-led, asynchronous study, lab/studio practice); fully-online and self-paced variants supported},
  capstone={End-to-end business analysis on the Northstar Mutual case with stakeholder-panel defense},
  alignment={BABOK Guide v3 (IIBA, 2015); Minimum Curriculum Quality Checklist (15 items)},
  header={Foundations of Practice --- BABOK v3},
  label={CURRICULUM HANDOUT \,/\, BABOK V3 EDITION},
  disclaimer={Designed for cohort delivery by an institutional Curriculum Council. Aligned with \emph{A Guide to the Business Analysis Body of Knowledge} (BABOK Guide), Version 3, IIBA, 2015. This handout is a curriculum design document; it is not a substitute for the BABOK itself, which remains the authoritative reference text.},
  pdfauthor={Curriculum Council (Lead Architect, Senior Instructors, External Practitioner Reviewer, Accessibility Reviewer, Learner Representative, IIBA-Credentialed Advisor)},
  pdfsubject={BABOK v3-Aligned Business Analysis Curriculum},
  pdfkeywords={BABOK, business analysis, IIBA, CCBA, CBAP, curriculum, BACCM, requirements, elicitation},
  pdftitle={Foundations of Practice: A BABOK v3-Aligned Business Analysis Curriculum}
}

% =====================================================================
\begin{document}

\makecurriculumtitle

\pagenumbering{roman}
\setcounter{page}{2}
\tableofcontents
\clearpage
\pagenumbering{arabic}
\setcounter{page}{1}

% =====================================================================
\FrontMatterChapter{Curriculum Overview}

\section*{1. Curriculum Identity}
\textbf{\large Foundations of Practice: A BABOK v3-Aligned Business Analysis Curriculum} \\
\textit{Program code BA-PRO-100 \textbar\ Version 1.0 \textbar\ Standards reference: BABOK Guide v3, IIBA, 2015.}

\section*{2. Curriculum Metadata}
\CurriculumMetadataTable

\section*{3. Audience and Entry Prerequisites}

\subsection*{3.1 Primary Audience}
The curriculum is designed for practicing business analysts seeking a structured, BABOK-aligned grounding; for cross-functional professionals (developers, architects, product managers, project managers, compliance specialists) who perform business analysis tasks; and for candidates preparing for IIBA's CCBA or CBAP certifications. The program assumes participants work, or expect to work, on initiatives that change processes, capabilities, or systems for organizational stakeholders.

\subsection*{3.2 Entry Prerequisites}
\begin{itemize}
  \item Two or more years of professional experience in a project, product, technology, operations, or process role.
  \item Basic literacy with workplace collaboration tools (document collaboration, video conferencing, ticket trackers).
  \item Reading-level English (or the language of delivery).
  \item Successful completion of a 30-minute diagnostic that screens for foundational concepts in projects, processes, and stakeholders. Learners who do not pass the diagnostic are routed to a 4-hour bridging primer before Module~01.
\end{itemize}

\subsection*{3.3 Recommended Duration and Delivery Format}
The default delivery is a 14-week blended cohort at approximately 20 hours/week, totaling ${\sim}$280 learning hours: ${\sim}$6~hours/week instructor-led synchronous, ${\sim}$8~hours/week asynchronous study, and ${\sim}$6~hours/week lab/studio practice. The learning platform hosts readings, recordings, artifact templates, peer-review workflows, captioned videos, and a discussion space. Variants for fully online, accelerated (8 weeks), and self-paced (16+ weeks) delivery are specified in Module Usability Package (\S11).

\section*{4. Instructional Philosophy}
The curriculum is outcomes-based, BABOK-aligned, and centered on the practitioner's working artifacts. Learning is sequenced from concept comprehension to applied performance, with most module time spent producing, critiquing, and revising the actual work products a business analyst is expected to deliver. Adult learning principles drive the design: relevance to participants' workplace context, immediate application, peer learning, reflective practice, and progressive autonomy. Every concept is introduced with a concrete scenario, practiced on a controlled exercise, applied to a richer case, and finally exercised under capstone conditions.

\section*{5. How the Curriculum Aligns with the BABOK}
The curriculum maps explicitly to all five structural elements of BABOK v3:

\begin{itemize}
  \item \textbf{Business Analysis Key Concepts} (BABOK \S2) anchor Module~01 and recur in every subsequent module through deliberate use of the Business Analysis Core Concept Model (\skill{BACCM}): change, need, solution, stakeholder, value, context.
  \item \textbf{The six knowledge areas} (BABOK \S3--\S8) each receive a dedicated module (M02--M07), sequenced to support cumulative learning rather than to mirror the BABOK's reference order.
  \item \textbf{Underlying Competencies} (BABOK \S9) are introduced in M01, threaded through every module via deliberate practice activities, and consolidated in M08.
  \item \textbf{Techniques} (BABOK \S10) are taught in context where they are first useful, deliberately reinforced in later modules, and mastered in the Integration Lab (M10) and the Capstone (M11). The technique integration plan is in \S9.
  \item \textbf{Perspectives} (BABOK \S11) --- Agile, Business Intelligence, Information Technology, Business Architecture, and Business Process Management --- are taught as distinct lenses on the same analysis work, applied in dedicated studio rotations in M09, and required in the Capstone.
\end{itemize}

\begin{curriculumnote}[Note on Context Defaults]
The original commissioning prompt left several context placeholders unfilled. Defaults applied: cross-industry coverage with worked examples from financial services, healthcare, regulated SaaS, and government modernization; intermediate learner level with a mandatory foundational module to normalize entry; default delivery length 14 weeks at ${\sim}$280 contact-equivalent hours. These defaults are stated explicitly so they can be confirmed or overridden before deployment.
\end{curriculumnote}

% =====================================================================
\FrontMatterChapter{Program-Level Learning Outcomes}

\section*{6. Program-Level Learning Outcomes (PLOs)}
On successful completion of the program, the learner will be able to perform the fifteen outcomes below. Each is structured as \emph{action verb + observable behavior + criterion} and is mapped to one or more BABOK areas.

\renewcommand{\arraystretch}{1.20}
\rowcolors{2}{afrowalt}{white}
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Program-Level Learning Outcomes (PLO-01 to PLO-15)\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{6pt}%
\begin{longtable}{@{}R{0.85cm} R{10.40cm} R{3.05cm}@{}}
\toprule
\sffamily\bfseries \# &
\sffamily\bfseries Outcome (action verb + observable behavior + criterion) &
\sffamily\bfseries BABOK Area(s) \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries \# &
\sffamily\bfseries Outcome &
\sffamily\bfseries BABOK Area(s) \\
\midrule
\endhead
\bottomrule
\endfoot

PLO-01 & \textbf{Apply} the BACCM to a given initiative by producing a one-page concept map identifying change, need, solution, stakeholder, value, and context, with all six concepts traced to specific evidence in the case. & Key Concepts \\

PLO-02 & \textbf{Construct} a fit-for-purpose business analysis approach (predictive, adaptive, or hybrid) by producing a Business Analysis Plan that specifies governance, information management, deliverables, stakeholder engagement, and performance measures defensible against given constraints. & BAPM \\

PLO-03 & \textbf{Plan and conduct} a multi-technique elicitation event (interview, workshop, observation, or document analysis) and produce confirmed elicitation results suitable for downstream analysis. & Elicitation \& Collaboration \\

PLO-04 & \textbf{Manage} requirements through their lifecycle by producing a traceability matrix, prioritization decisions, change assessments, and approvals consistent with a defined governance approach. & RLCM \\

PLO-05 & \textbf{Analyze} the current state, define a future state, and recommend a change strategy with justified options, risk analysis, and measurable business objectives. & Strategy Analysis \\

PLO-06 & \textbf{Specify} requirements and designs at appropriate levels of abstraction using a coherent set of textual, structured, and visual models that are verifiable, traceable, and validated against need. & RADD \\

PLO-07 & \textbf{Evaluate} a proposed or implemented solution against defined measures of value and recommend actions to increase realized value. & Solution Evaluation \\

PLO-08 & \textbf{Demonstrate} the underlying competencies of a professional business analyst --- analytical thinking, behavioral characteristics, business knowledge, communication, interaction, and tool use --- across observed performance tasks. & Underlying Competencies \\

PLO-09 & \textbf{Select and apply} appropriate BABOK techniques to specific BA tasks, justifying each choice in terms of context, stakeholders, and information needs. & Techniques \\

PLO-10 & \textbf{Adapt} business analysis work to the Agile, Business Intelligence, Information Technology, Business Architecture, and Business Process Management perspectives, demonstrating how the same BA task differs by perspective. & Perspectives \\

PLO-11 & \textbf{Facilitate} stakeholder collaboration and conflict resolution in workshops and reviews, producing recorded decisions and action items. & E\&C; UC \\

PLO-12 & \textbf{Communicate} business analysis information to technical and non-technical audiences in writing, in models, and in spoken presentations, calibrated to audience needs. & E\&C; UC \\

PLO-13 & \textbf{Practice} ethical, inclusive, and culturally aware business analysis, evidenced by stakeholder maps that account for marginalized voices, decision logs that disclose conflicts of interest, and reviewer-validated reflection on bias. & UC \\

PLO-14 & \textbf{Reflect} on professional practice using structured frameworks and produce a portfolio that evidences growth across all BABOK knowledge areas. & All \\

PLO-15 & \textbf{Perform} end-to-end business analysis on a realistic case in the capstone, integrating all knowledge areas, perspectives, and competencies, and defending recommendations to a stakeholder panel. & All \\

\end{longtable}

% =====================================================================
\FrontMatterChapter{Curriculum Architecture and Sequence}

\section*{7. Curriculum Map}
The curriculum is organized into eleven modules, M01--M11. Each module aligns to one or more BABOK knowledge areas, exercises a specific set of underlying competencies, introduces and reinforces a defined set of techniques, and is delivered through one or more BABOK perspectives.

\subsection*{7.1 Time Allocation by Activity Type}

\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,title=Time Allocation Across the Program]
\begin{tabular}{@{}R{8.0cm} r r@{}}
\toprule
\sffamily\bfseries Activity Type & \sffamily\bfseries Hours & \sffamily\bfseries \% of Program \\
\midrule
Direct instruction (lecture, demonstration, structured discussion)         &  56 & 20\% \\
Supervised lab/studio practice (artifact production, workshops, reviews)   & 112 & 40\% \\
Independent study (readings, micro-exercises, portfolio writing)           &  56 & 20\% \\
Capstone work (case analysis, deliverables, defense preparation)           &  56 & 20\% \\
\bottomrule
\end{tabular}
\end{tcolorbox}

Active and applied work (lab + capstone) accounts for 60\% of program hours, satisfying the active-engagement requirement of the Minimum Curriculum Quality Checklist.

\subsection*{7.2 Module Map (Landscape)}
The full module map is presented in landscape on the next page so that all eight columns remain legible.

\begin{landscape}
\thispagestyle{plain}
\setlength\LTleft{0pt}\setlength\LTright\fill
\renewcommand{\arraystretch}{1.15}
\footnotesize
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Curriculum Map --- Modules M01--M11\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{6pt}%
\begin{longtable}{@{}R{0.65cm} R{2.85cm} R{2.50cm} R{2.30cm} R{4.20cm} R{2.30cm} R{0.85cm} R{2.65cm}@{}}
\toprule
\sffamily\bfseries M\# &
\sffamily\bfseries Module &
\sffamily\bfseries BABOK Alignment &
\sffamily\bfseries Primary Competencies &
\sffamily\bfseries Major Techniques &
\sffamily\bfseries Perspectives &
\sffamily\bfseries Hrs &
\sffamily\bfseries Primary Assessment \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries M\# &
\sffamily\bfseries Module &
\sffamily\bfseries BABOK &
\sffamily\bfseries Competencies &
\sffamily\bfseries Techniques &
\sffamily\bfseries Perspectives &
\sffamily\bfseries Hrs &
\sffamily\bfseries Assessment \\
\midrule
\endhead
\bottomrule
\endfoot

M01 & Foundations of Business Analysis (BACCM and the Profession)
    & Key Concepts (\S2); intro to all six KAs; intro to UC (\S9)
    & Conceptual thinking; ethics; communication
    & Glossary; concept modelling; stakeholder list (intro)
    & All (introduced)
    & 16
    & BACCM concept map + reflective essay \\

M02 & Business Analysis Planning and Monitoring
    & BAPM (\S3)
    & Analytical thinking; methodology; tools
    & Plan BA approach; stakeholder analysis; RACI; metrics; risk analysis
    & Agile vs. predictive contrasts
    & 24
    & Business Analysis Plan artifact \\

M03 & Elicitation and Collaboration
    & E\&C (\S4)
    & Communication; interaction; behavioral
    & Interviews; workshops; observation; document analysis; brainstorming; focus groups; survey; collaborative games
    & Agile, IT
    & 24
    & Elicitation event package + recorded facilitation \\

M04 & Strategy Analysis
    & Strategy Analysis (\S6)
    & Analytical \& business thinking; business knowledge
    & SWOT; PESTLE; root-cause; capability analysis; business model canvas; scope modelling; decision analysis; risk
    & Business Architecture, BPM
    & 28
    & Future-state recommendation with options analysis \\

M05 & Requirements Analysis and Design Definition
    & RADD (\S7)
    & Conceptual thinking; modeling; tools
    & Process/data/decision/state modelling; user stories; use cases; functional decomposition; prototyping; data dictionary; NFR analysis; acceptance criteria
    & Agile, IT, BI
    & 36
    & Requirements \& design specification with V\&V evidence \\

M06 & Requirements Life Cycle Management
    & RLCM (\S5)
    & Analytical thinking; tools; communication
    & Traceability; prioritization (MoSCoW, Kano, Weighted); change management; backlog; configuration management; reviews
    & Agile, IT
    & 20
    & Traceability matrix + governance decision log \\

M07 & Solution Evaluation
    & Solution Evaluation (\S8)
    & Analytical \& systems thinking; business knowledge
    & KPIs; balanced scorecard; benchmarking; survey; observation; root-cause; risk; process analysis; decision modelling
    & BPM, BI, IT
    & 20
    & Solution evaluation report with recommendations \\

M08 & Underlying Competencies in Practice
    & Underlying Competencies (\S9)
    & All six competency categories
    & Workshop facilitation; reviews; collaborative games
    & All
    & 16
    & Performance task: facilitation + decision review (observed) \\

M09 & Perspectives Studio
    & Perspectives (\S11)
    & Cross-domain knowledge; tools; analytical thinking
    & Re-application of techniques from prior modules under each perspective
    & Agile, BI, IT, BA, BPM
    & 24
    & Five perspective-specific artifacts (one per perspective) \\

M10 & Integration \& Techniques Mastery Lab
    & Techniques (\S10), all KAs
    & Tools \& technology; analytical thinking; communication
    & Open lab using technique catalogue
    & All
    & 16
    & Technique-selection rationale + executed deliverable \\

M11 & Capstone
    & All
    & All
    & All
    & All required
    & 56
    & Capstone deliverables + stakeholder panel defense \\

\end{longtable}
\end{landscape}

\section*{8. Scope and Sequence Rationale}
The sequence is designed to mirror how a competent practitioner moves from understanding the BA's role, to organizing analysis work, to discovering and analyzing need, to specifying solutions, to managing requirements through change, to evaluating value, and finally to integrating it all under realistic conditions. The sequence intentionally departs from BABOK's reference numbering in two places, with explicit pedagogical justification.

\subsection*{8.1 Sequence Justification (M01 $\to$ M11)}
\begin{itemize}
  \item \textbf{M01 --- Foundations.} Learners need a shared conceptual vocabulary (\skill{BACCM}) and a clear picture of what business analysts do, before any knowledge-area work makes sense. Underlying Competencies are introduced here so they can be deliberately practiced in every subsequent module rather than tacked on at the end.
  \item \textbf{M02 --- Business Analysis Planning and Monitoring.} Once learners understand the BA role and the BACCM, the first practical question is ``how do you plan to do this work on a real initiative?'' BAPM precedes elicitation and analysis because every subsequent activity requires a planned approach, identified stakeholders, defined deliverables, and an information management strategy. Without these, downstream modules degrade into ungoverned exercises.
  \item \textbf{M03 --- Elicitation and Collaboration.} Elicitation comes next because learners now need to discover information to analyze. Placing E\&C immediately after planning is also pedagogically motivated: it is the most visible BA activity and provides early skill confidence. Most subsequent modules depend on elicitation as a feeder activity.
  \item \textbf{M04 --- Strategy Analysis.} Strategy Analysis is taught before Requirements Life Cycle Management --- a deliberate reorder relative to BABOK. The reason is pedagogical: RLCM is the discipline of \emph{managing} requirements information once you have requirements. Strategy Analysis is where requirements \emph{originate} (from need $\to$ future state $\to$ change strategy). Teaching SA before RLCM means learners produce real requirements they can then practice managing in M06.
  \item \textbf{M05 --- Requirements Analysis and Design Definition (RADD).} Once need is understood and a change strategy exists, learners model and specify requirements and designs. RADD is the largest module (36 hours) because it carries the heaviest cognitive load and the broadest technique surface.
  \item \textbf{M06 --- Requirements Life Cycle Management.} With elicited and modelled requirements in hand, learners now learn to trace, prioritize, govern change, approve, and maintain them through the lifecycle. RLCM is taught after RADD because it is meaningful only against actual requirements artifacts.
  \item \textbf{M07 --- Solution Evaluation.} The lifecycle closes with assessing whether the implemented solution delivers expected value. SE follows RLCM because it consumes the artifacts that RLCM has been managing and asks ``did this deliver?''
  \item \textbf{M08 --- Underlying Competencies in Practice.} UC is consolidated here as a deliberate, observed performance module. The competencies were practiced throughout the prior modules; M08 surfaces them, names them, and assesses them under direct observation.
  \item \textbf{M09 --- Perspectives Studio.} Once the learner has a complete BA toolkit, perspectives are taught as variations on it. The studio rotates the same kind of analysis work through five perspective lenses, so the learner sees first-hand how the same BACCM yields different artifacts and emphases.
  \item \textbf{M10 --- Integration \& Techniques Mastery Lab.} This lab gives learners a controlled environment to select techniques freely, justify the selection, and execute. It is the pre-capstone consolidation and identifies remaining gaps before the high-stakes capstone.
  \item \textbf{M11 --- Capstone.} The capstone is the integrated performance task. It demands every BA skill, every KA, multiple perspectives, defended in front of a stakeholder panel.
\end{itemize}

This ordering ensures \term{cumulative sequencing}: each module produces or strengthens an asset that the next module consumes or refines. Dependencies are visible in the alignment matrix (\S15).

\begin{curriculumnote}[Notation Convention]
Each module objective is written as: \textbf{action verb + observable performance + criterion}. Each is mapped to one or more PLOs. Module objective IDs use the form \tool{Mxx-Oy} (e.g., \tool{M01-O3} = Module 01, Objective 3).
\end{curriculumnote}

% =====================================================================
\FrontMatterChapter{Module Detail: Outcomes, Activities, and Assessment}

The remainder of this section presents each module with measurable learning outcomes, the BABOK alignment, key topics, required vocabulary, practical activities, techniques practiced, artifact deliverables, assessment methods, reflection prompts, accessibility considerations, and instructor notes.

% --------------------- MODULE 1 (M01) ---------------------
\chapter{Foundations of Business Analysis (BACCM and the Profession)}

\subsection*{Purpose}
To install a shared professional vocabulary, establish the \skill{BACCM} as the program's organizing concept model, and introduce the Underlying Competencies that will be deliberately practiced throughout.

\subsection*{BABOK Alignment}
Key Concepts (\S2) primary; Underlying Competencies (\S9) introductory; brief overview of all six knowledge areas (\S3--\S8).

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M01-O1.} \textbf{Define} each BACCM core concept (change, need, solution, stakeholder, value, context) and \textbf{explain} its interrelationships in writing of at least 200 words. ($\to$ PLO-01)
  \item \textbf{M01-O2.} \textbf{Map} the six BACCM concepts to a provided initiative case, citing specific case evidence for each concept. ($\to$ PLO-01)
  \item \textbf{M01-O3.} \textbf{Identify} the six BABOK knowledge areas and \textbf{describe} each in one sentence. ($\to$ PLO-01, PLO-09)
  \item \textbf{M01-O4.} \textbf{Describe} the six categories of underlying competencies and \textbf{self-assess} current proficiency in each on a 4-point rubric. ($\to$ PLO-08, PLO-14)
  \item \textbf{M01-O5.} \textbf{Articulate} at least two ethical considerations relevant to BA practice using examples. ($\to$ PLO-13)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
What business analysis is and is not; the BACCM and its application; the six knowledge areas at-a-glance; the role and accountabilities of the BA; ethics, conflicts of interest, and stakeholder representation; the BA's relationship to project management, product management, systems engineering, and product ownership; introduction to the underlying competencies.

\subsection*{Required Vocabulary}
\skill{Business analysis}; \skill{BACCM}; change; need; solution; stakeholder; value; context; knowledge area; technique; perspective; underlying competency; requirement; design; elicitation; specification; traceability; baseline.

\begin{labbox}[Module 1 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (small group, 60 min): Apply BACCM to a one-page case of a hospital scheduling change.
  \item \textbf{Activity A2} (individual, 30 min): Self-assessment against the underlying competency rubric.
  \item \textbf{Activity A3} (paired, 45 min): ``What I do / What a BA does'' interview with a peer; reconcile in writing.
  \item \textbf{Activity A4} (full group, 45 min): Ethics scenario discussion using a structured decision frame.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Glossary; concept modelling (introductory); stakeholder list (introductory).

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item BACCM concept map for the case (1 page).
  \item Personal underlying-competency self-assessment.
  \item Reflective entry in the program portfolio (300--500 words).
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} In-class concept-map peer review; ungraded BACCM quiz (10 items).\\[3pt]
\textbf{Summative.} BACCM concept map for the case + reflective essay, scored against rubric \tool{R-M01} (BACCM completeness; evidence quality; reflection depth; written communication).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Where in your current work do you already perform business analysis without naming it as such?
  \item Which of the underlying competencies will be hardest for you, and why?
  \item Identify one ethical edge case in your current work; what would the BACCM tell you to do?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Captioned recordings of all live sessions; large-text and high-contrast templates; readings provided in HTML and PDF; no time-limited live quizzes; alternative concept-map formats (text outline, visual, narrative); anonymous discussion option; a culturally varied case bank; explicit norms for turn-taking in group discussion.
\end{curriculumnote}

\begin{curriculumteachingnotes}
This module sets the cultural and conceptual tone. Spend disproportionate time on the BACCM and on what counts as defensible evidence. Resist the temptation to lecture through the KAs --- students will see them in depth from M02 forward. Prepare to challenge the most experienced participants by asking them to defend their assertions against BABOK terminology rather than personal vocabulary. \textbf{Estimated prep:} 4 hours first time, 1.5 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 2 (M02) ---------------------
\chapter{Business Analysis Planning and Monitoring}

\subsection*{Purpose}
To enable learners to plan their business analysis work appropriately for a given initiative --- identifying stakeholders, choosing an approach, governing information, defining performance measures, and adapting as conditions change.

\subsection*{BABOK Alignment}
BAPM (\S3) primary; selected techniques from \S10.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M02-O1.} \textbf{Construct} a Business Analysis Plan (governance, approach, information management, deliverables, stakeholder engagement, performance) defensible against the constraints of a given case. ($\to$ PLO-02)
  \item \textbf{M02-O2.} \textbf{Identify and analyze} stakeholders for a given initiative using a stakeholder list, RACI, and onion-diagram view. ($\to$ PLO-02, PLO-04)
  \item \textbf{M02-O3.} \textbf{Compare} predictive vs. adaptive vs. hybrid BA approaches and \textbf{justify} an approach choice for a given case. ($\to$ PLO-02, PLO-10)
  \item \textbf{M02-O4.} \textbf{Define} at least five business analysis performance metrics for a given initiative, each with a baseline, target, and source. ($\to$ PLO-02)
  \item \textbf{M02-O5.} \textbf{Analyze} at least three risks to the analysis work itself and propose mitigations. ($\to$ PLO-02)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Plan business analysis approach; plan stakeholder engagement; plan business analysis governance; plan business analysis information management; identify business analysis performance improvements; predictive/adaptive/hybrid; types of stakeholders (BABOK \S2.4); decision-making authority and approval flows; configuration baselines for analysis information.

\subsection*{Required Vocabulary}
Plan business analysis approach; \term{stakeholder}; stakeholder analysis; \tool{RACI}; engagement; governance; baseline; performance metric; risk; assumption; constraint; cadence; ceremony.

\begin{labbox}[Module 2 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (small group, 90 min): Build a stakeholder list and onion diagram for a public-sector grant management modernization case; defend choices.
  \item \textbf{Activity A2} (small group, 120 min): Workshop a Business Analysis Plan for the case, including approach selection.
  \item \textbf{Activity A3} (individual, 45 min): Define performance metrics and risks for analysis work.
  \item \textbf{Activity A4} (paired, 45 min): Peer review of plans using rubric \tool{R-M02}.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Stakeholder list, map, or personas; RACI; brainstorming; risk analysis and management; metrics and KPIs; document analysis; interviews (planning conversations); business cases (introductory); decision analysis (approach selection).

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Business Analysis Plan ($\approx$8--12 pages or equivalent structured artifact).
  \item Stakeholder analysis (list, RACI, onion or matrix view).
  \item Initial set of analysis performance metrics.
  \item Risk register for analysis work.
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} In-class plan critique using a structured rubric; checklist self-review.\\[3pt]
\textbf{Summative.} Business Analysis Plan artifact, scored against \tool{R-M02} (completeness; defensibility; stakeholder coverage; alignment to constraints; usability).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Whose voice is missing from your stakeholder map, and why?
  \item What would you change about your plan if the initiative shifted from predictive to agile delivery?
  \item Which performance metrics would actually change BA behavior, and which are vanity metrics?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Plan templates available in document, spreadsheet, and structured-text formats. Stakeholder identification deliberately includes prompts for marginalized and indirect stakeholders. Workshop activities use structured turn-taking. Deliverable length is bounded by completeness, not page count, to avoid penalizing concise writers.
\end{curriculumnote}

\begin{curriculumteachingnotes}
The most common failure in this module is treating planning as form-filling. Push learners to defend trade-offs (``you said this; what does that cost you?''). Use the stakeholder-onion diagram to surface assumptions about influence and proximity. For learners with PMP or PRINCE2 backgrounds, explicitly distinguish project planning from business analysis planning. \textbf{Estimated prep:} 6 hours first time, 2 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 3 (M03) ---------------------
\chapter{Elicitation and Collaboration}

\subsection*{Purpose}
To make learners competent in planning, conducting, confirming, and communicating elicitation, with deliberate practice in facilitation and collaboration.

\subsection*{BABOK Alignment}
E\&C (\S4) primary; techniques from \S10; underlying competencies (\S9) --- communication, interaction, behavioral characteristics.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M03-O1.} \textbf{Plan} an elicitation event with selected techniques justified by stakeholder availability, information needs, and risk. ($\to$ PLO-03, PLO-09)
  \item \textbf{M03-O2.} \textbf{Conduct} at least one observed elicitation activity (interview or workshop), demonstrating active listening, neutrality, and structured questioning. ($\to$ PLO-03, PLO-08, PLO-11)
  \item \textbf{M03-O3.} \textbf{Confirm} elicitation results with stakeholders and produce a record suitable for downstream analysis. ($\to$ PLO-03, PLO-04)
  \item \textbf{M03-O4.} \textbf{Communicate} elicitation results to a non-technical audience in plain language. ($\to$ PLO-12)
  \item \textbf{M03-O5.} \textbf{Manage} stakeholder collaboration over time, including handling conflict and disengagement. ($\to$ PLO-11, PLO-13)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Prepare for elicitation; conduct elicitation; confirm elicitation results; communicate business analysis information; manage stakeholder collaboration; active listening; questioning techniques; facilitation patterns; handling silence, dominance, and conflict; remote vs. in-person elicitation; cultural and linguistic considerations.

\subsection*{Required Vocabulary}
\term{Elicitation}; interview; workshop; observation; document analysis; brainstorming; focus group; survey; questionnaire; collaborative games; facilitation; active listening; open vs. closed question; scribing; minute-taking; stakeholder collaboration.

\begin{labbox}[Module 3 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (paired, 60 min): Structured interview practice with rotating roles (interviewer, stakeholder, observer); observer scores the interviewer with a behavioral rubric.
  \item \textbf{Activity A2} (small group, 180 min): Facilitate a 60-minute requirements workshop on the M02 case; recorded and reviewed.
  \item \textbf{Activity A3} (individual, 60 min): Document analysis on a real (sanitized) policy artifact.
  \item \textbf{Activity A4} (small group, 60 min): Survey design exercise: write 10 questions, peer-review for bias, leading-question structure, and accessibility.
  \item \textbf{Activity A5} (full group, 60 min): Conflict simulation --- facilitate a workshop in which two stakeholders disagree.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Interviews; workshops; observation; document analysis; brainstorming; focus groups; survey/questionnaire; collaborative games; mind mapping; concept modelling (light use); item tracking.

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Elicitation plan for an event.
  \item Confirmed elicitation results document.
  \item Recorded facilitation (audio or video, with consent).
  \item Communication brief for a non-technical audience ($\leq$2 pages).
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Peer behavioral rubric on facilitation practice; self-review of recordings.\\[3pt]
\textbf{Summative.} Elicitation event package + observed facilitation, scored against \tool{R-M03} (planning quality; technique selection; observed facilitation behaviors; quality of confirmed results; stakeholder communication).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item When did you stop listening and start preparing your next question?
  \item Whose perspective did your workshop privilege? Whose did it suppress?
  \item How would your facilitation change if conducted entirely asynchronously?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Multiple modalities for the observed facilitation (in-person, live virtual, asynchronous moderated discussion). Elicitation roleplays include scenarios with neuro-diverse, non-native-language, and remote stakeholders. Templates encourage written-first contribution to balance verbal-dominant cultures. Recording is opt-in with written-only alternatives offered. Captioning required for all recorded reviews.
\end{curriculumnote}

\begin{curriculumteachingnotes}
Facilitation is a craft, not a checklist. Budget time to debrief observed sessions in detail. Encourage learners to fail safely in the conflict simulation. Anchor every technique choice in stakeholder availability and information need; resist ``I'll do a workshop because I always do a workshop.'' \textbf{Estimated prep:} 8 hours first time, 3 hours per repeat (high because of recording reviews).
\end{curriculumteachingnotes}

% --------------------- MODULE 4 (M04) ---------------------
\chapter{Strategy Analysis}

\subsection*{Purpose}
To enable learners to analyze the current state, define a desired future state, recommend a change strategy, and assess risks --- producing the strategic context that downstream requirements and design will refine.

\subsection*{BABOK Alignment}
Strategy Analysis (\S6) primary; selected techniques from \S10; perspectives --- Business Architecture, BPM.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M04-O1.} \textbf{Analyze} a current state, capturing organizational context, capabilities, business processes, performance, infrastructure, and external factors. ($\to$ PLO-05)
  \item \textbf{M04-O2.} \textbf{Define} a future state with measurable business objectives, scope, value, and constraints. ($\to$ PLO-05)
  \item \textbf{M04-O3.} \textbf{Assess} risks to the future state and to the change itself. ($\to$ PLO-05)
  \item \textbf{M04-O4.} \textbf{Define} a change strategy with options, trade-offs, and a justified recommendation. ($\to$ PLO-05)
  \item \textbf{M04-O5.} \textbf{Trace} future-state objectives to expected value and to candidate solution approaches. ($\to$ PLO-05, PLO-04)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Analyze current state; define future state; assess risks; define change strategy; capability and process modelling; SWOT/PESTLE; root-cause analysis; business model canvas; investment analysis; risk analysis and management; financial analysis (introductory); decision modelling and decision analysis.

\subsection*{Required Vocabulary}
Current state; future state; \term{business need}; goal; objective (SMART); business case; capability; process; value; opportunity cost; transition state; change strategy; root cause; assumption; constraint; risk.

\begin{labbox}[Module 4 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (small group, 180 min): Current-state analysis for a regional credit union seeking to enter digital lending; produce capability map, process inventory, and SWOT.
  \item \textbf{Activity A2} (small group, 120 min): Future-state definition workshop with measurable objectives.
  \item \textbf{Activity A3} (paired, 90 min): Root-cause analysis on a recurring customer-service problem using fishbone and 5-Whys.
  \item \textbf{Activity A4} (small group, 120 min): Generate three change-strategy options; perform decision analysis; recommend one.
  \item \textbf{Activity A5} (individual, 45 min): Risk analysis for the recommended strategy.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
SWOT; PESTLE; business capability analysis; business model canvas; functional decomposition; root-cause analysis; risk analysis and management; decision analysis; decision modelling; financial analysis (introductory); scope modelling.

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Current-state assessment (capabilities, processes, performance, environment).
  \item Future-state definition with SMART objectives.
  \item Risk register.
  \item Change strategy with options analysis and recommendation.
  \item Updated business case (light).
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Peer review of options analysis; ungraded objective-quality rubric self-check.\\[3pt]
\textbf{Summative.} Future-state recommendation pack, scored against \tool{R-M04} (rigor of current-state evidence; quality of objectives; quality of options; defensibility of recommendation; risk treatment).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item What current-state evidence did you accept on assertion rather than verification?
  \item Are your objectives measurable, or do they merely sound measurable?
  \item Which option did you reject most quickly, and why? Was that bias or analysis?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Templates available in multiple formats; case studies span sectors and geographies (US, EU, sub-Saharan African retail banking, public health in South Asia); options analysis explicitly invites equity considerations among options.
\end{curriculumnote}

\begin{curriculumteachingnotes}
This module suffers most when learners skip current-state evidence and jump to ``what we should do.'' Police this hard. Use the decision-analysis activity as a vehicle for developing structured thinking. Be ready to challenge SMART objectives --- many learners write goals dressed as objectives. \textbf{Estimated prep:} 7 hours first time, 2.5 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 5 (M05) ---------------------
\chapter{Requirements Analysis and Design Definition (RADD)}

\subsection*{Purpose}
To enable learners to model, specify, verify, and validate requirements; define design options; and analyze potential value and recommend a solution.

\subsection*{BABOK Alignment}
RADD (\S7) primary; many techniques from \S10; perspectives --- Agile, IT, BI.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M05-O1.} \textbf{Model} requirements using process models, data models, behavior models, and rule models appropriate to the case. ($\to$ PLO-06, PLO-09)
  \item \textbf{M05-O2.} \textbf{Specify} requirements in textual, structured, and visual forms suitable for the audience. ($\to$ PLO-06, PLO-12)
  \item \textbf{M05-O3.} \textbf{Verify} requirements for quality and \textbf{validate} them against business objectives. ($\to$ PLO-06)
  \item \textbf{M05-O4.} \textbf{Define} design options and \textbf{analyze} potential value, including non-functional considerations. ($\to$ PLO-06)
  \item \textbf{M05-O5.} \textbf{Recommend} a solution and \textbf{trace} it to need. ($\to$ PLO-06, PLO-04)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Specify and model requirements; verify requirements; validate requirements; define requirements architecture; define design options; analyze potential value and recommend solution; functional vs. non-functional; user stories; use cases; process modelling (BPMN, UML AD); data modelling (ERD, class); state modelling; decision modelling and DMN; user interface modelling and prototyping; acceptance and evaluation criteria; data dictionary; requirements verification techniques (review, walkthrough, inspection).

\subsection*{Required Vocabulary}
Requirement (business, stakeholder, solution --- functional and non-functional, transition); design; model; specification; user story; use case; scenario; activity diagram; sequence diagram; class diagram; entity-relationship diagram; state machine; decision table; \tool{DMN}; \tool{BPMN}; non-functional requirement; constraint; acceptance criterion; verification; validation; requirements architecture; baseline.

\begin{labbox}[Module 5 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (individual, 120 min): Specify a stakeholder requirements set in textual form for the M04 case.
  \item \textbf{Activity A2} (small group, 180 min): Model a key business process (BPMN) and the relevant data (ERD); reconcile with the textual requirements.
  \item \textbf{Activity A3} (small group, 120 min): Write user stories and acceptance criteria for an Agile delivery slice.
  \item \textbf{Activity A4} (small group, 180 min): Construct a decision model (DMN) for an automated lending decision.
  \item \textbf{Activity A5} (small group, 120 min): Conduct a structured requirements walkthrough using a verification checklist.
  \item \textbf{Activity A6} (small group, 120 min): Define two design options and analyze potential value; recommend one.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Process modelling; data modelling; data flow diagrams; user stories; use cases and scenarios; functional decomposition; state modelling; decision modelling; non-functional requirements analysis; acceptance and evaluation criteria; data dictionary; prototyping; reviews; interface analysis; roles and permissions matrix; business rules analysis; concept modelling.

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Stakeholder and solution requirements specification (textual).
  \item Process model(s) and data model(s).
  \item User stories with acceptance criteria.
  \item Decision model.
  \item NFR list and rationale.
  \item Design options analysis with recommendation.
  \item Verification and validation evidence.
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Model peer-review using a structured walkthrough; in-class verification checklist.\\[3pt]
\textbf{Summative.} Requirements \& design specification pack, scored against \tool{R-M05} (model quality; specification clarity; coverage of NFRs; verification evidence; validation against objectives; traceability).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Which model did the most analytical work for you, and why?
  \item Where did your specification leak ambiguity?
  \item Which non-functional requirement is most likely to be ignored, and what are the consequences?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Modeling tools must include a no-cost option and a text-equivalent representation (e.g., \tool{Mermaid}, \tool{PlantUML}, structured text). All visual artifacts have textual descriptions. Walkthroughs use a structured speaking order to balance contributions. Keyboard-accessible diagramming required.
\end{curriculumnote}

\begin{curriculumteachingnotes}
This is the heaviest module --- protect time. The most common failure is producing pretty diagrams that contradict the text. Insist on reconciliation. The decision-modeling exercise is high-value for analytical thinking; do not skip it. For BAs from heavy-text backgrounds, push them into modeling; for BAs from architecture backgrounds, push them into specification quality. \textbf{Estimated prep:} 10 hours first time, 4 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 6 (M06) ---------------------
\chapter{Requirements Life Cycle Management}

\subsection*{Purpose}
To enable learners to manage requirements information across its life cycle: tracing, maintaining, prioritizing, assessing changes, and approving.

\subsection*{BABOK Alignment}
RLCM (\S5) primary; techniques from \S10.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M06-O1.} \textbf{Construct} a traceability matrix linking business objectives, requirements, designs, tests, and value measures. ($\to$ PLO-04)
  \item \textbf{M06-O2.} \textbf{Maintain} requirements through versioning, baselining, and configuration management. ($\to$ PLO-04)
  \item \textbf{M06-O3.} \textbf{Prioritize} requirements using two or more methods justified by case context. ($\to$ PLO-04, PLO-09)
  \item \textbf{M06-O4.} \textbf{Assess} a proposed change to requirements and \textbf{recommend} a disposition. ($\to$ PLO-04)
  \item \textbf{M06-O5.} \textbf{Approve} requirements through a structured review with documented stakeholder authorization. ($\to$ PLO-04, PLO-11)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Trace requirements; maintain requirements; prioritize requirements; assess requirements changes; approve requirements; backlog management; configuration management; reviews; item tracking.

\subsection*{Required Vocabulary}
\term{Traceability}; baseline; configuration item; version; backlog; prioritization (\tool{MoSCoW}, \tool{Kano}, Weighted Shortest Job First); change assessment; impact analysis; approval; disposition.

\begin{labbox}[Module 6 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (individual, 90 min): Build a traceability matrix for the M04/M05 outputs.
  \item \textbf{Activity A2} (small group, 90 min): Prioritize a 30-item backlog using MoSCoW, Kano, and a weighted method; reconcile.
  \item \textbf{Activity A3} (small group, 90 min): Assess a proposed scope change; produce a change-request package.
  \item \textbf{Activity A4} (paired, 60 min): Run an approval review using a structured approval rubric.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Backlog management; prioritization (MoSCoW, Kano, Weighted); item tracking; reviews; configuration management (process); estimation; risk analysis; decision analysis; metrics \& KPIs; document analysis.

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Traceability matrix.
  \item Prioritized backlog with rationale.
  \item Change-request package.
  \item Approval decision log.
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Peer review of traceability completeness.\\[3pt]
\textbf{Summative.} Traceability matrix + governance decision log, scored against \tool{R-M06} (completeness; consistency; defensibility of prioritization; quality of change assessment; governance fit).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Which prioritization method most challenged your assumptions about value?
  \item What did your traceability matrix reveal that the original requirements did not?
  \item Where would your governance approach fail under scale?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Tooling alternatives offered (spreadsheet, lightweight database, structured text). Decision-making activities use written-first protocols. Time-bounded but not time-pressured exercises.
\end{curriculumnote}

\begin{curriculumteachingnotes}
RLCM is dry on its surface but reveals the structural integrity of prior work. Tie every activity back to M04/M05 outputs so learners feel the cumulative payoff. Use prioritization debates as a vehicle for negotiation skill. \textbf{Estimated prep:} 5 hours first time, 1.5 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 7 (M07) ---------------------
\chapter{Solution Evaluation}

\subsection*{Purpose}
To enable learners to assess solution performance, identify limitations and constraints, and recommend actions to increase realized value.

\subsection*{BABOK Alignment}
Solution Evaluation (\S8) primary; perspectives --- BPM, BI, IT.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M07-O1.} \textbf{Measure} solution performance against defined value measures. ($\to$ PLO-07)
  \item \textbf{M07-O2.} \textbf{Analyze} performance measures and identify limitations of the solution and of the enterprise. ($\to$ PLO-07)
  \item \textbf{M07-O3.} \textbf{Assess} enterprise readiness, organizational change, and adoption. ($\to$ PLO-07)
  \item \textbf{M07-O4.} \textbf{Recommend} actions to increase solution value. ($\to$ PLO-07)
  \item \textbf{M07-O5.} \textbf{Trace} evaluation findings to the original business objectives. ($\to$ PLO-07, PLO-04)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Measure solution performance; analyze performance measures; assess solution limitations; assess enterprise limitations; recommend actions to increase solution value; KPIs; balanced scorecard; benchmarking; root-cause analysis; observation; interviews; survey; process analysis; risk analysis; decision modelling; metrics that drive (and distort) behavior.

\subsection*{Required Vocabulary}
Value; benefit; outcome; \tool{KPI}; balanced scorecard; benchmarking; adoption; enterprise readiness; constraint; limitation; recommended action; sunset; pivot; persevere.

\begin{labbox}[Module 7 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (small group, 120 min): Define performance measures for a solution from the M04/M05 case.
  \item \textbf{Activity A2} (small group, 120 min): Analyze given (synthetic) performance data; identify limitations.
  \item \textbf{Activity A3} (small group, 120 min): Assess organizational readiness using a structured tool.
  \item \textbf{Activity A4} (small group, 120 min): Recommend actions; defend in a stakeholder review.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Metrics and KPIs; balanced scorecard; benchmarking and market analysis; survey; observation; interviews; root-cause analysis; risk analysis; process analysis; decision modelling; financial analysis.

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Solution performance measurement plan.
  \item Performance analysis report.
  \item Enterprise readiness assessment.
  \item Recommendations memo with justification.
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Peer challenge of measures (do they actually evidence value?); checklist self-review.\\[3pt]
\textbf{Summative.} Solution evaluation report, scored against \tool{R-M07} (measure quality; analysis rigor; readiness assessment depth; recommendation defensibility; traceability to objectives).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Which of your measures could be gamed? How?
  \item What organizational signals would you treat as evidence of low adoption?
  \item When would you recommend pivoting away from a solution that is performing acceptably?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Synthetic datasets supplied in spreadsheet, CSV, and screen-reader-friendly tabular formats. Multiple cases drawn from contrasting sectors. Stakeholder review activity offers asynchronous alternative.
\end{curriculumnote}

\begin{curriculumteachingnotes}
Resist treating this module as project closeout. SE is ongoing. Push learners to identify enterprise limitations (not just solution limitations) --- this is where many BAs underperform. \textbf{Estimated prep:} 5 hours first time, 1.5 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 8 (M08) ---------------------
\chapter{Underlying Competencies in Practice}

\subsection*{Purpose}
To consolidate, observe, and assess the underlying competencies --- analytical thinking and problem solving, behavioral characteristics, business knowledge, communication skills, interaction skills, tools and technology --- that have been practiced throughout the program.

\subsection*{BABOK Alignment}
Underlying Competencies (\S9) primary.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M08-O1.} \textbf{Demonstrate} analytical thinking by structuring an unfamiliar problem in writing and orally within a 30-minute exercise. ($\to$ PLO-08)
  \item \textbf{M08-O2.} \textbf{Demonstrate} facilitation and interaction skills in an observed live workshop. ($\to$ PLO-08, PLO-11)
  \item \textbf{M08-O3.} \textbf{Communicate} the same business analysis information in three forms: an executive memo, a stakeholder briefing, and a developer-oriented specification. ($\to$ PLO-12)
  \item \textbf{M08-O4.} \textbf{Apply} ethical reasoning to a stakeholder conflict scenario and produce a written decision rationale. ($\to$ PLO-13)
  \item \textbf{M08-O5.} \textbf{Self-assess} growth across the six competency categories using documented evidence. ($\to$ PLO-14)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Each of the six BABOK competency categories with worked examples; deliberate-practice patterns; communication calibration to audience; decision frames for ethical conflict; evidence-based self-assessment.

\subsection*{Required Vocabulary}
\skill{Analytical thinking}; \skill{conceptual thinking}; \skill{systems thinking}; learning; problem solving; ethics; personal accountability; trustworthiness; adaptability; verbal/non-verbal/written communication; facilitation; leadership and influencing; teamwork; negotiation and conflict resolution; teaching; tools and technology.

\begin{labbox}[Module 8 Activities]
\begin{itemize}
  \item \textbf{Activity A1} (individual, 60 min): ``Structured problem-framing'' exercise on a novel scenario.
  \item \textbf{Activity A2} (small group, 120 min): Observed facilitation of a 45-minute decision workshop.
  \item \textbf{Activity A3} (individual, 90 min): Three-form communication exercise on the same content.
  \item \textbf{Activity A4} (small group, 90 min): Ethics roleplay with rotating roles.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Workshops; reviews; collaborative games; decision analysis; risk analysis; document analysis.

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Three communication artifacts on the same content.
  \item Recorded observed facilitation with debrief.
  \item Ethics decision rationale.
  \item Updated personal competency self-assessment with documented evidence.
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Peer feedback on facilitation behaviors using behavioral rubric.\\[3pt]
\textbf{Summative.} Performance task --- observed facilitation + decision review, scored against \tool{R-M08} (observed behaviors against rubric; communication calibration; ethical reasoning; evidence quality).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Which competency has grown most? Which has stalled?
  \item Whose feedback have you most resisted, and what does that tell you?
  \item What competency will you focus on after the program ends, and how?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Multiple modalities for observed facilitation. Communication exercise allows assistive tools. Roleplays are scripted to avoid harmful stereotypes; participants can opt out of any role with no penalty.
\end{curriculumnote}

\begin{curriculumteachingnotes}
This is the hardest module to instruct because the content is the learner. Make the rubric explicit and visible. Avoid praise inflation. Keep the focus on observable behavior, not on personality. \textbf{Estimated prep:} 6 hours first time, 2 hours per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 9 (M09) ---------------------
\chapter{Perspectives Studio}

\subsection*{Purpose}
To develop the learner's ability to apply business analysis across the five BABOK perspectives --- Agile, Business Intelligence, Information Technology, Business Architecture, and Business Process Management --- by performing the same kind of analysis work through each lens and contrasting the outputs.

\subsection*{BABOK Alignment}
Perspectives (\S11) primary; techniques (\S10) re-applied; all KAs touched.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M09-O1.} \textbf{Adapt} the same business analysis task to each of the five perspectives, producing an artifact appropriate to each. ($\to$ PLO-10)
  \item \textbf{M09-O2.} \textbf{Compare} how stakeholders, value, and deliverables differ across the perspectives for the same need. ($\to$ PLO-10)
  \item \textbf{M09-O3.} \textbf{Justify} technique selection for each perspective. ($\to$ PLO-09, PLO-10)
  \item \textbf{M09-O4.} \textbf{Reflect} on which perspective most closely matches the learner's current context and which is most growth-relevant. ($\to$ PLO-14)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
\begin{itemize}
  \item \textbf{Agile perspective:} Agile philosophies; product backlog; user stories at multiple horizons; Agile ceremonies; the BA in Agile; just-in-time analysis; cadence and flow; product-ownership relationship.
  \item \textbf{Business Intelligence perspective:} analytical needs; data sources, quality, and lineage; report and dashboard requirements; semantic layer; metric definitions; data governance; analytical use cases.
  \item \textbf{Information Technology perspective:} IT context; integration analysis; service-level requirements; security and identity requirements; data and interface specifications; transition requirements; legacy considerations.
  \item \textbf{Business Architecture perspective:} enterprise capability mapping; value streams; information concepts; business motivation; architecture roadmaps; alignment to enterprise frameworks.
  \item \textbf{Business Process Management perspective:} process discovery and modelling; process performance; continuous improvement; standardization vs. innovation; process governance.
\end{itemize}

\subsection*{Required Vocabulary}
Per perspective: backlog, sprint, product owner; data warehouse, data mart, ETL/ELT, lineage, semantic layer; service-level requirement, integration pattern, identity, transition; capability, value stream, information concept; process owner, process performance, RPA boundary.

\begin{labbox}[Module 9 Studio Rotations]
A single case (e.g., a multi-channel retailer modernizing membership) is analyzed across five 4-hour rotations, one per perspective. Each rotation produces a perspective-specific artifact:
\begin{itemize}
  \item \textbf{Agile:} refined backlog with epics, stories, acceptance criteria, and a release plan.
  \item \textbf{BI:} analytical use case with metric definitions, source-to-target mapping summary, and a dashboard wireframe.
  \item \textbf{IT:} integration map, NFR set, transition requirements list.
  \item \textbf{Business Architecture:} capability and value-stream maps with current/future heat maps.
  \item \textbf{BPM:} current process model with performance baseline and target future process.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Backlog management; user stories; data modelling; data flow; concept modelling; functional decomposition; interface analysis; non-functional requirements analysis; process modelling; process analysis; metrics and KPIs; business capability analysis.

\subsection*{Artifacts and Deliverables}
Five perspective-specific artifacts plus a comparative reflection ($\approx$1{,}500 words).

\begin{curriculumassessment}
\textbf{Formative.} Rotation-end peer critique using perspective-specific rubrics.\\[3pt]
\textbf{Summative.} Portfolio of five perspective artifacts + comparative reflection, scored against \tool{R-M09} (perspective fidelity; technique justification; comparative depth; reflection quality).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Which perspective most surprised you in what it surfaced?
  \item Where did the same stakeholder appear under different names or roles in different perspectives?
  \item Which perspective is most underdeveloped in your current organization, and what would change if it were strengthened?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
Templates and case materials available in plain-text alternatives. Studio rotations allow individual or paired work. Asynchronous alternative provided for any rotation if cohort scheduling is constrained.
\end{curriculumnote}

\begin{curriculumteachingnotes}
The most common failure is treating perspectives as superficial labels. Demand fidelity: an Agile artifact must look and behave like an Agile artifact, not a Gantt-chart wearing a hoodie. Bring guest practitioners where possible. \textbf{Estimated prep:} 8 hours first time, 3 hours per repeat (high because of five rubrics).
\end{curriculumteachingnotes}

% --------------------- MODULE 10 (M10) ---------------------
\chapter{Integration and Techniques Mastery Lab}

\subsection*{Purpose}
To consolidate technique mastery by giving learners an open lab in which they select techniques freely, justify selections, and execute --- under instructor coaching --- to identify and close residual gaps before the capstone.

\subsection*{BABOK Alignment}
Techniques (\S10) primary; all KAs.

\begin{outcomes}
On completion, the learner will:
\begin{enumerate}
  \item \textbf{M10-O1.} \textbf{Select} the most appropriate techniques for a given BA task and \textbf{justify} the selection in writing. ($\to$ PLO-09)
  \item \textbf{M10-O2.} \textbf{Execute} selected techniques to a publishable standard. ($\to$ PLO-06, PLO-09)
  \item \textbf{M10-O3.} \textbf{Identify} the learner's own remaining gaps and produce a personal pre-capstone study plan. ($\to$ PLO-14)
\end{enumerate}
\end{outcomes}

\subsection*{Key Topics}
Technique catalogue navigation; selection heuristics (stakeholder availability, information need, time budget, regulatory context); cross-technique reconciliation; self-diagnosis.

\subsection*{Required Vocabulary}
Technique; technique selection; reconciliation; stress-testing; gap analysis; learning plan.

\begin{labbox}[Module 10 Lab Activities]
\begin{itemize}
  \item \textbf{Activity A1} (individual, 60 min): Diagnostic --- given a BA task, select techniques and justify in writing.
  \item \textbf{Activity A2} (individual or paired, 360 min over the lab): Execute selected techniques on a provided case.
  \item \textbf{Activity A3} (individual, 90 min): Personal gap analysis and pre-capstone study plan.
\end{itemize}
\end{labbox}

\subsection*{Techniques Practiced}
Free selection from BABOK \S10 (50 techniques in catalogue; see \S10 of this curriculum for the integration plan).

\subsection*{Artifacts and Deliverables}
\begin{itemize}
  \item Technique-selection rationale.
  \item Executed deliverable(s).
  \item Personal pre-capstone study plan.
\end{itemize}

\begin{curriculumassessment}
\textbf{Formative.} Instructor coaching during execution.\\[3pt]
\textbf{Summative.} Technique-selection rationale + executed deliverable, scored against \tool{R-M10} (selection defensibility; execution quality; reconciliation across techniques; reflective gap analysis).
\end{curriculumassessment}

\begin{curriculumreflection}
\begin{itemize}
  \item Where do you reach for a familiar technique when a less familiar one would serve better?
  \item Which techniques have you avoided, and why?
  \item What is the smallest set of techniques that would cover 80\% of your work?
\end{itemize}
\end{curriculumreflection}

\begin{curriculumnote}[Accessibility \& Inclusion]
All technique resources available in multiple formats. Lab is structured as a series of short cycles to accommodate different attention patterns. Coaching available individually or in small groups.
\end{curriculumnote}

\begin{curriculumteachingnotes}
This lab succeeds or fails on the quality of coaching. Prepare to ask diagnostic questions (``why this technique, not that one?''; ``what would change your mind?'') rather than to teach. \textbf{Estimated prep:} 4 hours first time, 1 hour per repeat.
\end{curriculumteachingnotes}

% --------------------- MODULE 11 (M11) ---------------------
\chapter{Capstone}

\subsection*{Pointer}
Module 11 is the integrated end-to-end Capstone. The full capstone design --- case description, mandate, deliverables, scoring, and integrity safeguards --- is presented as a dedicated chapter (\textbf{Capstone Experience}) following the Assessment System chapter.

\begin{curriculumprerequisites}
Modules M01--M10 must be completed and the personal pre-capstone study plan from M10 must be on file with the instructor before the capstone window opens. The mid-capstone scripted change event (a regulatory update) will be released at the midpoint and must be addressed inside the capstone deliverables.
\end{curriculumprerequisites}

% =====================================================================
\FrontMatterChapter{Assessment System}

\section*{9. Assessment Design Principles}
The assessment system is designed for: (a) coverage of all program-level outcomes; (b) balance across formative and summative methods; (c) authenticity --- assessments reflect work BAs actually do; (d) constructive alignment --- every assessment ties to outcomes and to instruction; (e) defensible scoring via published rubrics; (f) accessibility --- multiple modalities, no time-pressure unless time-management is the construct under assessment; (g) feedback loops --- formative feedback always precedes summative judgment.

\section*{10. Assessment Portfolio}
The assessment portfolio is presented in landscape on the next page so all six columns remain legible.

\begin{landscape}
\thispagestyle{plain}
\setlength\LTleft{0pt}\setlength\LTright\fill
\renewcommand{\arraystretch}{1.15}
\footnotesize
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Assessment Portfolio --- IDs, Types, Coverage, Evidence, Scoring\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{6pt}%
\begin{longtable}{@{}R{2.55cm} R{2.30cm} R{3.40cm} R{2.55cm} R{4.30cm} R{3.50cm}@{}}
\toprule
\sffamily\bfseries Assessment ID &
\sffamily\bfseries Type &
\sffamily\bfseries What Is Assessed &
\sffamily\bfseries Outcome Alignment &
\sffamily\bfseries Evidence Expected &
\sffamily\bfseries Scoring Approach \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries ID &
\sffamily\bfseries Type &
\sffamily\bfseries What &
\sffamily\bfseries Aligns &
\sffamily\bfseries Evidence &
\sffamily\bfseries Scoring \\
\midrule
\endhead
\bottomrule
\endfoot

\tool{FC-M01..M10}
  & Formative checks (per module)
  & Conceptual fluency; vocabulary; basic technique recall
  & All module objectives
  & Quizzes (10--15 items), worked examples, in-class application
  & Ungraded with feedback; mastery threshold for self-monitoring \\

\tool{SA-M02}
  & Scenario analysis
  & Plan suitability under constraints
  & PLO-02
  & Written analysis of two contrasting scenarios
  & Analytical rubric R-SA \\

\tool{WS-M03}
  & Workshop performance
  & Facilitation behaviors; technique selection; conflict handling
  & PLO-03, PLO-08, PLO-11
  & Observed live (or recorded) workshop with debrief
  & Behavioral rubric R-WS \\

\tool{PE-M03..M07}
  & Practical exercises
  & Technique execution
  & Module objectives
  & Completed deliverables
  & Analytical rubric per module (R-Mxx) \\

\tool{AR-M02..M07}
  & Artifact reviews
  & Quality of produced artifacts
  & Module objectives, PLO-04 to PLO-07
  & Module artifact + structured review
  & Rubric per module + peer-review record \\

\tool{CB-M04, CB-M05}
  & Case-based assignments
  & Integrated application across multiple techniques
  & PLO-05, PLO-06
  & Written case response with artifacts
  & Analytical rubric R-CB \\

\tool{PT-M08}
  & Performance task
  & Underlying competencies under direct observation
  & PLO-08, PLO-11, PLO-12, PLO-13
  & Observed facilitation + three-form communication
  & Behavioral + analytical rubric R-PT \\

\tool{PS-M09}
  & Perspective studio portfolio
  & Adaptation across the five perspectives
  & PLO-10, PLO-09
  & Five perspective artifacts + comparative reflection
  & Analytical rubric R-M09 \\

\tool{TM-M10}
  & Techniques mastery lab
  & Free technique selection and execution
  & PLO-09
  & Selection rationale + executed deliverable + study plan
  & Analytical rubric R-M10 \\

\tool{CAP-M11}
  & Capstone
  & All outcomes integrated
  & All PLOs
  & End-to-end deliverables + stakeholder defense
  & Composite rubric R-CAP \\

\tool{SUM}
  & Summative evaluation
  & Demonstrated mastery across the program
  & All PLOs
  & Capstone outcome + portfolio evidence
  & Pass / Pass-with-distinction / Not yet, against R-SUM \\

\tool{RP-M01..M11}
  & Reflection portfolio
  & Metacognition; growth; identification of next learning
  & PLO-14
  & One portfolio entry per module
  & Analytical rubric R-RP (depth, evidence, action) \\

\end{longtable}
\end{landscape}

\section*{11. Frequency, Weighting, and Rubric Architecture}

\subsection*{11.1 Frequency and Weighting}
Formative assessment occurs continuously and is ungraded. Summative weighting is shown below.

\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,title=Summative Assessment Weighting]
\begin{tabularx}{\linewidth}{@{}R{8.0cm} c >{\raggedright\arraybackslash}X@{}}
\toprule
\sffamily\bfseries Component & \sffamily\bfseries Weight & \sffamily\bfseries Notes \\
\midrule
Capstone (M11)                                           & 40\% & Composite scoring against R-CAP. \\
Perspective studio portfolio (M09)                       & 15\% & Five perspective artifacts + comparative reflection. \\
Performance task (M08)                                   & 10\% & Observed facilitation + decision review. \\
Module artifact reviews (M02--M07, equal)                & 25\% & Six modules at $\approx$4.17\% each. \\
Reflection portfolio (RP-M01..M11)                       & 10\% & Aggregated portfolio depth, evidence, and action. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

A learner must achieve at least the ``Meets'' threshold on every PLO to pass the program.

\subsection*{11.2 Rubric Architecture}
All summative rubrics share a common structure: criterion $\to$ descriptors at four levels (\textbf{Not Yet}, \textbf{Approaching}, \textbf{Meets}, \textbf{Exceeds}) $\to$ evidence cues $\to$ score. Rubrics are published before the assessment is attempted. Inter-rater reliability is monitored across instructors via shared moderation sessions on a sample of submissions.

% =====================================================================
\FrontMatterChapter{Capstone Experience}

\section*{12. Capstone Overview}

\subsection*{12.1 Title and Case}
\textbf{Title.} \emph{Modernizing Member Services at Northstar Mutual.}

\textbf{Case description.} Northstar Mutual is a mid-sized member-owned financial cooperative serving 480{,}000 members across three regions. The Board has set a three-year strategic objective: ``Become the digital-first cooperative for working families.'' A new CEO has commissioned a portfolio of initiatives. The capstone learner is staffed as the lead business analyst on one initiative within the portfolio: replacing the current member-services platform with a modern omnichannel platform that integrates contact center, mobile, web, and branch.

The case provides realistic source materials: existing strategy slide deck, three quarters of contact-center metrics, current process documentation (partially out-of-date), an enterprise architecture inventory, two prior failed RFP responses, regulatory context (consumer financial protection), member satisfaction survey results, employee engagement data, vendor pitch decks, and a stakeholder roster of 22 named roles.

The case is deliberately ambiguous, contradictory in places, and richer than any single learner can fully exhaust --- replicating real conditions.

\textbf{Duration.} 56 hours over the final four weeks. Cohorts work individually with structured peer-review checkpoints; instructor-led panel defense at the end.

\subsection*{12.2 Capstone Learner Mandate}
Perform end-to-end business analysis on the initiative. The learner must, at minimum:

\begin{enumerate}
  \item \textbf{Identify need and context.} Synthesize the case materials into a stated business need, situated in the BACCM. Surface contradictions in source materials. ($\to$ PLO-01, PLO-05)
  \item \textbf{Analyze stakeholders.} Produce a stakeholder analysis (list, RACI, onion or matrix view) accounting for indirect, marginalized, and regulatory stakeholders. ($\to$ PLO-02, PLO-13)
  \item \textbf{Plan the analysis work.} Produce a Business Analysis Plan calibrated to a hybrid delivery context. Define performance measures for the analysis itself. ($\to$ PLO-02)
  \item \textbf{Conduct elicitation.} Plan and execute at least three elicitation activities of different techniques (e.g., simulated interview with role-player, document analysis, asynchronous workshop). Confirm and record results. ($\to$ PLO-03, PLO-09, PLO-11)
  \item \textbf{Analyze strategy.} Define current state, future state, and a change strategy with at least three options, decision analysis, and justified recommendation. Define SMART business objectives and trace expected value. ($\to$ PLO-05)
  \item \textbf{Model and specify requirements and designs.} Produce process, data, and decision models; functional and non-functional requirements; user stories with acceptance criteria; interface analysis; design options analysis; recommended solution. Include verification and validation evidence. ($\to$ PLO-06, PLO-09)
  \item \textbf{Manage requirements changes.} A scripted change event is injected mid-capstone (a regulatory update). The learner must assess impact, recommend disposition, update traceability, and re-baseline. ($\to$ PLO-04)
  \item \textbf{Evaluate solution options.} Define measures of value; assess solution fit; recommend value-realization actions for the recommended solution. ($\to$ PLO-07)
  \item \textbf{Apply at least three perspectives.} Produce capstone artifacts that explicitly demonstrate Agile, Information Technology, and one of (Business Architecture, BPM, BI) at the learner's choice. ($\to$ PLO-10)
  \item \textbf{Present recommendations.} Defend the work in a 45-minute panel: 20-minute presentation + 25-minute Q\&A from a panel composed of an instructor, a peer learner, and an external practitioner reviewer. ($\to$ PLO-12, PLO-15)
\end{enumerate}

\subsection*{12.3 Capstone Deliverables}

\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,title=Required Capstone Deliverables (C-01 to C-11)]
\begin{tabular}{@{}R{1.0cm} R{6.4cm} R{4.6cm} c@{}}
\toprule
\sffamily\bfseries ID & \sffamily\bfseries Deliverable & \sffamily\bfseries Format & \sffamily\bfseries Maps to Task \\
\midrule
\tool{C-01} & Need and context brief ($\leq$4 pages)                                  & Document             & 1  \\
\tool{C-02} & Stakeholder analysis pack                                                & Document + visual    & 2  \\
\tool{C-03} & Business Analysis Plan                                                   & Document             & 3  \\
\tool{C-04} & Elicitation event package (plans, results, confirmations)                & Document set         & 4  \\
\tool{C-05} & Strategy analysis pack (current/future/options/recommendation)           & Document + visuals   & 5  \\
\tool{C-06} & Requirements \& design specification (text, models, NFRs, V\&V)          & Document set + models& 6  \\
\tool{C-07} & Change-request response and updated traceability                         & Document + matrix    & 7  \\
\tool{C-08} & Solution evaluation plan and recommendations memo                        & Document             & 8  \\
\tool{C-09} & Three perspective-specific artifacts                                     & Per-perspective      & 9  \\
\tool{C-10} & Stakeholder defense pack (slides + handouts) and recorded defense        & Slides + recording   & 10 \\
\tool{C-11} & Reflection essay (1{,}500--2{,}000 words)                                & Document             & All\\
\bottomrule
\end{tabular}
\end{tcolorbox}

\subsection*{12.4 Capstone Scoring}
The capstone is scored against a composite rubric \tool{R-CAP} with weighted criteria:

\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,title=Capstone Composite Rubric R-CAP]
\begin{tabularx}{\linewidth}{@{}R{9.0cm} c >{\RaggedRight\arraybackslash}X@{}}
\toprule
\sffamily\bfseries Dimension & \sffamily\bfseries Weight & \sffamily\bfseries Excellent (full credit) means\dots \\
\midrule
Integration across knowledge areas                  & 20\% & Every KA contributes evidence to the recommendation; no KA is window dressing. \\
Rigor of analysis                                   & 20\% & Claims are evidenced; assumptions are surfaced; alternatives were considered. \\
Quality of artifacts                                & 20\% & Models reconcile with text; specifications are unambiguous; traceability is intact. \\
Perspective fidelity                                & 10\% & Each chosen perspective looks and behaves like itself, not a relabel. \\
Communication and defense                           & 15\% & Defense is responsive to Q\&A; presentation is calibrated to the panel. \\
Reflection                                          & 10\% & At least five named decisions traced to artifacts with rationale. \\
Ethics and inclusion in stakeholder treatment       &  5\% & Marginalized voices accounted for; conflicts of interest disclosed. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

External practitioner reviewers participate in scoring on at least 25\% of capstones to maintain external benchmarking.

\subsection*{12.5 Capstone Integrity and Authenticity Safeguards}
\begin{itemize}
  \item Mid-capstone scripted change event prevents pre-fabricated submissions.
  \item Defense Q\&A is unscripted and probes specific decisions in artifacts.
  \item Reflection essay must reference at least five named decisions with rationale traceable to artifacts.
  \item Originality declaration required.
\end{itemize}

% =====================================================================
\FrontMatterChapter{Techniques Integration Plan}

\section*{13. Technique Coverage Across the Program}
The BABOK lists 50 techniques (\S10). The plan below shows where major techniques are introduced (\textbf{I}), reinforced (\textbf{R}), and mastered (\textbf{M}) --- that is, where the learner is expected to perform the technique competently without scaffolding. All 50 techniques appear in the program at least at introduction; the table summarizes those given dedicated instructional time.

\begin{landscape}
\thispagestyle{plain}
\setlength\LTleft{0pt}\setlength\LTright\fill
\renewcommand{\arraystretch}{1.10}
\footnotesize
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Techniques Integration Plan --- I/R/M Across Modules\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{6pt}%
\begin{longtable}{@{}R{4.45cm} R{2.10cm} R{2.55cm} R{1.55cm} R{8.40cm}@{}}
\toprule
\sffamily\bfseries Technique &
\sffamily\bfseries Introduced &
\sffamily\bfseries Reinforced &
\sffamily\bfseries Mastered &
\sffamily\bfseries Example Learner Task \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries Technique &
\sffamily\bfseries Intro &
\sffamily\bfseries Reinforce &
\sffamily\bfseries Master &
\sffamily\bfseries Example Task \\
\midrule
\endhead
\bottomrule
\endfoot

Acceptance and evaluation criteria & M05 & M06, M09 & M11 & Write acceptance criteria for a user story slated for sprint commitment \\
Backlog management                 & M03, M06 & M09 & M11 & Refine and order a 30-item backlog using a chosen prioritization method \\
Balanced scorecard                 & M07 & M09 & M11 & Construct a scorecard that links operational metrics to strategic objectives \\
Benchmarking and market analysis   & M04 & M07 & M11 & Benchmark proposed digital onboarding against three competitor offerings \\
Brainstorming                      & M01 & M03, M04 & M11 & Facilitate a structured brainstorm to surface candidate solution options \\
Business capability analysis       & M04 & M09 & M11 & Map current capabilities, score maturity, identify capability gaps \\
Business cases                     & M02 & M04 & M11 & Build a defensible business case for a recommended option \\
Business model canvas              & M04 & M09 & M11 & Construct a business model canvas for a proposed product line \\
Business rules analysis            & M05 & M09 & M11 & Extract and structure business rules from policy documents \\
Collaborative games                & M03 & M08 & ---  & Run a collaborative game to surface latent stakeholder priorities \\
Concept modelling                  & M01 & M05 & M11 & Build a concept model that disambiguates ``member'' across departments \\
Data dictionary                    & M05 & M06 & M11 & Author a data dictionary for the in-scope domain entities \\
Data flow diagrams                 & M05 & M09 & M11 & Diagram data flow across legacy and proposed systems \\
Data mining                        & M07 & M09 & ---  & Identify patterns in synthetic contact-center data \\
Data modelling                     & M05 & M09 & M11 & Produce an ERD for the in-scope domain \\
Decision analysis                  & M02, M04 & M06, M08 & M11 & Score three change-strategy options against weighted criteria \\
Decision modelling                 & M05 & M07 & M11 & Construct a DMN model for an automated decision \\
Document analysis                  & M02 & M03, M07 & M11 & Synthesize regulatory and policy documents into requirements input \\
Estimation                         & M02 & M06 & M11 & Estimate analysis effort under three delivery scenarios \\
Financial analysis                 & M04 & M07 & M11 & Build a five-year cost-benefit projection \\
Focus groups                       & M03 & M07 & ---  & Plan and run a focus group on member experience \\
Functional decomposition           & M02, M05 & M09 & M11 & Decompose a high-level capability into sub-functions \\
Glossary                           & M01 & All & M11 & Maintain a project glossary across the program \\
Interface analysis                 & M05 & M09 & M11 & Document integration interfaces with non-functional needs \\
Interviews                         & M03 & M04, M07 & M11 & Conduct a structured interview with a difficult role-player stakeholder \\
Item tracking                      & M02 & M06 & M11 & Maintain an issue/risk/decision log across the capstone \\
Lessons learned                    & M07 & M11 & M11 & Lead a structured lessons-learned session post-capstone \\
Metrics and KPIs                   & M02 & M07 & M11 & Define KPIs for solution value realization \\
Mind mapping                       & M03 & M04 & ---  & Use mind mapping to organize an elicitation session output \\
Non-functional requirements analysis & M05 & M09 & M11 & Specify NFRs covering performance, security, accessibility, and resilience \\
Observation                        & M03 & M07 & M11 & Conduct an active observation of a contact-center workflow \\
Organizational modelling           & M02 & M04 & M11 & Map the organizational structure relevant to the change \\
Prioritization                     & M06 & M09 & M11 & Apply MoSCoW, Kano, and weighted methods; reconcile differences \\
Process analysis                   & M07 & M09 & M11 & Analyze a current process for waste, variation, and constraint \\
Process modelling                  & M04, M05 & M09 & M11 & Model current and proposed processes in BPMN \\
Prototyping                        & M05 & M09 & M11 & Produce a low-fidelity prototype of a key user interaction \\
Reviews                            & M02 & M05, M06, M08 & M11 & Lead a structured requirements walkthrough \\
Risk analysis and management       & M02 & M04, M07 & M11 & Build and maintain a risk register through the capstone \\
Roles and permissions matrix       & M05 & M09 & M11 & Define a roles-and-permissions matrix for a sensitive capability \\
Root-cause analysis                & M04 & M07 & M11 & Diagnose a recurring service failure using fishbone and 5-Whys \\
Scope modelling                    & M02, M04 & M09 & M11 & Define and visualize scope boundaries with assumptions and exclusions \\
Sequence diagrams                  & M05 & M09 & ---  & Diagram a key cross-system interaction \\
Stakeholder list, map, or personas & M02 & All & M11 & Produce stakeholder analysis, refresh as the case evolves \\
State modelling                    & M05 & M09 & ---  & Model the state lifecycle of a key domain entity \\
Survey or questionnaire            & M03 & M07 & M11 & Design a member survey instrument and analyze results \\
SWOT analysis                      & M04 & M09 & M11 & Conduct SWOT for the future-state operating model \\
Use cases and scenarios            & M05 & M09 & M11 & Author a use case set with main and alternative flows \\
User stories                       & M05 & M06, M09 & M11 & Write user stories with acceptance criteria for a release slice \\
Vendor assessment                  & M04 & M07 & M11 & Score vendor proposals against weighted criteria \\
Workshops                          & M03 & M04, M05, M08 & M11 & Plan and facilitate a multi-stakeholder requirements workshop \\

\end{longtable}
\end{landscape}

% =====================================================================
\FrontMatterChapter{Perspectives Integration Plan}

\section*{14. Perspectives Approach}
Perspectives are not taught as appendices. They are woven through the program as follows:

\begin{itemize}
  \item \textbf{Introduced} in M01 as one of the five structural elements of BABOK.
  \item \textbf{Threaded} through M02--M07: each module's instruction explicitly contrasts how the module's KA tasks differ across at least two perspectives.
  \item \textbf{Studio} in M09: each perspective is applied as a self-contained rotation against a shared case, producing a perspective-specific artifact and surfacing comparative differences.
  \item \textbf{Required} in the capstone: at least three perspectives must be evidenced in capstone artifacts with explicit fidelity.
\end{itemize}

\section*{15. Perspective Contrast --- Worked Example}
Consider a single business analysis task: \textbf{``Define requirements for the way members request a loan.''} The contrast below illustrates how the work differs by perspective. This contrast is used in M09 as the studio's reference example.

\begin{landscape}
\thispagestyle{plain}
\setlength\LTleft{0pt}\setlength\LTright\fill
\renewcommand{\arraystretch}{1.18}
\footnotesize
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Perspective Contrast --- ``Define Requirements for Member Loan Request''\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{6pt}%
\begin{longtable}{@{}R{2.40cm} R{3.60cm} R{3.60cm} R{3.60cm} R{3.60cm} R{3.60cm}@{}}
\toprule
\sffamily\bfseries Aspect &
\sffamily\bfseries Agile &
\sffamily\bfseries Business Intelligence &
\sffamily\bfseries Information Technology &
\sffamily\bfseries Business Architecture &
\sffamily\bfseries Business Process Mgmt \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries Aspect &
\sffamily\bfseries Agile &
\sffamily\bfseries BI &
\sffamily\bfseries IT &
\sffamily\bfseries BA &
\sffamily\bfseries BPM \\
\midrule
\endhead
\bottomrule
\endfoot

Primary unit of analysis
  & User-story slice ready for a sprint
  & Analytical use case and metric definition
  & Functional/non-functional requirement, integration spec
  & Capability and value-stream segment
  & Process step within an end-to-end process \\

Lead artifact
  & Refined backlog with acceptance criteria
  & Metric catalog + dashboard wireframe + source-to-target summary
  & Requirements \& integration specification
  & Capability heat map + value stream map
  & Current and future BPMN models with performance baseline \\

Stakeholder emphasis
  & Product owner; delivery team; user representative
  & Business decision-makers; data steward; analytics consumer
  & Architect; security; platform engineering; data
  & Enterprise architect; capability owner; strategy office
  & Process owner; operations leader; quality \\

Verification posture
  & Acceptance criteria validated incrementally during the sprint
  & Metric definitions reviewed by data governance
  & Walkthroughs against architectural standards and NFRs
  & Alignment to enterprise model
  & Process simulation or pilot vs. defined performance targets \\

Risk emphasis
  & Slice doesn't deliver thin vertical value
  & Metric is misdefined or gameable
  & Integration brittleness; security; resilience
  & Capability duplication or gap
  & Process variation; bottleneck; non-compliance \\

Typical failure mode
  & Backlog filled with technical tasks dressed as stories
  & Dashboards proliferate without trustworthy data
  & Specs over-engineered or under-specified
  & Capability map disconnected from operating reality
  & Process map approved that no one follows \\

\end{longtable}
\end{landscape}

\section*{16. Capstone Perspective Requirement}
The capstone requires at least three perspectives to be evidenced. Two are mandatory (\textbf{Agile} and \textbf{IT}) because they are most universally relevant to the case as written; the learner selects the third from \{Business Architecture, BPM, Business Intelligence\}.

% =====================================================================
\FrontMatterChapter{Instructor Usability Package}

\section*{17. Program-Level Teaching Notes}
\begin{itemize}
  \item Teach to the rubric, not around it. Every assessment rubric should be visible to learners before the work begins. Use rubrics as in-class diagnostic tools.
  \item Move the teaching weight toward observation and feedback, not delivery. The lecture-to-lab ratio is intentionally low (20:40 in \S7.1).
  \item Manage cognitive load. RADD (M05) is the heaviest module --- protect its time and avoid overscheduling the surrounding weeks.
  \item Resist over-customization. Encourage learners to use the program case for skill-building; let workplace transfer come through the reflection portfolio rather than by replacing the case.
  \item Be deliberate about the perspectives. Do not collapse them into a single ``modern delivery'' stew. They are real differences that the program tests on.
  \item Build a moderation habit. Score samples together with co-instructors monthly to maintain inter-rater reliability.
\end{itemize}

\section*{18. Facilitation Guidance}
\begin{itemize}
  \item \textbf{Cold-open with case.} Begin every session with a 5-minute case prompt. Lecture, if any, should answer questions the case has raised.
  \item \textbf{Use the 70/30 rule for activity.} Spend 70\% of synchronous time in activity, 30\% in instruction or debrief.
  \item \textbf{Explicitly name competencies during practice.} When a learner exhibits structured analytical thinking, name it. When facilitation falters, name the behavior, not the person.
  \item \textbf{Keep records.} Use the supplied facilitation log (per-session template) to note observed behaviors that will inform the M08 performance task.
\end{itemize}

\section*{19. Estimated Preparation Effort}

\clearpage
\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,
  title=Instructor Preparation Effort by Module]
\begin{tabularx}{\linewidth}{@{}R{6.0cm} c c >{\RaggedRight\arraybackslash}X@{}}
\toprule
\sffamily\bfseries Module & \sffamily\bfseries First-time Prep & \sffamily\bfseries Repeat Prep & \sffamily\bfseries Notes \\
\midrule
M01 --- Foundations                                  &  4 h & 1.5 h & Tone-setting; light prep for repeats. \\
M02 --- BAPM                                          &  6 h & 2 h   & Case selection drives prep. \\
M03 --- Elicitation \& Collaboration                  &  8 h & 3 h   & High because of recording reviews. \\
M04 --- Strategy Analysis                             &  7 h & 2.5 h & Case + decision-analysis prep. \\
M05 --- RADD                                          & 10 h & 4 h   & Heaviest module; protect prep. \\
M06 --- RLCM                                          &  5 h & 1.5 h & Rubric grading is the time sink. \\
M07 --- Solution Evaluation                           &  5 h & 1.5 h & Synthetic data setup. \\
M08 --- Underlying Competencies in Practice           &  6 h & 2 h   & Behavioral rubric calibration. \\
M09 --- Perspectives Studio                           &  8 h & 3 h   & Five rubrics; guest practitioners. \\
M10 --- Integration \& Techniques Mastery Lab         &  4 h & 1 h   & Coaching, not teaching. \\
M11 --- Capstone facilitation                         & 12 h & 5 h   & Panel coordination; external reviewers. \\
\midrule
\textbf{Total}                                       & \textbf{75 h} & \textbf{27 h} & \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\section*{20. Required Materials}
\begin{itemize}
  \item Learning platform (LMS) supporting captioned video, document hosting, discussion, peer review, rubric-based scoring, and timed quizzes.
  \item Modeling toolchain: at least one freely-available option (e.g., a \tool{Mermaid} or \tool{PlantUML}-capable editor) and one industry-standard option, instructor's choice.
  \item Northstar Mutual capstone case bundle (provided): strategy deck, contact-center metrics, process documentation, EA inventory, RFP responses, regulatory snapshot, survey data, employee engagement data, vendor decks, stakeholder roster, role-player briefing scripts.
  \item Per-module artifact templates (provided in document, spreadsheet, and structured-text formats).
  \item Rubric set (\tool{R-M01} through \tool{R-M11}, \tool{R-SA}, \tool{R-WS}, \tool{R-CB}, \tool{R-PT}, \tool{R-RP}, \tool{R-CAP}, \tool{R-SUM}).
  \item Observation log template; peer-review template; facilitation log template.
\end{itemize}

\section*{21. Optional Adaptations}
\begin{itemize}
  \item \textbf{Industry overlays:} each module includes optional case variants for regulated financial services, public-sector modernization, healthcare claims, and B2B SaaS.
  \item \textbf{Tooling overlays:} the program is tool-agnostic; instructors may localize tooling choices without changing assessment design.
  \item \textbf{Certification preparation overlays:} optional CCBA/CBAP study pacing guides keyed to BABOK chapter readings, scheduled across modules.
\end{itemize}

\section*{22. Delivery Variants}

\subsection*{22.1 Online Delivery Adaptations}
\begin{itemize}
  \item Replace in-room workshops with structured live virtual workshops ($\leq$8 learners per breakout; structured roles).
  \item Replace observed in-room facilitation with recorded virtual facilitation and asynchronous review.
  \item Use written-first protocols in synchronous discussion to balance participation across time zones and language preferences.
  \item Provide all materials in a screen-reader-tested LMS structure with captioned recordings within 24 hours.
\end{itemize}

\subsection*{22.2 Accelerated Cohort (8 Weeks)}
\begin{itemize}
  \item Compress to ${\sim}$35 hours per learner per week.
  \item Combine M01 and M02 into a single intensive week.
  \item Deliver M09 perspectives as parallel small-team rotations rather than sequential.
  \item Reduce capstone to 3 weeks with a smaller, more constrained case scope.
  \item Pre-work expectation: 20-hour primer covering BACCM and BABOK structure before week 1.
\end{itemize}

\subsection*{22.3 Self-Paced (16+ Weeks)}
\begin{itemize}
  \item Deliver instruction as captioned video micro-modules (15--25 minutes each).
  \item Replace synchronous workshops with cohort-rotation workshops scheduled monthly.
  \item Pair every learner with a coach for fortnightly 30-minute check-ins.
  \item Capstone defended in a regularly scheduled monthly panel.
\end{itemize}

% =====================================================================
\FrontMatterChapter{Equity, Accessibility, and Inclusion Review}

\section*{23. Different Experience Levels}
The mandatory entry diagnostic and bridging primer normalize entry. The foundational module (M01) re-establishes shared vocabulary regardless of background. Within modules, activities are tiered with stretch options for experienced learners (e.g., advanced modeling exercises in M05, perspective leadership in M09). Reflection prompts are calibrated to the learner's own starting point rather than to a fixed peer norm.

\section*{24. Different Learning Styles}
Every concept appears in at least two modalities (visual and textual; or visual, textual, and spoken). Activities span individual, paired, small-group, and full-cohort formats. Practice is embedded in every module; instruction-only sequences are bounded.

\section*{25. Different Communication Preferences}
Written-first discussion protocols are available in every synchronous activity. Asynchronous alternatives are provided for at least one major activity per module. Live presentation requirements (e.g., capstone defense) include a stated accommodation pathway (extended time, presentation format alternatives such as recorded defense with live Q\&A).

\section*{26. Accessibility Needs}
\begin{itemize}
  \item All recorded materials are captioned within 24 hours; transcripts available.
  \item All visual materials have textual alternatives.
  \item All templates available in document, spreadsheet, and structured-text (e.g., Markdown) formats.
  \item Modeling tools include a no-cost, keyboard-accessible option (e.g., \tool{Mermaid} or \tool{PlantUML}) so visual modeling never requires a pointing device.
  \item Time-bounded activities are not time-pressured unless time-management is the construct under assessment; extended time is available on request.
  \item Color is never the sole encoding in any model or visualization; high-contrast palettes are used by default.
  \item Reading load is bounded per module ($\leq$8 hours of reading).
\end{itemize}

\section*{27. Culturally Inclusive Examples}
The case bank intentionally spans sectors, geographies, and organizational types. Stakeholder rosters in cases include diverse names, roles, and contexts. Roleplay scripts are reviewed to avoid harmful stereotypes; participants can opt out of any role with no penalty. Examples avoid culturally specific idioms unless explicitly taught.

\section*{28. Multiple Professional Contexts}
The program is designed to serve learners working in regulated and unregulated industries; in product, project, and operational contexts; in consultancy and in-house roles; in startups and in enterprises. Module activities use cases drawn from across these contexts. The capstone is an enterprise case by design --- learners from other contexts apply techniques to a setting larger than their day-to-day, which is the point.

% =====================================================================
\FrontMatterChapter{Continuous Improvement Plan}

\section*{29. Governance}
A \textbf{Curriculum Council} owns the program. Composition: lead curriculum architect (chair), two senior instructors, one external practitioner reviewer (rotating annually), one accessibility reviewer, one learner representative (rotating per cohort), one IIBA-credentialed advisor.

\section*{30. Inputs to the Review Cycle}
\begin{itemize}
  \item \textbf{Stakeholder feedback inputs.} End-of-module surveys; end-of-program surveys; instructor debriefs; capstone external reviewer feedback; employer feedback (where available); IIBA standards updates.
  \item \textbf{Learner performance evidence.} Aggregate rubric scores by criterion; failure-mode analysis on items where learners cluster below ``Meets''; reflection portfolio thematic analysis; capstone defense Q\&A patterns.
  \item \textbf{Instructor observations.} Facilitation logs; moderation session notes; identified gaps in materials.
  \item \textbf{External signals.} Updates to BABOK; changes in widely-used industry frameworks (e.g., Agile, BPM, BI); shifts in tooling.
\end{itemize}

\section*{31. Revision Triggers}

\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,title=Revision Trigger Catalogue]
\begin{tabularx}{\linewidth}{@{}R{6.4cm} >{\RaggedRight\arraybackslash}X@{}}
\toprule
\sffamily\bfseries Trigger & \sffamily\bfseries Action \\
\midrule
BABOK new edition                                                                & Full alignment review; major version update planned within 6 months. \\
Cohort-level ``Not Yet'' rate $>$15\% on any PLO                                  & Targeted module review and instructional revision. \\
Three consecutive cohort surveys flag the same instructional issue               & Immediate corrective revision. \\
Accessibility complaint or audit finding                                         & Immediate corrective action; remediation within 30 days. \\
External practitioner feedback indicates real-world drift                        & Industry overlay review. \\
Persistent inter-rater reliability below threshold                               & Rubric revision and instructor moderation refresh. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\section*{32. Versioning and Documentation}
\textbf{Semantic versioning.} Major (X.0.0) for structural/PLO changes; minor (1.X.0) for module redesigns or new perspectives content; patch (1.0.X) for corrections, accessibility fixes, and template updates. A change log is maintained alongside the curriculum, identifying date, version, change, and rationale. Each version bundles the curriculum text, rubrics, templates, and case materials so that any cohort can be reproduced exactly.

\section*{33. Review Cycle Cadence}
\begin{itemize}
  \item \textbf{Continuous:} defect log open to instructors and learners.
  \item \textbf{Per cohort:} end-of-cohort retrospective produces a recommendations memo.
  \item \textbf{Quarterly:} Council reviews the recommendations memo, decides on patch-level changes.
  \item \textbf{Annually:} Council reviews aggregate evidence, decides on minor-version revision.
  \item \textbf{On BABOK update:} major-version review.
\end{itemize}

\section*{34. Stakeholder Validation for New Releases}
Before any major version release: at least one external practitioner review; at least one accessibility review; pilot with a small cohort with explicit feedback collection; sign-off by the Council recorded in the change log.

% =====================================================================
\FrontMatterChapter{Alignment Matrix}

\section*{35. Single Source of Alignment Truth}
The matrix below is the curriculum's single source of truth for alignment. Reading order: each row is a Program Learning Outcome with the modules where it is developed, the BABOK areas it draws on, the assessments that evidence it, the artifacts produced, the underlying competencies exercised, the techniques required, and the perspectives applied.

\begin{landscape}
\thispagestyle{plain}
\setlength\LTleft{0pt}\setlength\LTright\fill
\renewcommand{\arraystretch}{1.10}
\footnotesize
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Alignment Matrix --- PLO $\to$ Modules $\to$ BABOK $\to$ Assessments $\to$ Artifacts $\to$ Competencies $\to$ Techniques $\to$ Perspectives\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{5pt}%
\begin{longtable}{@{}R{2.40cm} R{2.50cm} R{1.85cm} R{2.30cm} R{3.05cm} R{2.30cm} R{3.10cm} R{2.40cm}@{}}
\toprule
\sffamily\bfseries PLO &
\sffamily\bfseries Modules &
\sffamily\bfseries BABOK &
\sffamily\bfseries Assessments &
\sffamily\bfseries Artifacts &
\sffamily\bfseries Competencies &
\sffamily\bfseries Techniques &
\sffamily\bfseries Perspectives \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries PLO &
\sffamily\bfseries Modules &
\sffamily\bfseries BABOK &
\sffamily\bfseries Assessments &
\sffamily\bfseries Artifacts &
\sffamily\bfseries Competencies &
\sffamily\bfseries Techniques &
\sffamily\bfseries Perspectives \\
\midrule
\endhead
\bottomrule
\endfoot

PLO-01 BACCM application
  & M01 (lead); recurring M02--M11
  & Key Concepts
  & FC-M01; reflection portfolio; capstone
  & BACCM concept maps; capstone need brief
  & Conceptual thinking; communication
  & Concept modelling; glossary
  & All (introduced) \\

PLO-02 Plan BA work
  & M02 (lead); revisited M11
  & BAPM
  & AR-M02; SA-M02; CAP-M11
  & BA Plan; stakeholder analysis; performance metrics; risk register
  & Analytical thinking; tools
  & Plan BA approach; stakeholder analysis; RACI; risk; metrics; decision analysis
  & Predictive vs. Agile contrast \\

PLO-03 Conduct elicitation
  & M03 (lead); revisited M07, M11
  & E\&C
  & WS-M03; AR-M03; CAP-M11
  & Elicitation plan; results; recorded facilitation; stakeholder briefing
  & Communication; interaction; behavioral
  & Interviews; workshops; observation; document analysis; surveys; brainstorming
  & Agile, IT \\

PLO-04 Manage requirements lifecycle
  & M06 (lead); revisited M11
  & RLCM
  & AR-M06; CAP-M11 (incl. change event)
  & Traceability matrix; prioritized backlog; change-request response; decision log
  & Analytical thinking; tools
  & Traceability; prioritization; item tracking; reviews
  & Agile, IT \\

PLO-05 Strategy analysis
  & M04 (lead); revisited M11
  & SA
  & AR-M04; CB-M04; CAP-M11
  & Current/future state; options analysis; recommendation; updated business case
  & Analytical \& business; problem-solving
  & SWOT; PESTLE; capability analysis; root-cause; decision analysis; financial
  & Business Architecture, BPM \\

PLO-06 Specify requirements \& designs
  & M05 (lead); revisited M09, M10, M11
  & RADD
  & AR-M05; CB-M05; PE-M05; TM-M10; CAP-M11
  & Requirements \& design spec; models; NFRs; V\&V evidence; design options
  & Conceptual thinking; modelling; tools
  & Process, data, decision, state modelling; user stories; use cases; NFR analysis; reviews; prototyping
  & Agile, IT, BI \\

PLO-07 Solution evaluation
  & M07 (lead); revisited M11
  & SE
  & AR-M07; CAP-M11
  & Performance plan; analysis; readiness; recommendations
  & Analytical \& systems; business knowledge
  & KPIs; balanced scorecard; benchmarking; root-cause; survey; observation
  & BPM, BI, IT \\

PLO-08 Underlying competencies
  & All modules; consolidated M08
  & Underlying Competencies
  & PT-M08; CAP-M11
  & Three-form communication; observed facilitation; competency self-assessment
  & All six categories
  & Workshops; reviews; collaborative games
  & All \\

PLO-09 Technique selection
  & M02--M07; M10 (lead); M11
  & Techniques
  & TM-M10; CAP-M11; module ARs
  & Selection rationale; executed deliverables
  & Tools \& technology; analytical
  & Free selection from \S10
  & All \\

PLO-10 Perspective adaptation
  & M02--M07 (threads); M09 (lead); M11
  & Perspectives
  & PS-M09; CAP-M11
  & Five perspective artifacts; capstone perspectives
  & Cross-domain knowledge; analytical
  & Perspective-appropriate techniques
  & Agile, BI, IT, BA, BPM \\

PLO-11 Facilitation \& conflict
  & M03 (lead); M08; M11
  & E\&C; UC
  & WS-M03; PT-M08; CAP-M11
  & Recorded facilitations; debriefs
  & Interaction; behavioral; communication
  & Workshops; collaborative games; reviews
  & Agile, IT \\

PLO-12 Communication
  & All modules; M08 (lead); M11
  & E\&C; UC
  & PT-M08; CAP-M11
  & Three-form communication; defense pack
  & Communication
  & Document analysis; reviews; concept modelling
  & All \\

PLO-13 Ethical, inclusive practice
  & M01; M02; M08; M11
  & UC
  & RP all; PT-M08; CAP-M11
  & Stakeholder maps; decision logs; ethics rationale
  & Ethics; behavioral
  & Stakeholder analysis; decision analysis
  & All \\

PLO-14 Reflection \& growth
  & All modules
  & UC; all KAs
  & RP-M01..M11; CAP-M11
  & Portfolio entries; pre-capstone study plan
  & Learning; personal accountability
  & ---
  & All \\

PLO-15 End-to-end performance
  & M11
  & All
  & CAP-M11
  & All capstone deliverables
  & All
  & All
  & At least three required \\

\end{longtable}
\end{landscape}

% =====================================================================
\FrontMatterChapter{Quality Gate Review}

\section*{36. Minimum Curriculum Quality Checklist Review}
Each minimum quality criterion is evaluated below. Status options: \textbf{Meets}, \textbf{Partially Meets}, \textbf{Does Not Yet Meet}.

\begin{landscape}
\thispagestyle{plain}
\setlength\LTleft{0pt}\setlength\LTright\fill
\renewcommand{\arraystretch}{1.15}
\footnotesize
\noindent\colorbox{afnavy}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{\raggedright\color{white}\sffamily\bfseries\large\, Quality Gate Review (Q1--Q15)\strut}}\par\nobreak\vspace{2pt}\nobreak
\setlength{\tabcolsep}{6pt}%
\begin{longtable}{@{}R{0.65cm} R{3.85cm} R{1.45cm} R{4.25cm} R{3.55cm} R{4.30cm}@{}}
\toprule
\sffamily\bfseries \# &
\sffamily\bfseries Criterion &
\sffamily\bfseries Status &
\sffamily\bfseries Evidence &
\sffamily\bfseries Gap (if any) &
\sffamily\bfseries Recommended Correction \\
\midrule
\endfirsthead
\toprule
\sffamily\bfseries \# &
\sffamily\bfseries Criterion &
\sffamily\bfseries Status &
\sffamily\bfseries Evidence &
\sffamily\bfseries Gap &
\sffamily\bfseries Correction \\
\midrule
\endhead
\bottomrule
\endfoot

Q1  & Clear, specific, measurable learning outcomes
    & Meets
    & 15 PLOs each with action verb + observable behavior + criterion (\S6). Module objectives identically structured. Alignment matrix (\S35).
    & None material.
    & --- \\

Q2  & Logical scope and sequence from foundational to advanced
    & Meets
    & Sequence rationale documented (\S8) with explicit dependencies and a justified departure from BABOK reference order.
    & None material.
    & --- \\

Q3  & Student-centered design
    & Meets
    & Diagnostic-and-primer pathway (\S3); reflection portfolio in every module; free technique selection (M10); capstone artifact ownership; multiple delivery variants (\S22).
    & Could deepen via optional specialization tracks.
    & Add a specialization track (e.g., regulated industries) in v1.1. \\

Q4  & Inclusive and equitable access
    & Meets
    & Dedicated equity, accessibility, and inclusion review (\S23--\S28); accessibility considerations per module; written-first protocols; multi-format templates; tool-agnostic design.
    & Cohort-language localization not yet addressed.
    & Plan localization workstream in next minor release. \\

Q5  & Real-world relevance
    & Meets
    & All modules produce practitioner artifacts; capstone is a realistic enterprise case with deliberate ambiguity (\S12); industry-overlay variants (\S21).
    & None material.
    & --- \\

Q6  & Engagement and active learning
    & Meets
    & 60\% of program time in active work (\S7.1); every module has 4--6 named activities; workshops, labs, studios.
    & Could add more cross-cohort collaborative work.
    & Pilot a paired-cohort exchange in M09 in v1.1. \\

Q7  & Critical thinking, creativity, communication, collaboration, digital literacy
    & Meets
    & Critical thinking developed in M04, M05, M07, M10; creativity in design options (M05), brainstorming (M03), perspective re-application (M09); communication in M03, M08, M11; collaboration in workshops across M03--M11; digital literacy in modeling toolchain, traceability tooling, capstone defense pack.
    & None material.
    & --- \\

Q8  & Flexible and adaptable structure
    & Meets
    & Default, online, accelerated, and self-paced variants (\S22); industry overlays; tool-agnostic.
    & None material.
    & --- \\

Q9  & Balanced assessment strategy
    & Meets
    & Formative, scenario, workshop, practical, artifact review, case-based, performance, capstone, summative --- all present and mapped in \S10.
    & None material.
    & --- \\

Q10 & Student reflection opportunities
    & Meets
    & Reflection prompts in every module; reflection portfolio assessment RP-M01..M11; capstone reflection essay (\S12).
    & None material.
    & --- \\

Q11 & Collaborative development and stakeholder validation
    & Meets
    & Curriculum Council composition includes external and learner representation (\S29); validation gates for new releases (\S34).
    & First-cohort pilot data not yet incorporated by definition.
    & Insert pilot cohort gate before public release. \\

Q12 & Continuous improvement and revision process
    & Meets
    & Defined inputs, triggers, versioning, cadence, validation (\S29--\S34).
    & None material.
    & --- \\

Q13 & Strong representation of multiple perspectives
    & Meets
    & Perspectives introduced (M01), threaded across M02--M07, applied in studio (M09), required in capstone (M11). Worked contrast in \S15.
    & None material.
    & --- \\

Q14 & Visible instruction--assessment alignment
    & Meets
    & Alignment matrix (\S35); each module objective maps to PLO; each assessment maps to PLOs; rubrics published in advance.
    & None material.
    & --- \\

Q15 & High instructor usability
    & Meets
    & Instructor package \S17--\S22: teaching notes, facilitation guidance, prep effort estimates per module, materials list, four delivery variants, optional adaptations.
    & Could provide annotated session-by-session lesson plans.
    & Develop annotated session plans as v1.0 instructor companion. \\

\end{longtable}
\end{landscape}

\section*{37. Critical Acceptance Criteria Check}

\begin{tcolorbox}[callout,colback=afpaper,colframe=afnavy,colbacktitle=afnavy,title=Critical Acceptance Criteria]
\begin{tabularx}{\linewidth}{@{}R{6.6cm} R{2.0cm} >{\RaggedRight\arraybackslash}X@{}}
\toprule
\sffamily\bfseries Criterion & \sffamily\bfseries Status & \sffamily\bfseries Evidence \\
\midrule
Observable and measurable outcomes
  & Meets
  & All 15 PLOs structured action-verb + behavior + criterion; module objectives same structure. \\
Cumulative sequencing
  & Meets
  & \S8 documents dependencies; modules consume prior modules' artifacts. \\
Inclusive and accessible design
  & Meets
  & \S23--\S28 review; per-module accessibility considerations; multi-format templates; written-first protocols; no-cost tool option; opt-out roles. \\
Active engagement through discussion, collaboration, inquiry, creation, application
  & Meets
  & 60\% active time; named activities in every module; workshops, labs, studios; capstone. \\
Ongoing and balanced assessment aligned to outcomes
  & Meets
  & \S10 portfolio + \S35 matrix prove balance and alignment. \\
Critical thinking, collaboration, communication, creativity, digital literacy
  & Meets
  & See Q7 above. \\
Documented stakeholder review, revision, and continuous improvement
  & Meets
  & \S29--\S34 governance, inputs, triggers, cadence, validation. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}

\section*{38. Overall Quality Gate Verdict}

\begin{tcolorbox}[callout,colback=afteal!8,colframe=afteal,colbacktitle=afteal,title=Overall Verdict]
\sffamily\bfseries\large Passes the Quality Gate.\\[2pt]
\normalfont All fifteen minimum criteria are at \textbf{Meets}. Three criteria carry non-blocking enhancement recommendations queued for v1.1: specialization track (Q3), paired-cohort exchange (Q6), annotated session plans (Q15). A pilot cohort gate is inserted before public release (Q11) to satisfy the stakeholder validation criterion with empirical evidence.
\end{tcolorbox}

% =====================================================================
\FrontMatterChapter{Appendix A --- Required and Recommended Reading}

\section*{A.1 Primary Reference}
\begin{itemize}
  \item \term{A Guide to the Business Analysis Body of Knowledge (BABOK Guide)}, Version 3. International Institute of Business Analysis (IIBA), 2015. Required for all learners; the curriculum maps explicitly to BABOK \S2--\S11.
\end{itemize}

\section*{A.2 Recommended Companions (Optional)}
\begin{itemize}
  \item Source materials supplied with each module: per-module reading list ($\leq$8 hours/module), drawn from open and licensed sources.
  \item CCBA / CBAP study companions (instructor-curated; pacing schedules in \S21).
  \item Modeling reference: \tool{BPMN 2.0} reference card; \tool{DMN 1.x} primer (instructor's choice).
\end{itemize}

\begin{infobox}[Reference Discipline]
The BABOK is the program's authoritative reference. Where the curriculum departs from BABOK ordering or grouping, the rationale is documented in \S8. Instructors should make this explicit to learners early so the program's pedagogy and the BABOK's reference structure are not confused.
\end{infobox}

% =====================================================================
\FrontMatterChapter{Appendix B --- Artifact Template Index}

The following artifact templates are bundled with the curriculum and supplied in document, spreadsheet, and structured-text formats so each learner can choose the modality that suits the task and the accommodations.

\renewcommand{\arraystretch}{1.18}
\rowcolors{2}{afrowalt}{white}
\begin{tabularx}{\linewidth}{@{}R{2.0cm} R{4.6cm} >{\RaggedRight\arraybackslash}X@{}}
\rowcolor{afnavy}
\multicolumn{3}{l}{\color{white}\sffamily\bfseries Artifact Template Index}\\
\toprule
\sffamily\bfseries Module & \sffamily\bfseries Template & \sffamily\bfseries Purpose \\
\midrule
M01 & BACCM concept map; competency self-assessment & Foundations and self-baseline \\
M02 & Business Analysis Plan; stakeholder list/RACI/onion; risk register; metrics catalogue & Planning and monitoring artifacts \\
M03 & Elicitation plan; results-confirmation record; communication brief & Elicitation and collaboration evidence \\
M04 & Current-state assessment; future-state SMART objectives; options analysis; change strategy & Strategy analysis pack \\
M05 & Requirements specification; BPMN/ERD/DMN models; user-story register; NFR list; V\&V record & Requirements and design \\
M06 & Traceability matrix; prioritized backlog; change-request package; approval decision log & Lifecycle management \\
M07 & Performance measurement plan; performance analysis report; readiness assessment; recommendations memo & Solution evaluation \\
M08 & Three-form communication set; ethics decision rationale; competency self-assessment update & Underlying competencies evidence \\
M09 & Five perspective-specific artifact templates; comparative reflection guide & Perspectives studio \\
M10 & Technique-selection rationale; pre-capstone study plan & Integration lab \\
M11 & Capstone deliverable templates C-01 through C-11; defense pack outline; reflection essay frame & Capstone \\
\bottomrule
\end{tabularx}

% =====================================================================
\FrontMatterChapter{Appendix C --- Instructor Quick Reference}

\renewcommand{\arraystretch}{1.18}
\rowcolors{2}{afrowalt}{white}
\begin{tabularx}{\linewidth}{@{}R{2.4cm} R{3.4cm} >{\RaggedRight\arraybackslash}X@{}}
\rowcolor{afnavy}
\multicolumn{3}{l}{\color{white}\sffamily\bfseries Instructor Quick Reference}\\
\toprule
\sffamily\bfseries Phase & \sffamily\bfseries Modules & \sffamily\bfseries Instructor Focus \\
\midrule
Foundations         & M01            & BACCM fluency; ethics tone; competency baseline.\\
Planning            & M02            & Defend trade-offs; police form-filling; surface stakeholder assumptions.\\
Discovery           & M03            & Facilitation craft; debrief recordings; resist tool-first thinking.\\
Strategy \& Need    & M04            & Police current-state evidence; challenge SMART objectives; structured decision analysis.\\
Specification       & M05            & Reconcile models with text; protect time; insist on V\&V.\\
Lifecycle           & M06            & Tie back to M04/M05 outputs; use prioritization debates as negotiation practice.\\
Evaluation          & M07            & Push past project closeout; identify enterprise (not just solution) limitations.\\
Competencies        & M08            & Behavioral focus; rubric visibility; avoid praise inflation.\\
Perspectives        & M09            & Demand fidelity; bring guest practitioners; avoid ``modern delivery stew''.\\
Integration Lab     & M10            & Coach by question; do not teach; protect coaching time.\\
Capstone            & M11            & Panel coordination; externals on $\geq$25\% of defenses; protect mid-cycle change event.\\
\bottomrule
\end{tabularx}

\vfill
\begin{infobox}[End of Curriculum]
This curriculum and its companion artifacts (rubrics, templates, case bundle, lesson plans) constitute Version 1.0. Revisions follow the cadence defined in \S33. Distribute alongside the cohort's signed program agreement, the published rubrics, and the capstone case bundle.
\end{infobox}

\end{document}
