\documentclass[11pt]{article}

% ---------- Safe, common packages ----------
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{setspace}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% ---------- Simple helpers ----------
\setlist[itemize]{itemsep=2pt, topsep=4pt}
\setlist[enumerate]{itemsep=4pt, topsep=6pt}
\newcommand{\Section}[1]{\vspace{0.6em}\noindent\textbf{\Large #1}\par\vspace{0.25em}}
\newcommand{\Subsection}[1]{\vspace{0.4em}\noindent\textbf{\large #1}\par\vspace{0.2em}}
\newcommand{\meta}[2]{\noindent\textbf{#1:} #2\par}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\begin{document}

\begin{center}
  {\LARGE \textbf{Creamy Potato Celery Soup}}\\[2pt]
  \small Updated: May 21, 2024 \quad\textbullet\quad Written by: Caroline Phelps
\end{center}

\Section{Overview}
Hearty yet refined, this velvety cream of potato and celery soup is a nostalgic, nourishing bowl designed for cozy nights. Blended potatoes create the creaminess (no heavy cream required), while pan-fried garlic and onions build depth. Finish with tender bites of celery and a sprinkle of scallions.

\Section{At a Glance}
\begin{tabularx}{\textwidth}{@{} l L @{}}
\toprule
\textbf{Yield} & 4 servings \\
\textbf{Diet} & Vegetarian (vegan option with oat milk) \\
\textbf{Prep Time} & 15 minutes \\
\textbf{Cook Time} & 35 minutes \\
\textbf{Total Time} & 50 minutes \\
\bottomrule
\end{tabularx}

\Section{Ingredients}
\begin{itemize}
  \item 2 cloves garlic, minced
  \item 1 medium onion, finely chopped
  \item 2 tablespoons extra-virgin olive oil
  \item 2 medium potatoes, peeled and roughly chopped (Yukon Gold preferred; Russet okay)
  \item 6 celery stalks, divided (4 for simmering, 2 for celery bits)
  \item 6 cups vegetable broth \emph{(or low-sodium chicken broth)}
  \item 1\,$\tfrac{1}{2}$ cups 2\% milk \emph{or} unsweetened oat milk (vegan)
  \item 4 green onions (scallions), thinly sliced, for garnish
  \item Salt and freshly ground black pepper, to taste
\end{itemize}

\Section{Expert Tip}
If your celery is fibrous, run a vegetable peeler from tip to base to remove the tough outer strings. The interior is tender and will blend silkier.

\Section{Instructions}
\begin{enumerate}
  \item \textbf{Boil the potatoes.} In a medium pot, cover the chopped potatoes with water and bring to a boil. Cook until fully tender, 7--10 minutes. Drain and set aside.
  \item \textbf{Sauté aromatics.} In a large pot over medium-high heat, warm the olive oil. Add garlic and onion; sauté 2--3 minutes until onions are translucent.
  \item \textbf{Add celery (first 4 stalks).} Roughly chop 4 stalks and add to the pot. Cook 2 minutes, stirring.
  \item \textbf{Simmer with broth.} Pour in the vegetable broth, stir, and bring to a boil. Reduce to a gentle simmer, cover, and cook 20 minutes until celery is tender.
  \item \textbf{Tender celery bits.} In a separate small pot, bring salted water to a boil. Slice the remaining 2 celery stalks into small pieces; simmer 10--15 minutes until tender but not falling apart. Drain and set aside.
  \item \textbf{Blend.} Transfer the simmered soup (broth, celery, onions, garlic, and cooked potatoes) to a blender in batches and blend until silky. Return to the pot.
  \item \textbf{Finish.} Stir in the milk (or oat milk) and the reserved tender celery bits. Warm gently. Do not boil after adding milk.
  \item \textbf{Season \& serve.} Season to taste with salt and pepper. Ladle into bowls and top with sliced green onions. Serve hot (crushed Saltines optional, highly encouraged).
\end{enumerate}

\Section{Recipe Variations}
\begin{itemize}
  \item \textbf{Make it spicy:} Add $\tfrac{1}{4}$ teaspoon red pepper flakes with the onions and garlic.
  \item \textbf{Use chicken broth:} Low-sodium chicken broth adds cozy, savory depth (great in winter).
  \item \textbf{Umami garnish:} Top with diced, caramelized mushrooms.
  \item \textbf{Crouton crunch:} Finish with store-bought or homemade croutons.
\end{itemize}

\Section{Storage}
\begin{itemize}
  \item Cool to room temperature; refrigerate in an airtight container up to 5 days.
  \item Reheat gently on the stovetop or in the microwave.
  \item \textbf{Freezing:} Dairy-based versions may separate or turn gummy when frozen. If using cow's milk, freezing is not recommended (though technically possible up to 3 months). Oat-milk versions generally freeze better.
\end{itemize}

\Section{What to Serve With Creamy Soups}
\begin{itemize}
  \item Niçoise salad
  \item Cucumber sandwiches with tzatziki
  \item Strawberry--tomatillo--watermelon salad
  \item Tuna sandwich
  \item Tempeh katsu sandwich
\end{itemize}

\Section{Frequently Asked Questions}
\Subsection{Why add milk at the end?}
Milk can scorch or boil over if heated too hard. Stirring it in at the end gives creaminess without burning or messy boil-overs.

\Subsection{Can I use an immersion blender?}
Yes. Blend in the pot for easy cleanup. For the silkiest texture, blend a little longer to catch any lingering bits.

\Section{Notes}
For a richer finish without cream, you can blend in a small knob of butter (or vegan butter) just before serving. Adjust thickness with additional broth or milk to taste.

\Section{Nutrition (Approximate, per serving; 4 servings)}
\begin{tabular}{@{}ll@{}}
\toprule
Calories & 227 \\
Sugar & 11.6 g \\
Sodium & 964.3 mg \\
Total Fat & 9.2 g \\
\quad Saturated Fat & 2.2 g \\
\quad Trans Fat & 0.1 g \\
Carbohydrates & 31.9 g \\
Fiber & 4.7 g \\
Protein & 6 g \\
Cholesterol & 7.4 mg \\
\bottomrule
\end{tabular}

\vfill
\begin{center}
  \footnotesize
  \textit{Description: Silky, comforting cream of potato and celery soup.} \\
  \textit{Source adaptation: Home kitchen notes.}
\end{center}

\end{document}

