\documentclass[11pt]{article}

% ---------- Safe, common packages ----------
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{setspace}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% ---------- Simple helpers ----------
\setlist[itemize]{itemsep=2pt, topsep=4pt}
\setlist[enumerate]{itemsep=4pt, topsep=6pt}
\newcommand{\Section}[1]{\vspace{0.6em}\noindent\textbf{\Large #1}\par\vspace{0.25em}}
\newcommand{\Subsection}[1]{\vspace{0.4em}\noindent\textbf{\large #1}\par\vspace{0.2em}}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\begin{document}

\begin{center}
  {\LARGE \textbf{The Best Cinnamon Rolls You'll Ever Eat}}\\[4pt]
  \small Big, fluffy, soft, and absolutely delicious. Options for overnight, make-ahead, and freezing.
\end{center}

\Section{Overview}
These cinnamon rolls are extra soft and fluffy thanks to bread flour, with a rich cinnamon--brown sugar swirl and a vanilla cream cheese frosting that melts into every layer.

\Section{At a Glance}
\begin{tabularx}{\textwidth}{@{} l L @{}}
\toprule
\textbf{Yield} & 9 rolls (1 \,9-inch round or square pan) \\
\textbf{Prep Time} & 2 hours (includes rises) \\
\textbf{Bake Time} & 20--25 minutes \\
\textbf{Total Time} & About 2 hours 20 minutes \\
\bottomrule
\end{tabularx}

\Section{Ingredients}

\Subsection{Dough}
\begin{itemize}
  \item $\tfrac{3}{4}$ cup (180 g) milk (whole or 2\% preferred), warmed to about $115^\circ$F
  \item 2\,$\tfrac{1}{4}$ teaspoons (one 1/4-oz packet) quick-rise/instant \emph{or} active dry yeast
  \item $\tfrac{1}{4}$ cup (50 g) granulated sugar
  \item 4 tablespoons (57 g) salted butter, melted and cooled
  \item 1 large egg \emph{plus} 1 large egg yolk, at room temperature
  \item 3 cups (360 g) bread flour, plus more as needed
  \item $\tfrac{3}{4}$ teaspoon kosher salt
  \item Extra-virgin olive oil, for greasing the bowl
\end{itemize}

\Subsection{Filling}
\begin{itemize}
  \item 4 tablespoons (57 g) salted butter, softened (room temperature)
  \item $\tfrac{2}{3}$ cup (142 g) packed dark brown sugar
  \item 1\,$\tfrac{1}{2}$ tablespoons ground cinnamon
\end{itemize}

\Subsection{Cream Cheese Frosting}
\begin{itemize}
  \item 4 oz (112 g) cream cheese, softened
  \item $\tfrac{3}{4}$ cup (85 g) powdered sugar
  \item 3 tablespoons salted butter, softened
  \item $\tfrac{1}{2}$ teaspoon vanilla extract
\end{itemize}

\Section{Instructions}

\begin{enumerate}
  \item \textbf{Warm milk \& start yeast.} Warm the milk to about $115^\circ$F (warm bath temperature). Add to the bowl of a stand mixer. Sprinkle yeast over the milk. \emph{If using active dry yeast:} add the sugar now and let sit until foamy, about 5 minutes.
  \item \textbf{Mix the dough.} Add the sugar (if not already added), melted butter, egg, and egg yolk. Mix briefly to combine. Remove the bowl and stir in flour and salt with a wooden spoon until a shaggy dough forms.
  \item \textbf{Knead.} Fit the dough hook and knead on medium speed until the dough is soft and slightly tacky, 8--10 minutes. If it clings to the bottom, add 1--3 tablespoons flour, 1 tbsp at a time. (\emph{By hand:} knead on a well-floured surface 8--10 minutes.)
  \item \textbf{First rise.} Lightly oil a large bowl. Shape the dough into a ball, place in the bowl, cover with plastic wrap and a warm towel, and let rise until doubled, 1 to 1\,$\tfrac{1}{2}$ hours. It’s ready when an indentation from your fingertip does not immediately spring back.
  \item \textbf{Roll out.} On a generously floured surface, roll the dough to a $10\times14$-inch rectangle.
  \item \textbf{Fill.} Spread the softened butter over the dough, leaving a $\tfrac{1}{4}$-inch margin along one short edge. In a small bowl, mix brown sugar and cinnamon, then sprinkle evenly over the butter and gently rub it in.
  \item \textbf{Roll \& trim.} Starting from the short side without the margin, roll up tightly. Pinch to seal the seam and place seam-side down. Trim about $\tfrac{1}{2}$ inch from both ends (little filling there).
  \item \textbf{Slice.} Using unflavored dental floss or a serrated knife, cut nine 1-inch rolls.
  \item \textbf{Second rise.} Line a 9-inch round or square pan with parchment. Arrange rolls cut-side up, cover, and let rise until puffed, doubled, and touching, 45--60 minutes. Meanwhile, preheat oven to $350^\circ$F (175$^\circ$C).
  \item \textbf{Bake.} Bake 20--25 minutes, until the edges are lightly golden. Do not overbake; this keeps the centers soft. Cool 10 minutes.
  \item \textbf{Frost.} Beat cream cheese, powdered sugar, butter, and vanilla (stand mixer with whisk or by hand) until smooth and fluffy, 1--2 minutes. Spread over warm rolls and serve.
\end{enumerate}

\Section{Key Tips for Perfect Rolls}
\begin{itemize}
  \item \textbf{Milk temperature:} $105$--$115^\circ$F. Too hot can kill yeast.
  \item \textbf{Use bread flour:} Higher protein means softer, fluffier rolls.
  \item \textbf{Room-temp ingredients:} Egg and filling butter at room temp. Use softened (not melted) butter in the filling so it stays put.
  \item \textbf{Rub in the sugar:} After sprinkling the brown sugar, gently rub it into the buttered dough.
  \item \textbf{Flour the surface \& pin:} Prevents sticking during rolling.
\end{itemize}

\Section{Overnight, Make-Ahead \& Freezing}

\Subsection{Overnight (Refrigerator) Method}
After step 8 (slicing), arrange rolls in the lined pan, cover tightly, and refrigerate overnight. Before baking, bring to room temp and let proof 45--60 minutes until puffy. Bake as directed.

\Subsection{Freeze Before Baking}
After slicing, place rolls in a pan, cover well, and freeze. Thaw in the refrigerator overnight, then proof in a warm spot 30--60 minutes before baking.

\Subsection{Freeze After Baking}
Cool completely. Freeze the whole pan or individual rolls (ideally before frosting). Thaw at room temp; warm at $350^\circ$F for 10--15 minutes or microwave individual rolls 20--30 seconds. Frost after reheating. (Frosting also freezes well.)

\Section{Storage}
Cover and keep at room temperature up to 2 hours, then refrigerate up to 5 days. Reheat gently to serve warm and gooey.

\Section{Yeast \& Flour Notes}
\begin{itemize}
  \item \textbf{Instant/quick-rise yeast:} No proofing required (though you may still sprinkle over warm milk).
  \item \textbf{Active dry yeast:} Proof in warm milk with sugar until foamy before mixing. Rise times may be longer.
  \item \textbf{All-purpose flour:} Works in a pinch, but bread flour yields the fluffiest texture.
\end{itemize}

\Section{Helpful Tools}
Stand mixer with dough hook (optional), wooden spoon, rolling pin, unflavored dental floss or serrated knife, 9-inch pan, parchment paper.

\vfill
\begin{center}
  \footnotesize
  \textit{Dairy-free option:} Use almond milk, vegan butter, and vegan cream cheese.\\
  \textit{Make it yours:} Add nuts, raisins, or warm spices (e.g., a pinch of nutmeg or pumpkin pie spice) to the filling.
\end{center}

\end{document}

