\documentclass[10pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[letterpaper,margin=0.55in]{geometry}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{ragged2e}
\usepackage{enumitem}
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.35em}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=0.15em,topsep=0.2em}
\emergencystretch=1.5em
\sloppy

\definecolor{TitleBlue}{HTML}{163A5F}
\definecolor{RuleBlue}{HTML}{DCE6F2}
\definecolor{BoxBlue}{HTML}{F5F8FC}
\definecolor{HeaderBlue}{HTML}{EAF1F8}
\definecolor{DoneBg}{HTML}{E8F5E9}
\definecolor{DoneFg}{HTML}{1B5E20}
\definecolor{ProgBg}{HTML}{E8F0FE}
\definecolor{ProgFg}{HTML}{174EA6}
\definecolor{NotBg}{HTML}{F3F4F6}
\definecolor{NotFg}{HTML}{374151}

\newcolumntype{L}[1]{>{\RaggedRight\arraybackslash}p{#1}}
\newcommand{\statusdone}{\colorbox{DoneBg}{\textcolor{DoneFg}{\strut\hspace{0.35em}Done\hspace{0.35em}}}}
\newcommand{\statusprogress}{\colorbox{ProgBg}{\textcolor{ProgFg}{\strut\hspace{0.35em}In Progress\hspace{0.35em}}}}
\newcommand{\statusnot}{\colorbox{NotBg}{\textcolor{NotFg}{\strut\hspace{0.35em}Not Started\hspace{0.35em}}}}
\newcommand{\labelcell}[1]{\textcolor{TitleBlue}{\textbf{#1}}}

\begin{document}

{\LARGE\bfseries\textcolor{TitleBlue}{AuditPack Startup Roadmap}}\\[-0.1em]
{\normalsize\textcolor{TitleBlue}{SBA-aligned execution view for AuditPack for GHAS --- audit-ready compliance evidence automation for regulated enterprises}}

\vspace{0.45em}
{\color{RuleBlue}\rule{\textwidth}{0.9pt}}

\vspace{0.2em}
\begingroup
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.22}
\small
\rowcolors{2}{white}{BoxBlue}
\begin{tabularx}{\textwidth}{L{0.18\textwidth}X L{0.20\textwidth}X}
\rowcolor{HeaderBlue}
\labelcell{Stage} & Validation + MVP definition + productized service pre-launch & \labelcell{Current Focus} & Build the CLI-first MVP, package the service offer, and secure the first paid GHAS audit-readiness engagement \\
\labelcell{Target Customer} & Heads of AppSec or Security Engineering at regulated SaaS companies using GitHub Enterprise Cloud and GHAS, especially lean teams under audit pressure & \labelcell{Revenue Model} & Fixed-price Audit-Readiness Sprint, quarterly evidence retainer, then subscription SaaS for continuous evidence generation \\
\end{tabularx}
\endgroup

\vspace{0.45em}
\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.16}
\begin{longtable}{@{}L{0.045\textwidth}L{0.155\textwidth}L{0.125\textwidth}L{0.395\textwidth}L{0.18\textwidth}@{}}
\rowcolor{HeaderBlue}
\textbf{Step} & \textbf{SBA Step} & \textbf{Status} & \textbf{What it means for AuditPack} & \textbf{Next action} \\
\midrule
\endfirsthead
\rowcolor{HeaderBlue}
\textbf{Step} & \textbf{SBA Step} & \textbf{Status} & \textbf{What it means for AuditPack} & \textbf{Next action} \\
\midrule
\endhead
1 & Conduct market research & \statusprogress & Validate that regulated-enterprise AppSec leaders will pay for faster, defensible GHAS evidence generation; confirm the most acute pain around evidence assembly, dismissal defense, control mapping, and auditor scrutiny. & Run focused discovery and outbound with target buyers; capture repeated buying triggers, objections, and minimum acceptable deliverables. \\

2 & Write your business plan & \statusprogress & Convert the current concept into an operator-grade launch plan for a service-first entry and SaaS follow-on, including ICP, offer structure, delivery workflow, pricing logic, and MVP scope. & Finalize a concise operating plan with offer definition, delivery model, milestones, startup costs, and first-customer path. \\

3 & Fund your business & \statusprogress & For AuditPack, this is primarily a bootstrap plan: use paid audit-readiness work to finance the CLI-first MVP and avoid premature outside capital before problem-solution fit is proven. & Set founder budget, revenue targets, and a simple cash plan tied to the first sprint sale and near-term build costs. \\

4 & Pick your business location & \statusnot & For a remote GHAS-focused service business, this means deciding the formation state, operating address model, and any state-specific implications for contracts, taxes, and future hiring. & Choose the home jurisdiction and document the remote-first operating model. \\

5 & Choose a business structure & \statusnot & Decide the legal wrapper that best supports early consulting revenue, liability separation, simple administration, and later transition toward a software product business. & Select the initial entity structure and record the decision rationale. \\

6 & Choose your business name & \statusprogress & The working name AuditPack for GHAS is established, but this step is not fully complete until naming risk is reduced across brand use, entity naming, domain availability, and trademark exposure. & Confirm the final operating name, domain plan, and basic clearance checks. \\

7 & Register your business & \statusnot & Make the business legally operable under the chosen structure and jurisdiction so contracts, invoicing, and customer onboarding can proceed on a formal footing. & Prepare and submit the required formation and registration filings. \\

8 & Get federal and state tax IDs & \statusnot & Obtain the EIN and any required state tax registrations needed to bank, invoice, and operate cleanly as a real business rather than an informal project. & Apply for the EIN and determine whether any state tax IDs are required. \\

9 & Apply for licenses and permits & \statusnot & AuditPack is likely a low-permit professional service business, but the founder still needs a defensible check of state and local requirements rather than assuming none apply. & Complete a jurisdiction-specific review and record any required licenses, permits, or exemptions. \\

10 & Open a business bank account & \statusnot & Separate founder and business cash handling so the first paid Audit-Readiness Sprint, recurring retainers, and future SaaS revenue can be tracked cleanly from day one. & Open the account once registration and tax ID prerequisites are in hand. \\
\bottomrule
\end{longtable}
\endgroup

\vspace{0.2em}
{\footnotesize\textit{Note: This roadmap mirrors the current SBA ten-step startup sequence and has been tailored to AuditPack using prior conversation context. Statuses are intentionally conservative and reflect validated progress rather than aspirational completion.}}

\end{document}
