\documentclass[10pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=0.65in]{geometry}
\usepackage[table]{xcolor}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{pdflscape}
\usepackage{enumitem}
\usepackage{url}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\renewcommand{\arraystretch}{1.18}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=3pt}

\definecolor{HeaderGray}{gray}{0.92}
\definecolor{RuleGray}{gray}{0.35}

\newcolumntype{L}[1]{>{\RaggedRight\arraybackslash}p{#1}}
\newcommand{\Status}[1]{\textbf{#1}}

\begin{document}

{\LARGE \textbf{AETHERFORGE ENGINE Startup Roadmap}}\\
{\large SBA-aligned founder handout using the live SBA 10-step startup sequence}

\textbf{Purpose.} This working roadmap translates the U.S. Small Business Administration's current ``10 steps to start your business'' sequence into founder actions for \textbf{AETHERFORGE ENGINE}. The business is positioned as a modular, performance-oriented engine and tooling stack for technically capable indie developers, small studios, and simulation-focused builders who want disciplined native desktop and WebAssembly delivery from a shared codebase. Current evidence points to a \textbf{concept validation / architecture definition / pre-MVP} stage: the engine direction, technical stack, and differentiation are defined, but the first commercial wedge, formal business setup, and proof of traction are still being established.

\begin{landscape}
\footnotesize
\begin{longtable}{L{0.45in}L{1.35in}L{1.95in}L{0.80in}L{1.85in}L{1.85in}}
\rowcolor{HeaderGray}
\textbf{Step} & \textbf{SBA Step} & \textbf{AETHERFORGE-Specific Interpretation} & \textbf{Current Status} & \textbf{Why This Status Fits} & \textbf{Next Concrete Action} \\
\toprule
\endfirsthead

\rowcolor{HeaderGray}
\textbf{Step} & \textbf{SBA Step} & \textbf{AETHERFORGE-Specific Interpretation} & \textbf{Current Status} & \textbf{Why This Status Fits} & \textbf{Next Concrete Action} \\
\toprule
\endhead

\midrule
\multicolumn{6}{r}{\textit{Continued on next page}}\\
\endfoot

\bottomrule
\endlastfoot

1 & Conduct market research & Validate which buyer will pay first: an indie studio building a browser-deployable game, a simulation-oriented technical team, or a developer-tools customer buying the engine foundation plus workflow. Confirm demand for the native + WASM value proposition, portability story, and ownership/control positioning. & \Status{In Progress} & The target customer, problem, and value proposition are already articulated, which is stronger than an idea-only stage. However, the founder's own stated priority is still to lock the first product wedge and exact initial buyer/use case, so market evidence is not yet complete. & Run a focused discovery sprint with a short interview guide, speak with 10--15 technically credible prospects, and decide on one initial paying wedge with explicit problem, buyer, and purchase trigger. \\

2 & Write your business plan & Convert the current concept into a practical founder plan: positioning, commercial wedge, product scope, pricing logic, go-to-market path, delivery model, milestones, and near-term financial assumptions. & \Status{In Progress} & There is already meaningful business-planning material: working idea, problem, target customer, value proposition, stage, focus, and immediate priorities are defined. What is missing is a consolidated operating plan that ties product scope, customer acquisition, pricing, and milestones into one business document. & Draft a lean operating plan with sections for wedge, MVP scope, packaging, pricing hypothesis, customer acquisition approach, 12-month milestone plan, and core technical/business risks. \\

3 & Fund your business & Decide how the engine will be financed through the pre-revenue phase: founder-funded, consulting-supported, services-led cash flow, grants, or outside capital later. Estimate startup costs for tools, hosting, build infrastructure, business formation, legal review, and runway. & \Status{Not Started} & No evidence shows a startup cost model, financing plan, runway estimate, or funding strategy for AETHERFORGE ENGINE. The current context is oriented around product and architecture, not capitalization. & Build a 12-month startup cost sheet and choose a funding approach for the first phase, with an explicit rule for what will be self-funded versus deferred. \\

4 & Pick your business location & Define the operating footprint for the business: likely a remote-first software company with digital distribution, but still requiring a legal home state, banking jurisdiction, tax nexus assumptions, and a practical business mailing setup. & \Status{Not Started} & The product is software-native and can be sold online, but there is no evidence that the business jurisdiction, operating location, or state-specific implications have been chosen. & Select the formation state and business address strategy, then record the tax, registration, and banking implications of that choice. \\

5 & Choose a business structure & Select the legal structure that matches a founder-led technical startup, liability needs, tax treatment, and the possibility of evolving from productized services into software tooling. & \Status{Not Started} & No evidence indicates that an LLC, corporation, or other structure has been selected or compared. This is still an open legal/operational decision. & Compare single-member LLC versus corporation for the intended path, then choose the structure that best fits early-stage liability protection, tax simplicity, and future flexibility. \\

6 & Choose your business name & Finalize and protect the AETHERFORGE ENGINE brand for use in company formation, website/domain footprint, source repositories, documentation, and future tooling products. & \Status{Done} & The name \textbf{AETHERFORGE ENGINE} is already selected and consistently used as the business identity. Prior context also showed this name as the favored engine brand direction, so the naming decision itself is effectively complete. & Perform trademark, domain, and handle clearance as a protection step, then standardize the naming system across the website, repositories, sample projects, and founder materials. \\

7 & Register your business & Legally form the entity and secure the registrations needed to operate under the AETHERFORGE ENGINE name. & \Status{Not Started} & There is no evidence that the business has been legally formed or that formal registration steps have begun. & File the formation paperwork, reserve or register the business name as needed, and create a registration checklist with filing confirmations and renewal dates. \\

8 & Get federal and state tax IDs & Obtain the EIN and any required state tax registrations needed to open accounts, handle taxes, and formalize operations. & \Status{Not Started} & No evidence indicates an EIN or state tax ID has been obtained. Those activities generally follow entity formation and banking readiness, neither of which is yet evidenced here. & After formation, apply for an EIN immediately and document any state tax registrations that are required for the chosen jurisdiction. \\

9 & Apply for licenses and permits & Determine whether any local business licenses, home-office registrations, assumed-name filings, or software-related compliance registrations are required for the chosen jurisdiction and operating model. & \Status{Not Started} & There is no evidence of a compliance check for permits or local licensing. Even if requirements are light for a software startup, they still need to be verified rather than assumed away. & Create a jurisdiction-specific checklist covering state, county, and city requirements, and explicitly mark each item as required, not required, or pending confirmation. \\

10 & Open a business bank account & Establish clean financial separation for founder spending, startup costs, revenue collection, contractor payments, and accounting. & \Status{Not Started} & No evidence suggests a dedicated business bank account has been opened. Because formal registration and EIN steps are not yet evidenced, banking is appropriately still pending. & Once formation and EIN are complete, open a business checking account and define a minimal chart-of-accounts plus bookkeeping workflow for startup expenses and revenue. \\

\end{longtable}
\end{landscape}

\normalsize
\textbf{Immediate Founder Focus}
\begin{itemize}
  \item \textbf{Lock the first commercial wedge.} Decide exactly who buys first, what job AETHERFORGE solves first, and what proof would make that buyer commit.
  \item \textbf{Build the technical spine.} Deliver the core runtime, thin platform layer, rendering abstraction, starter asset pipeline, build presets, and repeatable native + WASM workflow.
  \item \textbf{Produce a credible proof point.} Ship a polished sample project or vertical slice that demonstrates the engine's portability, rendering path, and developer experience.
\end{itemize}

\vspace{0.25em}
{\small \textit{Working founder note.} This roadmap is a live founder document aligned to the SBA's current 10-step startup sequence and tailored to AETHERFORGE ENGINE's present pre-MVP stage. Source sequence: U.S. Small Business Administration, \textit{10 steps to start your business}, accessed May 6, 2026, \url{https://www.sba.gov/business-guide/10-steps-start-your-business}.}

\end{document}



