
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{Administration and IT Security System}{Run the back office, protect company information, and maintain reliable digital operations.}
\sectionbar{Purpose}
Provide the administrative backbone for documents, scheduling, access, backups, and baseline cybersecurity protections.
\sectionbar{Core Capabilities}
\begin{itemize}
  \item Document management and retention
  \item Calendar, scheduling, and administrative coordination
  \item User access management
  \item Device, account, and software inventory
  \item Backup and recovery routines
  \item Cybersecurity awareness and incident escalation
\end{itemize}

\sectionbar{Minimum SOP Library}
\begin{itemize}[label={}]
  \item \checkbox Document naming and retention SOP
  \item \checkbox Access request and removal SOP
  \item \checkbox Backup verification SOP
  \item \checkbox Device onboarding SOP
  \item \checkbox Security incident escalation SOP
  \item \checkbox Vendor account review SOP
\end{itemize}

\sectionbar{Recommended Tooling Capabilities}
\begin{itemize}
  \item Cloud document repository
  \item Password manager
  \item Identity and access management features
  \item Backup solution
  \item Endpoint security tooling
  \item Ticket or request tracker
\end{itemize}

\sectionbar{KPI Scorecard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.34\textwidth}>{\raggedright\arraybackslash}p{0.56\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries KPI} & \textcolor{white}{\bfseries Why It Matters} \\
\toprule
Backup success rate & Shows whether the system is producing reliable, measurable business value. \\
Account review completion rate & Shows whether the system is producing reliable, measurable business value. \\
Security training completion & Shows whether the system is producing reliable, measurable business value. \\
Mean time to revoke access & Shows whether the system is producing reliable, measurable business value. \\
Device inventory accuracy & Shows whether the system is producing reliable, measurable business value. \\
Incident response time & Shows whether the system is producing reliable, measurable business value. \\
\bottomrule
\end{tabularx}}

\sectionbar{System Handoffs and Dependencies}
\begin{itemize}
  \item HR triggers onboarding and offboarding access changes.
  \item Operations depends on reliable systems and document access.
  \item Leadership receives risk and continuity reporting.
\end{itemize}

\sectionbar{Key Operating Risks}
\begin{itemize}
  \item Orphaned accounts expose data.
  \item Untested backups fail during incidents.
  \item Uncontrolled document storage creates compliance and confidentiality risk.
\end{itemize}

\sectionbar{Implementation Checklist}
\begin{itemize}[label={}]
  \item \checkbox Assign a named system owner and backup owner.
  \item \checkbox Document the current workflow and identify recurring decisions.
  \item \checkbox Create or update the minimum SOPs listed above.
  \item \checkbox Select the smallest tool set needed to execute the SOPs consistently.
  \item \checkbox Define KPI targets, review cadence, and evidence requirements.
  \item \checkbox Create handoff criteria for upstream and downstream systems.
  \item \checkbox Review the first month of KPI results and add improvement actions to the backlog.
\end{itemize}

\sectionbar{Definition of Done}
This system is operational when work can be performed consistently by trained staff, handoffs are explicit, KPIs are reviewed on a recurring cadence, and exceptions are escalated through a documented path.
\vfill\small Source basis: synthesized from the small business systems brief supplied by the requester.\par
\end{document}


