
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{HR and People Management System}{Hire, onboard, train, and retain the people needed to operate and scale the business.}
\sectionbar{Purpose}
Create consistent people processes for hiring, onboarding, payroll coordination, training, performance management, and compliance.
\sectionbar{Core Capabilities}
\begin{itemize}
  \item Role definition and hiring workflow
  \item Onboarding and training plans
  \item Payroll and benefits coordination
  \item Performance check-ins and feedback routines
  \item Policy acknowledgement and compliance tracking
  \item Offboarding and knowledge transfer
\end{itemize}

\sectionbar{Minimum SOP Library}
\begin{itemize}[label={}]
  \item \checkbox Hiring requisition SOP
  \item \checkbox Interview and selection SOP
  \item \checkbox Employee onboarding SOP
  \item \checkbox Training completion SOP
  \item \checkbox Performance review SOP
  \item \checkbox Offboarding SOP
\end{itemize}

\sectionbar{Recommended Tooling Capabilities}
\begin{itemize}
  \item Applicant tracking board
  \item Employee record repository
  \item Payroll system
  \item Training tracker
  \item Performance review template
  \item Policy acknowledgement form
\end{itemize}

\sectionbar{KPI Scorecard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.34\textwidth}>{\raggedright\arraybackslash}p{0.56\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries KPI} & \textcolor{white}{\bfseries Why It Matters} \\
\toprule
Time to fill & Shows whether the system is producing reliable, measurable business value. \\
New hire onboarding completion & Shows whether the system is producing reliable, measurable business value. \\
Training completion rate & Shows whether the system is producing reliable, measurable business value. \\
Retention rate & Shows whether the system is producing reliable, measurable business value. \\
Performance review completion & Shows whether the system is producing reliable, measurable business value. \\
Offboarding checklist completion & Shows whether the system is producing reliable, measurable business value. \\
\bottomrule
\end{tabularx}}

\sectionbar{System Handoffs and Dependencies}
\begin{itemize}
  \item Administration and IT Security executes access provisioning and removal.
  \item Operations receives trained staff and role capacity.
  \item Leadership receives workforce capacity and culture indicators.
\end{itemize}

\sectionbar{Key Operating Risks}
\begin{itemize}
  \item Unstructured onboarding slows productivity.
  \item Inconsistent role expectations create performance disputes.
  \item Incomplete offboarding creates security and knowledge-loss exposure.
\end{itemize}

\sectionbar{Implementation Checklist}
\begin{itemize}[label={}]
  \item \checkbox Assign a named system owner and backup owner.
  \item \checkbox Document the current workflow and identify recurring decisions.
  \item \checkbox Create or update the minimum SOPs listed above.
  \item \checkbox Select the smallest tool set needed to execute the SOPs consistently.
  \item \checkbox Define KPI targets, review cadence, and evidence requirements.
  \item \checkbox Create handoff criteria for upstream and downstream systems.
  \item \checkbox Review the first month of KPI results and add improvement actions to the backlog.
\end{itemize}

\sectionbar{Definition of Done}
This system is operational when work can be performed consistently by trained staff, handoffs are explicit, KPIs are reviewed on a recurring cadence, and exceptions are escalated through a documented path.
\vfill\small Source basis: synthesized from the small business systems brief supplied by the requester.\par
\end{document}