
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{KPI and Continuous Improvement Layer}{A practical measurement and review cadence for improving small business systems.}
\sectionbar{Purpose}
The KPI and continuous improvement layer converts operating data into decisions. It defines what is measured, who owns each metric, how often metrics are reviewed, and how improvement work is prioritized.
\sectionbar{Scorecard Design Rules}
\begin{itemize}
  \item Use a small number of actionable KPIs per system.
  \item Assign one accountable owner to every KPI.
  \item Define the calculation method and data source before reporting the KPI.
  \item Review trend direction, not isolated point-in-time numbers.
  \item Convert KPI gaps into specific improvement actions with owners and due dates.
\end{itemize}

\sectionbar{Core KPI Register}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}p{0.34\textwidth}>{\raggedright\arraybackslash}p{0.32\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries System} & \textcolor{white}{\bfseries Example KPIs} & \textcolor{white}{\bfseries Primary Review Question} \\
\toprule
Marketing & Qualified leads, lead source conversion, cost per lead & Are we attracting the right prospects at an acceptable cost? \\
Sales & Win rate, sales cycle time, average deal value & Are we converting qualified demand into profitable customers? \\
Operations & On-time delivery, cycle time, rework rate & Are we fulfilling commitments consistently and efficiently? \\
Financial Management & Cash on hand, AR aging, margin, budget variance & Are we protecting cash flow and profitability? \\
Administration and IT Security & Backup success, access review completion, incident response time & Are the back office and digital systems reliable and secure? \\
HR and People Management & Onboarding completion, training completion, retention & Do we have the people capacity to execute the plan? \\
Leadership and Strategy & Strategic initiative completion, scorecard health, risk status & Are we making disciplined decisions and improving the business? \\
\bottomrule
\end{tabularx}}

\sectionbar{Continuous Improvement Backlog}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.65\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Field} & \textcolor{white}{\bfseries Description} \\
\toprule
Issue or opportunity & Clear statement of the process gap, risk, or improvement. \\
Source KPI & Metric or observation that triggered the item. \\
System owner & Owner accountable for resolution. \\
Impact & Expected value to customer experience, cost, speed, quality, cash, or risk. \\
Effort & Estimated complexity and time required. \\
Priority & Ranked using impact, urgency, dependency, and risk. \\
Next action & Specific action with owner and date. \\
\bottomrule
\end{tabularx}}

\sectionbar{Review Cadence Checklist}
\begin{itemize}[label={}]
  \item \checkbox Weekly review captures blockers, handoff failures, customer issues, and urgent operational decisions.
  \item \checkbox Monthly review updates system KPIs and selects improvement actions.
  \item \checkbox Quarterly review validates strategy, budget, capacity, risk, and tooling decisions.
  \item \checkbox Annual review refreshes the operating model, owners, SOPs, tools, and strategic plan.
\end{itemize}

\end{document}