
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{Marketing and Lead Generation System}{Attract qualified prospects, build awareness, and create a predictable lead pipeline.}
\sectionbar{Purpose}
Create a repeatable mechanism for attracting the right market segment, capturing lead information, and preparing prospects for sales conversion.
\sectionbar{Core Capabilities}
\begin{itemize}
  \item Ideal customer profile and buyer persona definition
  \item Content calendar and campaign planning
  \item Website, landing page, and form management
  \item Lead capture, tagging, and source attribution
  \item Email nurture sequences and social media scheduling
  \item Campaign performance review and message testing
\end{itemize}

\sectionbar{Minimum SOP Library}
\begin{itemize}[label={}]
  \item \checkbox Campaign planning SOP
  \item \checkbox Lead capture and CRM entry SOP
  \item \checkbox Content review and publishing SOP
  \item \checkbox Lead source attribution SOP
  \item \checkbox Marketing performance review SOP
\end{itemize}

\sectionbar{Recommended Tooling Capabilities}
\begin{itemize}
  \item CRM or contact database
  \item Email marketing platform
  \item Website or landing page builder
  \item Content calendar
  \item Social media scheduling tool
  \item Analytics dashboard
\end{itemize}

\sectionbar{KPI Scorecard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.34\textwidth}>{\raggedright\arraybackslash}p{0.56\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries KPI} & \textcolor{white}{\bfseries Why It Matters} \\
\toprule
New qualified leads & Shows whether the system is producing reliable, measurable business value. \\
Lead source conversion rate & Shows whether the system is producing reliable, measurable business value. \\
Website form conversion rate & Shows whether the system is producing reliable, measurable business value. \\
Email open and click rates & Shows whether the system is producing reliable, measurable business value. \\
Campaign cost per lead & Shows whether the system is producing reliable, measurable business value. \\
Marketing qualified lead to sales qualified lead rate & Shows whether the system is producing reliable, measurable business value. \\
\bottomrule
\end{tabularx}}

\sectionbar{System Handoffs and Dependencies}
\begin{itemize}
  \item Sales pipeline receives qualified leads with source, persona, and campaign context.
  \item Leadership receives campaign performance and pipeline contribution metrics.
  \item Financial management receives budget and campaign spend data.
\end{itemize}

\sectionbar{Key Operating Risks}
\begin{itemize}
  \item Unqualified lead volume creates sales noise.
  \item No attribution prevents budget decisions.
  \item Inconsistent messaging weakens brand positioning.
\end{itemize}

\sectionbar{Implementation Checklist}
\begin{itemize}[label={}]
  \item \checkbox Assign a named system owner and backup owner.
  \item \checkbox Document the current workflow and identify recurring decisions.
  \item \checkbox Create or update the minimum SOPs listed above.
  \item \checkbox Select the smallest tool set needed to execute the SOPs consistently.
  \item \checkbox Define KPI targets, review cadence, and evidence requirements.
  \item \checkbox Create handoff criteria for upstream and downstream systems.
  \item \checkbox Review the first month of KPI results and add improvement actions to the backlog.
\end{itemize}

\sectionbar{Definition of Done}
This system is operational when work can be performed consistently by trained staff, handoffs are explicit, KPIs are reviewed on a recurring cadence, and exceptions are escalated through a documented path.
\vfill\small Source basis: synthesized from the small business systems brief supplied by the requester.\par
\end{document}


