
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{Operations and Fulfillment System}{Deliver products or services consistently, on time, and at the promised quality level.}
\sectionbar{Purpose}
Translate customer commitments into repeatable work execution, quality control, inventory or resource management, and customer delivery outcomes.
\sectionbar{Core Capabilities}
\begin{itemize}
  \item Work intake and prioritization
  \item Standard operating procedures for recurring work
  \item Project or order tracking
  \item Inventory, resource, or capacity management
  \item Quality assurance checkpoints
  \item Customer delivery confirmation and issue escalation
\end{itemize}

\sectionbar{Minimum SOP Library}
\begin{itemize}[label={}]
  \item \checkbox Order or service intake SOP
  \item \checkbox Fulfillment workflow SOP
  \item \checkbox Quality control SOP
  \item \checkbox Inventory or capacity review SOP
  \item \checkbox Customer delivery confirmation SOP
  \item \checkbox Issue escalation SOP
\end{itemize}

\sectionbar{Recommended Tooling Capabilities}
\begin{itemize}
  \item Project management board
  \item Task templates
  \item Inventory tracker
  \item Documented SOP repository
  \item Quality checklist
  \item Customer support ticket queue
\end{itemize}

\sectionbar{KPI Scorecard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.34\textwidth}>{\raggedright\arraybackslash}p{0.56\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries KPI} & \textcolor{white}{\bfseries Why It Matters} \\
\toprule
On-time delivery rate & Shows whether the system is producing reliable, measurable business value. \\
Rework rate & Shows whether the system is producing reliable, measurable business value. \\
Cycle time & Shows whether the system is producing reliable, measurable business value. \\
Inventory accuracy or capacity utilization & Shows whether the system is producing reliable, measurable business value. \\
Customer satisfaction score & Shows whether the system is producing reliable, measurable business value. \\
Escalation rate & Shows whether the system is producing reliable, measurable business value. \\
\bottomrule
\end{tabularx}}

\sectionbar{System Handoffs and Dependencies}
\begin{itemize}
  \item Sales provides delivery scope, promises, and customer expectations.
  \item Financial management uses delivery milestones for billing.
  \item Leadership uses operational KPIs for capacity and growth decisions.
\end{itemize}

\sectionbar{Key Operating Risks}
\begin{itemize}
  \item Undocumented processes depend on individual memory.
  \item Overcommitment harms delivery quality.
  \item No quality gates increase rework and customer churn.
\end{itemize}

\sectionbar{Implementation Checklist}
\begin{itemize}[label={}]
  \item \checkbox Assign a named system owner and backup owner.
  \item \checkbox Document the current workflow and identify recurring decisions.
  \item \checkbox Create or update the minimum SOPs listed above.
  \item \checkbox Select the smallest tool set needed to execute the SOPs consistently.
  \item \checkbox Define KPI targets, review cadence, and evidence requirements.
  \item \checkbox Create handoff criteria for upstream and downstream systems.
  \item \checkbox Review the first month of KPI results and add improvement actions to the backlog.
\end{itemize}

\sectionbar{Definition of Done}
This system is operational when work can be performed consistently by trained staff, handoffs are explicit, KPIs are reviewed on a recurring cadence, and exceptions are escalated through a documented path.
\vfill\small Source basis: synthesized from the small business systems brief supplied by the requester.\par
\end{document}