
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{Sales and Lead Conversion System}{Convert qualified opportunities into customers through a managed funnel and consistent follow-up.}
\sectionbar{Purpose}
Move prospects from inquiry to purchase using a structured pipeline, qualification rules, pricing discipline, and documented follow-up routines.
\sectionbar{Core Capabilities}
\begin{itemize}
  \item Lead qualification and opportunity scoring
  \item Defined sales stages and exit criteria
  \item Proposal, quote, and pricing workflow
  \item Follow-up cadence and objection handling
  \item Customer handoff to operations
  \item Win/loss review and sales forecast hygiene
\end{itemize}

\sectionbar{Minimum SOP Library}
\begin{itemize}[label={}]
  \item \checkbox Lead qualification SOP
  \item \checkbox Sales pipeline update SOP
  \item \checkbox Proposal and quote approval SOP
  \item \checkbox Follow-up cadence SOP
  \item \checkbox Closed-won handoff SOP
  \item \checkbox Win/loss review SOP
\end{itemize}

\sectionbar{Recommended Tooling Capabilities}
\begin{itemize}
  \item CRM pipeline board
  \item Quote or proposal templates
  \item Calendar scheduling tool
  \item Email templates
  \item E-signature tool
  \item Sales forecast dashboard
\end{itemize}

\sectionbar{KPI Scorecard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.34\textwidth}>{\raggedright\arraybackslash}p{0.56\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries KPI} & \textcolor{white}{\bfseries Why It Matters} \\
\toprule
Lead-to-opportunity conversion rate & Shows whether the system is producing reliable, measurable business value. \\
Opportunity win rate & Shows whether the system is producing reliable, measurable business value. \\
Average sales cycle time & Shows whether the system is producing reliable, measurable business value. \\
Average deal value & Shows whether the system is producing reliable, measurable business value. \\
Follow-up SLA compliance & Shows whether the system is producing reliable, measurable business value. \\
Forecast accuracy & Shows whether the system is producing reliable, measurable business value. \\
\bottomrule
\end{tabularx}}

\sectionbar{System Handoffs and Dependencies}
\begin{itemize}
  \item Marketing supplies source-attributed qualified leads.
  \item Operations receives closed-won handoff details and delivery commitments.
  \item Financial management receives bookings, invoices, and revenue forecast data.
\end{itemize}

\sectionbar{Key Operating Risks}
\begin{itemize}
  \item Poor qualification creates delivery strain.
  \item Unapproved discounts reduce margin.
  \item Weak handoff creates customer dissatisfaction.
\end{itemize}

\sectionbar{Implementation Checklist}
\begin{itemize}[label={}]
  \item \checkbox Assign a named system owner and backup owner.
  \item \checkbox Document the current workflow and identify recurring decisions.
  \item \checkbox Create or update the minimum SOPs listed above.
  \item \checkbox Select the smallest tool set needed to execute the SOPs consistently.
  \item \checkbox Define KPI targets, review cadence, and evidence requirements.
  \item \checkbox Create handoff criteria for upstream and downstream systems.
  \item \checkbox Review the first month of KPI results and add improvement actions to the backlog.
\end{itemize}

\sectionbar{Definition of Done}
This system is operational when work can be performed consistently by trained staff, handoffs are explicit, KPIs are reviewed on a recurring cadence, and exceptions are escalated through a documented path.
\vfill\small Source basis: synthesized from the small business systems brief supplied by the requester.\par
\end{document}