
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{Small Business Systems Implementation Checklist}{A step-by-step checklist for standing up the core operating model.}
\sectionbar{Phase 1 - Baseline the Business}
\begin{itemize}[label={}]
  \item \checkbox List the seven core systems and assign a temporary owner for each.
  \item \checkbox Inventory all recurring processes currently performed by the business.
  \item \checkbox Identify current tools, spreadsheets, shared folders, forms, and templates.
  \item \checkbox Identify customer-facing handoffs and internal handoffs.
  \item \checkbox Identify the top five process pain points, delays, quality issues, or risks.
  \item \checkbox Create an initial KPI list for each system.
\end{itemize}

\sectionbar{Phase 2 - Document Minimum SOPs}
\begin{itemize}[label={}]
  \item \checkbox Create SOP templates for recurring work.
  \item \checkbox Write minimum SOPs for marketing lead capture, sales handoff, fulfillment, invoicing, access management, onboarding, and KPI review.
  \item \checkbox Define inputs, outputs, owners, decision points, and evidence for each SOP.
  \item \checkbox Review each SOP with the people who perform the work.
  \item \checkbox Approve SOPs and store them in a controlled repository.
\end{itemize}

\sectionbar{Phase 3 - Stabilize Tooling}
\begin{itemize}[label={}]
  \item \checkbox Map each SOP to the tools required to execute it.
  \item \checkbox Remove duplicate tools where one tool can support the workflow.
  \item \checkbox Define system owners and access owners for all tools.
  \item \checkbox Confirm backup, export, and recovery approach for critical data.
  \item \checkbox Create a tool renewal and access review schedule.
\end{itemize}

\sectionbar{Phase 4 - Operationalize KPIs}
\begin{itemize}[label={}]
  \item \checkbox Define calculation method and data source for each KPI.
  \item \checkbox Assign KPI owners and update cadence.
  \item \checkbox Create a monthly scorecard.
  \item \checkbox Use KPI gaps to create improvement backlog items.
  \item \checkbox Review trends monthly and reset targets quarterly.
\end{itemize}

\sectionbar{Phase 5 - Improve and Automate}
\begin{itemize}[label={}]
  \item \checkbox Rank improvement backlog items by impact, risk, urgency, and effort.
  \item \checkbox Identify automation candidates only after workflows are stable.
  \item \checkbox Pilot automation on one low-risk workflow before scaling.
  \item \checkbox Define monitoring, exception handling, and rollback for each automation.
  \item \checkbox Review automations during the quarterly operating model review.
\end{itemize}

\sectionbar{Completion Criteria}
\begin{itemize}[label={}]
  \item \checkbox Every core system has an owner.
  \item \checkbox Every critical recurring process has an SOP.
  \item \checkbox Every business-critical tool has an owner and access review process.
  \item \checkbox Every system has a small KPI set and review cadence.
  \item \checkbox Handoffs between marketing, sales, operations, finance, HR, administration, and leadership are documented.
  \item \checkbox A continuous improvement backlog is active and reviewed monthly.
\end{itemize}

\end{document}