
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{Small Business Core Systems Operating Model}{A practical governance model for marketing, sales, operations, finance, administration, HR, strategy, SOPs, tooling, KPIs, and continuous improvement.}
\sectionbar{Executive Summary}
A small business becomes easier to operate and scale when its recurring work is organized into explicit systems. The core model in this bundle treats the business as a set of connected operating systems: marketing creates demand, sales converts demand, operations fulfills commitments, financial management protects cash flow, administration and IT security keep the back office reliable, HR manages people capacity, and leadership aligns the whole organization through strategy and KPIs.
\sectionbar{Operating Principles}
\begin{itemize}
  \item Make recurring work visible before trying to automate it.
  \item Document minimum viable SOPs before adding complex tooling.
  \item Assign one accountable owner for every system, KPI, and recurring review.
  \item Connect systems through clear handoffs rather than informal messages.
  \item Use KPIs to improve decisions, not to create reporting theater.
  \item Treat security, cash flow, and customer experience as operating constraints, not afterthoughts.
\end{itemize}

\sectionbar{Core Systems Map}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}>{\raggedright\arraybackslash}p{0.28\textwidth}>{\raggedright\arraybackslash}p{0.20\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries System} & \textcolor{white}{\bfseries Primary Objective} & \textcolor{white}{\bfseries Primary Output} & \textcolor{white}{\bfseries Downstream Consumer} \\
\toprule
Marketing and Lead Generation & Attract qualified prospects & Qualified leads with source context & Sales \\
Sales and Lead Conversion & Convert leads into customers & Closed-won customers and delivery commitments & Operations and Finance \\
Operations and Fulfillment & Deliver the promised product or service & Completed work, quality evidence, customer confirmation & Finance and Leadership \\
Financial Management & Protect cash flow and profitability & Books, invoices, budget, cash forecast & Leadership \\
Administration and IT Security & Maintain reliable back-office and secure systems & Documents, access, backups, admin coordination & All systems \\
HR and People Management & Build workforce capacity & Hired, trained, managed, and offboarded people & Operations and Leadership \\
Leadership and Strategy & Set direction and governance cadence & Strategic priorities, scorecard, decisions, risks & All systems \\
\bottomrule
\end{tabularx}}

\sectionbar{System Operating Model}
\subsectionbar{Marketing and Lead Generation System}
\textbf{Purpose.} Create a repeatable mechanism for attracting the right market segment, capturing lead information, and preparing prospects for sales conversion.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Ideal customer profile and buyer persona definition
  \item Content calendar and campaign planning
  \item Website, landing page, and form management
  \item Lead capture, tagging, and source attribution
\end{itemize}
 & \begin{itemize}
  \item Campaign planning SOP
  \item Lead capture and CRM entry SOP
  \item Content review and publishing SOP
  \item Lead source attribution SOP
\end{itemize}
 & \begin{itemize}
  \item New qualified leads
  \item Lead source conversion rate
  \item Website form conversion rate
  \item Email open and click rates
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\subsectionbar{Sales and Lead Conversion System}
\textbf{Purpose.} Move prospects from inquiry to purchase using a structured pipeline, qualification rules, pricing discipline, and documented follow-up routines.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Lead qualification and opportunity scoring
  \item Defined sales stages and exit criteria
  \item Proposal, quote, and pricing workflow
  \item Follow-up cadence and objection handling
\end{itemize}
 & \begin{itemize}
  \item Lead qualification SOP
  \item Sales pipeline update SOP
  \item Proposal and quote approval SOP
  \item Follow-up cadence SOP
\end{itemize}
 & \begin{itemize}
  \item Lead-to-opportunity conversion rate
  \item Opportunity win rate
  \item Average sales cycle time
  \item Average deal value
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\subsectionbar{Operations and Fulfillment System}
\textbf{Purpose.} Translate customer commitments into repeatable work execution, quality control, inventory or resource management, and customer delivery outcomes.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Work intake and prioritization
  \item Standard operating procedures for recurring work
  \item Project or order tracking
  \item Inventory, resource, or capacity management
\end{itemize}
 & \begin{itemize}
  \item Order or service intake SOP
  \item Fulfillment workflow SOP
  \item Quality control SOP
  \item Inventory or capacity review SOP
\end{itemize}
 & \begin{itemize}
  \item On-time delivery rate
  \item Rework rate
  \item Cycle time
  \item Inventory accuracy or capacity utilization
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\subsectionbar{Financial Management System}
\textbf{Purpose.} Maintain accurate books, manage receivables and payables, forecast cash needs, and create decision-ready financial visibility.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Bookkeeping and chart of accounts governance
  \item Invoicing and accounts receivable tracking
  \item Accounts payable and expense control
  \item Budgeting and cash flow forecasting
\end{itemize}
 & \begin{itemize}
  \item Monthly close SOP
  \item Invoice creation and collection SOP
  \item Expense approval SOP
  \item Cash flow forecast SOP
\end{itemize}
 & \begin{itemize}
  \item Cash on hand
  \item Accounts receivable aging
  \item Gross margin
  \item Net profit margin
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\subsectionbar{Administration and IT Security System}
\textbf{Purpose.} Provide the administrative backbone for documents, scheduling, access, backups, and baseline cybersecurity protections.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Document management and retention
  \item Calendar, scheduling, and administrative coordination
  \item User access management
  \item Device, account, and software inventory
\end{itemize}
 & \begin{itemize}
  \item Document naming and retention SOP
  \item Access request and removal SOP
  \item Backup verification SOP
  \item Device onboarding SOP
\end{itemize}
 & \begin{itemize}
  \item Backup success rate
  \item Account review completion rate
  \item Security training completion
  \item Mean time to revoke access
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\subsectionbar{HR and People Management System}
\textbf{Purpose.} Create consistent people processes for hiring, onboarding, payroll coordination, training, performance management, and compliance.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Role definition and hiring workflow
  \item Onboarding and training plans
  \item Payroll and benefits coordination
  \item Performance check-ins and feedback routines
\end{itemize}
 & \begin{itemize}
  \item Hiring requisition SOP
  \item Interview and selection SOP
  \item Employee onboarding SOP
  \item Training completion SOP
\end{itemize}
 & \begin{itemize}
  \item Time to fill
  \item New hire onboarding completion
  \item Training completion rate
  \item Retention rate
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\subsectionbar{Leadership and Strategy System}
\textbf{Purpose.} Define strategic priorities, operating cadence, core values, decision rights, and KPI review routines that keep the business aligned.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.31\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Capabilities} & \textcolor{white}{\bfseries Minimum SOPs} & \textcolor{white}{\bfseries KPIs} \\
\toprule
\begin{itemize}
  \item Mission, vision, and core values definition
  \item Annual and quarterly planning
  \item KPI scorecard governance
  \item Decision rights and accountability model
\end{itemize}
 & \begin{itemize}
  \item Quarterly planning SOP
  \item Weekly leadership review SOP
  \item KPI scorecard update SOP
  \item Decision log SOP
\end{itemize}
 & \begin{itemize}
  \item Revenue growth
  \item Profitability trend
  \item Customer retention
  \item Operational throughput
\end{itemize}
 \\
\bottomrule
\end{tabularx}}

\sectionbar{SOP and Tooling Layer}
The SOP and tooling layer is the execution layer that turns strategy into repeatable work. SOPs define what must happen, who performs the work, what evidence is produced, and when escalation is required. Tools should support the SOP, not replace it.
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.48\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Layer} & \textcolor{white}{\bfseries Minimum Standard} & \textcolor{white}{\bfseries Evidence} \\
\toprule
SOP inventory & Each recurring process has an owner, trigger, steps, inputs, outputs, and review date. & SOP register \\
Tool inventory & Each tool has a business owner, access owner, data purpose, and renewal date. & Tool register \\
Templates & Common work uses approved templates for consistency. & Template library \\
Automation & Automations are applied only after the manual workflow is stable and measured. & Automation backlog and change log \\
Access control & Access is granted by role and removed during offboarding. & Access review record \\
\bottomrule
\end{tabularx}}

\sectionbar{KPI and Continuous Improvement Layer}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.42\textwidth}>{\raggedright\arraybackslash}p{0.30\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Cadence} & \textcolor{white}{\bfseries Purpose} & \textcolor{white}{\bfseries Required Output} \\
\toprule
Weekly & Review short-cycle execution and blockers. & Updated action list and issue owner \\
Monthly & Review system-level KPIs and process health. & Scorecard, trend notes, improvement priorities \\
Quarterly & Review strategy, capacity, risk, budget, and growth assumptions. & Quarterly objectives and system improvement roadmap \\
Annually & Refresh the operating model, roles, tools, and strategic plan. & Annual plan and operating model revision \\
\bottomrule
\end{tabularx}}

\sectionbar{30-60-90 Day Implementation Roadmap}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.32\textwidth}>{\raggedright\arraybackslash}p{0.40\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Timeframe} & \textcolor{white}{\bfseries Primary Focus} & \textcolor{white}{\bfseries Deliverables} \\
\toprule
Days 1-30 & Inventory and baseline the business systems. & System owner list, SOP register, tool register, KPI draft, top process risks \\
Days 31-60 & Standardize the highest-value workflows. & Approved SOPs, handoff templates, meeting cadence, dashboard draft \\
Days 61-90 & Operationalize review and improvement. & KPI scorecard, improvement backlog, automation candidates, quarterly review package \\
\bottomrule
\end{tabularx}}

\sectionbar{Maturity Model}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.35\textwidth}>{\raggedright\arraybackslash}p{0.37\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Level} & \textcolor{white}{\bfseries Description} & \textcolor{white}{\bfseries Expected Condition} \\
\toprule
1 - Ad hoc & Work depends on individual memory. & Few SOPs, inconsistent tools, reactive decision-making. \\
2 - Defined & Major workflows are documented. & SOPs exist for critical processes, but metrics and ownership are uneven. \\
3 - Managed & Owners, KPIs, and reviews are operating. & Recurring work is measured and handoffs are reliable. \\
4 - Optimized & Continuous improvement is routine. & Automation, dashboards, and periodic redesign improve speed, quality, and margin. \\
\bottomrule
\end{tabularx}}

\sectionbar{Reference Links Supplied With Source Brief}
\begin{enumerate}[leftmargin=1.2em]
\item \url{https://www.systemsandteams.com/10-essential-small-business-systems-every-owner-needs}
\item \url{https://www.finsync.com/blog/6-core-business-processes-every-small-business-owner-should-know/}
\item \url{https://cdn2.hubspot.net/hubfs/5078742/The-7-Systems-That-Every-Business-Must-Have-To-Be-Successful-V4.pdf}
\item \url{https://www.youtube.com/watch?v=vIKJE1Mw41g}
\item \url{https://www.emyth.com/inside/7-essential-business-systems}
\item \url{https://www.linkedin.com/pulse/four-essential-systems-every-business-needs-thrive-hancott-cgma-hxhde}
\item \url{https://founderscpa.com/blog/7-business-systems-every-growing-startup-needs-to-define}
\item \url{https://incredibleoneenterprises.com/7-systems-every-entrepreneur-must-have-in-place-in-order-to-build-their-business/}
\item \url{https://www.systemsandteams.com/simple-ways-to-build-systems-for-your-small-business}
\item \url{https://www.eposnow.com/us/resources/small-business-must-haves/}
\item \url{https://www.caflou.com/10-essential-management-tools-every-small-business-needs}
\item \url{https://treefrogmarketing.com/4-systems-every-small-business-must-have-to-increase-efficiency-growth/}
\item \url{https://www.thinkwisesoftware.com/blog/3-types-of-core-applications-when-should-you-use-low-code}
\item \url{https://kraftbusiness.com/blog/cyber-security-training-for-small-business/}
\item \url{https://hildachuks.com/essential-business-tools}
\item \url{https://www.investglass.com/top-onboarding-solutions-for-the-public-sector-streamlining-success/}
\item \url{https://www.youtube.com/watch?v=zTFgClXTaRc}
\item \url{https://thebusinessmagnate.com/business-management-strategies-for-beginners/}
\end{enumerate}
\end{document}