
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{D9EAF7}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D7DCE2}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.18}
\newcommand{\docTitle}[2]{%
  {\Huge\bfseries\color{HeaderBlue} #1}\par
  {\large #2}\par\vspace{0.4em}\hrule\vspace{1.0em}}
\newcommand{\sectionbar}[1]{\vspace{0.6em}{\Large\bfseries\color{HeaderBlue} #1}\par\vspace{0.25em}\hrule\vspace{0.45em}}
\newcommand{\subsectionbar}[1]{\vspace{0.4em}{\bfseries\color{HeaderBlue} #1}\par\vspace{0.15em}}
\newcommand{\checkbox}{\ensuremath{\square}\hspace{0.4em}}
\newcommand{\tighttable}[1]{\rowcolors{2}{LightGray}{white}#1\rowcolors{2}{}{}}
\sloppy
\pagestyle{plain}

\begin{document}
\docTitle{SOP and Tooling Layer Guide}{A control layer for turning business routines into documented, measurable, and automatable workflows.}
\sectionbar{Purpose}
The SOP and tooling layer standardizes how work is performed across the business. Its goal is to reduce dependency on individual memory, clarify ownership, and create a stable foundation for automation and delegation.
\sectionbar{SOP Standard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth}>{\raggedright\arraybackslash}p{0.65\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries SOP Field} & \textcolor{white}{\bfseries Required Content} \\
\toprule
Process name & Clear name that matches the recurring business activity. \\
Owner & Role accountable for keeping the SOP current. \\
Trigger & Event that starts the process. \\
Inputs & Information, documents, tools, or approvals required before work begins. \\
Steps & Action-oriented numbered instructions. \\
Decision points & Branching rules and escalation conditions. \\
Outputs & Evidence, records, deliverables, or status changes produced. \\
KPIs & Measures used to confirm the process is working. \\
Review cadence & Frequency for validating and updating the SOP. \\
\bottomrule
\end{tabularx}}

\sectionbar{Tool Governance Standard}
\tighttable{\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.28\textwidth}>{\raggedright\arraybackslash}p{0.62\textwidth}}
\rowcolor{HeaderBlue}
\textcolor{white}{\bfseries Control} & \textcolor{white}{\bfseries Minimum Requirement} \\
\toprule
Business owner & Every tool has one accountable owner. \\
Access owner & Access is approved, reviewed, and removed by an assigned role. \\
Data purpose & The business data stored in the tool is defined. \\
Backup or export path & Critical data can be recovered or exported. \\
Integration owner & Integrations have an owner and documented failure path. \\
Renewal review & Cost, usage, and continued fit are reviewed before renewal. \\
\bottomrule
\end{tabularx}}

\sectionbar{SOP Template}
\begin{itemize}[label={}]
  \item \checkbox Process name and version are documented.
  \item \checkbox Owner and backup owner are assigned.
  \item \checkbox Trigger and scope are clear.
  \item \checkbox Prerequisites are listed before procedure steps.
  \item \checkbox Procedure steps are numbered and action-oriented.
  \item \checkbox Decision points and escalation criteria are explicit.
  \item \checkbox Outputs and required evidence are defined.
  \item \checkbox KPIs and review cadence are listed.
  \item \checkbox Last reviewed date and next review date are recorded.
\end{itemize}

\sectionbar{Automation Readiness Criteria}
\begin{itemize}[label={}]
  \item \checkbox The manual workflow is stable and documented.
  \item \checkbox Inputs and outputs are structured enough to be validated.
  \item \checkbox The exception path is known.
  \item \checkbox A human owner remains accountable for outcomes.
  \item \checkbox The automation has a monitoring and rollback approach.
  \item \checkbox Security and access implications have been reviewed.
\end{itemize}

\end{document}