\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[hidelinks]{hyperref}
\usepackage{titlesec}
\usepackage{fancyhdr}

\setlength{\headheight}{14pt}
\pagestyle{fancy}
\fancyhf{}
\lhead{Open-Source Deployment Orchestration}
\rhead{Checklist}
\cfoot{\thepage}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.55em}
\setlist[itemize]{leftmargin=1.25em}
\setlist[enumerate]{leftmargin=1.4em}
\titleformat{\section}{\large\bfseries}{\thesection}{0.6em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.6em}{}
\emergencystretch=2em
\urlstyle{same}

\newcommand{\checkbox}{\makebox[1.6em][l]{\(\square\)}}
\newcommand{\tool}[1]{\textbf{#1}}

\title{Open-Source Deployment Orchestration\\\large Selection and Review Checklist}
\author{}
\date{May 2026}

\begin{document}
\maketitle

\textbf{How to use this checklist.} Work from top to bottom. The first sections determine the \emph{tool class}; the later sections check implementation fit. The goal is to avoid selecting a tool whose documented scope does not match the actual control problem.\cite{KubernetesOverview,ArgoCDHome,AnsiblePlaybooks,AirflowHome,TemporalWhatIs,KubeflowIntro}

\section{Step 1: Identify the primary orchestration problem}
\checkbox We need a control plane that schedules and runs containerized or legacy application workloads.\\
\hspace*{1.2em}Likely class: \tool{Kubernetes}, \tool{Nomad}, or \tool{Docker Swarm}.\cite{KubernetesOverview,NomadWhatIs,DockerSwarm}

\checkbox We already have Kubernetes and want Git to be the source of truth for deployed state.\\
\hspace*{1.2em}Likely class: \tool{Argo CD}.\cite{ArgoCDHome,ArgoCDDeclarative}

\checkbox We need procedural rollout automation across hosts, services, or tiers.\\
\hspace*{1.2em}Likely class: \tool{Ansible}.\cite{AnsiblePlaybooks,AnsibleRolling}

\checkbox We need scheduled workflows, task dependencies, retries, and monitoring for data pipelines.\\
\hspace*{1.2em}Likely class: \tool{Airflow}, \tool{Prefect}, or \tool{Dagster}.\cite{AirflowHome,PrefectQuickstart,DagsterOverview}

\checkbox We need durable, long-running, stateful business-process execution that survives crashes and resumes correctly.\\
\hspace*{1.2em}Likely class: \tool{Temporal}.\cite{TemporalWhatIs,TemporalWorkflowExecution}

\checkbox We need ML- or AI-oriented workflow orchestration, training pipelines, or an AI platform stack.\\
\hspace*{1.2em}Likely class: \tool{Flyte}, \tool{Kubeflow}, or \tool{Metaflow}.\cite{FlyteOSS,KubeflowIntro,MetaflowWhatIs}

\section{Step 2: Eliminate common misclassification errors}
\checkbox Do \emph{not} classify \tool{Argo CD} as a general-purpose runtime scheduler. It is a Kubernetes GitOps continuous delivery controller.\cite{ArgoCDHome}

\checkbox Do \emph{not} classify \tool{Ansible} as a cluster scheduler. Its documented model is playbook-driven orchestration and deployment automation across remote systems.\cite{AnsiblePlaybooks,AnsibleRolling}

\checkbox Do \emph{not} classify \tool{Temporal} as primarily a batch DAG engine. Its documented model is durable workflow execution for application logic.\cite{TemporalWhatIs,TemporalUnderstanding}

\checkbox Do \emph{not} assume all ML tools are equivalent to data-pipeline tools. \tool{Flyte}, \tool{Kubeflow}, and \tool{Metaflow} emphasize AI and ML workflows.\cite{FlyteOSS,KubeflowIntro,MetaflowWhatIs}

\section{Step 3: Confirm infrastructure assumptions}
\checkbox Kubernetes is already present or explicitly desired.\\
\hspace*{1.2em}Relevant tools: \tool{Kubernetes}, \tool{Argo CD}, \tool{Flyte}, \tool{Kubeflow}.\cite{KubernetesOverview,ArgoCDHome,FlyteOSS,KubeflowIntro}

\checkbox We need one orchestrator for containers and non-containerized or legacy workloads.\\
\hspace*{1.2em}Relevant tool: \tool{Nomad}.\cite{NomadWhatIs}

\checkbox We prefer the Docker-native cluster model and accept a narrower feature set than Kubernetes ecosystems commonly provide.\\
\hspace*{1.2em}Relevant tool: \tool{Docker Swarm}.\cite{DockerSwarm,DockerSwarmConcepts}

\checkbox We need agentless host automation using declarative playbooks.\\
\hspace*{1.2em}Relevant tool: \tool{Ansible}.\cite{AnsiblePlaybooks}

\section{Step 4: Confirm workflow model}
\checkbox The work is naturally represented as a DAG of tasks with schedules and dependencies.\\
\hspace*{1.2em}Relevant tools: \tool{Airflow}, \tool{Prefect}, \tool{Dagster}.\cite{AirflowArchitecture,PrefectQuickstart,DagsterOverview}

\checkbox The work needs durable state across failures, timers, callbacks, retries, and long waits.\\
\hspace*{1.2em}Relevant tool: \tool{Temporal}.\cite{TemporalWhatIs,TemporalWorkflowExecution}

\checkbox The work needs ML-specific constructs, experiment-to-production flow support, or AI platform composition.\\
\hspace*{1.2em}Relevant tools: \tool{Flyte}, \tool{Kubeflow}, \tool{Metaflow}.\cite{FlyteOSS,KubeflowIntro,MetaflowWhatIs}

\section{Step 5: Fit-for-purpose checklist by tool class}

\subsection{Platform and workload orchestrators}
\checkbox We need workload placement, scheduling, scaling, and service management.\cite{KubernetesOverview,NomadWhatIs,DockerSwarm}

\checkbox We are comfortable operating a cluster control plane.\cite{KubernetesOverview,DockerSwarm}

\checkbox We have a clear plan for service discovery, secrets, observability, and rollout policy.

\subsection{GitOps and Kubernetes delivery controllers}
\checkbox Desired state will be stored and reviewed in Git.\cite{ArgoCDHome,ArgoCDDeclarative}

\checkbox Drift detection and reconciliation are required.\cite{ArgoCDHome}

\checkbox The deployment target is Kubernetes.\cite{ArgoCDHome}

\subsection{Configuration and rollout orchestration}
\checkbox The deployment process includes ordered host actions, maintenance windows, or multi-tier coordination.\cite{AnsibleRolling}

\checkbox Teams can maintain playbooks and inventory cleanly.\cite{AnsiblePlaybooks}

\subsection{Data workflow orchestration}
\checkbox The core value comes from dependency management, schedules, backfills, retries, and execution observability.\cite{AirflowHome,PrefectQuickstart,DagsterOverview}

\checkbox Data lineage or asset-centric reasoning is important.\\
\hspace*{1.2em}Especially relevant for \tool{Dagster}.\cite{DagsterOverview}

\subsection{Durable application workflows}
\checkbox The workflow spans external systems, long waits, or application-level compensation logic.\cite{TemporalWhatIs,TemporalUnderstanding}

\checkbox Reliability requirements justify a durable execution model.\cite{TemporalWhatIs}

\subsection{ML and AI workflow platforms}
\checkbox Reproducibility, training workflows, or AI platform modularity are first-order concerns.\cite{FlyteOSS,KubeflowIntro,MetaflowWhatIs}

\checkbox The team accepts the surrounding platform model, especially if Kubernetes is part of the solution.\cite{FlyteOSS,KubeflowIntro}

\section{Step 6: Final decision gate}
Approve the selection only if all of the following are true:
\begin{itemize}
    \item The selected tool's \emph{documented scope} matches the actual orchestration problem.\cite{KubernetesOverview,ArgoCDHome,AnsiblePlaybooks,AirflowHome,TemporalWhatIs,KubeflowIntro}
    \item Required infrastructure assumptions are acceptable.\cite{NomadWhatIs,DockerSwarm,FlyteOSS}
    \item The workflow model matches the execution model documented by the tool.\cite{AirflowArchitecture,TemporalWorkflowExecution,MetaflowWhatIs}
    \item The team can operate the tool and its ecosystem sustainably.
\end{itemize}

\begin{thebibliography}{99}
\bibitem{KubernetesOverview} Kubernetes Documentation. \emph{Overview}. Available: \url{https://kubernetes.io/docs/concepts/overview/}
\bibitem{ArgoCDHome} Argo CD Documentation. \emph{Argo CD - Declarative GitOps CD for Kubernetes}. Available: \url{https://argo-cd.readthedocs.io/en/stable/}
\bibitem{ArgoCDDeclarative} Argo CD Documentation. \emph{Declarative Setup}. Available: \url{https://argo-cd.readthedocs.io/en/stable/operator-manual/declarative-setup/}
\bibitem{NomadWhatIs} HashiCorp Nomad Documentation. \emph{What is Nomad?}. Available: \url{https://developer.hashicorp.com/nomad/docs/what-is-nomad}
\bibitem{DockerSwarm} Docker Docs. \emph{Swarm mode}. Available: \url{https://docs.docker.com/engine/swarm/}
\bibitem{DockerSwarmConcepts} Docker Docs. \emph{Swarm mode key concepts}. Available: \url{https://docs.docker.com/engine/swarm/key-concepts/}
\bibitem{AnsiblePlaybooks} Ansible Community Documentation. \emph{Working with playbooks}. Available: \url{https://docs.ansible.com/projects/ansible/latest/playbook_guide/playbooks.html}
\bibitem{AnsibleRolling} Ansible Community Documentation. \emph{Playbook Example: Continuous Delivery and Rolling Upgrades}. Available: \url{https://docs.ansible.com/projects/ansible/latest/playbook_guide/guide_rolling_upgrade.html}
\bibitem{AirflowHome} Apache Airflow Documentation. \emph{What is Airflow?}. Available: \url{https://airflow.apache.org/docs/apache-airflow/stable/index.html}
\bibitem{AirflowArchitecture} Apache Airflow Documentation. \emph{Architecture Overview}. Available: \url{https://airflow.apache.org/docs/apache-airflow/stable/core-concepts/overview.html}
\bibitem{PrefectQuickstart} Prefect Documentation. \emph{Quickstart}. Available: \url{https://docs.prefect.io/v3/get-started/quickstart}
\bibitem{DagsterOverview} Dagster Documentation. \emph{Overview}. Available: \url{https://docs.dagster.io/}
\bibitem{TemporalWhatIs} Temporal Documentation. \emph{What is Temporal?}. Available: \url{https://docs.temporal.io/temporal}
\bibitem{TemporalWorkflowExecution} Temporal Documentation. \emph{Workflow Execution overview}. Available: \url{https://docs.temporal.io/workflow-execution}
\bibitem{TemporalUnderstanding} Temporal Documentation. \emph{Understanding Temporal}. Available: \url{https://docs.temporal.io/evaluate/understanding-temporal}
\bibitem{FlyteOSS} Flyte OSS Documentation. \emph{User guide / Flyte OSS}. Available: \url{https://docs.flyte.org/}
\bibitem{KubeflowIntro} Kubeflow Documentation. \emph{Introduction}. Available: \url{https://www.kubeflow.org/docs/started/introduction/}
\bibitem{MetaflowWhatIs} Metaflow Documentation. \emph{What is Metaflow}. Available: \url{https://docs.metaflow.org/introduction/what-is-metaflow}
\end{thebibliography}

\end{document}
