\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{titlesec}
\usepackage{fancyhdr}

\setlength{\headheight}{14pt}
\pagestyle{fancy}
\fancyhf{}
\lhead{Open-Source Deployment Orchestration}
\rhead{Handout}
\cfoot{\thepage}
\setlist[itemize]{leftmargin=1.25em}
\setlist[enumerate]{leftmargin=1.4em}
\titleformat{\section}{\large\bfseries\color{black}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\normalsize\bfseries\color{black}}{\thesubsection}{0.6em}{}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.65em}
\emergencystretch=2em
\urlstyle{same}

\title{Open-Source Deployment Orchestration Tools\\\large Citation-Clean Classification Handout}
\author{}
\date{May 2026}

\begin{document}
\maketitle

\textbf{Purpose.} This handout provides a cleaner, documentation-grounded classification of common open-source deployment orchestration tools. It separates tools by \emph{what they actually orchestrate}: infrastructure and application workloads, Kubernetes delivery state, configuration-driven rollout automation, data workflows, durable application workflows, and machine learning workflows.

\section{Why the taxonomy matters}
A flat list of ``orchestration tools'' is often misleading because these systems operate at different layers of the stack. Kubernetes and Nomad schedule and manage workloads. Argo CD continuously reconciles Kubernetes clusters to a declarative desired state stored in Git. Ansible coordinates actions across hosts. Airflow, Prefect, and Dagster run workflow graphs. Temporal provides durable execution for long-running application processes. Flyte, Kubeflow, and Metaflow emphasize data science and machine learning workflow needs.\cite{KubernetesOverview,ArgoCDHome,NomadWhatIs,DockerSwarm,AnsiblePlaybooks,AirflowHome,PrefectQuickstart,DagsterOverview,TemporalWhatIs,FlyteOSS,KubeflowIntro,MetaflowWhatIs}

\section{Recommended classification}

\subsection{1. Platform and workload orchestrators}
These tools schedule or coordinate the execution of application workloads.
\begin{itemize}
    \item \textbf{Kubernetes}: an open-source platform for managing containerized workloads and services with declarative configuration and automation.\cite{KubernetesOverview}
    \item \textbf{HashiCorp Nomad}: a flexible workload orchestrator for containerized and legacy applications using a unified workflow.\cite{NomadWhatIs}
    \item \textbf{Docker Swarm}: Docker Engine swarm mode for managing a cluster of Docker daemons and deploying services across them.\cite{DockerSwarm,DockerSwarmConcepts}
\end{itemize}

\textbf{Use this class when} you need a scheduler and execution control plane for running application workloads.

\subsection{2. Kubernetes GitOps and continuous delivery controllers}
These tools do not replace the cluster orchestrator; they manage desired state and delivery into Kubernetes.
\begin{itemize}
    \item \textbf{Argo CD}: a declarative GitOps continuous delivery tool for Kubernetes that continuously compares live cluster state to the target state stored in Git and syncs drift when needed.\cite{ArgoCDHome,ArgoCDDeclarative}
\end{itemize}

\textbf{Use this class when} Git should be the source of truth for Kubernetes application state.

\subsection{3. Configuration and rollout orchestration}
These tools coordinate actions across machines, tiers, and environments.
\begin{itemize}
    \item \textbf{Ansible}: uses playbooks to manage configurations and deployments and can coordinate multi-tier rollouts, rolling updates, and delegated actions across systems such as monitoring servers and load balancers.\cite{AnsiblePlaybooks,AnsibleRolling}
\end{itemize}

\textbf{Use this class when} the problem is host automation, procedural rollout coordination, or environment bootstrapping rather than cluster scheduling.

\subsection{4. Data and batch workflow orchestration}
These tools orchestrate ordered units of work, dependencies, schedules, retries, and observability around data flows.
\begin{itemize}
    \item \textbf{Apache Airflow}: a platform for developing, scheduling, and monitoring batch-oriented workflows represented as DAGs.\cite{AirflowHome,AirflowArchitecture}
    \item \textbf{Prefect}: a workflow orchestration tool for building, deploying, running, and monitoring data pipelines, with explicit handling of dependencies and failures.\cite{PrefectQuickstart}
    \item \textbf{Dagster}: a data orchestrator designed for data engineers, with integrated lineage, observability, a declarative programming model, and strong testability.\cite{DagsterOverview}
\end{itemize}

\textbf{Use this class when} you are orchestrating data pipelines, scheduled jobs, and data dependencies.

\subsection{5. Durable application workflow orchestration}
This class is different from batch DAG orchestration. It focuses on long-running, stateful application processes that must survive failures.
\begin{itemize}
    \item \textbf{Temporal}: a platform for durable workflow execution that preserves workflow state and progress across crashes, outages, and retries.\cite{TemporalWhatIs,TemporalWorkflowExecution}
\end{itemize}

\textbf{Use this class when} business processes span long durations, external calls, retries, compensation logic, or microservice boundaries.

\subsection{6. Machine learning and AI workflow platforms}
These tools support ML and AI lifecycle concerns such as training workflows, reproducibility, platform composition, and operationalization.
\begin{itemize}
    \item \textbf{Flyte}: a free and open-source platform for orchestrating AI workflows and a Kubernetes-native workflow orchestrator.\cite{FlyteOSS}
    \item \textbf{Kubeflow}: a foundation of tools for AI platforms on Kubernetes, covering stages of the AI lifecycle through modular Kubernetes-native components.\cite{KubeflowIntro,KubeflowPipelines}
    \item \textbf{Metaflow}: a human-friendly Python library for developing, deploying, and operating data-intensive applications, especially data science, ML, and AI projects.\cite{MetaflowWhatIs,MetaflowWhy}
\end{itemize}

\textbf{Use this class when} reproducible training, experiment-to-production flows, or AI-platform integration are central requirements.

\section{What this taxonomy fixes}
\begin{itemize}
    \item It avoids treating \textbf{Argo CD} as a general runtime orchestrator alongside Kubernetes or Nomad. Argo CD assumes Kubernetes and manages reconciliation to a desired state.\cite{ArgoCDHome}
    \item It avoids treating \textbf{Ansible} as a cluster scheduler. Its documented model is playbook-driven orchestration of remote systems and coordinated rollouts.\cite{AnsiblePlaybooks,AnsibleRolling}
    \item It separates \textbf{Temporal} from DAG-first data orchestration engines because Temporal centers durable workflow execution for application logic, not primarily scheduled batch dataflow graphs.\cite{TemporalWhatIs,TemporalUnderstanding}
    \item It distinguishes \textbf{ML workflow platforms} from general data orchestration because Flyte, Kubeflow, and Metaflow emphasize AI and ML workflow development and operations.\cite{FlyteOSS,KubeflowIntro,MetaflowWhatIs}
\end{itemize}

\section{Practical selection guidance}
Choose by the \emph{primary control problem}:
\begin{enumerate}
    \item Need to schedule and run application workloads? Start with \textbf{Kubernetes}, \textbf{Nomad}, or \textbf{Docker Swarm}.\cite{KubernetesOverview,NomadWhatIs,DockerSwarm}
    \item Need Git-driven reconciliation into Kubernetes? Add \textbf{Argo CD}.\cite{ArgoCDHome}
    \item Need host-level rollout automation and procedural change coordination? Use \textbf{Ansible}.\cite{AnsiblePlaybooks}
    \item Need scheduled data pipelines and dependency graphs? Use \textbf{Airflow}, \textbf{Prefect}, or \textbf{Dagster}.\cite{AirflowHome,PrefectQuickstart,DagsterOverview}
    \item Need reliable long-running business workflows with durable state? Use \textbf{Temporal}.\cite{TemporalWhatIs}
    \item Need ML- and AI-oriented workflow orchestration? Evaluate \textbf{Flyte}, \textbf{Kubeflow}, and \textbf{Metaflow}.\cite{FlyteOSS,KubeflowIntro,MetaflowWhatIs}
\end{enumerate}

\section{Bottom line}
The term \emph{deployment orchestration} is too broad to be useful without scope. A precise classification should identify whether the tool is orchestrating \emph{runtime workloads}, \emph{Kubernetes desired state}, \emph{host automation}, \emph{data workflows}, \emph{durable business processes}, or \emph{ML workflows}. Once that boundary is explicit, tool selection becomes substantially clearer.\cite{KubernetesOverview,ArgoCDHome,AnsiblePlaybooks,AirflowHome,TemporalWhatIs,KubeflowIntro}

\begin{thebibliography}{99}
\bibitem{KubernetesOverview} Kubernetes Documentation. \emph{Overview}. Available: \url{https://kubernetes.io/docs/concepts/overview/}
\bibitem{ArgoCDHome} Argo CD Documentation. \emph{Argo CD - Declarative GitOps CD for Kubernetes}. Available: \url{https://argo-cd.readthedocs.io/en/stable/}
\bibitem{ArgoCDDeclarative} Argo CD Documentation. \emph{Declarative Setup}. Available: \url{https://argo-cd.readthedocs.io/en/stable/operator-manual/declarative-setup/}
\bibitem{NomadWhatIs} HashiCorp Nomad Documentation. \emph{What is Nomad?}. Available: \url{https://developer.hashicorp.com/nomad/docs/what-is-nomad}
\bibitem{DockerSwarm} Docker Docs. \emph{Swarm mode}. Available: \url{https://docs.docker.com/engine/swarm/}
\bibitem{DockerSwarmConcepts} Docker Docs. \emph{Swarm mode key concepts}. Available: \url{https://docs.docker.com/engine/swarm/key-concepts/}
\bibitem{AnsiblePlaybooks} Ansible Community Documentation. \emph{Working with playbooks}. Available: \url{https://docs.ansible.com/projects/ansible/latest/playbook_guide/playbooks.html}
\bibitem{AnsibleRolling} Ansible Community Documentation. \emph{Playbook Example: Continuous Delivery and Rolling Upgrades}. Available: \url{https://docs.ansible.com/projects/ansible/latest/playbook_guide/guide_rolling_upgrade.html}
\bibitem{AirflowHome} Apache Airflow Documentation. \emph{What is Airflow?}. Available: \url{https://airflow.apache.org/docs/apache-airflow/stable/index.html}
\bibitem{AirflowArchitecture} Apache Airflow Documentation. \emph{Architecture Overview}. Available: \url{https://airflow.apache.org/docs/apache-airflow/stable/core-concepts/overview.html}
\bibitem{PrefectQuickstart} Prefect Documentation. \emph{Quickstart}. Available: \url{https://docs.prefect.io/v3/get-started/quickstart}
\bibitem{DagsterOverview} Dagster Documentation. \emph{Overview}. Available: \url{https://docs.dagster.io/}
\bibitem{TemporalWhatIs} Temporal Documentation. \emph{What is Temporal?}. Available: \url{https://docs.temporal.io/temporal}
\bibitem{TemporalWorkflowExecution} Temporal Documentation. \emph{Workflow Execution overview}. Available: \url{https://docs.temporal.io/workflow-execution}
\bibitem{TemporalUnderstanding} Temporal Documentation. \emph{Understanding Temporal}. Available: \url{https://docs.temporal.io/evaluate/understanding-temporal}
\bibitem{FlyteOSS} Flyte OSS Documentation. \emph{User guide / Flyte OSS}. Available: \url{https://docs.flyte.org/}
\bibitem{KubeflowIntro} Kubeflow Documentation. \emph{Introduction}. Available: \url{https://www.kubeflow.org/docs/started/introduction/}
\bibitem{KubeflowPipelines} Kubeflow Documentation. \emph{Kubeflow Pipelines Overview}. Available: \url{https://www.kubeflow.org/docs/components/pipelines/overview/}
\bibitem{MetaflowWhatIs} Metaflow Documentation. \emph{What is Metaflow}. Available: \url{https://docs.metaflow.org/introduction/what-is-metaflow}
\bibitem{MetaflowWhy} Metaflow Documentation. \emph{Why Metaflow}. Available: \url{https://docs.metaflow.org/introduction/why-metaflow}
\end{thebibliography}

\end{document}



