\documentclass[11pt]{article}

\usepackage[letterpaper,margin=0.78in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xurl}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage[most]{tcolorbox}

\definecolor{accent}{HTML}{1F4E79}
\definecolor{lightaccent}{HTML}{EAF2F8}
\definecolor{midgray}{HTML}{666666}
\definecolor{lightgray}{HTML}{F5F7FA}
\definecolor{warn}{HTML}{8A4B08}

\hypersetup{
  colorlinks=true,
  linkcolor=accent,
  urlcolor=accent,
  citecolor=accent,
  pdftitle={Meeting with a Director of Franchise Sales: Candidate Positioning Handout},
  pdfauthor={Prepared with ChatGPT},
  pdfsubject={Franchise candidate meeting preparation},
  pdfkeywords={franchise, franchise disclosure document, FDD, SBA, FTC, due diligence}
}

\setlength{\headheight}{14pt}
\pagestyle{fancy}
\fancyhf{}
\lhead{Director of Franchise Sales Meeting Prep}
\rhead{Franchise Candidate Handout}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\section}{\Large\bfseries\color{accent}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries\color{accent}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{accent}}{}{0em}{}
\setlist[itemize]{leftmargin=1.35em,itemsep=0.2em,topsep=0.25em}
\setlist[enumerate]{leftmargin=1.45em,itemsep=0.2em,topsep=0.25em}
\renewcommand{\arraystretch}{1.22}
\emergencystretch=2em
\sloppy

\newtcolorbox{keybox}[1]{
  colback=lightaccent,
  colframe=accent,
  title=\textbf{#1},
  arc=2mm,
  boxrule=0.6pt,
  left=1.1mm,
  right=1.1mm,
  top=0.9mm,
  bottom=0.9mm
}

\newtcolorbox{scriptbox}[1]{
  colback=lightgray,
  colframe=midgray,
  title=\textbf{#1},
  arc=1.5mm,
  boxrule=0.45pt,
  left=1.1mm,
  right=1.1mm,
  top=0.9mm,
  bottom=0.9mm
}

\newtcolorbox{warningbox}[1]{
  colback=orange!7,
  colframe=warn,
  title=\textbf{#1},
  arc=1.5mm,
  boxrule=0.45pt,
  left=1.1mm,
  right=1.1mm,
  top=0.9mm,
  bottom=0.9mm
}

\newcommand{\placeholder}[1]{\textcolor{accent}{\texttt{[#1]}}}

\begin{document}

\begin{center}
  {\LARGE\bfseries Meeting with a Director of Franchise Sales}\\[0.35em]
  {\large Candidate Positioning, Qualification, and Due-Diligence Handout}\\[0.4em]
  {\small Prepared for direct, professional franchise-discovery conversations}
\end{center}

\vspace{0.5em}

\begin{keybox}{Core meeting objective}
The meeting is a two-way qualification conversation. Your job is to communicate your goals, liquid capital, operating readiness, market interest, and decision timeline while evaluating whether the franchise system is a disciplined fit for your risk profile, target market, and ownership model.
\end{keybox}

\section{What the Director Needs to Learn Quickly}

A Director of Franchise Sales is usually trying to determine whether you are a serious, financially qualified candidate who fits the brand's operating model. Prepare to communicate the following without rambling:

\begin{longtable}{>{\raggedright\arraybackslash}p{0.28\linewidth}>{\raggedright\arraybackslash}p{0.66\linewidth}}
\toprule
\textbf{Topic} & \textbf{What to communicate} \\
\midrule
Ownership goal & Whether you are pursuing owner-operator, semi-absentee, area development, or eventual multi-unit ownership. \\
Liquid capital & A clear range, not a vague statement. Example: ``I am evaluating brands that fit a liquid-capital range of \placeholder{amount}.'' \\
Net worth and financing & Whether you expect to self-fund, use SBA-backed financing, bring partners, or combine financing sources. \\
Market focus & Your target geography, preferred territory, and reason the market is attractive. \\
Timeline & Whether you are exploring, actively qualifying, or ready to move through FDD review and discovery. \\
Operating fit & Relevant experience in operations, technology, sales, hiring, budgeting, compliance, customer service, or team leadership. \\
Due-diligence posture & You are serious but disciplined: you intend to review the FDD, speak with franchisees, evaluate startup costs, and validate support claims. \\
\bottomrule
\end{longtable}

\section{Regulatory and Diligence Baseline}

\begin{itemize}
  \item The Federal Trade Commission's Franchise Rule requires franchisors to provide prospective franchisees with a disclosure document containing 23 specific items of information about the franchise offer, officers, and other franchisees.\cite{ftc-rule}
  \item The FTC states that you must receive the Franchise Disclosure Document (FDD) at least 14 days before you are asked to sign a contract or pay money to the franchisor or its affiliate.\cite{ftc-consumer-guide,ftc-deep-dive}
  \item You should treat the FDD as a diligence instrument, not a formality. The FTC advises prospective franchise buyers to investigate before investing and to ask questions while reviewing the FDD and attached documents.\cite{ftc-consumer-guide,ftc-deep-dive}
  \item If SBA-backed financing is part of your plan, confirm whether the brand appears in the SBA Franchise Directory. SBA states that the directory contains franchises and other brands reviewed and found eligible for SBA financial assistance.\cite{sba-directory}
  \item Independent validation matters. The International Franchise Association recommends analyzing the FDD, speaking with current and former franchisees, and seeking professional advice as part of franchise diligence.\cite{ifa-diligence}
\end{itemize}

\section{Recommended Opening Script}

\begin{scriptbox}{Concise opening}
``Thank you for taking the time to speak with me. I am evaluating franchise opportunities where my professional background, available capital, and long-term ownership goals align with a proven operating system. I am interested in \placeholder{Brand Name} because of its market positioning, growth model, and potential fit for the \placeholder{target market} market.''
\end{scriptbox}

\begin{scriptbox}{Candidate positioning}
``My background is in \placeholder{professional background}, and I am looking for a franchise opportunity where I can be an engaged \placeholder{owner-operator / executive operator / multi-unit developer}. I want to understand whether my goals, capital position, and target market align with what your strongest franchisees are doing.''
\end{scriptbox}

\subsection{Use this structure when introducing yourself}

\begin{enumerate}
  \item \textbf{One-sentence background:} your professional history and operating strengths.
  \item \textbf{One-sentence capital posture:} liquid capital range and financing strategy.
  \item \textbf{One-sentence ownership goal:} single unit, multi-unit, or area development.
  \item \textbf{One-sentence diligence standard:} you are reviewing the brand through FDD, franchisee validation, support model, and market fit.
\end{enumerate}

\section{Financial Qualification Statement}

Be direct about your readiness without over-disclosing unnecessary personal detail.

\begin{scriptbox}{Capital and readiness statement}
``I understand that franchise qualification depends heavily on liquid capital, net worth, financing readiness, and the full initial investment range. I am prepared to discuss my liquid-capital range of \placeholder{amount}, my funding strategy of \placeholder{self-fund / SBA loan / partner capital / combination}, and my expected timeline of \placeholder{timeframe}. I want to confirm whether I meet your financial profile before moving deeper into the process.''
\end{scriptbox}

\subsection{Numbers to know before the call}

\begin{longtable}{>{\raggedright\arraybackslash}p{0.38\linewidth}>{\raggedright\arraybackslash}p{0.56\linewidth}}
\toprule
\textbf{Metric} & \textbf{Your prepared answer} \\
\midrule
Liquid capital available & \placeholder{cash, marketable securities, immediately available funds} \\
Estimated net worth & \placeholder{range} \\
Financing plan & \placeholder{cash / SBA / HELOC / investor / retirement rollover / other} \\
Target investment range & \placeholder{maximum all-in startup investment} \\
Working capital cushion & \placeholder{months of operating cushion beyond initial investment} \\
Target market & \placeholder{city, county, region, or territory} \\
Ownership model & \placeholder{owner-operator, semi-absentee, multi-unit, area developer} \\
Decision timeline & \placeholder{exploring, 3--6 months, 6--12 months, ready now} \\
\bottomrule
\end{longtable}

\section{How to Explain Why the Brand Fits}

Avoid generic praise such as ``I like the concept.'' Tie your interest to operating logic.

\begin{scriptbox}{Brand-fit statement}
``What attracts me to your brand is not just the concept, but the operating system behind it. I am looking for a model with strong unit economics, repeatable training, clear local marketing support, disciplined field support, and a franchisee profile that matches how I want to operate.''
\end{scriptbox}

\begin{scriptbox}{More advanced version}
``I am not looking to buy a logo. I am looking to buy into a system with disciplined execution, strong franchisee support, realistic startup assumptions, and a credible path to profitability. That is why I want to understand how your best operators succeed and where new franchisees typically struggle.''
\end{scriptbox}

\section{Strategic Questions to Ask}

Ask questions that sound like an investor and operator, not a casual buyer. The goal is to test system quality, support quality, economics, territory strength, and candidate fit.

\subsection{Brand and market fit}

\begin{itemize}
  \item ``What makes your top-performing franchisees successful compared with average performers?''
  \item ``What are the most common reasons new franchisees struggle in the first 12 to 24 months?''
  \item ``How do you evaluate whether a territory is strong enough to support a new location?''
  \item ``What are you seeing in the \placeholder{Atlanta / Georgia / target market} market specifically?''
  \item ``Which competitors do you most often displace, and what is your strongest differentiation against them?''
\end{itemize}

\subsection{FDD and economics}

\begin{itemize}
  \item ``Which FDD items do you recommend I pay closest attention to when evaluating your system?''
  \item ``Does your FDD include Item 19 financial performance representations, and how should I interpret them conservatively?''
  \item ``What are the most commonly underestimated startup or operating costs?''
  \item ``What working-capital cushion do you recommend beyond the stated initial investment range?''
  \item ``How do royalties, brand-fund contributions, technology fees, training fees, and renewal costs change over time?''
\end{itemize}

\subsection{Training, launch, and field support}

\begin{itemize}
  \item ``What does onboarding look like from signing through opening?''
  \item ``How much support do franchisees receive with site selection, hiring, launch marketing, vendor setup, and first-year operations?''
  \item ``What does ongoing field support look like after opening?''
  \item ``How do you measure whether a new franchisee is on track during the first 90, 180, and 365 days?''
\end{itemize}

\subsection{Financing and SBA readiness}

\begin{itemize}
  \item ``Is your brand listed in the SBA Franchise Directory?''
  \item ``Do franchisees commonly use SBA-backed financing, and do you have lender relationships familiar with your model?''
  \item ``What financing mistakes do candidates make during the approval process?''
\end{itemize}

\subsection{Candidate fit}

\begin{itemize}
  \item ``Based on what I have shared, do I match the financial, operational, and cultural profile you look for in a franchisee?''
  \item ``What would you need to see from me to move to the next stage?''
  \item ``Are there any concerns about my fit at this stage?''
\end{itemize}

\section{What Not to Say}

\begin{longtable}{>{\raggedright\arraybackslash}p{0.38\linewidth}>{\raggedright\arraybackslash}p{0.56\linewidth}}
\toprule
\textbf{Avoid saying} & \textbf{Say this instead} \\
\midrule
``How much money can I make?'' & ``What performance ranges are disclosed in the FDD, and what assumptions should I use when building a conservative pro forma?'' \\
``Is this passive income?'' & ``What level of owner involvement do your strongest franchisees typically have during the first year?'' \\
``Can I negotiate the franchise agreement?'' & ``Which terms are standard across the system, and are there any items that commonly vary by market, development schedule, or ownership structure?'' \\
``I love the brand and just want to get started.'' & ``I am interested in the brand, but I want to validate the economics, support model, territory strategy, and franchisee experience before moving forward.'' \\
``I do not know my budget yet.'' & ``I am targeting an all-in investment range of \placeholder{amount} and want to confirm whether that aligns with your current requirements.'' \\
\bottomrule
\end{longtable}

\section{Closing the Meeting}

End by forcing clarity on fit and next steps.

\begin{scriptbox}{Professional close}
``This has been helpful. Based on what we discussed, I would like to understand the next stage in your process. What materials should I review, what qualifications do you need from me, and what is the expected timeline from this initial conversation to discovery day, FDD review, and possible franchise agreement?''
\end{scriptbox}

\begin{scriptbox}{Direct fit question}
``Based on what I have shared, do you see me as a strong candidate for your brand? Are there any concerns I should address before moving forward?''
\end{scriptbox}

\section{Post-Meeting Follow-Up Email Template}

\begin{scriptbox}{Follow-up email}
\textbf{Subject:} Follow-up on \placeholder{Brand Name} franchise discussion\\[0.4em]
\textbf{Body:}\\
Hello \placeholder{Name},\\[0.25em]
Thank you for taking the time to speak with me about \placeholder{Brand Name}. I appreciated the overview of the franchise model, candidate profile, training process, and market opportunity.\\[0.25em]
Based on our conversation, I remain interested in evaluating the opportunity further. My next diligence priorities are reviewing the FDD, understanding the full investment range and working-capital expectations, validating the support model, and speaking with current franchisees where appropriate.\\[0.25em]
Please send the next-step materials and any qualification documents you need from me. I am also happy to provide additional detail on my liquid capital, financing plan, target market, and ownership timeline.\\[0.25em]
Best,\\
\placeholder{Your Name}
\end{scriptbox}

\section{Pre-Call Checklist}

\begin{itemize}
  \item Confirm your liquid-capital range and maximum all-in investment range.
  \item Decide whether your target model is owner-operator, semi-absentee, multi-unit, or area development.
  \item Identify your target market and why it makes sense demographically and operationally.
  \item Review the franchisor's website, current territories, public growth announcements, and available franchise materials.
  \item Prepare at least three brand-specific questions that cannot be answered by the website.
  \item Confirm whether you want SBA-backed financing and whether the brand appears in the SBA Franchise Directory.
  \item Prepare a post-meeting diligence plan: FDD review, franchisee validation calls, professional review, financing review, and conservative pro forma.
\end{itemize}

\section{Red Flags to Watch For}

\begin{warningbox}{Warning signs during the sales conversation}
\begin{itemize}
  \item Pressure to pay quickly before you have reviewed the FDD and agreements.
  \item Vague answers about startup costs, working capital, training, field support, or franchisee performance.
  \item Overemphasis on brand excitement without clear operating expectations.
  \item Resistance to franchisee validation calls or professional review.
  \item Claims of easy passive income without clear labor, management, and execution requirements.
  \item Unit economics discussed without clear disclosure boundaries or FDD support.
\end{itemize}
\end{warningbox}

\section{Best Overall Script}

\begin{keybox}{Use this when you need a polished summary}
``I am looking for a franchise opportunity where my capital, professional background, and long-term ownership goals align with a proven operating system. I am interested in \placeholder{Brand Name} because I see potential fit in \placeholder{target market}, but I want to evaluate the opportunity carefully through the FDD, franchisee validation, unit economics, training model, and territory strategy. My goal today is to understand whether I am a strong candidate for your system and whether your system is the right fit for me.''
\end{keybox}

\renewcommand{\refname}{\color{accent}References}
\begin{thebibliography}{9}

\bibitem{ftc-rule}
Federal Trade Commission.
\textit{Franchise Rule}.
\url{https://www.ftc.gov/legal-library/browse/rules/franchise-rule}

\bibitem{ftc-consumer-guide}
Federal Trade Commission.
\textit{A Consumer's Guide to Buying a Franchise}.
\url{https://www.ftc.gov/business-guidance/resources/consumers-guide-buying-franchise}

\bibitem{ftc-deep-dive}
Federal Trade Commission.
\textit{Franchise Fundamentals: Taking a Deep Dive into the Franchise Disclosure Document}.
\url{https://www.ftc.gov/business-guidance/blog/2023/05/franchise-fundamentals-taking-deep-dive-franchise-disclosure-document}

\bibitem{sba-directory}
U.S. Small Business Administration.
\textit{SBA Franchise Directory}.
\url{https://www.sba.gov/business-guide/plan-your-business/buy-existing-business-or-franchise/sba-franchise-directory}

\bibitem{ifa-diligence}
International Franchise Association.
\textit{Diligence in Franchising}.
\url{https://www.franchise.org/franchising-overview/diligence-in-franchising/}

\end{thebibliography}

\vfill
\noindent\textcolor{midgray}{\small This handout is for meeting preparation and business diligence. It is not legal, tax, accounting, or financial advice. Review the FDD and franchise agreement with qualified professionals before signing or paying money.}

\end{document}
