
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.72in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{pdflscape}
\usepackage{listings}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=blue,citecolor=black}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{EAF3F8}
\definecolor{LightGray}{HTML}{F5F7FA}
\definecolor{RuleGray}{HTML}{B7C3D0}
\definecolor{RiskAmber}{HTML}{FFF4D6}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.45em}
\setlist[itemize]{leftmargin=1.2em,itemsep=0.18em,topsep=0.2em}
\setlist[enumerate]{leftmargin=1.4em,itemsep=0.18em,topsep=0.2em}
\pagestyle{fancy}
\setlength{\headheight}{14pt}
\fancyhf{}
\lhead{Gauntlt DevSecOps Security Testing}
\rhead{\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\sloppy
\lstset{basicstyle=\ttfamily\small,breaklines=true,columns=fullflexible,keepspaces=true,frame=single,framerule=0.3pt,rulecolor=\color{RuleGray},backgroundcolor=\color{LightGray}}
\newcommand{\TitleBlock}[2]{%
  \begin{center}
  {\LARGE\bfseries\color{HeaderBlue} #1}\\[0.35em]
  {\large #2}\\[0.4em]
  {\small Prepared for DevSecOps implementation planning | Updated May 6, 2026}
  \end{center}
  \vspace{0.4em}\hrule\vspace{0.8em}
}
\newsavebox{\gauntltbox}
\newenvironment{Callout}[1]{%
  \par\vspace{0.3em}\noindent\begin{lrbox}{\gauntltbox}\begin{minipage}{0.96\linewidth}\textbf{#1}\par\smallskip}
  {\end{minipage}\end{lrbox}\fcolorbox{RuleGray}{LightBlue}{\usebox{\gauntltbox}}\par\vspace{0.4em}}
\newenvironment{RiskBox}[1]{%
  \par\vspace{0.3em}\noindent\begin{lrbox}{\gauntltbox}\begin{minipage}{0.96\linewidth}\textbf{#1}\par\smallskip}
  {\end{minipage}\end{lrbox}\fcolorbox{RuleGray}{RiskAmber}{\usebox{\gauntltbox}}\par\vspace{0.4em}}
\newcommand{\checkbox}{\raisebox{0.2ex}{\fbox{\rule{0pt}{0.9ex}\rule{0.9ex}{0pt}}}}
\begin{document}

\TitleBlock{Gauntlt CI/CD Integration Architecture Guide}{Reference pattern for security-as-code DAST checks}

\section{Architecture Objective}
The objective is to add repeatable, readable, and auditable dynamic security checks to a CI/CD pipeline using Gauntlt as the orchestration layer. Gauntlt should execute against an authorized target environment and produce evidence that can be consumed by developers, AppSec, and release governance.

\section{Reference Architecture}
\begin{enumerate}
  \item \textbf{Developer commits code}. Source changes and attack files are versioned together.
  \item \textbf{CI builds and tests}. Unit, integration, SAST, SCA, and packaging stages run first.
  \item \textbf{CD deploys to a test environment}. The target is a controlled staging or ephemeral environment.
  \item \textbf{Gauntlt runner executes attack files}. The runner invokes external tools such as \texttt{curl}, \texttt{nmap}, \texttt{sqlmap}, \texttt{dirb}, and \texttt{SSLyze}.
  \item \textbf{Results are evaluated}. Pass/fail output becomes a release signal.
  \item \textbf{Evidence is retained}. Pipeline logs, attack files, output, and target details are archived.
  \item \textbf{Failures create work}. Blocking findings are routed to the owning team with reproducible evidence.
\end{enumerate}

\section{Component Responsibilities}
\begin{tabularx}{\linewidth}{@{}L{0.23\linewidth}Y@{}}
\toprule
\textbf{Component} & \textbf{Responsibility} \\
\midrule
Source repository & Stores application code, attack files, pipeline definitions, and remediation history. \\
CI runner & Builds the application, prepares test dependencies, and runs Gauntlt. \\
Deployment platform & Deploys the candidate release into an approved test environment. \\
Gauntlt & Reads \texttt{.attack} files, invokes adapters/tools, and evaluates pass/fail expectations. \\
Security tools & Perform the actual scan, request, probe, or validation logic. \\
Artifact store & Keeps reports, logs, target metadata, and evidence bundles. \\
Issue tracker & Tracks remediation, risk acceptance, and closure evidence. \\
\bottomrule
\end{tabularx}

\section{Control Flow}
\begin{enumerate}
  \item Pipeline retrieves attack files from the same repository or a centrally approved security test repository.
  \item Runner validates that Gauntlt and required adapters are installed.
  \item Runner executes approved attack suites based on application type and environment.
  \item Gauntlt invokes command-line tools and parses outputs based on scenario assertions.
  \item Pipeline fails if a mandatory attack fails.
  \item Results are attached to the pipeline run and linked to any remediation ticket.
\end{enumerate}

\section{Recommended Attack Categories}
\begin{longtable}{@{}L{0.24\linewidth}L{0.25\linewidth}L{0.41\linewidth}@{}}
\toprule
\textbf{Category} & \textbf{Example Tool} & \textbf{Pipeline Gate Example} \\
\midrule
Expected ports & \texttt{nmap} & Only 443 is externally reachable for the web tier. \\
HTTP security headers & \texttt{curl} & Required headers are present on authenticated and unauthenticated routes. \\
TLS configuration & \texttt{SSLyze} & Deprecated protocols are disabled and certificate validation passes. \\
Directory exposure & \texttt{dirb} or equivalent & Sensitive directories are not discoverable. \\
SQL injection regression & \texttt{sqlmap} & Known protected endpoints do not exhibit SQL injection behavior. \\
Custom controls & Shell or Python script & Organization-specific checks return expected status. \\
\bottomrule
\end{longtable}

\section{Design Decisions}
\begin{itemize}
  \item \textbf{Runner isolation}: execute tests from a dedicated runner identity with minimal permissions.
  \item \textbf{Target isolation}: prefer staging or ephemeral environments for active checks.
  \item \textbf{Failure policy}: distinguish blocking failures from warning-only observations.
  \item \textbf{Evidence policy}: retain attack file, command output, exit status, target version, and exception records.
  \item \textbf{Maintenance policy}: periodically review Gauntlt compatibility and adapter health because Gauntlt is a legacy framework.
\end{itemize}

\begin{RiskBox}{Safety Constraint}
Do not run dynamic attacks against systems outside the approved scope. Even simple probes can trigger monitoring, rate limits, incident response, or availability concerns.
\end{RiskBox}


\section*{References}
\begin{itemize}
  \item Gauntlt official site: \url{http://gauntlt.org/}
  \item Gauntlt GitHub repository: \url{https://github.com/gauntlt/gauntlt}
  \item Gauntlt Starter Kit: \url{https://github.com/gauntlt/gauntlt-starter-kit}
  \item Gauntlt Demo: \url{https://github.com/gauntlt/gauntlt-demo}
\end{itemize}

\end{document}



