
% =====================================================================
% Study Plan — Cloud Computing Security (User Story Cards — Full)
% =====================================================================
\documentclass[11pt,letterpaper]{article}

% --- Encoding & layout ---
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[letterpaper,margin=1in]{geometry}
\usepackage{parskip}

% --- Structure & visuals ---
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue!60!black, urlcolor=blue!60!black, citecolor=blue!60!black}
\usepackage{enumitem}
\usepackage{array}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[most]{tcolorbox}
\usepackage{tikz}
\usepackage{textcomp} % for \textless \textgreater

% ---------------------------------------------------------------------
% Style tokens
% ---------------------------------------------------------------------
\definecolor{CardBorder}{RGB}{210,210,210}
\definecolor{CardStripeA}{RGB}{27,78,138}   % deep blue
\definecolor{CardStripeB}{RGB}{140,180,220} % light blue
\definecolor{TagBack}{gray}{0.95}
\definecolor{TagFrame}{gray}{0.70}
\definecolor{Muted}{gray}{0.35}

% --- Tcolorbox style for the story card (with left gradient stripe) ---
\tcbset{
  cloudcard/.style={
    enhanced,
    breakable,
    colback=white,
    colframe=CardBorder,
    boxrule=0.35pt,
    arc=2mm,
    left=4mm, right=4mm, top=2mm, bottom=2mm,
    fonttitle=\bfseries,
    titlerule=0pt,
    coltitle=black,
    boxed title style={colback=white, colframe=white, boxrule=0pt},
    attach boxed title to top left={xshift=0mm, yshift=-1mm},
    overlay unbroken={\fill[left color=CardStripeA,right color=CardStripeB] (frame.north west) rectangle ([xshift=2mm]frame.south west);},
    overlay first={\fill[left color=CardStripeA,right color=CardStripeB] (frame.north west) rectangle ([xshift=2mm]frame.south west);},
    overlay middle={\fill[left color=CardStripeA,right color=CardStripeB] (frame.north west) rectangle ([xshift=2mm]frame.south west);},
    overlay last={\fill[left color=CardStripeA,right color=CardStripeB] (frame.north west) rectangle ([xshift=2mm]frame.south west);},
  }
}

% --- Pill/Tag style ---
\newtcbox{\pill}{on line, colback=TagBack, colframe=TagFrame, boxrule=0.3pt,
  top=1pt, bottom=1pt, left=4pt, right=4pt, arc=2pt}

% --- Two-column key/value table (no tabularx) ---
\newenvironment{kvtable}{%
  \begin{tabular}{@{}>{\bfseries\arraybackslash}p{0.26\linewidth} p{0.70\linewidth}@{}}
}{\end{tabular}}

% --- BDD table (no tabularx) ---
\newenvironment{bdd}{%
  \begin{tabular}{@{}>{\bfseries\arraybackslash}p{0.26\linewidth} p{0.70\linewidth}@{}}
}{\end{tabular}}

% --- Checklist style (checkbox squares) ---
\newlist{checklist}{itemize}{1}
\setlist[checklist]{label=$\square$, left=1.6em, itemsep=0.25em, topsep=0.25em}

% --- Footnote line for DoR / DoD ---
\newcommand{\DoRDoD}[2]{%
  \vspace{0.35em}
  {\footnotesize\color{Muted}%
    \textit{Definition of Ready:} #1 \; \textbullet\; 
    \textit{Definition of Done:} #2%
  }%
}

% --- Story Card environment ---
\newenvironment{storycard}[2]{%
  \begin{tcolorbox}[cloudcard, title={#1 --- #2}]%
}{\end{tcolorbox}}

% --- Helper label ---
\newcommand{\sectionlabel}[1]{\vspace{0.35em}\textbf{#1}\par}

% =====================================================================
\begin{document}

\begin{center}
{\LARGE \textbf{Study Plan --- Cloud Computing Security (User Story Cards)}}\\[4pt]
{\large Aligned to ``Cloud Computing Security: Foundations and Challenges (2nd ed.)''}\\[2pt]
\textit{One story card per chapter with example tasks and BDD acceptance criteria.}
\end{center}

\vspace{0.5em}
\hrule
\vspace{0.75em}

\tableofcontents
\newpage

% ===============================
\section{Section I — Introduction (Ch. 1--5)}
% ===============================

\begin{storycard}{CCS-1}{Ch. 1 — Foundations \& Essentials}
\begin{kvtable}
Epic / Feature & Cloud Foundations\\
Business Value & Establish shared vocabulary and ownership to reduce onboarding confusion and security drift.\\
Priority / Estimate & Priority: Must \quad SP: 3\\
Persona & Developer starting on a cloud project\\
Dependencies & Sandbox account, org security policy, provider docs\\
Assumptions / Risks & Teams may interpret shared responsibility differently; risk of inconsistent defaults\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a developer, I want to document cloud foundations and shared responsibility so that I make secure choices and keep main green.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}\;\pill{Accessibility}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Shared understanding is captured\\
Given & A sandbox and policy references\\
When & I create a one-page shared responsibility matrix and a glossary of essential characteristics\\
Then & The documents are linked in onboarding and referenced by future stories\\
\end{bdd}
\DoRDoD{Persona clear; AC drafted; Dependencies identified.}{Matrix and glossary committed; reviewed by AppSec; referenced in README.}
\sectionlabel{Tasks}
\begin{checklist}
\item Write \texttt{docs/shared-responsibility-matrix.md}.
\item Capture five essential characteristics of cloud with examples.
\item Add link in \texttt{README.md} and onboarding checklist.
\item Open issues for disagreements or unclear ownership.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-2}{Ch. 2 — Overview of Cloud Computing}
\begin{kvtable}
Epic / Feature & Cloud Foundations\\
Business Value & Reduce design churn by listing provider primitives and portability considerations.\\
Priority / Estimate & Priority: Must \quad SP: 3\\
Persona & Solution architect\\
Dependencies & Access to AWS, Azure, and GCP docs\\
Assumptions / Risks & Feature parity differs; lock-in may be acceptable for value\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a solution architect, I want a lift-and-shift bill of materials so that teams can compare provider primitives and portability risks.}
\sectionlabel{Non-Functional}\pill{Performance}\;\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Cross-provider comparison documented\\
Given & A reference three-tier app\\
When & I create a bill of materials for AWS, Azure, and GCP with security notes\\
Then & Tradeoffs and lock-in vectors are captured with mitigation ideas\\
\end{bdd}
\DoRDoD{Reference app defined; providers chosen.}{BOM spreadsheet checked in; reviewed by platform and security.}
\sectionlabel{Tasks}
\begin{checklist}
\item Produce \texttt{bom/cloud-bom.xlsx} with compute, storage, network, IAM.
\item Annotate security defaults and required hardening per service.
\item Note portability blockers and alternatives.
\item File follow-up ADR if a single provider is selected.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-3}{Ch. 3 — Security Baselines}
\begin{kvtable}
Epic / Feature & Baseline Security\\
Business Value & Create fast feedback on misconfigurations and reduce time to first fix.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Cloud security engineer\\
Dependencies & Sandbox account/project, Prowler or ScoutSuite, read-only creds\\
Assumptions / Risks & Findings volume may be high; prioritize by risk\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a cloud security engineer, I want to run a baseline scan so that we triage top risks and create a remediation backlog.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}\;\pill{Privacy}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Baseline executed and triaged\\
Given & Read-only scanner access\\
When & I run Prowler or ScoutSuite and export results\\
Then & Top ten findings have owners, due dates, and tickets\\
\end{bdd}
\DoRDoD{Scanner configured; scope agreed.}{Report stored; backlog created; first PR opened for a fix.}
\sectionlabel{Tasks}
\begin{checklist}
\item Execute scan and export JSON plus HTML.
\item Create \texttt{baseline-findings.md} with risk ranking.
\item Open issues for top findings; tag teams.
\item Document suppressions and rationale.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-4}{Ch. 4 — Privacy \& Trust Baselines}
\begin{kvtable}
Epic / Feature & Data Protection\\
Business Value & Reduce regulatory risk through inventory, tagging, and residency controls.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Data protection officer / privacy engineer\\
Dependencies & Data catalog access, tagging standard\\
Assumptions / Risks & Unknown data flows; legacy datasets may be unlabeled\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a privacy engineer, I want to inventory and tag datasets so that location, residency, and access are controlled.}
\sectionlabel{Non-Functional}\pill{Privacy}\;\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Inventory completed for one workload\\
Given & Access to data catalog and cloud storage\\
When & I tag datasets by sensitivity and residency\\
Then & Policies and controls reference the tags for enforcement\\
\end{bdd}
\DoRDoD{Scope selected; tagging keys agreed.}{Inventory CSV committed; policies updated; gaps tracked.}
\sectionlabel{Tasks}
\begin{checklist}
\item Export dataset list and classify sensitivity.
\item Tag objects and databases by owner and residency.
\item Map to retention and legal requirements.
\item Open tasks to fix unlabeled datasets.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-5}{Ch. 5 — IaaS Focus}
\begin{kvtable}
Epic / Feature & Compute, Storage, Network Hardening\\
Business Value & Prevent common misconfigurations at the IaaS layer.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Platform engineer\\
Dependencies & Terraform repo, VPC/VNet, KMS, IAM\\
Assumptions / Risks & Breaking changes if defaults tighten without comms\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a platform engineer, I want least-privilege IAM and default-deny networking so that IaaS resources are secure by default.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Secure defaults enforced\\
Given & A Terraform baseline module\\
When & I add network policies and least-privilege roles\\
Then & New resources inherit secure defaults and tests verify enforcement\\
\end{bdd}
\DoRDoD{Module owners onboard; tests planned.}{Module released; pipelines pass; docs updated.}
\sectionlabel{Tasks}
\begin{checklist}
\item Harden security groups and route tables to default deny.
\item Create least-privilege role for compute and storage access.
\item Add unit tests for denial of wildcard permissions.
\item Publish module usage guide.
\end{checklist}
\end{storycard}

% ===============================
\section{Section II — Risk Analysis \& Responsibility (Ch. 6--9)}
% ===============================

\begin{storycard}{CCS-6}{Ch. 6 — Risk \& Trust Assessment Schemes}
\begin{kvtable}
Epic / Feature & Risk Management\\
Business Value & Make risk-driven decisions using a consistent scheme.\\
Priority / Estimate & Priority: Must \quad SP: 3\\
Persona & Risk analyst\\
Dependencies & Risk register template, stakeholder access\\
Assumptions / Risks & Disagreement on scoring methods\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a risk analyst, I want a risk register using a simple method so that leaders can compare and prioritize risks.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Register populated for one project\\
Given & A template and stakeholder interviews\\
When & I capture top risks with likelihood and impact\\
Then & Owners and treatments are assigned and accepted\\
\end{bdd}
\DoRDoD{Template chosen; scope set.}{Register committed; review notes added; next review scheduled.}
\sectionlabel{Tasks}
\begin{checklist}
\item Create \texttt{risk-register.xlsx}.
\item Schedule 30-minute interviews with owners.
\item Draft KRIs for top items.
\item Publish review cadence.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-7}{Ch. 7 — Managing Risk in the Cloud}
\begin{kvtable}
Epic / Feature & Risk Treatment\\
Business Value & Reduce probability or impact through concrete control choices.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Security program manager\\
Dependencies & Control catalog, platform team availability\\
Assumptions / Risks & Limited capacity; tradeoffs required\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security program manager, I want risk treatments tied to controls so that we see measurable reduction in risk.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Treatments committed\\
Given & A prioritized risk list\\
When & I link risks to controls and create implementation tickets\\
Then & KRIs and due dates are visible on the roadmap\\
\end{bdd}
\DoRDoD{Owners identified; catalog agreed.}{Tickets created; dashboard shows KRIs; exec summary published.}
\sectionlabel{Tasks}
\begin{checklist}
\item Map each top risk to controls and team.
\item Create epics with AC and success criteria.
\item Add KRI queries to dashboard.
\item Send monthly status note.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-8}{Ch. 8 — Cloud Security Risk Management}
\begin{kvtable}
Epic / Feature & Policy as Code\\
Business Value & Prevent misconfigurations from merging to main.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & DevOps engineer\\
Dependencies & CI pipeline, IaC repo, OPA or Checkov\\
Assumptions / Risks & False positives can slow teams\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a DevOps engineer, I want policy-as-code in CI so that insecure IaC cannot be merged.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & CI blocks insecure patterns\\
Given & A repo with Terraform modules\\
When & I add checks for public buckets and wildcard IAM\\
Then & Failing PRs show clear messages and remediation links\\
\end{bdd}
\DoRDoD{Rules selected; threshold agreed.}{Checks enforced; docs added; exceptions process documented.}
\sectionlabel{Tasks}
\begin{checklist}
\item Add OPA or Checkov job to CI.
\item Write two guardrails: storage public access and IAM wildcards.
\item Add test fixtures that fail until fixed.
\item Document remediation steps.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-9}{Ch. 9 — Risk Mitigation Methods}
\begin{kvtable}
Epic / Feature & Mitigation Patterns\\
Business Value & Choose effective mitigations for data and perimeter.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Security architect\\
Dependencies & KMS, WAF, segmentation capability\\
Assumptions / Risks & Cost and latency tradeoffs\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security architect, I want to compare mitigation options so that we pick the best fit per risk.}
\sectionlabel{Non-Functional}\pill{Performance}\;\pill{Security}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Decision record created\\
Given & Candidate mitigations and constraints\\
When & I evaluate tokenization, KMS, WAF, and micro-segmentation\\
Then & An ADR records the chosen approach and rationale\\
\end{bdd}
\DoRDoD{Alternatives listed; constraints known.}{ADR merged; next steps filed as tickets.}
\sectionlabel{Tasks}
\begin{checklist}
\item Draft ADR comparing options.
\item Measure expected latency impact.
\item Validate cost estimates.
\item Get sign-off from stakeholders.
\end{checklist}
\end{storycard}

% ===============================
\section{Section III — Securing Cloud Infrastructure (Ch. 10--16)}
% ===============================

\begin{storycard}{CCS-10}{Ch. 10 — Access Policy Specification \& Enforcement}
\begin{kvtable}
Epic / Feature & Identity and Access Management\\
Business Value & Reduce privilege and prevent policy sprawl.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & IAM engineer\\
Dependencies & Policy engine, repo access\\
Assumptions / Risks & Legacy policies contain wildcards\\
\end{kvtable}
\sectionlabel{Story}
\textit{As an IAM engineer, I want least-privilege policies with automated checks so that access stays minimal and auditable.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Wildcards prevented\\
Given & Policy files in version control\\
When & I add a rule that fails \texttt{"Action: *"} or \texttt{"Resource: *"}\\
Then & PRs fail with a clear message and link to examples\\
\end{bdd}
\DoRDoD{Policy style guide drafted.}{Rules enabled; examples published; old policies queued for refactor.}
\sectionlabel{Tasks}
\begin{checklist}
\item Create policy lints for wildcards and unused permissions.
\item Build reusable role templates.
\item Add presubmit unit tests.
\item Document escalation and exceptions.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-11}{Ch. 11 — Cryptographic Key Management}
\begin{kvtable}
Epic / Feature & Data Encryption\\
Business Value & Protect sensitive data at rest and in transit with managed lifecycle.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Security engineer\\
Dependencies & KMS, secrets store, CI access\\
Assumptions / Risks & Rotation can disrupt apps if not coordinated\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security engineer, I want envelope encryption with automated rotation so that key compromise risk is minimized.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}\;\pill{Privacy}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Rotation executed without downtime\\
Given & Apps using KMS for envelope encryption\\
When & I rotate keys and re-encrypt materials in staging\\
Then & Metrics show no errors and runbooks are updated\\
\end{bdd}
\DoRDoD{Apps instrumented; staging ready.}{Rotation proved in staging; prod schedule approved; runbook linked.}
\sectionlabel{Tasks}
\begin{checklist}
\item Implement envelope encryption example.
\item Add rotation job and alarms.
\item Create decrypt fallback and test.
\item Publish runbook with rollback plan.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-12}{Ch. 12 — Distributed Access Control}
\begin{kvtable}
Epic / Feature & Service Authorization\\
Business Value & Enforce consistent policies across microservices.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Platform engineer\\
Dependencies & OPA sidecar or service mesh, K8s cluster\\
Assumptions / Risks & Policy changes must be versioned and tested\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a platform engineer, I want sidecar policy enforcement so that services follow the same authorization rules.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Policy enforced at runtime\\
Given & A sample API deployed on Kubernetes\\
When & I deny requests lacking a required claim\\
Then & Access logs and metrics confirm policy decisions\\
\end{bdd}
\DoRDoD{Sidecar pattern approved.}{Policy repo created; sample deployed; dashboards live.}
\sectionlabel{Tasks}
\begin{checklist}
\item Deploy OPA sidecar with Rego bundle.
\item Write allow/deny rule based on JWT claim.
\item Emit decision logs to SIEM.
\item Add rollout and rollback steps.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-13}{Ch. 13 — User-Side Key Controls}
\begin{kvtable}
Epic / Feature & Client-Side Encryption\\
Business Value & Reduce provider breach blast radius.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Client app developer\\
Dependencies & Crypto library, performance test tool\\
Assumptions / Risks & UX latency and key handling complexity\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a client developer, I want to encrypt data before upload so that exposure risk is reduced.}
\sectionlabel{Non-Functional}\pill{Performance}\;\pill{Security}\;\pill{Privacy}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Client-side encryption prototype\\
Given & A demo dataset and upload path\\
When & I enable client encryption and measure overhead\\
Then & Latency impact and throughput are documented with limits\\
\end{bdd}
\DoRDoD{Test dataset ready.}{Prototype merged; thresholds set; backlog items filed.}
\sectionlabel{Tasks}
\begin{checklist}
\item Build client encryption function.
\item Add perf test for 1 MB and 10 MB files.
\item Document key custody options.
\item Decide thresholds and guardrails.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-14}{Ch. 14 — Security Essentials \& Reference Architectures}
\begin{kvtable}
Epic / Feature & Reference Architecture\\
Business Value & Provide a secure, repeatable pattern for web workloads.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Solutions architect\\
Dependencies & Diagramming tool, baseline modules\\
Assumptions / Risks & Pattern must be simple yet adaptable\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a solutions architect, I want a secure reference architecture so that teams can adopt consistent controls quickly.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Reference published\\
Given & Common requirements for a web app\\
When & I include logging, monitoring, KMS, WAF, and CIEM\\
Then & The diagram and bill of materials are reviewed and adopted\\
\end{bdd}
\DoRDoD{Scope finalized.}{Drawio file and BOM committed; approval recorded.}
\sectionlabel{Tasks}
\begin{checklist}
\item Draw diagram with shared services.
\item List components and controls.
\item Map to controls framework.
\item Create adoption checklist.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-15}{Ch. 15 — Architecture \& Security Concepts}
\begin{kvtable}
Epic / Feature & Zero Trust \& Immutable Infra\\
Business Value & Reduce lateral movement and configuration drift.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Platform architect\\
Dependencies & Image builder, autoscaling, CI/CD\\
Assumptions / Risks & Rebuild cadence must align with release cycles\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a platform architect, I want immutable images with micro-segmentation so that drift and attack paths are minimized.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Mutable to immutable migration\\
Given & A mutable VM pattern\\
When & I move to image-based deployments with autoscaling\\
Then & Patching occurs via rebuild and health checks stay green\\
\end{bdd}
\DoRDoD{Current pattern documented.}{New pattern deployed in staging; runbook updated.}
\sectionlabel{Tasks}
\begin{checklist}
\item Create hardened base image.
\item Update pipeline for image promotion.
\item Add health checks and rollbacks.
\item Retire in-place patching runbooks.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-16}{Ch. 16 — Secure Cloud Architecture}
\begin{kvtable}
Epic / Feature & Threat-Driven Design\\
Business Value & Align controls with threats and evidence.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Security architect\\
Dependencies & Threat modeling tool, control catalog\\
Assumptions / Risks & Over-design if threats not prioritized\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security architect, I want a threat model linked to controls so that architectural choices are justified.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & STRIDE model created\\
Given & The reference architecture\\
When & I document threats and mitigations\\
Then & Missing controls generate backlog items with owners\\
\end{bdd}
\DoRDoD{Architecture stable.}{Threat model published; issues created and prioritized.}
\sectionlabel{Tasks}
\begin{checklist}
\item Run a STRIDE session.
\item Map threats to controls.
\item Create tickets for gaps.
\item Review annually or after changes.
\end{checklist}
\end{storycard}

% ===============================
\section{Section IV — OS \& Network Security (Ch. 17--18)}
% ===============================

\begin{storycard}{CCS-17}{Ch. 17 — Locking Down Cloud Servers}
\begin{kvtable}
Epic / Feature & Host Hardening\\
Business Value & Reduce attack surface of compute instances.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Systems engineer\\
Dependencies & CIS benchmark, Ansible, osquery\\
Assumptions / Risks & Compatibility issues with legacy apps\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a systems engineer, I want hardened server images validated by scans so that host-level risk is reduced.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Hardened image baseline\\
Given & A base operating system image\\
When & I apply Ansible hardening and validate with osquery\\
Then & Compliance score meets target and drift alerts are enabled\\
\end{bdd}
\DoRDoD{Baseline chosen.}{Image published; score documented; rollout plan approved.}
\sectionlabel{Tasks}
\begin{checklist}
\item Build Ansible role for hardening.
\item Add osquery pack for controls.
\item Measure compliance score.
\item Define rollout and fallback.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-18}{Ch. 18 — Third-Party Provider Integrity}
\begin{kvtable}
Epic / Feature & Vendor Security\\
Business Value & Reduce supply-chain risk by verifying provider controls.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & TPRM analyst\\
Dependencies & Vendor portal, questionnaire, evidence store\\
Assumptions / Risks & Evidence may be incomplete or outdated\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a TPRM analyst, I want to evaluate a SaaS vendor so that integrity and compliance claims are validated.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Privacy}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Vendor assessed\\
Given & A completed questionnaire and shared evidence\\
When & I review SOC 2, ISO certs, pen test summaries\\
Then & Gaps and compensating controls are recorded with owners\\
\end{bdd}
\DoRDoD{Vendor identified.}{Assessment logged; renewal date tracked; follow-ups filed.}
\sectionlabel{Tasks}
\begin{checklist}
\item Collect attestations and reports.
\item Map to our control set.
\item Record gaps and compensations.
\item Schedule next review.
\end{checklist}
\end{storycard}

% ===============================
\section{Section V — Compliance Requirements (Ch. 19--28)}
% ===============================

\begin{storycard}{CCS-19}{Ch. 19 — Negotiating Security Requirements}
\begin{kvtable}
Epic / Feature & Security Addendum\\
Business Value & Contractualize minimum controls and reporting duties.\\
Priority / Estimate & Priority: Must \quad SP: 3\\
Persona & Security lead / legal partner\\
Dependencies & Contract template, DPA, counsel review\\
Assumptions / Risks & Negotiations may extend timelines\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security lead, I want measurable security clauses so that vendor obligations are enforceable.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Privacy}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Addendum executed\\
Given & A vendor contract in negotiation\\
When & I add breach notice timing, logging, crypto, and SRT clauses\\
Then & The executed contract contains measurable commitments\\
\end{bdd}
\DoRDoD{Template aligned with legal.}{Signed addendum archived; obligations tracked.}
\sectionlabel{Tasks}
\begin{checklist}
\item Draft 10 key clauses.
\item Align with DPA terms.
\item Review with counsel.
\item Track obligations in register.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-20}{Ch. 20 — Legal Compliance for Personal Data}
\begin{kvtable}
Epic / Feature & Privacy Compliance\\
Business Value & Demonstrate lawful processing and accountability.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & Privacy engineer\\
Dependencies & ROPA template, data map, DPO review\\
Assumptions / Risks & Data lineage unknown for some fields\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a privacy engineer, I want a record of processing and DFD so that obligations and flows are clear.}
\sectionlabel{Non-Functional}\pill{Privacy}\;\pill{Security}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & ROPA completed\\
Given & Access to systems and owners\\
When & I capture purposes, lawful basis, retention, and transfers\\
Then & Data flows and controls are documented and approved\\
\end{bdd}
\DoRDoD{Scope bounded; owners engaged.}{ROPA and DFD in repo; review sign-off captured.}
\sectionlabel{Tasks}
\begin{checklist}
\item Build \texttt{privacy/ropa.xlsx}.
\item Draw data flow diagram.
\item Validate retention policies.
\item Add cross-border transfer notes.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-21}{Ch. 21 — Integrity Assurance for Data Outsourcing}
\begin{kvtable}
Epic / Feature & Data Integrity\\
Business Value & Detect tampering and ensure recoverability.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Storage engineer\\
Dependencies & Object lock, versioning, checksum pipeline\\
Assumptions / Risks & Immutability might affect lifecycle costs\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a storage engineer, I want object immutability and verification so that outsourced data integrity is assured.}
\sectionlabel{Non-Functional}\pill{Reliability}\;\pill{Security}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Immutability and verification enabled\\
Given & A critical bucket\\
When & I enable object lock and periodic checksum verification\\
Then & Evidence of integrity is logged and alerts fire on mismatch\\
\end{bdd}
\DoRDoD{Bucket identified.}{Policies active; verification job scheduled; alerts tested.}
\sectionlabel{Tasks}
\begin{checklist}
\item Turn on versioning and object lock.
\item Create checksum job.
\item Store proofs and logs.
\item Test corruption scenario.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-22}{Ch. 22 — Secure Computation Outsourcing}
\begin{kvtable}
Epic / Feature & Confidential Computing\\
Business Value & Protect workloads from host compromise.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Platform engineer\\
Dependencies & Confidential VM or enclave offering\\
Assumptions / Risks & Limited tooling; higher cost\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a platform engineer, I want to deploy a confidential VM demo so that sensitive code and data run in a protected environment.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Performance}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Demo deployed\\
Given & Access to confidential VM/Enclave service\\
When & I run a sample workload with attestation\\
Then & Attestation evidence is captured and documented\\
\end{bdd}
\DoRDoD{Service quota available.}{Demo works; evidence stored; decision matrix written.}
\sectionlabel{Tasks}
\begin{checklist}
\item Launch confidential instance.
\item Run attestation example.
\item Capture measurements and logs.
\item Write decision matrix.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-23}{Ch. 23 — Computation Over Encrypted Data}
\begin{kvtable}
Epic / Feature & Searchable Encryption / FHE Survey\\
Business Value & Enable limited queries without decrypting data.\\
Priority / Estimate & Priority: Could \quad SP: 2\\
Persona & Research engineer\\
Dependencies & Sample dataset, library support\\
Assumptions / Risks & Performance and complexity constraints\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a research engineer, I want to prototype encrypted search so that feasibility and limits are documented.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Performance}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Prototype results recorded\\
Given & A text dataset\\
When & I run encrypted keyword search\\
Then & Latency, correctness, and limits are summarized\\
\end{bdd}
\DoRDoD{Dataset ready.}{PoC code and report committed; go/no-go noted.}
\sectionlabel{Tasks}
\begin{checklist}
\item Choose library and scheme.
\item Index and query dataset.
\item Measure latency and size.
\item Summarize findings.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-24}{Ch. 24 — Trusted Computing Technology}
\begin{kvtable}
Epic / Feature & Platform Trust\\
Business Value & Validate boot and workload integrity.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Platform engineer\\
Dependencies & vTPM, secure boot, measurement service\\
Assumptions / Risks & Hardware support varies by provider\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a platform engineer, I want to verify secure boot and vTPM so that platform trust is established.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Attestation enabled\\
Given & A VM image and policy\\
When & I enable secure boot and verify measurements\\
Then & Evidence is stored and non-compliant boots alert\\
\end{bdd}
\DoRDoD{Image pipeline documented.}{Attestation evidence archived; alert tested.}
\sectionlabel{Tasks}
\begin{checklist}
\item Enable secure boot and vTPM.
\item Capture PCR measurements.
\item Store evidence in repo.
\item Add alert for failures.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-25}{Ch. 25 — Trusted Security Tech: Survey \& Gaps}
\begin{kvtable}
Epic / Feature & Capability Heatmap\\
Business Value & Identify maturity and gaps for trusted tech.\\
Priority / Estimate & Priority: Could \quad SP: 2\\
Persona & Security architect\\
Dependencies & Stakeholder input\\
Assumptions / Risks & Divergent views on maturity\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security architect, I want a capability heatmap so that prioritization of trust tech investments is clear.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Heatmap published\\
Given & A list of capabilities\\
When & I score maturity and document gaps\\
Then & Roadmap items and owners are assigned\\
\end{bdd}
\DoRDoD{Capabilities enumerated.}{Heatmap slide shared; roadmap updated.}
\sectionlabel{Tasks}
\begin{checklist}
\item Define scoring rubric.
\item Collect scores from owners.
\item Aggregate and visualize.
\item File roadmap items.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-26}{Ch. 26 — Trusted Computing Proposals}
\begin{kvtable}
Epic / Feature & End-to-End Trust Chain\\
Business Value & Prove device to workload trust with attestations.\\
Priority / Estimate & Priority: Could \quad SP: 2\\
Persona & Security engineer\\
Dependencies & Attestation service, identity provider\\
Assumptions / Risks & Complexity of chain-of-trust proofs\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security engineer, I want an attestation flow design so that trust decisions can be automated.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Flow documented\\
Given & Components for device, boot, and workload\\
When & I design sequence of attestations\\
Then & Verification steps and failure modes are defined\\
\end{bdd}
\DoRDoD{Components inventoried.}{Sequence diagram and notes committed.}
\sectionlabel{Tasks}
\begin{checklist}
\item Draft sequence diagram.
\item List verification artifacts.
\item Define failure responses.
\item Review with platform team.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-27}{Ch. 27 — Government Certification \& Accreditation}
\begin{kvtable}
Epic / Feature & FedRAMP Readiness (example)\\
Business Value & Understand inheritance and required artifacts.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Compliance lead\\
Dependencies & Control catalog, SSP template\\
Assumptions / Risks & Scope must be tightly defined\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a compliance lead, I want a control inheritance map so that authorization scope and responsibilities are clear.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Inheritance mapped\\
Given & A SaaS in scope\\
When & I mark provider vs customer controls\\
Then & The SSP references the map and gaps are tracked\\
\end{bdd}
\DoRDoD{Boundary defined.}{Matrix committed; SSP section drafted; gaps logged.}
\sectionlabel{Tasks}
\begin{checklist}
\item Build inheritance matrix.
\item Tag inherited controls.
\item Draft SSP outline.
\item Create gap tickets.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-28}{Ch. 28 — Government Regulations \& Compliance Risks}
\begin{kvtable}
Epic / Feature & Regulatory Risk Log\\
Business Value & Avoid surprise obligations and penalties.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Compliance analyst\\
Dependencies & Legal counsel, records policy\\
Assumptions / Risks & Regulations evolve; periodic review needed\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a compliance analyst, I want a regulatory risk log so that export, retention, and sector rules are addressed.}
\sectionlabel{Non-Functional}\pill{Privacy}\;\pill{Security}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Risks recorded with treatments\\
Given & A list of applicable regulations\\
When & I log risks and proposed mitigations\\
Then & Owners and deadlines are tracked\\
\end{bdd}
\DoRDoD{Sources identified.}{Log published; review cadence set; items assigned.}
\sectionlabel{Tasks}
\begin{checklist}
\item List applicable regs per region.
\item Identify records and retention needs.
\item Document export-control flags.
\item Assign owners and dates.
\end{checklist}
\end{storycard}

% ===============================
\section{Section VI — Disaster Recovery (Ch. 29--30)}
% ===============================

\begin{storycard}{CCS-29}{Ch. 29 — Regions, Zones, \& Trust Boundaries}
\begin{kvtable}
Epic / Feature & Multi-Region Planning\\
Business Value & Improve resilience and clarify trust boundaries.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Site reliability engineer\\
Dependencies & Provider region matrix, latency data\\
Assumptions / Risks & Cost vs availability tradeoffs\\
\end{kvtable}
\sectionlabel{Story}
\textit{As an SRE, I want a region selection ADR so that latency, compliance, and trust boundaries are explicit.}
\sectionlabel{Non-Functional}\pill{Reliability}\;\pill{Security}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Regions selected\\
Given & Latency and compliance constraints\\
When & I pick primary and failover regions\\
Then & ADR documents tradeoffs and dependencies\\
\end{bdd}
\DoRDoD{Constraints collected.}{ADR merged; DNS and data plans documented.}
\sectionlabel{Tasks}
\begin{checklist}
\item Measure latency from users.
\item Check data residency needs.
\item Choose primary and secondary.
\item Document dependencies and costs.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-30}{Ch. 30 — Availability, Recovery, \& Auditing}
\begin{kvtable}
Epic / Feature & DR Runbook and Evidence\\
Business Value & Prove recoverability and auditing across sites.\\
Priority / Estimate & Priority: Must \quad SP: 5\\
Persona & SRE lead\\
Dependencies & Backup system, failover tooling, logging\\
Assumptions / Risks & Testing may cause temporary disruption\\
\end{kvtable}
\sectionlabel{Story}
\textit{As an SRE lead, I want a game day failover with evidence so that RPO and RTO are validated.}
\sectionlabel{Non-Functional}\pill{Reliability}\;\pill{Security}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Game day executed\\
Given & A DR runbook and staging environment\\
When & I fail over traffic and restore data\\
Then & RPO/RTO targets are met and audit artifacts stored\\
\end{bdd}
\DoRDoD{Runbook drafted; window approved.}{Postmortem completed; evidence archived; action items filed.}
\sectionlabel{Tasks}
\begin{checklist}
\item Dry-run backup restore.
\item Switch traffic in staging.
\item Collect logs and screenshots.
\item Write postmortem with follow-ups.
\end{checklist}
\end{storycard}

% ===============================
\section{Section VII — Advanced Cloud Security (Ch. 31--34)}
% ===============================

\begin{storycard}{CCS-31}{Ch. 31 — Advanced Security Architecture}
\begin{kvtable}
Epic / Feature & Service Mesh and Zero Trust\\
Business Value & Strong identity boundaries between services.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Platform engineer\\
Dependencies & Istio or Linkerd, PKI\\
Assumptions / Risks & mTLS introduces complexity\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a platform engineer, I want mTLS via a service mesh so that service-to-service trust is explicit and auditable.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}\;\pill{Performance}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Mesh policy enforced\\
Given & A cluster and sample services\\
When & I require authenticated identities for calls\\
Then & Unauthorized calls fail and metrics show encrypted traffic\\
\end{bdd}
\DoRDoD{PKI ready; cluster available.}{Policies applied; dashboards in place; rollback steps documented.}
\sectionlabel{Tasks}
\begin{checklist}
\item Install mesh and issue certificates.
\item Enforce mTLS and authZ policy.
\item Expose metrics and logs.
\item Document rollout plan.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-32}{Ch. 32 — Side-Channel Attacks \& Defenses}
\begin{kvtable}
Epic / Feature & Side-Channel Awareness\\
Business Value & Reduce risk from timing and cache leakage.\\
Priority / Estimate & Priority: Could \quad SP: 2\\
Persona & Security researcher\\
Dependencies & Benchmark harness, controlled environment\\
Assumptions / Risks & Synthetic tests may not reflect production\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security researcher, I want to measure a simple timing side-channel and mitigation so that risk is understood.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Performance}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Signal measured and reduced\\
Given & A controlled test harness\\
When & I demonstrate a cache-timing signal and apply a mitigation\\
Then & The signal-to-noise ratio decreases per target threshold\\
\end{bdd}
\DoRDoD{Harness prepared.}{Notebook and results committed; mitigation guidance added.}
\sectionlabel{Tasks}
\begin{checklist}
\item Implement timing measurement.
\item Capture baseline signal.
\item Apply mitigation and remeasure.
\item Document recommendations.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-33}{Ch. 33 — Critical Analysis of Threat Models}
\begin{kvtable}
Epic / Feature & Meta Threat Modeling\\
Business Value & Expose assumptions about control planes and multi-tenancy.\\
Priority / Estimate & Priority: Should \quad SP: 3\\
Persona & Red team lead\\
Dependencies & Access to assumptions and architecture docs\\
Assumptions / Risks & Sensitive topics require careful handling\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a red team lead, I want to challenge threat model assumptions so that blind spots in cloud control planes are addressed.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Assumptions documented and tested\\
Given & An existing threat model\\
When & I write a red-team hypothesis targeting meta-control plane risks\\
Then & Detection ideas and mitigations are proposed and tracked\\
\end{bdd}
\DoRDoD{Model available.}{Hypothesis published; action items filed; follow-up scheduled.}
\sectionlabel{Tasks}
\begin{checklist}
\item List explicit assumptions.
\item Draft hypothesis and tests.
\item Propose mitigations and detections.
\item Track actions to closure.
\end{checklist}
\end{storycard}

\begin{storycard}{CCS-34}{Ch. 34 — Future Directions, Risks \& Challenges}
\begin{kvtable}
Epic / Feature & Forward-Looking Roadmap\\
Business Value & Prepare for PQ crypto, confidential ML, and SBOM/SLSA.\\
Priority / Estimate & Priority: Could \quad SP: 2\\
Persona & Security strategist\\
Dependencies & Crypto inventory, CI pipeline, artifact signing\\
Assumptions / Risks & Changing standards and vendor support\\
\end{kvtable}
\sectionlabel{Story}
\textit{As a security strategist, I want a forward-looking roadmap so that the program is ready for emerging risks and controls.}
\sectionlabel{Non-Functional}\pill{Security}\;\pill{Reliability}
\sectionlabel{Acceptance Criteria (BDD)}
\begin{bdd}
Scenario & Roadmap published\\
Given & Current state inventory\\
When & I add PQ readiness, confidential ML, and SLSA milestones\\
Then & Owners and dates are assigned with review cadence\\
\end{bdd}
\DoRDoD{Inventory complete.}{Roadmap committed; review dates on calendar; scorecard added.}
\sectionlabel{Tasks}
\begin{checklist}
\item Build crypto inventory and PQ plan.
\item Define model confidentiality needs.
\item Add SBOM and SLSA targets.
\item Create quarterly scorecard.
\end{checklist}
\end{storycard}

% =====================================================================
\end{document}
