
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\emergencystretch=3em
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.25}
\newcommand{\checkbox}{$\Box$\hspace{0.4em}}
\newcommand{\ok}{\textbf{Success criteria:} }
\newcommand{\sourcebasis}{\section*{Official Source Basis}\begin{itemize}
\item Authentik Docker Compose installation: \url{https://docs.goauthentik.io/install-config/install/docker-compose/}
\item Authentik reverse proxy guidance: \url{https://docs.goauthentik.io/install-config/reverse-proxy/}
\item Authentik configuration reference: \url{https://docs.goauthentik.io/install-config/configuration/}
\item Authentik first steps: \url{https://docs.goauthentik.io/install-config/first-steps/}
\item Authentik backup and restore: \url{https://docs.goauthentik.io/sys-mgmt/ops/backup-restore/}
\item Authentik upgrade guidance: \url{https://docs.goauthentik.io/install-config/upgrade/}
\item Authentik monitoring: \url{https://docs.goauthentik.io/sys-mgmt/ops/monitoring/}
\item Authentik example flows: \url{https://docs.goauthentik.io/add-secure-apps/flows-stages/flow/examples/flows/}
\item Authentik service accounts: \url{https://docs.goauthentik.io/sys-mgmt/service-accounts/}
\item Authentik certificates: \url{https://docs.goauthentik.io/sys-mgmt/certificates/}
\end{itemize}}

\title{Authentik Implementation Checklist}\author{Prepared from official Authentik documentation}\date{May 6, 2026}\begin{document}\maketitle

\section*{Purpose}
This checklist turns the official Docker Compose installation, configuration, first steps, backup, upgrade, and monitoring guidance into an implementation-ready sequence.

\section*{Phase 1 - Plan the Deployment}
\begin{itemize}
\item \checkbox Confirm the deployment is suitable for Docker Compose: test setup, lab, homelab, or small-scale production setup.
\item \checkbox Allocate a host with at least 2 CPU cores and 2 GB of RAM.
\item \checkbox Install Docker Compose v2 or Podman with Compose support.
\item \checkbox Select a fully qualified domain name for the Authentik service.
\item \checkbox Select the reverse proxy pattern: Nginx, Caddy, Traefik, or another proxy capable of HTTPS and WebSocket forwarding.
\item \checkbox Define backup ownership for PostgreSQL and static directories before installation.
\end{itemize}

\section*{Phase 2 - Install Authentik with Docker Compose}
\begin{itemize}
\item \checkbox Create a dedicated installation directory.
\item \checkbox Download the official \texttt{compose.yml} file from Authentik documentation.
\item \checkbox Generate \texttt{PG\_PASS} and \texttt{AUTHENTIK\_SECRET\_KEY} in the local \texttt{.env} file.
\item \checkbox Add optional error reporting configuration only if the organization accepts the operational behavior.
\item \checkbox Add optional \texttt{COMPOSE\_PORT\_HTTP} and \texttt{COMPOSE\_PORT\_HTTPS} variables when exposing different ports.
\item \checkbox Run \texttt{docker compose pull}.
\item \checkbox Run \texttt{docker compose up -d}.
\item \checkbox Access \texttt{http://<server>:9000/if/flow/initial-setup/} and include the trailing slash.
\item \checkbox Set the default \texttt{akadmin} password during initial setup.
\end{itemize}

\section*{Phase 3 - Configure the Reverse Proxy}
\begin{itemize}
\item \checkbox Enforce HTTPS at the proxy.
\item \checkbox Pass \texttt{X-Forwarded-Proto} upstream.
\item \checkbox Pass \texttt{X-Forwarded-For} upstream.
\item \checkbox Pass \texttt{Host} upstream.
\item \checkbox Support \texttt{Connection: Upgrade} and \texttt{Upgrade: WebSocket} for WebSocket endpoints.
\item \checkbox Configure trusted proxy CIDRs when the proxy is not connecting from a private IP range covered by Authentik defaults.
\item \checkbox Use a certificate matching the Authentik FQDN, preferably issued by a publicly trusted certificate authority.
\end{itemize}

\section*{Phase 4 - Configure Identity Objects}
\begin{itemize}
\item \checkbox Log in as an administrator and open the Admin interface.
\item \checkbox Create the first application and provider pair.
\item \checkbox Select the provider type required by the target application, such as OAuth2/OpenID Connect, SAML, LDAP, or another supported provider.
\item \checkbox Record client IDs, client secrets, redirect URIs, slugs, and signing keys where applicable.
\item \checkbox Create the first non-admin user.
\item \checkbox Use bindings and policies to control application visibility and access.
\end{itemize}

\section*{Phase 5 - Harden, Monitor, and Operate}
\begin{itemize}
\item \checkbox Configure MFA-sensitive authentication flows when the application risk profile requires it.
\item \checkbox Decide whether to retain the Docker socket mount, use a Docker socket proxy, or manually deploy outposts.
\item \checkbox Configure monitoring for server, worker, outpost, and metrics endpoints.
\item \checkbox Configure recurring PostgreSQL backups.
\item \checkbox Back up static directories used by the deployment.
\item \checkbox Before every upgrade, read release notes, back up PostgreSQL, retrieve the current Compose file, run \texttt{docker compose pull}, and run \texttt{docker compose up -d}.
\item \checkbox Verify the running version in \texttt{Dashboards > Overview} after upgrades.
\end{itemize}

\sourcebasis

\end{document}
