
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\emergencystretch=3em
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.25}
\newcommand{\checkbox}{$\Box$\hspace{0.4em}}
\newcommand{\ok}{\textbf{Success criteria:} }
\newcommand{\sourcebasis}{\section*{Official Source Basis}\begin{itemize}
\item Authentik Docker Compose installation: \url{https://docs.goauthentik.io/install-config/install/docker-compose/}
\item Authentik reverse proxy guidance: \url{https://docs.goauthentik.io/install-config/reverse-proxy/}
\item Authentik configuration reference: \url{https://docs.goauthentik.io/install-config/configuration/}
\item Authentik first steps: \url{https://docs.goauthentik.io/install-config/first-steps/}
\item Authentik backup and restore: \url{https://docs.goauthentik.io/sys-mgmt/ops/backup-restore/}
\item Authentik upgrade guidance: \url{https://docs.goauthentik.io/install-config/upgrade/}
\item Authentik monitoring: \url{https://docs.goauthentik.io/sys-mgmt/ops/monitoring/}
\item Authentik example flows: \url{https://docs.goauthentik.io/add-secure-apps/flows-stages/flow/examples/flows/}
\item Authentik service accounts: \url{https://docs.goauthentik.io/sys-mgmt/service-accounts/}
\item Authentik certificates: \url{https://docs.goauthentik.io/sys-mgmt/certificates/}
\end{itemize}}

\title{Authentik Self-Hosting Architecture Guide}\author{Prepared from official Authentik documentation}\date{May 6, 2026}\begin{document}\maketitle

\section*{Purpose}
This guide defines a small-scale self-hosted Authentik architecture using Docker Compose, a reverse proxy, persistent storage, PostgreSQL, and operational guardrails. It is intended for lab, homelab, and small-scale production deployments that follow the official Authentik Docker Compose installation path.

\section*{Recommended Logical Architecture}
\begin{longtable}{>{\raggedright\arraybackslash}p{0.23\linewidth}>{\raggedright\arraybackslash}p{0.36\linewidth}>{\raggedright\arraybackslash}p{0.32\linewidth}}
\textbf{Layer} & \textbf{Primary responsibility} & \textbf{Design guidance} \\
\hline
Client access & Browser, users, and administrators access Authentik through a fully qualified domain name. & Put Authentik behind a reverse proxy that terminates HTTPS and passes the required forwarding and WebSocket headers. \\
Reverse proxy & TLS termination, HTTPS redirect, upstream routing, and WebSocket support. & Pass \texttt{X-Forwarded-Proto}, \texttt{X-Forwarded-For}, \texttt{Host}, and WebSocket upgrade headers. Use modern TLS and disable protocols older than TLS 1.3 where feasible. \\
Authentik server & Serves the web interface, application endpoints, API traffic, and identity-provider flows. & The default Docker Compose deployment exposes HTTP on 9000 and HTTPS on 9443 internally. Exposed ports can be controlled with \texttt{COMPOSE\_PORT\_HTTP} and \texttt{COMPOSE\_PORT\_HTTPS}. \\
Authentik worker & Runs background tasks and watches filesystem-based blueprints and certificates. & Keep the worker healthy and monitored. Treat Docker socket access as privileged; use a socket proxy or manually deploy outposts when reducing Docker socket exposure. \\
PostgreSQL & Persistent store for users, policies, configuration, providers, flows, and other Authentik data. & Back up the PostgreSQL database using native PostgreSQL backup tools before upgrades and on a regular operational schedule. \\
Static storage & Stores local files such as application icons, flow backgrounds, uploaded files, CSV reports, filesystem certificates, custom templates, and blueprints. & Back up \texttt{/data}, \texttt{/certs}, \texttt{/custom-templates}, and \texttt{/blueprints} when those paths are used. Use S3 storage only when the deployment intentionally moves file storage out of the container filesystem. \\
Applications and providers & Define how external applications use Authentik through OIDC/OAuth2, SAML, LDAP, proxy providers, or other supported integration patterns. & Every application added to Authentik requires a provider that defines the protocol and protocol-specific settings. \\
Users, groups, and policies & Control who can access which applications and under what conditions. & Use bindings and policies intentionally. Apply MFA-sensitive flows where the risk profile requires additional verification. \\
Monitoring & Health checks, readiness checks, worker checks, outpost checks, and Prometheus metrics. & Monitor liveness, readiness, worker health, outpost ping, and metrics on port 9300 when appropriate. \\
\end{longtable}

\section*{Deployment Boundary}
The Docker Compose installation method is appropriate for test setups and small-scale production setups. For high-availability requirements, use an architecture that separates shared database responsibilities and supports active-passive failover or equivalent operational controls.

\section*{Security Boundary}
The reverse proxy is part of the trust boundary. Only trusted proxy CIDRs should be permitted to set proxy headers such as \texttt{X-Forwarded-For}. Authentik defaults include private and loopback network ranges; adjust \texttt{AUTHENTIK\_LISTEN\_\_TRUSTED\_PROXY\_CIDRS} when the proxy path requires it.

\section*{Operational Boundary}
Before upgrading Authentik, read the release notes, back up PostgreSQL, and upgrade in the required major/minor sequence. Authentik does not support downgrading. The Authentik server and all Authentik outposts must run matching versions after an upgrade.

\sourcebasis

\end{document}
