
\documentclass[11pt]{article}
\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\emergencystretch=3em
\setlist[itemize]{leftmargin=1.2em,itemsep=2pt,topsep=2pt}
\setlist[enumerate]{leftmargin=1.4em,itemsep=2pt,topsep=2pt}
\renewcommand{\arraystretch}{1.25}
\newcommand{\checkbox}{$\Box$\hspace{0.4em}}
\newcommand{\ok}{\textbf{Success criteria:} }
\newcommand{\sourcebasis}{\section*{Official Source Basis}\begin{itemize}
\item Authentik Docker Compose installation: \url{https://docs.goauthentik.io/install-config/install/docker-compose/}
\item Authentik reverse proxy guidance: \url{https://docs.goauthentik.io/install-config/reverse-proxy/}
\item Authentik configuration reference: \url{https://docs.goauthentik.io/install-config/configuration/}
\item Authentik first steps: \url{https://docs.goauthentik.io/install-config/first-steps/}
\item Authentik backup and restore: \url{https://docs.goauthentik.io/sys-mgmt/ops/backup-restore/}
\item Authentik upgrade guidance: \url{https://docs.goauthentik.io/install-config/upgrade/}
\item Authentik monitoring: \url{https://docs.goauthentik.io/sys-mgmt/ops/monitoring/}
\item Authentik example flows: \url{https://docs.goauthentik.io/add-secure-apps/flows-stages/flow/examples/flows/}
\item Authentik service accounts: \url{https://docs.goauthentik.io/sys-mgmt/service-accounts/}
\item Authentik certificates: \url{https://docs.goauthentik.io/sys-mgmt/certificates/}
\end{itemize}}

\title{Authentik Self-Hosting Full LaTeX Handout}\author{Prepared from official Authentik documentation}\date{May 6, 2026}\begin{document}\maketitle

\section*{Executive Summary}
Authentik can be self-hosted as a centralized identity provider for SSO and MFA-sensitive access patterns. For small-scale deployments, the official Docker Compose path provides the core stack, while the reverse proxy, backup, upgrade, monitoring, and identity-policy configuration determine whether the deployment is operationally safe.

\section*{Reference Architecture}
\begin{itemize}
\item \textbf{Users and administrators} access Authentik through a domain name and HTTPS reverse proxy.
\item \textbf{Reverse proxy} terminates TLS, redirects HTTP to HTTPS, forwards required proxy headers, and supports WebSocket upgrades.
\item \textbf{Authentik server} handles the web UI, API, and authentication flows.
\item \textbf{Authentik worker} runs background tasks and watches blueprints and certificates.
\item \textbf{PostgreSQL} stores persistent Authentik data, including users, policies, and configuration.
\item \textbf{Static directories} preserve local data such as uploaded files, icons, certificates, custom templates, and blueprints when those features are used.
\item \textbf{Applications and providers} connect external applications to Authentik through the selected protocol.
\item \textbf{Outposts} extend Authentik into proxy or LDAP patterns and must stay version-aligned with the core instance.
\end{itemize}

\section*{Minimum Deployment Requirements}
\begin{longtable}{>{\raggedright\arraybackslash}p{0.28\linewidth}>{\raggedright\arraybackslash}p{0.62\linewidth}}
\textbf{Requirement} & \textbf{Minimum or required action} \\
\hline
Compute & At least 2 CPU cores and 2 GB of RAM. \\
Runtime & Docker Compose v2 or Podman with Compose support. \\
Installation scope & Docker Compose is suitable for test setups and small-scale production setups. \\
Secrets & Generate \texttt{PG\_PASS} and \texttt{AUTHENTIK\_SECRET\_KEY} for a fresh install. \\
Network & Use ports 9000 and 9443 internally by default, or set \texttt{COMPOSE\_PORT\_HTTP} and \texttt{COMPOSE\_PORT\_HTTPS}. \\
TLS & Use a reverse proxy with modern TLS and a certificate matching the Authentik FQDN. \\
Proxy headers & Forward protocol, client IP, host, and WebSocket upgrade headers. \\
Backup & Back up PostgreSQL and local static directories as applicable. \\
Monitoring & Monitor server liveness, readiness, worker health, outpost health, and metrics. \\
Upgrade & Back up first, read release notes, follow upgrade sequencing, and match outpost versions. \\
\end{longtable}

\section*{Docker Compose Installation Procedure}
\subsection*{1. Prepare the directory}
\begin{verbatim}
sudo mkdir -p /opt/authentik
sudo chown "$USER":"$USER" /opt/authentik
cd /opt/authentik
\end{verbatim}

\subsection*{2. Download the Compose file}
\begin{verbatim}
wget https://docs.goauthentik.io/compose.yml
\end{verbatim}
Or:
\begin{verbatim}
curl -O https://docs.goauthentik.io/compose.yml
\end{verbatim}

\subsection*{3. Generate required secrets}
\begin{verbatim}
echo "PG_PASS=$(openssl rand -base64 36 | tr -d '\n')" >> .env
echo "AUTHENTIK_SECRET_KEY=$(openssl rand -base64 60 | tr -d '\n')" >> .env
\end{verbatim}

\subsection*{4. Optionally set exposed ports}
\begin{verbatim}
cat >> .env <<'EOF'
COMPOSE_PORT_HTTP=80
COMPOSE_PORT_HTTPS=443
EOF
\end{verbatim}

\subsection*{5. Pull and start}
\begin{verbatim}
docker compose pull
docker compose up -d
\end{verbatim}

\subsection*{6. Complete first setup}
Open:
\begin{verbatim}
http://<your server's IP or hostname>:9000/if/flow/initial-setup/
\end{verbatim}
Set the default \texttt{akadmin} password. The trailing slash is required.

\section*{Reverse Proxy Requirements}
A reverse proxy must support WebSockets and forward the required upstream headers. HTTP/1.0 reverse proxies are not supported for Authentik because WebSockets and protocol upgrades are required for outpost communication.

\begin{longtable}{>{\raggedright\arraybackslash}p{0.33\linewidth}>{\raggedright\arraybackslash}p{0.57\linewidth}}
\textbf{Header or setting} & \textbf{Reason} \\
\hline
\texttt{X-Forwarded-Proto} & Communicates whether Authentik and proxy providers are being served over HTTPS. \\
\texttt{X-Forwarded-For} & Preserves client IP information for logging and IP-sensitive policy behavior. \\
\texttt{Host} & Required for security checks, WebSocket handshake, and outpost/proxy-provider communication. \\
\texttt{Connection: Upgrade} and \texttt{Upgrade: WebSocket} & Required for WebSocket endpoints under HTTP/1.1. \\
Trusted proxy CIDRs & Required when Authentik should accept proxy headers from a non-default proxy network. \\
\end{longtable}

\section*{First Application and User}
After installation, create an application and provider pair from the Admin interface. Every application added to Authentik requires a provider that defines the protocol between the application and Authentik. For OAuth2/OpenID Connect integrations, preserve the client ID, client secret, slug, redirect URI, and related protocol settings required by the service provider. Then create a non-admin user under \texttt{Directory > Users} and apply bindings or policies where access should be restricted.

\section*{Security Hardening Checklist}
\begin{itemize}
\item \checkbox Enforce HTTPS at the reverse proxy.
\item \checkbox Pass all required proxy and WebSocket headers.
\item \checkbox Configure trusted proxy CIDRs when required by the proxy network design.
\item \checkbox Use an external certificate matching the FQDN, preferably from a publicly trusted CA.
\item \checkbox Protect \texttt{.env} and any secrets used for PostgreSQL and Authentik.
\item \checkbox Evaluate whether the Docker socket mount should remain, be replaced with a Docker socket proxy, or be removed in favor of manual outpost deployment.
\item \checkbox Use MFA-capable flows for high-risk applications and administrative access patterns.
\item \checkbox Use bindings and policies to restrict application access.
\item \checkbox Apply least privilege to service accounts and prefer expiring tokens where feasible.
\item \checkbox Avoid trace logging except during tightly controlled troubleshooting because it can expose sensitive information.
\end{itemize}

\section*{Backup and Restore Scope}
\begin{longtable}{>{\raggedright\arraybackslash}p{0.25\linewidth}>{\raggedright\arraybackslash}p{0.35\linewidth}>{\raggedright\arraybackslash}p{0.30\linewidth}}
\textbf{Component} & \textbf{Purpose} & \textbf{Backup note} \\
\hline
PostgreSQL & Persistent Authentik data, users, policies, and configuration. & Back up with PostgreSQL-native tools. Exclude \texttt{template0} and \texttt{template1}. \\
\texttt{/data} & Icons, flow backgrounds, uploaded files, and CSV reports. & Required if not using S3 external storage. \\
\texttt{/certs} & TLS certificates in filesystem storage. & Back up if the filesystem copies are relied upon. \\
\texttt{/custom-templates} & Custom UI changes. & Required when default appearance has been modified. \\
\texttt{/blueprints} & Custom blueprints. & Optional but recommended when custom blueprints are used. \\
\end{longtable}

\section*{Monitoring Checklist}
\begin{itemize}
\item \checkbox Server liveness: \texttt{/-/health/live/} returns HTTP 200 while Authentik is running.
\item \checkbox Server readiness: \texttt{/-/health/ready/} returns HTTP 200 when PostgreSQL connectivity is available.
\item \checkbox Worker health: run \texttt{ak healthcheck} inside the worker container.
\item \checkbox Outpost health: send HTTP requests to \texttt{/outpost.goauthentik.io/ping} on the outpost.
\item \checkbox Metrics: scrape Prometheus metrics on port 9300 from intended monitoring networks only.
\end{itemize}

\section*{Upgrade Procedure}
\begin{enumerate}
\item Read release notes for the target version.
\item Back up PostgreSQL.
\item Follow the required major/minor upgrade sequence; do not skip directly from an older major version to the newest version.
\item Download the current Compose file.
\item Run \texttt{docker compose pull} and \texttt{docker compose up -d}.
\item Upgrade any outposts with the core Authentik instance.
\item Verify the running version in \texttt{Dashboards > Overview}.
\end{enumerate}

\section*{Operational Warnings}
\begin{itemize}
\item Authentik does not support downgrading; keep a database backup before upgrades.
\item All internal operations use UTC. Do not mount \texttt{/etc/timezone} or \texttt{/etc/localtime} into Authentik containers.
\item The default Compose file mounts the Docker socket into the worker container for automatic outpost management; treat that as privileged access.
\item Trace logging can expose sensitive information, including session cookies.
\end{itemize}

\sourcebasis

\end{document}
