\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[dvipsnames]{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{array}

\hypersetup{
    colorlinks=true,
    linkcolor=MidnightBlue,
    urlcolor=MidnightBlue,
    pdftitle={Calibre Workflow for Converting DRM-Free EPUB, AZW3, and MOBI Files to PDF},
    pdfauthor={OpenAI},
    pdfsubject={LaTeX handout},
    pdfcreator={ChatGPT}
}

\pagestyle{fancy}
\fancyhf{}
\lhead{\textit{DRM-Free Ebook Conversion Workflow}}
\rhead{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\section}{\large\bfseries\color{MidnightBlue}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\normalsize\bfseries\color{MidnightBlue}}{\thesubsection}{0.6em}{}

\setlist[itemize]{leftmargin=1.4em}
\setlist[enumerate]{leftmargin=1.6em}

\begin{document}

\begin{center}
    {\LARGE \textbf{Calibre Workflow for Converting DRM-Free EPUB, AZW3, and MOBI Files to PDF}}\\[0.75em]
    {\large A concise, practical handout}\\[1em]
\end{center}

\noindent\textbf{Purpose.} This handout provides a clean, lawful workflow for using \textbf{Calibre} to convert \textbf{DRM-free} ebook files into PDF format.

\medskip

\noindent\textbf{Scope.} This workflow applies to:
\begin{itemize}
    \item \textbf{EPUB}
    \item \textbf{AZW3}
    \item \textbf{MOBI}
\end{itemize}

\noindent Use this process only for files you are authorized to convert.

\section{Recommended Conversion Order}

For the best output quality, use this priority order when multiple source formats are available:

\begin{enumerate}
    \item \textbf{EPUB} $\rightarrow$ PDF
    \item \textbf{AZW3} $\rightarrow$ PDF
    \item \textbf{MOBI} $\rightarrow$ PDF
\end{enumerate}

\noindent EPUB often produces the cleanest PDF because it typically preserves structure better for conversion.

\section{Workflow}

\subsection{1. Install Calibre}

Download and install \textbf{Calibre} on your computer. Once installed, launch the application and create or open your Calibre library.

\subsection{2. Add DRM-Free Ebook Files}

Import your ebook files into Calibre using one of the following methods:

\begin{itemize}
    \item Drag and drop the file into the Calibre library window
    \item Click \textbf{Add books} and select the file manually
\end{itemize}

\noindent Supported inputs for this workflow are DRM-free \texttt{.epub}, \texttt{.azw3}, and \texttt{.mobi} files.

\subsection{3. Review and Clean Metadata}

Before conversion, select the book and review its metadata. At minimum, confirm:

\begin{itemize}
    \item Title
    \item Author
    \item Cover image
    \item Series name and number, if applicable
\end{itemize}

\noindent Clean metadata improves file organization and makes the resulting PDF easier to identify and archive.

\subsection{4. Start the Conversion}

Select the book in Calibre, then choose:

\begin{center}
    \fbox{\textbf{Convert books} $\rightarrow$ \textbf{Convert individually}}
\end{center}

\noindent In the conversion dialog, set the \textbf{Output format} to \textbf{PDF}.

\subsection{5. Review Conversion Settings}

Before running the conversion, check the most important settings.

\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.24\textwidth} X}
\toprule
\textbf{Setting Area} & \textbf{What to Review} \\
\midrule
Page Setup & Choose page size and profile appropriate for how the PDF will be read or printed. \\
Look \& Feel & Review font sizing, spacing, and general presentation. \\
Structure Detection & Improve heading recognition and chapter breaks where possible. \\
Heuristic Processing & Use only when the source formatting is poor or inconsistent. \\
PDF Output & Review layout, scaling, and output-specific formatting controls. \\
\bottomrule
\end{tabularx}

\medskip

\noindent Default settings are often acceptable for a first pass. If the source ebook has unusual formatting, expect to make adjustments and reconvert.

\subsection{6. Run the Conversion}

Click \textbf{OK} to begin conversion. Calibre will generate a PDF version of the selected ebook and add it to the book's available formats.

\subsection{7. Inspect the Result}

After conversion, open the PDF and inspect the following:

\begin{itemize}
    \item Chapter headings
    \item Page breaks
    \item Margins
    \item Font size and readability
    \item Images, tables, or diagrams
\end{itemize}

\noindent If the output does not look right, return to the conversion settings and adjust page size, margins, or formatting options, then reconvert.

\subsection{8. Save or Export the PDF}

When the output is satisfactory, use \textbf{Save to disk} to export the PDF to a location outside the Calibre library if needed.

\section{Practical Notes}

\begin{itemize}
    \item PDF is a \textbf{fixed-layout} format, while EPUB, AZW3, and MOBI are generally \textbf{reflowable} formats.
    \item Because of that difference, a converted PDF may not perfectly match the reading experience of the original ebook.
    \item Image-heavy, heavily styled, or unusually structured ebooks may require more tuning.
    \item For simple text-centric books, Calibre's default settings often work well.
\end{itemize}

\section{Quick Reference}

\begin{center}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.28\textwidth} X}
\toprule
\textbf{Task} & \textbf{Action} \\
\midrule
Import ebook & \textbf{Add books} or drag and drop the file into Calibre \\
Edit metadata & Confirm title, author, cover, and series information \\
Convert to PDF & \textbf{Convert books} $\rightarrow$ \textbf{Convert individually} \\
Choose output format & Select \textbf{PDF} \\
Check output & Review headings, margins, page breaks, and readability \\
Export result & Use \textbf{Save to disk} if needed \\
\bottomrule
\end{tabularx}
\end{center}

\section{Summary}

A clean Calibre workflow for DRM-free ebook conversion is straightforward: import the ebook, clean the metadata, convert to PDF, review the result, and adjust settings if necessary. For the best results, prefer EPUB when available, then AZW3, and use MOBI as a fallback.

\end{document}
