\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{xurl}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{ragged2e}
\usepackage{colortbl}
\usepackage{etoolbox}

\setlength{\parindent}{0pt}
\setlength{\headheight}{14pt}
\emergencystretch=2em
\Urlmuskip=0mu plus 2mu
\renewcommand{\arraystretch}{1.25}
\pagestyle{fancy}
\fancyhf{}
\lhead{SEI Views and Beyond and ISO/IEC/IEEE 42010:2022}
\rhead{Handout}
\cfoot{\thepage}

\definecolor{titleblue}{HTML}{17365D}
\definecolor{rulegray}{HTML}{D9E2F3}
\definecolor{tablehead}{HTML}{EAF0FB}
\definecolor{accent}{HTML}{2F5597}

\titleformat{\section}{\large\bfseries\color{titleblue}}{}{0pt}{}
\titleformat{\subsection}{\normalsize\bfseries\color{titleblue}}{}{0pt}{}

\newcolumntype{Y}{>{\RaggedRight\arraybackslash}X}
\newcolumntype{P}[1]{>{\RaggedRight\arraybackslash}p{#1}}

\begin{document}

{\LARGE\bfseries\color{titleblue}SEI Views and Beyond Templates\par}
{\LARGE\bfseries\color{titleblue}Aligned to ISO/IEC/IEEE 42010:2022\par}
\vspace{0.5em}
{\large Clean reference handout for architecture documentation\par}
\vspace{0.75em}
\noindent{\color{rulegray}\rule{\textwidth}{1pt}}
\vspace{0.5em}

\textbf{Purpose.} This handout converts the source markdown into a tighter, project-usable summary of how to document software architecture with the SEI \emph{Views and Beyond} (V\&B) approach while staying aligned to ISO/IEC/IEEE 42010:2022.

\textbf{Core idea.} ISO/IEC/IEEE 42010:2022 defines what an architecture description must contain, including requirements around architecture descriptions, viewpoints, and model kinds. It does not prescribe a single method, notation, or document format. SEI V\&B supplies a practical multi-view documentation method with reusable templates and view-oriented organization. Used together, they give you both a compliance-oriented structure and a concrete documentation approach.

\section{How the two approaches fit together}

ISO/IEC/IEEE 42010:2022 specifies requirements for the structure and expression of an architecture description and distinguishes the architecture of an entity from the architecture description that expresses it. The standard also defines requirements for architecture viewpoints and model kinds, while explicitly not mandating a particular process, notation, technique, tool, or recording medium.

SEI V\&B treats architecture documentation as the selection of relevant views, the documentation of each selected view, and the addition of information that applies across views. This makes V\&B a strong implementation approach for producing architecture descriptions that satisfy the standard's conceptual expectations.

\section{Working rule for an ISO 42010-aligned V\&B document}

For each architecture view, document two things together:
\begin{enumerate}[leftmargin=1.5em]
    \item the \textbf{viewpoint framing}, including stakeholders, concerns, and model kinds; and
    \item the \textbf{view content}, including the representation, element catalog, relationships, and rationale relevant to that view.
\end{enumerate}

This pairing is what closes the gap between a practical V\&B template and an ISO 42010-aligned architecture description.

\section{Primary SEI view type templates}

\subsection{1. Module view type}

\textbf{Use when the primary concern is} modifiability, reuse, responsibility assignment, implementation structure, or work breakdown.

\begin{longtable}{@{}P{0.26\textwidth}P{0.69\textwidth}@{}}
\rowcolor{tablehead}
\textbf{Template element} & \textbf{What to capture} \\
\toprule
\endfirsthead
\rowcolor{tablehead}
\textbf{Template element} & \textbf{What to capture} \\
\toprule
\endhead
Viewpoint metadata & Stakeholders such as developers, maintainers, and technical leads; concerns such as impact of change, reuse, decomposition, layering, and responsibility allocation; model kinds used to express the module structure. \\
Primary presentation & A decomposition, layered, uses, or generalization view rendered in an appropriate notation, such as a package diagram, class diagram, or architecture-specific schematic. \\
Element catalog & Definitions of the modules, layers, packages, subsystems, or other structural units; responsibilities; externally visible interfaces; and constraints that govern use. \\
Context information & Boundaries and dependencies involving external libraries, frameworks, legacy systems, or platform services that affect the module structure. \\
Mappings & Cross-references from module elements to related runtime elements, deployment elements, code repositories, ownership assignments, or requirements where useful. \\
\bottomrule
\end{longtable}

\subsection{2. Component-and-connector (C\&C) view type}

\textbf{Use when the primary concern is} runtime behavior, communication, availability, performance, security, fault handling, or data movement.

\begin{longtable}{@{}P{0.26\textwidth}P{0.69\textwidth}@{}}
\rowcolor{tablehead}
\textbf{Template element} & \textbf{What to capture} \\
\toprule
\endfirsthead
\rowcolor{tablehead}
\textbf{Template element} & \textbf{What to capture} \\
\toprule
\endhead
Viewpoint metadata & Stakeholders such as operations teams, SREs, performance engineers, and security reviewers; concerns such as latency, throughput, resilience, trust boundaries, and data protection; model kinds used for runtime structure and interaction. \\
Primary presentation & A runtime diagram showing active elements and their interactions, for example client-server, shared data, event-driven, message bus, or publish-subscribe styles. \\
Element catalog & Descriptions of runtime components and connectors, including protocols, message types, interaction semantics, important properties, and operational assumptions. \\
Variability or scaling notes & How runtime elements replicate, fail over, scale out, or change shape across environments, load profiles, or operational modes. \\
Design rationale & Why specific communication patterns, synchronization strategies, or trust-boundary decisions were chosen, especially where quality attributes are affected. \\
\bottomrule
\end{longtable}

\subsection{3. Allocation view type}

\textbf{Use when the primary concern is} deployment, infrastructure, environment mapping, cost, geography, compliance, operational support, or team ownership.

\begin{longtable}{@{}P{0.26\textwidth}P{0.69\textwidth}@{}}
\rowcolor{tablehead}
\textbf{Template element} & \textbf{What to capture} \\
\toprule
\endfirsthead
\rowcolor{tablehead}
\textbf{Template element} & \textbf{What to capture} \\
\toprule
\endhead
Viewpoint metadata & Stakeholders such as platform engineers, infrastructure owners, compliance leads, and service owners; concerns such as placement, residency, installation, operational responsibility, and environmental constraints. \\
Primary presentation & A mapping of software elements to non-software structures such as deployment nodes, cloud services, environments, organizations, teams, sites, or hardware platforms. \\
Element catalog & Execution environments, platforms, clusters, accounts, regions, data stores, networks, or teams, plus the software elements allocated to each. \\
Correspondence rules & Explicit mapping rules that show how module or runtime elements correspond to deployment or organizational elements. \\
Constraints & Regulatory, legal, physical, budgetary, operational, or residency constraints that limit where or how software may be allocated. \\
\bottomrule
\end{longtable}

\section{What must exist beyond individual views}

A strong V\&B document does not stop at isolated diagrams. It also includes information that applies across views so the full architecture description reads as one coherent package rather than a loose collection of snapshots.

\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Documentation roadmap:} how to read the document, who each section is for, and how the views relate.
    \item \textbf{Architecture background:} scope, problem statement, context, assumptions, driving quality attributes, and major constraints.
    \item \textbf{Cross-view rationale:} architecture decisions, tradeoffs, and reasoning that affect more than one view.
    \item \textbf{Consistency rules:} naming, correspondence rules, and traceability expectations that keep views aligned.
\end{itemize}

\section{Minimal compliance checklist for project teams}

Use this checklist when adapting a V\&B template for a live project:
\begin{enumerate}[leftmargin=1.5em]
    \item Identify the entity of interest and the scope of the architecture description.
    \item Name the stakeholders and their concerns.
    \item Define the viewpoint for each planned view, including the model kinds used.
    \item Select only the views that materially address those concerns.
    \item Populate each view with a primary representation, an element catalog, and supporting rationale.
    \item Add the cross-view material: roadmap, background, rationale, and correspondence information.
    \item Check that the document is internally consistent across module, runtime, and allocation structures.
\end{enumerate}

\section{Practical usage note}

In practice, a project does not become ISO 42010-aligned merely by drawing more diagrams. It becomes aligned when each view is explicitly connected to stakeholders, concerns, and defined viewpoint guidance, and when the document includes the cross-view material needed to explain why the architecture looks the way it does.

\section{References}

\begin{enumerate}[leftmargin=1.5em]
    \item ISO, \emph{ISO/IEC/IEEE 42010:2022 Software, systems and enterprise --- Architecture description}, standard overview page. \url{https://www.iso.org/standard/74393.html}
    \item ISO Architecture Working Group, \emph{ISO/IEC/IEEE 42010 Conceptual Model}. \url{https://www.iso-architecture.org/42010/cm/}
    \item Paul Clements, \emph{Comparing the SEI's Views and Beyond Approach for Documenting Software Architectures with ANSI-IEEE 1471-2000}, CMU/SEI-2005-TN-017. \url{https://www.sei.cmu.edu/documents/2072/2005_004_001_14498.pdf}
    \item Carnegie Mellon Software Engineering Institute, \emph{Documenting Software Architectures: Views and Beyond, 2nd Edition}. \url{https://www.sei.cmu.edu/library/documenting-software-architectures-views-and-beyond-second-edition/}
    \item ISO Architecture Working Group, \emph{Architecture Description Template}. \url{https://www.iso-architecture.org/42010/templates/42010-ad-template.pdf}
\end{enumerate}

\end{document}



