\documentclass[11pt]{article}

% ---------- Safe, common packages ----------
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{setspace}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% ---------- Simple helpers ----------
\setlist[itemize]{itemsep=2pt, topsep=4pt}
\setlist[enumerate]{itemsep=4pt, topsep=6pt}
\newcommand{\Section}[1]{\vspace{0.6em}\noindent\textbf{\Large #1}\par\vspace{0.25em}}
\newcommand{\Subsection}[1]{\vspace{0.4em}\noindent\textbf{\large #1}\par\vspace{0.2em}}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\begin{document}

\begin{center}
  {\LARGE \textbf{Roasted Celery and Potato Soup}}\\[4pt]
  \small by Kylie Perrotti \quad\textbullet\quad Posted on: August 19, 2022 \quad\textbullet\quad Vegan \\
  \small Rating: 4.40/5 (86 votes)
\end{center}

\Section{Overview}
This budget-friendly vegan soup roasts celery for deep flavor, then blends it with tender potatoes and broth for creaminess without dairy. A handful of elbows adds body and comfort. Finish with herbs and a drizzle of olive oil.

\Section{At a Glance}
\begin{tabularx}{\textwidth}{@{} l L @{}}
\toprule
\textbf{Yield} & Serves 4 \\
\textbf{Prep Time} & 10 minutes \\
\textbf{Cook Time} & 45 minutes \\
\textbf{Total Time} & 55 minutes \\
\textbf{Diet} & Vegan \\
\bottomrule
\end{tabularx}

\Section{Equipment}
\begin{itemize}
  \item Large pot
  \item Baking sheet
  \item Food processor or blender
\end{itemize}

\Section{Ingredients}
\begin{itemize}
  \item 2 tablespoons olive oil, divided
  \item 1 stalk celery (about 10 ribs), root end trimmed; cut into thirds
  \item 1 pound Yukon Gold potatoes, peeled and diced
  \item 5 ounces cremini mushrooms, sliced
  \item 2 teaspoons dry thyme
  \item 7 cups water or vegetable stock
  \item 1 cup elbow pasta
  \item Salt and black pepper, to taste
\end{itemize}

\Subsection{For serving (optional)}
Paprika or red pepper flakes; fresh dill or parsley; extra-virgin olive oil.

\Section{Instructions}
\begin{enumerate}
  \item \textbf{Preheat the oven.} Heat oven to 400$^\circ$F (200$^\circ$C).
  \item \textbf{Roast the celery.} Arrange celery on a baking sheet. Drizzle with 1 tablespoon oil; season with salt and pepper. Roast 30 minutes.
  \item \textbf{Cook mushrooms and potatoes.} In a large pot, heat remaining 1 tablespoon oil over medium heat. Add mushrooms; cook 8--10 minutes until browned and their liquid evaporates. Season with salt and pepper. Add diced potatoes, more salt and pepper, and dry thyme; cook 1 minute until fragrant.
  \item \textbf{Simmer the soup.} Pour in water or vegetable stock. Bring to a boil, then reduce heat and simmer 20 minutes, or until potatoes are fork-tender.
  \item \textbf{Blend celery and potatoes.} Using a slotted spoon, transfer about half of the cooked potatoes to a food processor. Add the roasted celery and a few ladles of hot broth. Blend until smooth.
  \item \textbf{Cook the pasta and finish.} Return the pot to a boil. Stir in the elbow pasta and the celery-potato purée. Cook, stirring occasionally, until pasta is al dente. Taste and adjust seasoning with salt and pepper.
  \item \textbf{Serve.} Ladle into bowls. Garnish with dill or parsley, a pinch of paprika or red pepper flakes, and a drizzle of extra-virgin olive oil, if desired.
\end{enumerate}

\Section{Notes \& Variations}
\begin{itemize}
  \item Add aromatics like a caramelized onion or minced garlic with the mushrooms.
  \item Stir in diced carrots or sweet potatoes with the potatoes.
  \item For protein, add a drained can of beans (cannellini or chickpeas) with the pasta.
  \item Not vegan? Finish with a splash of cream at the end of cooking.
  \item Texture: Blend more for silkier soup or less to keep it rustic and chunky.
\end{itemize}

\Section{Nutrition (Approximate per serving; 4 servings)}
\begin{tabular}{@{}ll@{}}
\toprule
Calories & 287 kcal \\
Carbohydrates & 48 g \\
Protein & 8 g \\
Fat & 8 g \\
Sodium & 12 mg \\
Fiber & 4 g \\
Sugar & 2 g \\
Vitamin C & 22 mg \\
\bottomrule
\end{tabular}

\vfill
\begin{center}
  \footnotesize
  \textit{Tip: For even deeper flavor, deglaze the mushroom pan with a splash of dry white wine before adding the potatoes.}
\end{center}

\end{document}

