\documentclass[10pt]{article}
\usepackage[letterpaper,landscape,margin=0.45in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{ragged2e}
\usepackage{titlesec}
\usepackage{enumitem}
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\titleformat{\section}{\large\bfseries}{}{0pt}{}
\titlespacing*{\section}{0pt}{6pt}{3pt}
\renewcommand{\arraystretch}{1.18}
\definecolor{SBMBlue}{HTML}{1F4E79}
\definecolor{SBMLight}{HTML}{EAF2F8}
\definecolor{SBMRow}{HTML}{F7F9FB}
\definecolor{SBMGray}{HTML}{5B6570}
\newcolumntype{L}[1]{>{\RaggedRight\arraybackslash}p{#1}}
\newcolumntype{Y}{>{\RaggedRight\arraybackslash}X}
\newcommand{\statusdone}{\fcolorbox{black!20}{green!10}{\strut Done}}
\newcommand{\statusprogress}{\fcolorbox{black!20}{yellow!15}{\strut In Progress}}
\newcommand{\statusnotstarted}{\fcolorbox{black!20}{red!8}{\strut Not Started}}
\begin{document}
\footnotesize

{\Huge\bfseries Smart Building Manager (SBM) Startup Roadmap}\\[-1pt]
{\large SBA-aligned founder handout using the current SBA 10-step startup sequence, tailored to SBM's present maturity}

\vspace{4pt}
{\color{SBMBlue}\rule{\textwidth}{1.2pt}}

\textbf{Executive Summary.} Smart Building Manager (SBM) is still in validation and MVP definition: the concept, target customer, and first priorities are visible, but most company-formation and launch mechanics remain open. This handout mirrors the live SBA 10-step sequence and translates each step into concrete founder work for SBM.

\vspace{2pt}
\rowcolors{2}{SBMRow}{white}
\setlength{\tabcolsep}{3pt}
\scriptsize
\begin{tabularx}{\textwidth}{L{0.055\textwidth}L{0.16\textwidth}YL{0.11\textwidth}Y}
\toprule
\rowcolor{SBMBlue!12}
\textbf{SBA Step \#} & \textbf{SBA Step Name} & \textbf{SBM Interpretation / What This Means for SBM} & \textbf{Current Status} & \textbf{Near-Term Action} \\
\midrule
1 & Conduct market research & Validate which operational pain should anchor the first commercial wedge: asset visibility, energy monitoring, workflow automation, governance, or a tighter sub-problem inside smart-building operations. Confirm who buys, who uses, and what measurable outcome matters most. & \statusprogress & Interview at least 10 target operators or facilities leaders, compare current tools and spreadsheets, and convert findings into one ranked problem statement with a quantified success metric for the MVP. \\
2 & Write your business plan & Convert the working idea into an operator-grade business plan that states the wedge, customer profile, problem, offering, deployment model, pricing hypothesis, implementation model, and first-year go-to-market assumptions. & \statusprogress & Expand the existing startup working file into a concise business plan with sections for market, offering, revenue model, launch assumptions, delivery model, risks, and a 12-month milestone view. \\
3 & Fund your business & Determine how SBM will finance the greenfield build, early cloud costs, prototyping, integrations, pilot delivery, and founder operating runway before recurring revenue exists. & \statusnotstarted & Build a bottom-up startup cost model for the first 12 months, decide founder-funded versus outside capital versus services-led bootstrapping, and set a minimum runway target tied to pilot milestones. \\
4 & Pick your business location & Decide SBM's legal and operating home base, initial selling geography, and whether the business will run as a remote-first software and services company with any on-site implementation footprint for buildings. & \statusnotstarted & Select the initial state of formation and operating address, define the first target region for pilots, and document how location affects taxes, contracts, insurance, and any field-service obligations. \\
5 & Choose a business structure & Finalize how SBM should exist as a legal entity and how that structure relates to the broader AdjecTex/parent-company context already discussed in prior planning. & \statusprogress & Review the intended ownership model, choose the target entity structure for launch, and document founder control, liability boundaries, IP ownership, and how future SaaS revenue and services revenue will be booked. \\
6 & Choose your business name & The business already has a working operating name --- Smart Building Manager --- and the name is consistently used in planning, architecture, and business discussions. The remaining work is validation and brand hygiene, not ideation. & \statusdone & Run a formal name-screening pass: domain availability, state availability, trademark conflict review, and a short brand standard covering how SBM appears in customer-facing materials. \\
7 & Register your business & Make the chosen structure real by filing the entity, locking the official business name, and preparing the documents needed for contracts, banking, and tax setup. & \statusnotstarted & Once structure and location are chosen, file formation documents, reserve or register the name as needed, and create a formation checklist with owners, filing dates, and required records. \\
8 & Get federal and state tax IDs & Obtain the identifiers needed to open accounts, invoice customers, run payroll if needed later, and keep finance and compliance operations clean from the start. & \statusnotstarted & Apply for the EIN after entity registration, verify whether a state tax ID is required in the formation state, and store all tax setup artifacts in a founder operations folder. \\
9 & Apply for licenses and permits & Determine whether SBM needs any business, reseller, low-voltage, installation, privacy, or jurisdiction-specific permits for software, services, hardware resale, or on-site smart-building work. & \statusnotstarted & Create a compliance matrix by state and service type that distinguishes pure SaaS, implementation services, managed services, and any future hardware or installation activity. \\
10 & Open a business bank account & Separate founder and company finances early so customer payments, tooling spend, cloud spend, taxes, and implementation costs are traceable and operationally clean. & \statusnotstarted & Open a business checking account immediately after formation and EIN issuance, define the chart of spending categories, and connect bookkeeping so product, cloud, services, and overhead costs can be tracked separately. \\
\bottomrule
\end{tabularx}
\footnotesize

\vspace{2pt}
{\scriptsize\textbf{Assumptions and status basis.}
Statuses are conservative and based on evidence visible in prior conversation context plus the uploaded SBM startup working file. ``Done'' is used only where the record clearly supports completion. The SBA step order and step names follow the live SBA page reviewed on May 6, 2026; SBM-specific interpretations reflect the current stated stage of validation / MVP definition, the security-first smart-building platform concept, and the present focus on wedge selection, MVP scope, data model, and pilot discovery.}

\end{document}
