\documentclass[11pt]{article}

% --- Page + typography ---
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{parskip}

% --- Lists, tables, links ---
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{hyperref}
\usepackage{nicefrac}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=black
}
\setlist[itemize]{topsep=4pt,itemsep=2pt,leftmargin=1.2em}
\setlist[enumerate]{topsep=6pt,itemsep=4pt,leftmargin=1.4em}

\title{\vspace{-1em}\textbf{Chef Salad}}
\date{} % no date
\author{} % no author

\begin{document}
\maketitle

\begin{center}
\itshape
A hearty main-meal salad that brings together crisp lettuces, classic deli meats, cheese, tomatoes, cucumbers, and seasoned croutons. Finish with your favorite dressing.
\end{center}

\vspace{0.5em}

% --- At-a-glance ---
\begin{tabularx}{\textwidth}{@{}>{\bfseries}l X @{} >{\bfseries}l X@{}}
Course & Salad & Cuisine & American \\
Prep Time & 10 minutes & Servings & 4 \\
Calories (per serving) & 371 kcal &  & \\
\end{tabularx}

\section*{Ingredients}

\textbf{For the Salad}
\begin{itemize}
  \item 1 medium head romaine lettuce, chopped
  \item 2 to 3 cups iceberg lettuce, chopped into bite-size pieces
  \item 4 hard-boiled eggs, peeled and quartered lengthwise
  \item 4 slices crispy cooked bacon
  \item 4 ounces deli-style sliced Swiss cheese, cut into julienne strips
  \item 4 ounces deli-style sliced ham, cut into julienne strips
  \item 4 ounces deli-style sliced turkey, cut into julienne strips
  \item 1 cup sliced English cucumber (about one-third of an English cucumber)
  \item 1 cup halved grape tomatoes
  \item 1 cup seasoned croutons
\end{itemize}

\textbf{For Serving}
\begin{itemize}
  \item \nicefrac{2}{3} to 1 cup salad dressing (ranch, chipotle ranch, avocado ranch, bleu cheese, Italian, or honey mustard)
  \item Kosher salt, to taste
  \item Freshly ground black pepper, to taste
\end{itemize}

\section*{Instructions}
\begin{enumerate}
  \item Add the romaine and iceberg lettuce to a large serving bowl.
  \item Top the lettuce with the eggs, bacon, Swiss cheese, ham, turkey, cucumbers, tomatoes, and croutons.
  \item Drizzle with your chosen dressing and toss gently to coat. Season with kosher salt and freshly ground black pepper to taste. Serve immediately.
\end{enumerate}

\section*{Ingredient Notes \& Substitutions}
\begin{itemize}
  \item \textbf{Lettuce:} Romaine is sturdy and holds up well; iceberg adds crunch. Use either or a mix.
  \item \textbf{Eggs:} Steam or pressure-cook eggs for easy peeling.
  \item \textbf{Meat:} Bacon, ham, and turkey are classic; chicken or roast beef also work well.
  \item \textbf{Cheese:} Swiss, cheddar, bleu, mozzarella, or provolone are all good choices.
  \item \textbf{Cucumber:} English cucumbers have tender skins and usually don’t need peeling. If using regular cucumbers, peel before slicing.
  \item \textbf{Tomatoes:} Grape, cherry, Campari, heirloom, or any ripe, sun-kissed tomatoes are great.
  \item \textbf{Croutons:} Homemade croutons elevate the salad significantly.
  \item \textbf{Dressings:} Ranch and bleu cheese are classic; thousand island, Italian, creamy poppyseed, avocado ranch, and honey mustard vinaigrette also pair well.
\end{itemize}

\section*{Preparation Tips}
\begin{itemize}
  \item Add-ins: olives, pepperoncini, avocado, red onion, and carrots all fit naturally.
  \item Make-ahead: assemble up to 1 day in advance without croutons or dressing. Cover tightly and refrigerate; add croutons and dressing just before serving.
  \item Customize: this salad is flexible—adjust meats, cheeses, and vegetables to suit your taste.
\end{itemize}

\section*{Nutrition (Per Serving)}
\begin{tabularx}{\textwidth}{@{}>{\bfseries}l >{\raggedleft\arraybackslash}X >{\bfseries}l >{\raggedleft\arraybackslash}X@{}}
Calories & 371 kcal & Carbohydrates & 15 g \\
Protein & 21 g & Fat (total) & 25 g \\
Saturated Fat & 8 g & Polyunsaturated Fat & 6 g \\
Monounsaturated Fat & 8 g & Trans Fat & 0.03 g \\
Cholesterol & 233 mg & Sodium & 833 mg \\
Potassium & 448 mg & Fiber & 2 g \\
Sugar & 7 g & Vitamin A & 1541 IU \\
Vitamin C & 8 mg & Calcium & 159 mg \\
Iron & 2 mg & & \\
\end{tabularx}

\vfill
\noindent\footnotesize
\textit{Notes:} Nutrition values are approximate and will vary with specific ingredients and dressing choices.

\end{document}

