\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{caption}
\captionsetup{justification=raggedright,singlelinecheck=false}

% Improve line-breaking without suppressing warnings.
\setlength{\emergencystretch}{2em}

\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{MISRA C++:2023 (C++17) - Guidelines (Paraphrased Headlines)}
\author{}
\date{\today}

\begin{document}
\maketitle

\raggedright


\section*{What this is}
This document lists every MISRA C++:2023 guideline (C++17) as a \textbf{paraphrased headline requirement}.
It includes the guideline identifier and category (Mandatory/Required/Advisory), and (for Rules) the guideline attributes of decidability and analysis scope.

\textbf{Important:} The official MISRA wording is proprietary; therefore this document does \emph{not} reproduce the standard's verbatim rule statements.
Use your MISRA document as the source of truth for canonical phrasing and interpretation.

\section*{Counts}
\begin{itemize}
  \item Directives: 4
  \item Rules: 175
  \item Mandatory: 5 \quad Required: 124 \quad Advisory: 50
\end{itemize}

\section{Mandatory}

\subsection{Rules}

{\small\setlength{\tabcolsep}{2pt}\renewcommand{\arraystretch}{1.05}%
\begin{longtable}{@{}P{2.2cm}P{1.7cm}P{2.1cm}P{3.0cm}P{6.7cm}@{}}
\caption{MISRA C++:2023 Rules (Mandatory) --- paraphrased headlines}\\
\toprule
ID & Category & Decidability & Scope & Requirement (paraphrased) \\
\midrule
\endfirsthead
\toprule
ID & Category & Decidability & Scope & Requirement (paraphrased) \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\emph{Continued on next page}} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
Rule 6.8.2 & Mandatory & Decidable & Single Translation Unit & Guideline intent: A function must not return a reference or a pointer to a local variable with automatic storage duration. \\
Rule 8.18.1 & Mandatory & Undecidable & System & Guideline intent: An object or subobject must not be copied to an overlapping object. \\
Rule 11.6.2 & Mandatory & Undecidable & System & Guideline intent: The value of an object must not be read before it has been set. \\
Rule 25.5.2 & Mandatory & Decidable & Single Translation Unit & Guideline intent: The pointers returned by the C++ Standard Library functions localeconv, getenv, setlocale or strerror must only be used as if they have pointer to const-quali9ed type. \\
Rule 25.5.3 & Mandatory & Undecidable & System & Guideline intent: The pointer returned by the C++ Standard Library functions asctime, ctime, gmtime, localtime, localeconv, getenv, setlocale or strerror must not be used following a subsequent call to the same function. \\
\end{longtable}}

\section{Required}

\subsection{Directives}

{\small\setlength{\tabcolsep}{2pt}\renewcommand{\arraystretch}{1.05}%
\begin{longtable}{@{}P{2.2cm}P{1.7cm}P{11.8cm}@{}}
\caption{MISRA C++:2023 Directives (Required) --- paraphrased headlines}\\
\toprule
ID & Category & Requirement (paraphrased) \\
\midrule
\endfirsthead
\toprule
ID & Category & Requirement (paraphrased) \\
\midrule
\endhead
\midrule
\multicolumn{3}{r}{\emph{Continued on next page}} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
Dir 0.3.2 & Required & Prohibited: function call is not permitted to violate the function’s preconditions General principles. \\
Dir 15.8.1 & Required & Required: User-provided copy assignment operators and move assignment operators is required to handle self-assignment Overloading. \\
\end{longtable}}

\subsection{Rules}

{\small\setlength{\tabcolsep}{2pt}\renewcommand{\arraystretch}{1.05}%
\begin{longtable}{@{}P{2.2cm}P{1.7cm}P{2.1cm}P{3.0cm}P{6.7cm}@{}}
\caption{MISRA C++:2023 Rules (Required) --- paraphrased headlines}\\
\toprule
ID & Category & Decidability & Scope & Requirement (paraphrased) \\
\midrule
\endfirsthead
\toprule
ID & Category & Decidability & Scope & Requirement (paraphrased) \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\emph{Continued on next page}} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
Rule 0.0.1 & Required & Decidable & Single Translation Unit & Prohibited: function is not permitted to contain unreachable statements. \\
Rule 0.1.2 & Required & Decidable & Single Translation Unit & Required: value returned by a function is required to be used. \\
Rule 0.2.2 & Required & Decidable & Single Translation Unit & Required: named function parameter is required to be used at least once. \\
Rule 4.1.1 & Required & Undecidable & System & Required: program is required to conform to ISO/IEC 14882:2017 (C++17). \\
Rule 4.1.3 & Required & Undecidable & System & Required: There is required to be no occurrence of unde9ned or critical unspeci9ed behaviour. \\
Rule 4.6.1 & Required & Undecidable & System & Required: Operations on a memory location is required to be sequenced appropriately Lexical conventions. \\
Rule 5.7.1 & Required & Decidable & Single Translation Unit & Prohibited: character sequence /* is not permitted to be used within a C-style comment 226 Licensed to: Jordan Suber 25 Nov 2024. \\
Rule 5.7.3 & Required & Decidable & Single Translation Unit & Prohibited: Line-splicing is not permitted to be used in // comments. \\
Rule 5.10.1 & Required & Decidable & Single Translation Unit & Required: User-de9ned identi9ers is required to have an appropriate form. \\
Rule 5.13.1 & Required & Decidable & Single Translation Unit & Required: Within character literals and non raw-string literals, \textbackslash{} is required to only be used to form a de9ned escape sequence or universal character name. \\
Rule 5.13.2 & Required & Decidable & Single Translation Unit & Required: Octal escape sequences, hexadecimal escape sequences and universal character names is required to be terminated. \\
Rule 5.13.3 & Required & Decidable & Single Translation Unit & Prohibited: Octal constants is not permitted to be used. \\
Rule 5.13.4 & Required & Decidable & Single Translation Unit & Required: Unsigned integer literals is required to be appropriately su[xed. \\
Rule 5.13.5 & Required & Decidable & Single Translation Unit & Prohibited: lowercase form of L is not permitted to be used as the 9rst character in a literal su[x. \\
Rule 5.13.6 & Required & Decidable & Single Translation Unit & Prohibited: integer-literal of type long long is not permitted to use a single L or l in any su[x. \\
Rule 5.13.7 & Required & Decidable & Single Translation Unit & Prohibited: String literals with diCerent encoding pre9xes is not permitted to be concatenated Basic concepts. \\
Rule 6.0.1 & Required & Decidable & Single Translation Unit & Prohibited: Block scope declarations is not permitted to be visually ambiguous. \\
Rule 6.0.4 & Required & Decidable & Single Translation Unit & Prohibited: identi9er main is not permitted to be used for a function other than the global function main. \\
Rule 6.2.1 & Required & Decidable & System & Prohibited: one-de9nition rule is not permitted to be violated. \\
Rule 6.2.2 & Required & Decidable & System & Required: All declarations of a variable or function is required to have the same type. \\
Rule 6.2.3 & Required & Decidable & System & Required: source code used to implement an entity is required to appear only once. \\
Rule 6.2.4 & Required & Decidable & Single Translation Unit & Prohibited: header file is not permitted to contain de9nitions of functions or objects that are non-inline and have external linkage. \\
Rule 6.4.1 & Required & Decidable & Single Translation Unit & Prohibited: variable declared in an inner scope is not permitted to hide a variable declared in an outer scope. \\
Rule 6.4.2 & Required & Decidable & Single Translation Unit & Prohibited: Derived classes is not permitted to conceal functions that are inherited from their bases. \\
Rule 6.4.3 & Required & Decidable & Single Translation Unit & Prohibited: name that is present in a dependent base is not permitted to be resolved by unquali9ed lookup. \\
Rule 6.7.1 & Required & Decidable & Single Translation Unit & Prohibited: Local variables is not permitted to have static storage duration. \\
Rule 6.7.2 & Required & Decidable & Single Translation Unit & Prohibited: Global variables is not permitted to be used. \\
Rule 6.8.1 & Required & Undecidable & System & Prohibited: object is not permitted to be accessed outside of its lifetime. \\
Rule 6.8.3 & Required & Decidable & Single Translation Unit & Prohibited: assignment operator is not permitted to assign the address of an object with automatic storage duration to an object with a greater lifetime. \\
Rule 6.9.1 & Required & Decidable & Single Translation Unit & Required: same type aliases is required to be used in all declarations of the same entity. \\
Rule 7.0.1 & Required & Decidable & Single Translation Unit & Required: There is required to be no conversion from type bool. \\
Rule 7.0.2 & Required & Decidable & Single Translation Unit & Required: There is required to be no conversion to type bool. \\
Rule 7.0.3 & Required & Decidable & Single Translation Unit & Prohibited: numerical value of a character is not permitted to be used. \\
Rule 7.0.4 & Required & Decidable & Single Translation Unit & Required: operands of bitwise operators and shift operators is required to be appropriate. \\
Rule 7.0.5 & Required & Decidable & Single Translation Unit & Prohibited: Integral promotion and the usual arithmetic conversions is not permitted to change the signedness or the type category of an operand. \\
Rule 7.0.6 & Required & Decidable & Single Translation Unit & Required: Assignment between numeric types is required to be appropriate. \\
Rule 7.11.1 & Required & Decidable & Single Translation Unit & Required: nullptr is required to be the only form of the null-pointer-constant. \\
Rule 7.11.2 & Required & Decidable & Single Translation Unit & Prohibited: array passed as a function argument is not permitted to decay to a pointer. \\
Rule 7.11.3 & Required & Decidable & Single Translation Unit & Required: conversion from function type to pointer-to-function type is required to only occur in appropriate contexts. \\
Rule 8.1.1 & Required & Decidable & Single Translation Unit & Prohibited: non-transient lambda is not permitted to implicitly capture this. \\
Rule 8.2.1 & Required & Decidable & Single Translation Unit & Required: virtual base class is required to only be cast to a derived class by means of dynamic\_cast. \\
Rule 8.2.2 & Required & Decidable & Single Translation Unit & Prohibited: C-style casts and functional notation casts is not permitted to be used. \\
Rule 8.2.3 & Required & Decidable & Single Translation Unit & Prohibited: cast is not permitted to remove any const or volatile quali9cation from the type accessed via a pointer or by reference. \\
Rule 8.2.4 & Required & Decidable & Single Translation Unit & Prohibited: Casts is not permitted to be performed between a pointer to function and any other type. \\
Rule 8.2.5 & Required & Decidable & Single Translation Unit & Prohibited: reinterpret\_cast is not permitted to be used. \\
Rule 8.2.6 & Required & Decidable & Single Translation Unit & Prohibited: object with integral, enumerated, or pointer to void type is not permitted to be cast to a pointer type. \\
Rule 8.2.8 & Required & Decidable & Single Translation Unit & Prohibited: object pointer type is not permitted to be cast to an integral type other than std::uintptr\_t or std::intptr\_t. \\
Rule 8.2.9 & Required & Decidable & Single Translation Unit & Prohibited: operand to typeid is not permitted to be an expression of polymorphic class type. \\
Rule 8.2.10 & Required & Undecidable & System & Prohibited: Functions is not permitted to call themselves, either directly or indirectly. \\
Rule 8.2.11 & Required & Decidable & Single Translation Unit & Required: argument passed via ellipsis is required to have an appropriate type. \\
Rule 8.7.1 & Required & Undecidable & System & Prohibited: Pointer arithmetic is not permitted to form an invalid pointer. \\
Rule 8.7.2 & Required & Undecidable & System & Required: Subtraction between pointers is required to only be applied to pointers that address elements of the same array. \\
Rule 8.9.1 & Required & Undecidable & System & Prohibited: built-in relational operators >, >=, < and <= is not permitted to be applied to objects of pointer type, except where they point to elements of the same array. \\
Rule 9.2.1 & Required & Decidable & Single Translation Unit & Prohibited: explicit type conversion is not permitted to be an expression statement. \\
Rule 9.3.1 & Required & Decidable & Single Translation Unit & Required: body of an iteration-statement or a selection-statement is required to be a compound-statement. \\
Rule 9.4.1 & Required & Decidable & Single Translation Unit & Required: All if ... else if constructs is required to be terminated with an else statement. \\
Rule 9.4.2 & Required & Decidable & Single Translation Unit & Required: structure of a switch statement is required to be appropriate. \\
Rule 9.5.2 & Required & Decidable & Single Translation Unit & Required: for-range-initializer is required to contain at most one function call. \\
Rule 9.6.2 & Required & Decidable & Single Translation Unit & Required: goto statement is required to reference a label in a surrounding block. \\
Rule 9.6.3 & Required & Decidable & Single Translation Unit & Required: goto statement is required to jump to a label declared later in the function body. \\
Rule 9.6.4 & Required & Undecidable & System & Guideline intent: comply with the corresponding MISRA requirement. \\
Rule 9.6.5 & Required & Decidable & Single Translation Unit & Required: function with non-void return type is required to return a value on all paths Declarations. \\
Rule 10.1.2 & Required & Decidable & Single Translation Unit & Required: volatile quali9er is required to be used appropriately. \\
Rule 10.2.1 & Required & Decidable & Single Translation Unit & Required: enumeration is required to be de9ned with an explicit underlying type. \\
Rule 10.2.3 & Required & Decidable & Single Translation Unit & Prohibited: numeric value of an unscoped enumeration with no 9xed underlying type is not permitted to be used. \\
Rule 10.4.1 & Required & Decidable & Single Translation Unit & Prohibited: asm declaration is not permitted to be used Declarators. \\
Rule 11.6.3 & Required & Decidable & Single Translation Unit & Required: Within an enumerator list, the value of an implicitly-speci9ed enumeration constant is required to be unique. \\
Rule 12.2.2 & Required & Decidable & Single Translation Unit & Required: bit-9eld is required to have an appropriate type. \\
Rule 12.2.3 & Required & Decidable & Single Translation Unit & Prohibited: named bit-9eld with signed integer type is not permitted to have a length of one bit. \\
Rule 12.3.1 & Required & Decidable & Single Translation Unit & Prohibited: union keyword is not permitted to be used Derived classes. \\
Rule 13.1.2 & Required & Decidable & Single Translation Unit & Prohibited: accessible base class is not permitted to be both virtual and non-virtual in the same hierarchy. \\
Rule 13.3.1 & Required & Decidable & Single Translation Unit & Required: User-declared member functions is required to use the virtual, override and 9nal speci9ers appropriately. \\
Rule 13.3.2 & Required & Decidable & Single Translation Unit & Prohibited: Parameters in an overriding virtual function is not permitted to specify diCerent default arguments. \\
Rule 13.3.3 & Required & Decidable & System & Required: parameters in all declarations or overrides of a function is required to either be unnamed or have identical names. \\
Rule 13.3.4 & Required & Decidable & Single Translation Unit & Required: comparison of a potentially virtual pointer to member function is required to only be with nullptr Member access control. \\
Rule 15.0.1 & Required & Decidable & Single Translation Unit & Required: Special member functions is required to be provided appropriately. \\
Rule 15.1.1 & Required & Undecidable & System & Prohibited: object’s dynamic type is not permitted to be used from within its constructor or destructor. \\
Rule 15.1.3 & Required & Decidable & Single Translation Unit & Required: Conversion operators and constructors that are callable with a single argument is required to be explicit. \\
Rule 15.1.5 & Required & Decidable & Single Translation Unit & Required: class is required to only de9ne an initializer-list constructor when it is the only constructor. \\
Rule 16.5.1 & Required & Decidable & Single Translation Unit & Prohibited: logical AND and logical OR operators is not permitted to be overloaded. \\
Rule 16.5.2 & Required & Decidable & Single Translation Unit & Prohibited: address-of operator is not permitted to be overloaded. \\
Rule 17.8.1 & Required & Decidable & Single Translation Unit & Prohibited: Function templates is not permitted to be explicitly specialized Exception handling. \\
Rule 18.1.1 & Required & Decidable & Single Translation Unit & Prohibited: exception object is not permitted to have pointer type. \\
Rule 18.1.2 & Required & Decidable & Single Translation Unit & Required: empty throw is required to only occur within the compound-statement of a catch handler. \\
Rule 18.3.2 & Required & Decidable & Single Translation Unit & Required: exception of class type is required to be caught by const reference or reference. \\
Rule 18.3.3 & Required & Decidable & Single Translation Unit & Prohibited: Handlers for a function-try-block of a constructor or destructor is not permitted to refer to non-static members from their class or its bases. \\
Rule 18.4.1 & Required & Decidable & Single Translation Unit & Required: Exception-unfriendly functions is required to be noexcept. \\
Rule 19.0.1 & Required & Decidable & Single Translation Unit & Required: line whose 9rst token is \# is required to be a valid preprocessing directive. \\
Rule 19.0.2 & Required & Decidable & Single Translation Unit & Prohibited: Function-like macros is not permitted to be de9ned. \\
Rule 19.1.1 & Required & Decidable & Single Translation Unit & Required: de9ned preprocessor operator is required to be used appropriately. \\
Rule 19.1.2 & Required & Decidable & Single Translation Unit & Required: All \#else, \#elif and \#endif preprocessor directives is required to reside in the same file as the \#if, \#ifdef or \#ifndef directive to which they are related. \\
Rule 19.1.3 & Required & Decidable & Single Translation Unit & Required: All identi9ers used in the controlling expression of \#if or \#elif preprocessing directives is required to be de9ned prior to evaluation. \\
Rule 19.2.1 & Required & Decidable & Single Translation Unit & Required: Precautions is required to be taken in order to prevent the contents of a header file being included more than once. \\
Rule 19.2.2 & Required & Decidable & Single Translation Unit & Required: \#include directive is required to be followed by either a <filename> or "filename" sequence. \\
Rule 19.2.3 & Required & Decidable & Single Translation Unit & Prohibited: ' or " or \textbackslash{} characters and the /* or // character sequences is not permitted to occur in a header file name. \\
Rule 19.3.2 & Required & Decidable & Single Translation Unit & Prohibited: macro parameter immediately following a \# operator is not permitted to be immediately followed by a \#\# operator. \\
Rule 19.3.3 & Required & Decidable & Single Translation Unit & Prohibited: argument to a mixed-use macro parameter is not permitted to be subject to further expansion. \\
Rule 19.3.4 & Required & Decidable & Single Translation Unit & Required: Parentheses is required to be used to ensure macro arguments are expanded appropriately. \\
Rule 19.3.5 & Required & Decidable & Single Translation Unit & Prohibited: Tokens that look like a preprocessing directive is not permitted to occur within a macro argument. \\
Rule 21.2.1 & Required & Decidable & Single Translation Unit & Prohibited: library functions atof, atoi, atol and atoll from <cstdlib> is not permitted to be used. \\
Rule 21.2.2 & Required & Decidable & Single Translation Unit & Prohibited: string handling functions from <cstring>, <cstdlib>, <cwchar> and <cinttypes> is not permitted to be used. \\
Rule 21.2.3 & Required & Decidable & Single Translation Unit & Prohibited: library function system from <cstdlib> is not permitted to be used. \\
Rule 21.2.4 & Required & Decidable & Single Translation Unit & Prohibited: macro oCsetof is not permitted to be used. \\
Rule 21.6.2 & Required & Decidable & Single Translation Unit & Required: Dynamic memory is required to be managed automatically. \\
Rule 21.6.3 & Required & Decidable & Single Translation Unit & Prohibited: Advanced memory management is not permitted to be used. \\
Rule 21.6.4 & Required & Decidable & System & Required: If a project de9nes either a sized or unsized version of a global operator delete, then both is required to be de9ned. \\
Rule 21.6.5 & Required & Decidable & Single Translation Unit & Prohibited: pointer to an incomplete class type is not permitted to be deleted. \\
Rule 21.10.1 & Required & Decidable & Single Translation Unit & Prohibited: features of <cstdarg> is not permitted to be used. \\
Rule 21.10.2 & Required & Decidable & Single Translation Unit & Prohibited: standard header file <csetjmp> is not permitted to be used. \\
Rule 21.10.3 & Required & Decidable & Single Translation Unit & Prohibited: facilities provided by the standard header file <csignal> is not permitted to be used. \\
Rule 22.3.1 & Required & Decidable & Single Translation Unit & Prohibited: assert macro is not permitted to be used with a constant-expression. \\
Rule 22.4.1 & Required & Decidable & Single Translation Unit & Required: literal value zero is required to be the only value assigned to errno. \\
Rule 24.5.1 & Required & Decidable & Single Translation Unit & Prohibited: character handling functions from <cctype> and <cwctype> is not permitted to be used. \\
Rule 24.5.2 & Required & Decidable & Single Translation Unit & Prohibited: C++ Standard Library functions memcpy, memmove and memcmp from <cstring> is not permitted to be used. \\
Rule 25.5.1 & Required & Decidable & Single Translation Unit & Prohibited: setlocale and std::locale::global functions is not permitted to be called. \\
Rule 28.3.1 & Required & Undecidable & System & Prohibited: Predicates is not permitted to have persistent side eCects. \\
Rule 28.6.1 & Required & Decidable & Single Translation Unit & Required: argument to std::move is required to be a non-const lvalue. \\
Rule 28.6.2 & Required & Decidable & Single Translation Unit & Required: Forwarding references and std::forward is required to be used together. \\
Rule 28.6.3 & Required & Decidable & Single Translation Unit & Prohibited: object is not permitted to be used while in a potentially moved-from state. \\
Rule 28.6.4 & Required & Decidable & Single Translation Unit & Required: result of std::remove, std::remove\_if, std::unique and empty is required to be used. \\
Rule 30.0.1 & Required & Decidable & Single Translation Unit & Prohibited: C Library input/output functions is not permitted to be used. \\
Rule 30.0.2 & Required & Undecidable & System & Required: Reads and writes on the same file stream is required to be separated by a positioning operation. \\
\end{longtable}}

\section{Advisory}

\subsection{Directives}

{\small\setlength{\tabcolsep}{2pt}\renewcommand{\arraystretch}{1.05}%
\begin{longtable}{@{}P{2.2cm}P{1.7cm}P{11.8cm}@{}}
\caption{MISRA C++:2023 Directives (Advisory) --- paraphrased headlines}\\
\toprule
ID & Category & Requirement (paraphrased) \\
\midrule
\endfirsthead
\toprule
ID & Category & Requirement (paraphrased) \\
\midrule
\endhead
\midrule
\multicolumn{3}{r}{\emph{Continued on next page}} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
Dir 0.3.1 & Advisory & Recommended: Floating-point arithmetic ought to be used appropriately. \\
Dir 5.7.2 & Advisory & Discouraged: Sections of code should be avoided; do not be “commented out”. \\
\end{longtable}}

\subsection{Rules}

{\small\setlength{\tabcolsep}{2pt}\renewcommand{\arraystretch}{1.05}%
\begin{longtable}{@{}P{2.2cm}P{1.7cm}P{2.1cm}P{3.0cm}P{6.7cm}@{}}
\caption{MISRA C++:2023 Rules (Advisory) --- paraphrased headlines}\\
\toprule
ID & Category & Decidability & Scope & Requirement (paraphrased) \\
\midrule
\endfirsthead
\toprule
ID & Category & Decidability & Scope & Requirement (paraphrased) \\
\midrule
\endhead
\midrule
\multicolumn{5}{r}{\emph{Continued on next page}} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
Rule 0.0.2 & Advisory & Undecidable & System & Discouraged: Controlling expressions should be avoided; do not be invariant. \\
Rule 0.1.1 & Advisory & Undecidable & System & Discouraged: value should be avoided; do not be unnecessarily written to a local object. \\
Rule 0.2.1 & Advisory & Decidable & Single Translation Unit & Recommended: Variables with limited visibility ought to be used at least once. \\
Rule 0.2.3 & Advisory & Decidable & Single Translation Unit & Recommended: Types with limited visibility ought to be used at least once. \\
Rule 0.2.4 & Advisory & Decidable & System & Recommended: Functions with limited visibility ought to be used at least once. \\
Rule 4.1.2 & Advisory & Decidable & Single Translation Unit & Discouraged: Deprecated features should be avoided; do not be used. \\
Rule 5.0.1 & Advisory & Decidable & Single Translation Unit & Discouraged: Trigraph-like sequences should be avoided; do not be used. \\
Rule 6.0.2 & Advisory & Decidable & Single Translation Unit & Recommended: When an array with external linkage is declared, its size ought to be explicitly speci9ed. \\
Rule 6.0.3 & Advisory & Decidable & Single Translation Unit & Recommended: only declarations in the global namespace ought to be main, namespace declarations and extern "C" declarations. \\
Rule 6.5.1 & Advisory & Decidable & Single Translation Unit & Recommended: function or object with external linkage ought to be introduced in a header file. \\
Rule 6.5.2 & Advisory & Decidable & Single Translation Unit & Recommended: Internal linkage ought to be speci9ed appropriately. \\
Rule 6.8.4 & Advisory & Decidable & Single Translation Unit & Recommended: Member functions returning references to their object ought to be ref- quali9ed appropriately. \\
Rule 6.9.2 & Advisory & Decidable & Single Translation Unit & Discouraged: names of the standard signed integer types and standard unsigned integer types should be avoided; do not be used Standard conversions. \\
Rule 8.0.1 & Advisory & Decidable & Single Translation Unit & Recommended: Parentheses ought to be used to make the meaning of an expression appropriately explicit. \\
Rule 8.1.2 & Advisory & Decidable & Single Translation Unit & Recommended: Variables ought to be captured explicitly in a non-transient lambda. \\
Rule 8.2.7 & Advisory & Decidable & Single Translation Unit & Discouraged: cast should be avoided; do not convert a pointer type to an integral type. \\
Rule 8.3.1 & Advisory & Decidable & Single Translation Unit & Discouraged: built-in unary - operator should be avoided; do not be applied to an expression of unsigned type. \\
Rule 8.3.2 & Advisory & Decidable & Single Translation Unit & Discouraged: built-in unary + operator should be avoided; do not be used. \\
Rule 8.14.1 & Advisory & Undecidable & System & Discouraged: right-hand operand of a logical \&\& or || operator should be avoided; do not contain persistent side eCects. \\
Rule 8.18.2 & Advisory & Decidable & Single Translation Unit & Discouraged: result of an assignment operator should be avoided; do not be used. \\
Rule 8.19.1 & Advisory & Decidable & Single Translation Unit & Discouraged: comma operator should be avoided; do not be used. \\
Rule 8.20.1 & Advisory & Decidable & Single Translation Unit & Discouraged: unsigned arithmetic operation with constant operands should be avoided; do not wrap. \\
Rule 9.5.1 & Advisory & Decidable & Single Translation Unit & Recommended: Legacy for statements ought to be simple. \\
Rule 9.6.1 & Advisory & Decidable & Single Translation Unit & Discouraged: goto statement should be avoided; do not be used. \\
Rule 10.0.1 & Advisory & Decidable & Single Translation Unit & Discouraged: declaration should be avoided; do not declare more than one variable or member variable. \\
Rule 10.1.1 & Advisory & Decidable & Single Translation Unit & Recommended: target type of a pointer or lvalue reference parameter ought to be const-quali9ed appropriately. \\
Rule 10.2.2 & Advisory & Decidable & Single Translation Unit & Discouraged: Unscoped enumerations should be avoided; do not be declared. \\
Rule 10.3.1 & Advisory & Decidable & Single Translation Unit & Recommended: There ought to be no unnamed namespaces in header files. \\
Rule 11.3.1 & Advisory & Decidable & Single Translation Unit & Discouraged: Variables of array type should be avoided; do not be declared. \\
Rule 11.3.2 & Advisory & Decidable & Single Translation Unit & Recommended: declaration of an object ought to contain no more than two levels of pointer indirection. \\
Rule 11.6.1 & Advisory & Decidable & Single Translation Unit & Recommended: All variables ought to be initialized. \\
Rule 12.2.1 & Advisory & Decidable & Single Translation Unit & Discouraged: Bit-9elds should be avoided; do not be declared. \\
Rule 13.1.1 & Advisory & Decidable & Single Translation Unit & Discouraged: Classes should be avoided; do not be inherited virtually. \\
Rule 14.1.1 & Advisory & Decidable & Single Translation Unit & Recommended: Non-static data members ought to be either all private or all public Special member functions. \\
Rule 15.0.2 & Advisory & Decidable & Single Translation Unit & Recommended: User-provided copy and move member functions of a class ought to have appropriate signatures. \\
Rule 15.1.2 & Advisory & Decidable & Single Translation Unit & Recommended: All constructors of a class ought to explicitly initialize all of its virtual base classes and immediate base classes. \\
Rule 15.1.4 & Advisory & Decidable & Single Translation Unit & Recommended: All direct, non-static data members of a class ought to be initialized before the class object is accessible. \\
Rule 16.6.1 & Advisory & Decidable & Single Translation Unit & Recommended: Symmetrical operators ought to only be implemented as non-member functions Templates. \\
Rule 18.3.1 & Advisory & Decidable & Single Translation Unit & Recommended: There ought to be at least one exception handler to catch all otherwise unhandled exceptions. \\
Rule 18.5.1 & Advisory & Undecidable & System & Discouraged: noexcept function should be avoided; do not attempt to propagate an exception to the calling function. \\
Rule 18.5.2 & Advisory & Decidable & Single Translation Unit & Discouraged: Program-terminating functions should be avoided; do not be used Preprocessing directives. \\
Rule 19.0.3 & Advisory & Decidable & Single Translation Unit & Recommended: \#include directives ought to only be preceded by preprocessor directives or comments. \\
Rule 19.0.4 & Advisory & Decidable & Single Translation Unit & Recommended: \#undef ought to only be used for macros de9ned previously in the same file. \\
Rule 19.3.1 & Advisory & Decidable & Single Translation Unit & Discouraged: \# and \#\# preprocessor operators should be avoided; do not be used. \\
Rule 19.6.1 & Advisory & Decidable & Single Translation Unit & Discouraged: \#pragma directive and the \_Pragma operator should be avoided; do not be used. \\
Rule 21.6.1 & Advisory & Undecidable & Single Translation Unit & Discouraged: Dynamic memory should be avoided; do not be used. \\
Rule 23.11.1 & Advisory & Decidable & Single Translation Unit & Discouraged: raw pointer constructors of std::shared\_ptr and std::unique\_ptr should be avoided; do not be used. \\
Rule 26.3.1 & Advisory & Decidable & Single Translation Unit & Discouraged: std::vector should be avoided; do not be specialized with bool. \\
\end{longtable}}

\end{document}
