\documentclass[11pt]{article}

% ---------- Common, stable packages ----------
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{amsmath}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% ---------- Helpers ----------
\setlist[itemize]{itemsep=2pt, topsep=4pt}
\setlist[enumerate]{itemsep=4pt, topsep=6pt}
\newcommand{\Section}[1]{\vspace{0.6em}\noindent\textbf{\Large #1}\par\vspace{0.25em}}
\newcommand{\Subsection}[1]{\vspace{0.4em}\noindent\textbf{\large #1}\par\vspace{0.2em}}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\begin{document}

\begin{center}
  {\LARGE \textbf{Eggplant Parmesan}}\\[4pt]
  \small Baked, crispy eggplant layered with herbed crumbs, tomato sauce, and cheeses; finished with fresh basil.
\end{center}

\Section{Overview}
A lighter, baked take on the classic: panko\,+\,Parmesan--crusted eggplant rounds are oven--crisped (not fried), then layered with marinara and fresh mozzarella and baked until bubbling.

\Section{At a Glance}
\begin{tabularx}{\textwidth}{@{} l L @{}}
\toprule
\textbf{Yield} & Serves 6--8 \\
\textbf{Prep Time} & 20 minutes \\
\textbf{Cook Time} & 40 minutes \\
\textbf{Total Time} & 1 hour \\
\textbf{Method} & Bake breaded eggplant; layer with sauces/cheeses; bake and broil to finish \\
\bottomrule
\end{tabularx}

\Section{Ingredients}

\Subsection{Main}
\begin{itemize}
  \item 2 large eggplants, cut into $\tfrac{1}{4}$-inch--thick rounds
  \item 2 large eggs, beaten
  \item $\tfrac{1}{4}$ cup almond milk
  \item 1\,$\tfrac{1}{2}$ cups panko breadcrumbs
  \item 1\,$\tfrac{1}{4}$ cups grated Parmesan cheese, divided
  \item 2 teaspoons dried oregano
  \item 2 tablespoons fresh thyme leaves
  \item $\tfrac{1}{2}$ teaspoon red pepper flakes
  \item $\tfrac{1}{2}$ teaspoon fine sea salt, plus more for sprinkling
  \item Freshly ground black pepper
  \item Extra-virgin olive oil, for drizzling
  \item 28 ounces marinara sauce (homemade or good-quality jarred)
  \item 2 large fresh mozzarella balls, thinly sliced
  \item $\tfrac{1}{3}$ cup fresh basil leaves, for topping
\end{itemize}

\Section{Equipment}
2 rimmed baking sheets; parchment paper; 8\,$\times$\,12- or 9\,$\times$\,13-inch baking dish; 2 shallow dishes; tongs.

\Section{Instructions}
\begin{enumerate}
  \item \textbf{Heat oven.} Preheat to $400^\circ$F (200$^\circ$C). Line two baking sheets with parchment.%
  \footnote{If your oven has convection, use it here for extra browning and crunch.}
  \item \textbf{Set up dredge.} In one shallow dish, whisk eggs with almond milk. In another, combine panko, 1 cup Parmesan, oregano, thyme, red pepper flakes, $\tfrac{1}{2}$ teaspoon salt, and several grinds of pepper.
  \item \textbf{Bread eggplant.} Working one slice at a time, dip eggplant into egg mixture (allow excess to drip), then press into the panko mixture to coat well on all sides. Arrange in a single layer on the prepared sheets.
  \item \textbf{Bake until crisp.} Drizzle the tops lightly with olive oil. Bake 16--18 minutes, until tender and golden brown, rotating sheets once.
  \item \textbf{Layer.} Spread $\tfrac{1}{2}$ cup marinara in the bottom of an 8\,$\times$\,12- or 9\,$\times$\,13-inch baking dish. Add half of the baked eggplant in an even layer. Spoon over 1 cup marinara, then half the mozzarella slices. Season lightly with salt and pepper. Repeat with remaining eggplant, remaining marinara, and remaining mozzarella. Sprinkle with the remaining $\tfrac{1}{4}$ cup Parmesan and a light drizzle of olive oil; add a few pinches of sea salt.
  \item \textbf{Bake \& broil.} Bake at $400^\circ$F for about 20 minutes, until bubbling. Switch to broil for 2--4 minutes to brown the cheese in spots (watch closely).
  \item \textbf{Finish.} Let rest 5--10 minutes. Top with torn basil leaves, slice, and serve.
\end{enumerate}

\Section{Pro Tips}
\begin{itemize}
  \item \textbf{Even slices} ensure even cooking; aim for consistent $\tfrac{1}{4}$-inch rounds and uniform mozzarella thickness.
  \item \textbf{Grate your Parmesan} for best flavor (Parmigiano Reggiano or pecorino).
  \item \textbf{Don’t skimp on crumbs:} a generous panko\,/\,Parmesan\,/\,herb coating drives texture and flavor.
  \item \textbf{Olive oil matters:} a light drizzle before baking promotes deep golden, crisp results.
  \item \textbf{Season between layers} (a pinch of salt and pepper over sauces/cheeses) for a well-seasoned final bake.
\end{itemize}

\Section{Serving Suggestions}
Hearty enough to stand alone, or pair with pasta (or zucchini noodles/spaghetti squash), crusty bread, and simple veggie sides like roasted broccoli/cauliflower/tomatoes, grilled zucchini, or a bright salad (Caprese, chopped Italian, Caesar, or panzanella).

\vfill
\begin{center}
  \footnotesize
  \textit{Make-ahead:} Bake eggplant rounds up to 1 day in advance; cool, refrigerate, and assemble/bake just before serving.
\end{center}

\end{document}

