
\documentclass[11pt]{article}

\usepackage[letterpaper,margin=0.72in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{bookmark}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{calc}
\usepackage{amsmath,amssymb}
\usepackage{textcomp}
\usepackage{upquote}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{etoolbox}

\definecolor{CIIBlue}{HTML}{1F4E79}
\definecolor{CIILightBlue}{HTML}{EAF3F8}
\definecolor{CIIGold}{HTML}{B8860B}
\definecolor{CIIGoldLight}{HTML}{FFF8E6}
\definecolor{CIIGreen}{HTML}{2E6F40}
\definecolor{CIIPurple}{HTML}{5B4B8A}
\definecolor{CIIRule}{HTML}{B7C9D6}
\definecolor{CIIGrey}{HTML}{F7F7F7}
\definecolor{CIIDarkGrey}{HTML}{444444}

\hypersetup{
  colorlinks=true,
  linkcolor=CIIBlue,
  urlcolor=CIIBlue,
  citecolor=CIIBlue,
  pdftitle={C Interfaces and Implementations Algorithms as C-like Pseudocode},
  pdfauthor={Jordan Suber / ChatGPT},
  pdfsubject={Pseudocode handout based on David R. Hanson's C Interfaces and Implementations},
  pdfkeywords={C, interfaces, implementations, reusable software, pseudocode, algorithms, ADT, memory management, hash tables, strings, arbitrary precision arithmetic, threads}
}

\pagestyle{fancy}
\fancyhf{}
\lhead{\textit{C Interfaces and Implementations} Pseudocode}
\rhead{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\headheight}{14pt}

\titleformat{\section}{\Large\bfseries\color{CIIBlue}}{}{0pt}{}
\titleformat{\subsection}{\large\bfseries\color{CIIBlue}}{}{0pt}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{CIIDarkGrey}}{}{0pt}{}
\titlespacing*{\section}{0pt}{1.35em}{0.5em}
\titlespacing*{\subsection}{0pt}{0.95em}{0.35em}
\titlespacing*{\subsubsection}{0pt}{0.75em}{0.25em}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{2}
\setlist[itemize]{leftmargin=1.4em,itemsep=0.2em,topsep=0.2em}
\setlist[enumerate]{leftmargin=1.6em,itemsep=0.25em,topsep=0.25em}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.52em}
\setlength{\LTpre}{0.35em}
\setlength{\LTpost}{0.35em}
\emergencystretch=4em
\sloppy

\lstdefinestyle{ciicode}{
  language=C,
  basicstyle=\ttfamily\scriptsize,
  keywordstyle=\bfseries\color{CIIBlue},
  commentstyle=\itshape\color{CIIDarkGrey},
  stringstyle=\color{CIIGreen},
  numbers=left,
  numberstyle=\tiny\color{CIIDarkGrey},
  stepnumber=1,
  numbersep=0.7em,
  breaklines=true,
  breakatwhitespace=false,
  columns=fullflexible,
  keepspaces=true,
  tabsize=4,
  showstringspaces=false,
  frame=single,
  framerule=0.3pt,
  rulecolor=\color{CIIRule},
  backgroundcolor=\color{CIIGrey},
  xleftmargin=1.0em,
  framexleftmargin=0.8em,
  framexrightmargin=0.35em,
  framesep=0.42em,
  aboveskip=0.55em,
  belowskip=0.55em,
  morekeywords={type,bool,true,false,NULL,size_t,uintptr_t,int64_t,uint64_t,BigInt,XP,AP,MP,Thread,Sem,Chan,Except,Atom,Text,Arena,Table,Set,List,Seq,Ring,Bit,foreach,repeat,until,continue,break,raise,assert,alloc,free,copy,zero,length,push,pop,enqueue,dequeue,hash,resize,canonicalize,normalize,swap,rotate,signal,wait,yield,dispatch,block,unblock}
}
\lstset{style=ciicode}

\newtcolorbox{scopebox}{
  enhanced,
  colback=CIILightBlue,
  colframe=CIIBlue,
  boxrule=0.5pt,
  arc=2mm,
  left=1em,
  right=1em,
  top=0.7em,
  bottom=0.7em
}

\newtcolorbox{methodbox}{
  enhanced,
  colback=CIIGrey,
  colframe=CIIRule,
  boxrule=0.4pt,
  arc=2mm,
  left=0.9em,
  right=0.9em,
  top=0.6em,
  bottom=0.6em
}

\newtcolorbox{warningbox}{
  enhanced,
  colback=CIIGoldLight,
  colframe=CIIGold,
  boxrule=0.45pt,
  arc=2mm,
  left=0.9em,
  right=0.9em,
  top=0.6em,
  bottom=0.6em
}

% Pandoc compatibility helpers
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\providecommand{\passthrough}[1]{#1}
\AtBeginEnvironment{longtable}{\small}
\renewcommand{\arraystretch}{1.14}

\begin{document}
\hypersetup{pageanchor=false}
\pagenumbering{gobble}

\begin{titlepage}
  \centering
  \vspace*{1.5cm}
  {\Huge\bfseries\color{CIIBlue}C Interfaces and Implementations\par}
  \vspace{0.35cm}
  {\LARGE Algorithms as C-like Pseudocode\par}
  \vspace{0.25cm}
  {\large Based on David R. Hanson, \textit{C Interfaces and Implementations: Techniques for Creating Reusable Software}\par}
  \vspace{0.95cm}
  \begin{scopebox}
  This handout reorganizes the book's literate C implementations into implementation-oriented pseudocode. It is designed for studying reusable C interface design, abstract data type implementation, memory-management discipline, string processing, arithmetic modules, and cooperative threading mechanics.
  \end{scopebox}
  \vspace{0.75cm}
  \begin{methodbox}
  \textbf{Scope:} systematic extraction from the uploaded book into independent C-like pseudocode.\\
  \textbf{Primary modules:} atoms, exceptions, assertions, memory management, arenas, lists, tables, sets, arrays, sequences, rings, bit vectors, formatting, low-level strings, high-level text, extended precision arithmetic, arbitrary precision arithmetic, multiple precision arithmetic, threads, semaphores, and channels.\\
  \textbf{Boundary:} the algorithms are rewritten as study-grade pseudocode rather than copied source listings.
  \end{methodbox}
  \vfill
  {\large Prepared May 29, 2026\par}
\end{titlepage}

\hypersetup{pageanchor=true}
\pagenumbering{roman}
\tableofcontents
\clearpage
\pagenumbering{arabic}

\begin{warningbox}
\textbf{Implementation note.} The original text is a literate-programming treatment of production-quality C interfaces. This handout keeps the exported names where they clarify orientation, but rewrites routine bodies in compact pseudocode. Use the official interface contracts and platform constraints before translating these sketches into production code.
\end{warningbox}

\section{C Interfaces and Implementations --- Algorithms as C-like Pseudocode}\label{c-interfaces-and-implementations-algorithms-as-c-like-pseudocode}

\textbf{Source analyzed:} David R. Hanson, \emph{C Interfaces and Implementations: Techniques for Creating Reusable Software}.

\textbf{Purpose:} Systematically extract the algorithms embodied in the book's literate C implementations and rewrite them as C-like pseudocode suitable for study, reimplementation, and architecture review.

\textbf{Copyright-safe scope:} This is an independent algorithmic reconstruction. It preserves the public interface names where they are useful for orientation, but it does not reproduce the book's source listings verbatim.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Reading Model and Pseudocode Conventions}\label{reading-model-and-pseudocode-conventions}

The book is organized around reusable C interfaces and their implementations. The algorithms therefore fall into four overlapping classes:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Interface contract algorithms} --- argument validation, checked runtime errors, exception raising, ownership rules, and representation invariants.
\item
  \textbf{Abstract data type algorithms} --- stack, list, table, set, array, sequence, ring, and bit-vector operations.
\item
  \textbf{Systems-support algorithms} --- memory allocation, arena allocation, exception frames, formatted output, and cooperative threads.
\item
  \textbf{Numeric/string algorithms} --- atoms, low-level string slicing, high-level text values, and extended/arbitrary/multiple precision arithmetic.
\end{enumerate}

The pseudocode uses the following conventions:

\begin{lstlisting}[language=C]
assert(condition);          // checked runtime error if false
raise(Exception, file,line); // nonlocal exception transfer
alloc(n);                  // allocate n bytes or raise allocation failure
free(p);                   // deallocate p when ownership is held
copy(dst, src, n);          // copy n bytes
zero(dst, n);               // write n zero bytes
length(x);                 // logical length, not necessarily byte length
\end{lstlisting}

Opaque ADTs are written as \passthrough{\lstinline!T!}. A \passthrough{\lstinline!Node!} is an internal representation record. \passthrough{\lstinline!HashTable!} means an array of bucket chains unless noted otherwise. Most algorithms assume ordinary C-style pointer semantics but are written to expose intent rather than exact syntax.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. Complete Algorithm Inventory}\label{complete-algorithm-inventory}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Area
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Interfaces / Modules
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Core algorithms extracted
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Foundations & \passthrough{\lstinline!Arith!}, \passthrough{\lstinline!Stack!} & Euclidean-style division/modulus semantics, stack push/pop/free \\
Atoms & \passthrough{\lstinline!Atom!} & hash-consing byte strings, integer-to-atom conversion, atom length lookup \\
Exceptions & \passthrough{\lstinline!Except!}, \passthrough{\lstinline!Assert!} & exception frame stack, raise/reraise, TRY/EXCEPT/FINALLY control transfer, assertion failure \\
Memory & \passthrough{\lstinline!Mem!}, checking \passthrough{\lstinline!Mem!}, \passthrough{\lstinline!Arena!} & allocation wrappers, first-fit debug allocator, descriptor hash table, arena chunk allocation/free \\
Lists & \passthrough{\lstinline!List!} & linked-list construction, traversal, append, reverse, map, to-array \\
Tables & \passthrough{\lstinline!Table!} & chained hash table lookup, insert/update, remove, map, resize-like bucket allocation \\
Sets & \passthrough{\lstinline!Set!} & chained hash set membership, insertion, removal, union/intersection/difference/minus \\
Arrays & \passthrough{\lstinline!Array!}, \passthrough{\lstinline!ArrayRep!} & dynamic fixed-element array allocation, resize, get/put/copy \\
Sequences & \passthrough{\lstinline!Seq!} & deque over circular buffer, grow-on-demand, high/low insertion/removal \\
Rings & \passthrough{\lstinline!Ring!} & circular doubly linked list, indexed access by shortest direction, rotate, insert/remove \\
Bit vectors & \passthrough{\lstinline!Bit!} & bit addressing, range set/clear/not, bit count, set algebra over machine words \\
Formatting & \passthrough{\lstinline!Fmt!} & printf-like format parser, converter registry, width/precision/padding emitters \\
Low-level strings & \passthrough{\lstinline!Str!} & normalized slice indexing, search/match, concatenation, mapping, reverse, duplication \\
High-level strings & \passthrough{\lstinline!Text!} & length-bearing text slices, atomized/copying text creation, save/restore pool, search/map \\
Extended precision & \passthrough{\lstinline!XP!} & unsigned digit-vector arithmetic, add/sub/mul/div, conversion, bit shifts \\
Arbitrary precision & \passthrough{\lstinline!AP!} & signed big integers layered on XP, arithmetic, parsing/formatting, exponentiation \\
Multiple precision & \passthrough{\lstinline!MP!} & fixed-width signed/unsigned multiword arithmetic, bitwise ops, shifts, division, conversion \\
Threads & \passthrough{\lstinline!Thread!}, \passthrough{\lstinline!Sem!}, \passthrough{\lstinline!Chan!} & cooperative thread creation, ready/wait queues, semaphore wait/signal, rendezvous channels \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. Foundation Algorithms}\label{foundation-algorithms}

\subsubsection{\texorpdfstring{3.1 \texttt{Arith\_max} and \texttt{Arith\_min}}{3.1 Arith\_max and Arith\_min}}\label{arith_max-and-arith_min}

\begin{lstlisting}[language=C]
int Arith_max(int x, int y) {
    return (x > y) ? x : y;
}

int Arith_min(int x, int y) {
    return (x > y) ? y : x;
}
\end{lstlisting}

\subsubsection{3.2 Floor-style integer division}\label{floor-style-integer-division}

The intent is to provide stable division semantics independent of whether the host C implementation truncates negative integer division toward zero or toward negative infinity.

\begin{lstlisting}[language=C]
bool division_truncates_toward_zero() {
    return (-13 / 5) == -2;
}

bool signs_differ(int x, int y) {
    return (x < 0) != (y < 0);
}

int Arith_div(int x, int y) {
    assert(y != 0); // conceptual; book treats y=0 as unchecked like C division

    int q = x / y;
    int r = x % y;

    if (division_truncates_toward_zero() && signs_differ(x, y) && r != 0)
        return q - 1;
    else
        return q;
}
\end{lstlisting}

\subsubsection{3.3 Matched modulo}\label{matched-modulo}

\begin{lstlisting}[language=C]
int Arith_mod(int x, int y) {
    assert(y != 0);

    int r = x % y;
    if (division_truncates_toward_zero() && signs_differ(x, y) && r != 0)
        return r + y;
    else
        return r;
}
\end{lstlisting}

\subsubsection{3.4 Floor and ceiling}\label{floor-and-ceiling}

\begin{lstlisting}[language=C]
int Arith_floor(int x, int y) {
    return Arith_div(x, y);
}

int Arith_ceiling(int x, int y) {
    assert(y != 0);
    return Arith_div(x, y) + ((x % y) != 0);
}
\end{lstlisting}

\subsubsection{3.5 Stack ADT}\label{stack-adt}

\textbf{Representation invariant:} \passthrough{\lstinline!Stack.count!} equals the number of nodes reachable from \passthrough{\lstinline!Stack.head!}; the top of stack is \passthrough{\lstinline!head!}.

\begin{lstlisting}[language=C]
typedef struct StackNode {
    void *value;
    StackNode *next;
} StackNode;

typedef struct Stack {
    int count;
    StackNode *head;
} *Stack_T;

Stack_T Stack_new(void) {
    Stack_T s = alloc(sizeof *s);
    s->count = 0;
    s->head = NULL;
    return s;
}

bool Stack_empty(Stack_T s) {
    assert(s != NULL);
    return s->count == 0;
}

void Stack_push(Stack_T s, void *x) {
    assert(s != NULL);

    StackNode *n = alloc(sizeof *n);
    n->value = x;
    n->next = s->head;
    s->head = n;
    s->count++;
}

void *Stack_pop(Stack_T s) {
    assert(s != NULL);
    assert(s->count > 0);

    StackNode *n = s->head;
    void *x = n->value;
    s->head = n->next;
    s->count--;
    free(n);
    return x;
}

void Stack_free(Stack_T *sp) {
    assert(sp != NULL && *sp != NULL);

    StackNode *n = (*sp)->head;
    while (n != NULL) {
        StackNode *next = n->next;
        free(n);
        n = next;
    }
    free(*sp);
    *sp = NULL;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. Atom Algorithms}\label{atom-algorithms}

An atom is an interned immutable byte sequence. Equal sequences map to the same stored address, so equality can be tested by pointer comparison after interning.

\subsubsection{4.1 Atom table representation}\label{atom-table-representation}

\begin{lstlisting}[language=C]
typedef struct AtomEntry {
    AtomEntry *next;
    int len;
    char *bytes;      // null-terminated copy, even for arbitrary byte sequences
} AtomEntry;

AtomEntry *buckets[ATOM_BUCKET_COUNT];
unsigned long scatter[256]; // byte-to-random-number table
\end{lstlisting}

\subsubsection{4.2 Hash a byte sequence}\label{hash-a-byte-sequence}

\begin{lstlisting}[language=C]
unsigned long Atom_hash(const char *s, int len) {
    unsigned long h = 0;
    for (int i = 0; i < len; i++)
        h = (h << 1) + scatter[(unsigned char)s[i]];
    return h;
}
\end{lstlisting}

\subsubsection{4.3 Intern a byte sequence}\label{intern-a-byte-sequence}

\begin{lstlisting}[language=C]
const char *Atom_new(const char *s, int len) {
    assert(s != NULL);
    assert(len >= 0);

    unsigned long h = Atom_hash(s, len) % ATOM_BUCKET_COUNT;

    for (AtomEntry *p = buckets[h]; p != NULL; p = p->next) {
        if (p->len == len && bytes_equal(p->bytes, s, len))
            return p->bytes;
    }

    AtomEntry *p = alloc(sizeof *p + len + 1);
    p->len = len;
    p->bytes = storage_after_record(p);
    copy(p->bytes, s, len);
    p->bytes[len] = '\0';

    p->next = buckets[h];
    buckets[h] = p;
    return p->bytes;
}
\end{lstlisting}

\subsubsection{4.4 Intern a null-terminated string}\label{intern-a-null-terminated-string}

\begin{lstlisting}[language=C]
const char *Atom_string(const char *s) {
    assert(s != NULL);
    return Atom_new(s, strlen(s));
}
\end{lstlisting}

\subsubsection{4.5 Convert integer to atom}\label{convert-integer-to-atom}

\begin{lstlisting}[language=C]
const char *Atom_int(long n) {
    char buf[MAX_LONG_DECIMAL_CHARS];
    char *p = end_of(buf);
    unsigned long magnitude;

    if (n == LONG_MIN)
        magnitude = (unsigned long)LONG_MAX + 1;
    else if (n < 0)
        magnitude = (unsigned long)(-n);
    else
        magnitude = (unsigned long)n;

    do {
        *--p = (char)('0' + (magnitude % 10));
        magnitude /= 10;
    } while (magnitude > 0);

    if (n < 0)
        *--p = '-';

    return Atom_new(p, end_of(buf) - p);
}
\end{lstlisting}

\subsubsection{4.6 Find atom length}\label{find-atom-length}

\begin{lstlisting}[language=C]
int Atom_length(const char *atom) {
    assert(atom != NULL);

    for each bucket in buckets
        for (AtomEntry *p = bucket; p != NULL; p = p->next)
            if (p->bytes == atom)
                return p->len;

    assert(false); // pointer was not an atom
    return 0;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. Exception and Assertion Algorithms}\label{exception-and-assertion-algorithms}

\subsubsection{5.1 Exception frame stack}\label{exception-frame-stack}

\begin{lstlisting}[language=C]
typedef struct Except_Frame {
    Except_Frame *prev;
    JumpBuffer env;
    const char *file;
    int line;
    const Except_T *exception;
} Except_Frame;

Except_Frame *Except_stack = NULL;
\end{lstlisting}

\subsubsection{5.2 Push and pop exception frames}\label{push-and-pop-exception-frames}

\begin{lstlisting}[language=C]
void push_frame(Except_Frame *f) {
    f->prev = Except_stack;
    Except_stack = f;
}

void pop_frame(void) {
    Except_stack = Except_stack->prev;
}
\end{lstlisting}

\subsubsection{5.3 Raise an exception}\label{raise-an-exception}

\begin{lstlisting}[language=C]
void Except_raise(const Except_T *e, const char *file, int line) {
    assert(e != NULL);

    Except_Frame *f = Except_stack;
    if (f == NULL) {
        print_uncaught_exception(e, file, line);
        abort_program();
    }

    f->exception = e;
    f->file = file;
    f->line = line;
    pop_frame();
    long_jump(f->env, EXCEPT_RAISED);
}
\end{lstlisting}

\subsubsection{5.4 TRY / EXCEPT control-flow algorithm}\label{try-except-control-flow-algorithm}

\begin{lstlisting}[language=C]
TRY {
    Except_Frame f;
    int state;

    push_frame(&f);
    state = set_jump(f.env);

    if (state == EXCEPT_ENTERED) {
        run_try_body();
        if (state == EXCEPT_ENTERED)
            pop_frame();
    } else if (f.exception == &SomeException) {
        state = EXCEPT_HANDLED;
        run_matching_handler();
    } else {
        state = EXCEPT_HANDLED;
        run_else_handler_or_reraise();
    }

    if (state == EXCEPT_RAISED)
        Except_raise(f.exception, f.file, f.line);
}
\end{lstlisting}

\subsubsection{5.5 TRY / FINALLY control-flow algorithm}\label{try-finally-control-flow-algorithm}

\begin{lstlisting}[language=C]
TRY_FINALLY {
    Except_Frame f;
    int state;

    push_frame(&f);
    state = set_jump(f.env);

    if (state == EXCEPT_ENTERED) {
        run_try_body();
        if (state == EXCEPT_ENTERED)
            pop_frame();
    }

    run_finally_body();

    if (state == EXCEPT_RAISED)
        Except_raise(f.exception, f.file, f.line);
}
\end{lstlisting}

\subsubsection{5.6 Assertions}\label{assertions}

\begin{lstlisting}[language=C]
void assert_algorithm(bool condition) {
    if (!condition)
        raise(Assert_Failed, current_file(), current_line());
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Memory Management Algorithms}\label{memory-management-algorithms}

\subsection{\texorpdfstring{6.1 Production \texttt{Mem} wrapper algorithms}{6.1 Production Mem wrapper algorithms}}\label{production-mem-wrapper-algorithms}

The production implementation deliberately delegates raw memory management to the C runtime, while enforcing better contracts.

\begin{lstlisting}[language=C]
void *Mem_alloc(long nbytes, const char *file, int line) {
    assert(nbytes > 0);
    void *p = malloc(nbytes);
    if (p == NULL)
        raise(Mem_Failed, file, line);
    return p;
}

void *Mem_calloc(long count, long nbytes, const char *file, int line) {
    assert(count > 0);
    assert(nbytes > 0);
    void *p = calloc(count, nbytes);
    if (p == NULL)
        raise(Mem_Failed, file, line);
    return p;
}

void Mem_free(void *p, const char *file, int line) {
    if (p != NULL)
        free(p);
}

void *Mem_resize(void *p, long nbytes, const char *file, int line) {
    assert(p != NULL);
    assert(nbytes > 0);
    void *q = realloc(p, nbytes);
    if (q == NULL)
        raise(Mem_Failed, file, line);
    return q;
}
\end{lstlisting}

\subsection{6.2 Checking allocator model}\label{checking-allocator-model}

The checking allocator maintains a descriptor table. Every address ever returned has a descriptor. Free blocks remain known so double frees and foreign pointers can be detected.

\begin{lstlisting}[language=C]
typedef struct Descriptor {
    Descriptor *free_next;    // non-null when block is free
    Descriptor *hash_next;    // next descriptor in bucket
    const void *ptr;
    long size;
    const char *file;
    int line;
} Descriptor;

Descriptor *htab[DESCRIPTOR_BUCKETS];
Descriptor freelist_sentinel;
\end{lstlisting}

\subsubsection{6.2.1 Hash and find descriptor}\label{hash-and-find-descriptor}

\begin{lstlisting}[language=C]
unsigned descriptor_hash(const void *p) {
    return (((uintptr_t)p) >> ALIGN_SHIFT) & (DESCRIPTOR_BUCKETS - 1);
}

Descriptor *find_descriptor(const void *p) {
    unsigned h = descriptor_hash(p);
    for (Descriptor *d = htab[h]; d != NULL; d = d->hash_next)
        if (d->ptr == p)
            return d;
    return NULL;
}
\end{lstlisting}

\subsubsection{6.2.2 Validate allocated pointer}\label{validate-allocated-pointer}

\begin{lstlisting}[language=C]
Descriptor *require_live_block(void *p, const char *file, int line) {
    if (!is_aligned_for_any_type(p))
        raise(Assert_Failed, file, line);

    Descriptor *d = find_descriptor(p);
    if (d == NULL || d->free_next != NULL)
        raise(Assert_Failed, file, line);

    return d;
}
\end{lstlisting}

\subsubsection{6.2.3 Allocate descriptor records}\label{allocate-descriptor-records}

\begin{lstlisting}[language=C]
Descriptor *descriptor_alloc(void *ptr, long size, const char *file, int line) {
    static Descriptor *available = NULL;
    static int remaining = 0;

    if (remaining == 0) {
        available = malloc(DESCRIPTORS_PER_CHUNK * sizeof *available);
        if (available == NULL)
            return NULL;
        remaining = DESCRIPTORS_PER_CHUNK;
    }

    Descriptor *d = available++;
    remaining--;

    d->ptr = ptr;
    d->size = size;
    d->file = file;
    d->line = line;
    d->free_next = NULL;
    d->hash_next = NULL;
    return d;
}
\end{lstlisting}

\subsubsection{6.2.4 First-fit checked allocation}\label{first-fit-checked-allocation}

\begin{lstlisting}[language=C]
void *Mem_alloc_checked(long nbytes, const char *file, int line) {
    assert(nbytes > 0);
    nbytes = round_up_to_alignment(nbytes);

    while (true) {
        Descriptor *prev = &freelist_sentinel;
        for (Descriptor *d = prev->free_next; d != &freelist_sentinel; d = d->free_next) {
            if (d->size > nbytes) {
                // allocate from the high end of this free block
                char *base = (char *)d->ptr;
                char *result = base + d->size - nbytes;
                d->size -= nbytes;

                Descriptor *newd = descriptor_alloc(result, nbytes, file, line);
                if (newd == NULL)
                    raise(Mem_Failed, file, line);

                insert_descriptor_in_hash_table(newd);
                return result;
            }
        }

        // no existing free block: obtain a new large chunk and add it to free list
        long chunk_size = NALLOC + nbytes;
        void *chunk = malloc(chunk_size);
        if (chunk == NULL)
            raise(Mem_Failed, file, line);

        Descriptor *d = descriptor_alloc(chunk, chunk_size, file, line);
        if (d == NULL)
            raise(Mem_Failed, file, line);
        insert_descriptor_in_hash_table(d);
        add_descriptor_to_free_list(d);
    }
}
\end{lstlisting}

\subsubsection{6.2.5 Checked free}\label{checked-free}

\begin{lstlisting}[language=C]
void Mem_free_checked(void *p, const char *file, int line) {
    if (p == NULL)
        return;

    Descriptor *d = require_live_block(p, file, line);
    add_descriptor_to_free_list(d);
}
\end{lstlisting}

\subsubsection{6.2.6 Checked resize}\label{checked-resize}

\begin{lstlisting}[language=C]
void *Mem_resize_checked(void *p, long nbytes, const char *file, int line) {
    assert(p != NULL);
    assert(nbytes > 0);

    Descriptor *old = require_live_block(p, file, line);
    void *q = Mem_alloc_checked(nbytes, file, line);
    copy(q, p, min(nbytes, old->size));
    Mem_free_checked(p, file, line);
    return q;
}
\end{lstlisting}

\subsubsection{6.2.7 Checked calloc}\label{checked-calloc}

\begin{lstlisting}[language=C]
void *Mem_calloc_checked(long count, long nbytes, const char *file, int line) {
    assert(count > 0);
    assert(nbytes > 0);

    long total = count * nbytes;
    void *p = Mem_alloc_checked(total, file, line);
    zero(p, total);
    return p;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. Arena Allocation Algorithms}\label{arena-allocation-algorithms}

An arena is a collection of chunks. Individual objects are allocated linearly from the current chunk; freeing an arena frees all chunks at once.

\subsubsection{7.1 Representation}\label{representation}

\begin{lstlisting}[language=C]
typedef struct Chunk {
    Chunk *prev;
    char *avail;
    char *limit;
} Chunk;

typedef struct Arena {
    Chunk *current;
} *Arena_T;

Chunk *arena_free_chunks;
\end{lstlisting}

\subsubsection{7.2 Create arena}\label{create-arena}

\begin{lstlisting}[language=C]
Arena_T Arena_new(void) {
    Arena_T a = malloc(sizeof *a);
    if (a == NULL)
        raise(Arena_NewFailed, current_file(), current_line());
    a->current = NULL;
    return a;
}
\end{lstlisting}

\subsubsection{7.3 Allocate from arena}\label{allocate-from-arena}

\begin{lstlisting}[language=C]
void *Arena_alloc(Arena_T a, long nbytes, const char *file, int line) {
    assert(a != NULL);
    assert(nbytes > 0);

    nbytes = round_up_to_alignment(nbytes);

    if (a->current == NULL || bytes_available(a->current) < nbytes) {
        long payload = max(nbytes, ARENA_CHUNK_SIZE);
        Chunk *c = get_chunk_from_freelist_or_malloc(payload);
        if (c == NULL)
            raise(Arena_Failed, file, line);

        c->prev = a->current;
        c->avail = first_payload_byte(c);
        c->limit = c->avail + payload;
        a->current = c;
    }

    void *p = a->current->avail;
    a->current->avail += nbytes;
    return p;
}
\end{lstlisting}

\subsubsection{7.4 Arena calloc}\label{arena-calloc}

\begin{lstlisting}[language=C]
void *Arena_calloc(Arena_T a, long count, long nbytes, const char *file, int line) {
    assert(count > 0);
    assert(nbytes > 0);
    long total = count * nbytes;
    void *p = Arena_alloc(a, total, file, line);
    zero(p, total);
    return p;
}
\end{lstlisting}

\subsubsection{7.5 Free all allocations in an arena}\label{free-all-allocations-in-an-arena}

\begin{lstlisting}[language=C]
void Arena_free(Arena_T a) {
    assert(a != NULL);

    Chunk *c = a->current;
    while (c != NULL) {
        Chunk *prev = c->prev;
        recycle_or_release_chunk(c);
        c = prev;
    }
    a->current = NULL;
}
\end{lstlisting}

\subsubsection{7.6 Dispose arena object}\label{dispose-arena-object}

\begin{lstlisting}[language=C]
void Arena_dispose(Arena_T *ap) {
    assert(ap != NULL && *ap != NULL);
    Arena_free(*ap);
    free(*ap);
    *ap = NULL;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{8. Linked List Algorithms}\label{linked-list-algorithms}

A list is a simple singly linked chain of arbitrary pointers.

\begin{lstlisting}[language=C]
typedef struct ListNode {
    void *value;
    ListNode *next;
} *List_T;
\end{lstlisting}

\subsubsection{8.1 Push and pop}\label{push-and-pop}

\begin{lstlisting}[language=C]
List_T List_push(List_T list, void *x) {
    List_T n = alloc(sizeof *n);
    n->value = x;
    n->next = list;
    return n;
}

List_T List_pop(List_T list, void **x) {
    if (list == NULL)
        return NULL;

    List_T next = list->next;
    if (x != NULL)
        *x = list->value;
    free(list);
    return next;
}
\end{lstlisting}

\subsubsection{8.2 Append two lists}\label{append-two-lists}

\begin{lstlisting}[language=C]
List_T List_append(List_T list, List_T tail) {
    if (list == NULL)
        return tail;

    List_T p = list;
    while (p->next != NULL)
        p = p->next;
    p->next = tail;
    return list;
}
\end{lstlisting}

\subsubsection{8.3 Copy}\label{copy}

\begin{lstlisting}[language=C]
List_T List_copy(List_T list) {
    List_T head = NULL;
    List_T *last = &head;

    for (List_T p = list; p != NULL; p = p->next) {
        *last = alloc(sizeof **last);
        (*last)->value = p->value;
        (*last)->next = NULL;
        last = &(*last)->next;
    }
    return head;
}
\end{lstlisting}

\subsubsection{8.4 Reverse in place}\label{reverse-in-place}

\begin{lstlisting}[language=C]
List_T List_reverse(List_T list) {
    List_T prev = NULL;
    while (list != NULL) {
        List_T next = list->next;
        list->next = prev;
        prev = list;
        list = next;
    }
    return prev;
}
\end{lstlisting}

\subsubsection{8.5 Length, map, free, and to-array}\label{length-map-free-and-to-array}

\begin{lstlisting}[language=C]
int List_length(List_T list) {
    int n = 0;
    for (; list != NULL; list = list->next)
        n++;
    return n;
}

void List_map(List_T list, void apply(void **x, void *cl), void *cl) {
    assert(apply != NULL);
    for (; list != NULL; list = list->next)
        apply(&list->value, cl);
}

void List_free(List_T *listp) {
    assert(listp != NULL);
    while (*listp != NULL)
        *listp = List_pop(*listp, NULL);
}

void **List_toArray(List_T list, void *end) {
    int n = List_length(list);
    void **a = alloc((n + 1) * sizeof a[0]);
    int i = 0;
    for (List_T p = list; p != NULL; p = p->next)
        a[i++] = p->value;
    a[i] = end;
    return a;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{9. Table Algorithms}\label{table-algorithms}

A table maps keys to values using client-supplied equality and hash functions. Buckets are separate chains.

\subsubsection{9.1 Representation}\label{representation-1}

\begin{lstlisting}[language=C]
typedef struct Binding {
    Binding *next;
    const void *key;
    void *value;
} Binding;

typedef struct Table {
    int size;                 // number of buckets
    int length;               // number of bindings
    int (*cmp)(const void *, const void *);
    unsigned (*hash)(const void *);
    Binding **buckets;
} *Table_T;
\end{lstlisting}

\subsubsection{9.2 Create table}\label{create-table}

\begin{lstlisting}[language=C]
Table_T Table_new(int hint, cmp_fn cmp, hash_fn hash) {
    assert(hint >= 0);

    Table_T t = alloc(sizeof *t);
    t->size = choose_prime_bucket_count(hint);
    t->length = 0;
    t->cmp = cmp != NULL ? cmp : pointer_compare;
    t->hash = hash != NULL ? hash : pointer_hash;
    t->buckets = calloc(t->size, sizeof t->buckets[0]);
    return t;
}
\end{lstlisting}

\subsubsection{9.3 Search bucket}\label{search-bucket}

\begin{lstlisting}[language=C]
Binding **Table_find(Table_T t, const void *key) {
    assert(t != NULL);
    assert(key != NULL);

    unsigned h = t->hash(key) % t->size;
    Binding **pp = &t->buckets[h];
    while (*pp != NULL && t->cmp(key, (*pp)->key) != 0)
        pp = &(*pp)->next;
    return pp;
}
\end{lstlisting}

\subsubsection{9.4 Get, put, remove}\label{get-put-remove}

\begin{lstlisting}[language=C]
void *Table_get(Table_T t, const void *key) {
    Binding **pp = Table_find(t, key);
    return (*pp != NULL) ? (*pp)->value : NULL;
}

void *Table_put(Table_T t, const void *key, void *value) {
    Binding **pp = Table_find(t, key);

    if (*pp != NULL) {
        void *old = (*pp)->value;
        (*pp)->value = value;
        return old;
    }

    Binding *b = alloc(sizeof *b);
    b->key = key;
    b->value = value;
    b->next = *pp;
    *pp = b;
    t->length++;
    return NULL;
}

void *Table_remove(Table_T t, const void *key) {
    Binding **pp = Table_find(t, key);
    if (*pp == NULL)
        return NULL;

    Binding *victim = *pp;
    void *value = victim->value;
    *pp = victim->next;
    free(victim);
    t->length--;
    return value;
}
\end{lstlisting}

\subsubsection{9.5 Map, to-array, free}\label{map-to-array-free}

\begin{lstlisting}[language=C]
void Table_map(Table_T t, void apply(const void *key, void **value, void *cl), void *cl) {
    assert(t != NULL && apply != NULL);

    for (int i = 0; i < t->size; i++)
        for (Binding *b = t->buckets[i]; b != NULL; b = b->next)
            apply(b->key, &b->value, cl);
}

void **Table_toArray(Table_T t, void *end) {
    assert(t != NULL);
    void **a = alloc((2 * t->length + 1) * sizeof a[0]);
    int j = 0;

    for each binding b in t
        a[j++] = (void *)b->key, a[j++] = b->value;
    a[j] = end;
    return a;
}

void Table_free(Table_T *tp) {
    assert(tp != NULL && *tp != NULL);

    for each bucket in (*tp)->buckets
        free all Binding records in bucket;
    free((*tp)->buckets);
    free(*tp);
    *tp = NULL;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{10. Set Algorithms}\label{set-algorithms}

A set is a hash table containing members but no values.

\subsubsection{10.1 Representation}\label{representation-2}

\begin{lstlisting}[language=C]
typedef struct Member {
    Member *next;
    const void *value;
} Member;

typedef struct Set {
    int size;
    int length;
    cmp_fn cmp;
    hash_fn hash;
    Member **buckets;
} *Set_T;
\end{lstlisting}

\subsubsection{10.2 Membership and insertion}\label{membership-and-insertion}

\begin{lstlisting}[language=C]
Member **Set_find(Set_T s, const void *member) {
    assert(s != NULL);
    assert(member != NULL);

    unsigned h = s->hash(member) % s->size;
    Member **pp = &s->buckets[h];
    while (*pp != NULL && s->cmp(member, (*pp)->value) != 0)
        pp = &(*pp)->next;
    return pp;
}

int Set_member(Set_T s, const void *member) {
    return *Set_find(s, member) != NULL;
}

void Set_put(Set_T s, const void *member) {
    Member **pp = Set_find(s, member);
    if (*pp == NULL) {
        Member *m = alloc(sizeof *m);
        m->value = member;
        m->next = *pp;
        *pp = m;
        s->length++;
    }
}
\end{lstlisting}

\subsubsection{10.3 Remove}\label{remove}

\begin{lstlisting}[language=C]
void *Set_remove(Set_T s, const void *member) {
    Member **pp = Set_find(s, member);
    if (*pp == NULL)
        return NULL;

    Member *victim = *pp;
    void *value = (void *)victim->value;
    *pp = victim->next;
    free(victim);
    s->length--;
    return value;
}
\end{lstlisting}

\subsubsection{10.4 Set algebra}\label{set-algebra}

All binary set operations require compatible comparison and hash functions.

\begin{lstlisting}[language=C]
Set_T Set_union(Set_T a, Set_T b) {
    require_compatible(a, b);
    Set_T r = Set_new(a->length + b->length, a->cmp, a->hash);

    for each member x in a
        Set_put(r, x);
    for each member x in b
        Set_put(r, x);
    return r;
}

Set_T Set_inter(Set_T a, Set_T b) {
    require_compatible(a, b);
    Set_T r = Set_new(min(a->length, b->length), a->cmp, a->hash);

    Set_T small = a->length <= b->length ? a : b;
    Set_T large = small == a ? b : a;
    for each member x in small
        if (Set_member(large, x))
            Set_put(r, x);
    return r;
}

Set_T Set_minus(Set_T a, Set_T b) {
    require_compatible(a, b);
    Set_T r = Set_new(a->length, a->cmp, a->hash);

    for each member x in a
        if (!Set_member(b, x))
            Set_put(r, x);
    return r;
}

Set_T Set_diff(Set_T a, Set_T b) {
    // symmetric difference: elements in exactly one set
    require_compatible(a, b);
    Set_T r = Set_new(a->length + b->length, a->cmp, a->hash);

    for each member x in a
        if (!Set_member(b, x)) Set_put(r, x);
    for each member x in b
        if (!Set_member(a, x)) Set_put(r, x);
    return r;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{11. Dynamic Array Algorithms}\label{dynamic-array-algorithms}

\subsubsection{11.1 Representation}\label{representation-3}

\begin{lstlisting}[language=C]
typedef struct Array {
    int length;   // number of elements
    int size;     // bytes per element
    char *array;  // raw storage
} *Array_T;
\end{lstlisting}

\subsubsection{11.2 Initialize representation}\label{initialize-representation}

\begin{lstlisting}[language=C]
void ArrayRep_init(Array_T a, int length, int size, void *storage) {
    assert(a != NULL);
    assert(length >= 0);
    assert(size > 0);

    a->length = length;
    a->size = size;
    a->array = storage;
}
\end{lstlisting}

\subsubsection{11.3 New, access, update}\label{new-access-update}

\begin{lstlisting}[language=C]
Array_T Array_new(int length, int size) {
    assert(length >= 0);
    assert(size > 0);

    Array_T a = alloc(sizeof *a);
    a->length = length;
    a->size = size;
    a->array = (length > 0) ? calloc(length, size) : NULL;
    return a;
}

void *Array_get(Array_T a, int i) {
    assert(a != NULL);
    assert(0 <= i && i < a->length);
    return a->array + i * a->size;
}

void *Array_put(Array_T a, int i, void *elem) {
    assert(a != NULL && elem != NULL);
    assert(0 <= i && i < a->length);
    copy(a->array + i * a->size, elem, a->size);
    return elem;
}
\end{lstlisting}

\subsubsection{11.4 Resize, copy, free}\label{resize-copy-free}

\begin{lstlisting}[language=C]
void Array_resize(Array_T a, int new_length) {
    assert(a != NULL);
    assert(new_length >= 0);

    long old_bytes = a->length * a->size;
    long new_bytes = new_length * a->size;

    if (new_length == 0) {
        free(a->array);
        a->array = NULL;
    } else {
        a->array = resize(a->array, new_bytes);
        if (new_bytes > old_bytes)
            zero(a->array + old_bytes, new_bytes - old_bytes);
    }
    a->length = new_length;
}

Array_T Array_copy(Array_T a, int new_length) {
    assert(a != NULL);
    assert(new_length >= 0);

    Array_T b = Array_new(new_length, a->size);
    copy(b->array, a->array, min(a->length, new_length) * a->size);
    return b;
}

void Array_free(Array_T *ap) {
    assert(ap != NULL && *ap != NULL);
    free((*ap)->array);
    free(*ap);
    *ap = NULL;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{12. Sequence Algorithms}\label{sequence-algorithms}

A sequence is a deque implemented with a circular buffer.

\subsubsection{12.1 Representation}\label{representation-4}

\begin{lstlisting}[language=C]
typedef struct Seq {
    int length;
    int capacity;
    int head;       // physical index of logical element 0
    void **array;
} *Seq_T;

int phys(Seq_T s, int logical_index) {
    return (s->head + logical_index) % s->capacity;
}
\end{lstlisting}

\subsubsection{12.2 Grow circular buffer}\label{grow-circular-buffer}

\begin{lstlisting}[language=C]
void Seq_expand_if_full(Seq_T s) {
    if (s->length < s->capacity)
        return;

    int newcap = max(2 * s->capacity, MIN_SEQ_CAPACITY);
    void **b = alloc(newcap * sizeof b[0]);

    for (int i = 0; i < s->length; i++)
        b[i] = s->array[phys(s, i)];

    free(s->array);
    s->array = b;
    s->capacity = newcap;
    s->head = 0;
}
\end{lstlisting}

\subsubsection{12.3 Add/remove at high and low ends}\label{addremove-at-high-and-low-ends}

\begin{lstlisting}[language=C]
void *Seq_addhi(Seq_T s, void *x) {
    assert(s != NULL);
    Seq_expand_if_full(s);
    s->array[phys(s, s->length)] = x;
    s->length++;
    return x;
}

void *Seq_addlo(Seq_T s, void *x) {
    assert(s != NULL);
    Seq_expand_if_full(s);
    s->head = (s->head - 1 + s->capacity) % s->capacity;
    s->array[s->head] = x;
    s->length++;
    return x;
}

void *Seq_remhi(Seq_T s) {
    assert(s != NULL);
    assert(s->length > 0);
    int i = phys(s, s->length - 1);
    void *x = s->array[i];
    s->length--;
    return x;
}

void *Seq_remlo(Seq_T s) {
    assert(s != NULL);
    assert(s->length > 0);
    void *x = s->array[s->head];
    s->head = (s->head + 1) % s->capacity;
    s->length--;
    return x;
}
\end{lstlisting}

\subsubsection{12.4 Indexed get/put}\label{indexed-getput}

\begin{lstlisting}[language=C]
void *Seq_get(Seq_T s, int i) {
    assert(s != NULL);
    assert(0 <= i && i < s->length);
    return s->array[phys(s, i)];
}

void *Seq_put(Seq_T s, int i, void *x) {
    assert(s != NULL);
    assert(0 <= i && i < s->length);
    void **slot = &s->array[phys(s, i)];
    void *old = *slot;
    *slot = x;
    return old;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{13. Ring Algorithms}\label{ring-algorithms}

A ring is a circular doubly linked list with a sentinel or head pointer. Rotating changes the logical head without moving nodes.

\subsubsection{13.1 Representation}\label{representation-5}

\begin{lstlisting}[language=C]
typedef struct RingNode {
    RingNode *left;
    RingNode *right;
    void *value;
} RingNode;

typedef struct Ring {
    int length;
    RingNode *head;
} *Ring_T;
\end{lstlisting}

\subsubsection{13.2 Locate node by shortest traversal}\label{locate-node-by-shortest-traversal}

\begin{lstlisting}[language=C]
RingNode *Ring_node(Ring_T r, int i) {
    assert(r != NULL);
    assert(0 <= i && i < r->length);

    if (i <= r->length / 2) {
        RingNode *p = r->head;
        for (int k = 0; k < i; k++) p = p->right;
        return p;
    } else {
        RingNode *p = r->head;
        for (int k = r->length; k > i; k--) p = p->left;
        return p;
    }
}
\end{lstlisting}

\subsubsection{13.3 Add and remove nodes}\label{add-and-remove-nodes}

\begin{lstlisting}[language=C]
void *Ring_add(Ring_T r, int pos, void *x) {
    assert(r != NULL);
    assert(-r->length <= pos && pos <= r->length + 1);

    RingNode *n = alloc(sizeof *n);
    n->value = x;

    if (r->length == 0) {
        n->left = n->right = n;
        r->head = n;
    } else {
        RingNode *anchor = choose_insertion_anchor(r, pos);
        link_n_before_or_after_anchor(n, anchor, pos);
        if (pos <= 0)
            r->head = n;
    }
    r->length++;
    return x;
}

void *Ring_remove(Ring_T r, int i) {
    assert(r != NULL);
    assert(r->length > 0);
    RingNode *n = Ring_node(r, normalized_index(r, i));
    void *x = n->value;

    if (r->length == 1) {
        r->head = NULL;
    } else {
        n->left->right = n->right;
        n->right->left = n->left;
        if (n == r->head)
            r->head = n->right;
    }
    r->length--;
    free(n);
    return x;
}
\end{lstlisting}

\subsubsection{13.4 Rotation}\label{rotation}

\begin{lstlisting}[language=C]
void Ring_rotate(Ring_T r, int n) {
    assert(r != NULL);
    if (r->length == 0)
        return;

    n = n % r->length;
    if (n < 0) n += r->length;

    for (int i = 0; i < n; i++)
        r->head = r->head->right;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{14. Bit Vector Algorithms}\label{bit-vector-algorithms}

\subsubsection{14.1 Representation and addressing}\label{representation-and-addressing}

\begin{lstlisting}[language=C]
typedef unsigned long Word;

typedef struct Bit {
    int length;
    int nwords;
    Word *words;
} *Bit_T;

int word_index(int bit) { return bit / BITS_PER_WORD; }
int bit_offset(int bit) { return bit % BITS_PER_WORD; }
Word bit_mask(int bit) { return (Word)1 << bit_offset(bit); }
\end{lstlisting}

\subsubsection{14.2 Get and put one bit}\label{get-and-put-one-bit}

\begin{lstlisting}[language=C]
int Bit_get(Bit_T b, int n) {
    assert(b != NULL);
    assert(0 <= n && n < b->length);
    return (b->words[word_index(n)] & bit_mask(n)) != 0;
}

int Bit_put(Bit_T b, int n, int value) {
    assert(b != NULL);
    assert(0 <= n && n < b->length);
    assert(value == 0 || value == 1);

    Word *w = &b->words[word_index(n)];
    Word m = bit_mask(n);
    int old = ((*w & m) != 0);
    if (value) *w |= m;
    else       *w &= ~m;
    return old;
}
\end{lstlisting}

\subsubsection{14.3 Range modification}\label{range-modification}

\begin{lstlisting}[language=C]
void Bit_set(Bit_T b, int lo, int hi) {
    assert_valid_range(b, lo, hi);
    for (int i = lo; i <= hi; i++)
        b->words[word_index(i)] |= bit_mask(i);
}

void Bit_clear(Bit_T b, int lo, int hi) {
    assert_valid_range(b, lo, hi);
    for (int i = lo; i <= hi; i++)
        b->words[word_index(i)] &= ~bit_mask(i);
}

void Bit_not(Bit_T b, int lo, int hi) {
    assert_valid_range(b, lo, hi);
    for (int i = lo; i <= hi; i++)
        b->words[word_index(i)] ^= bit_mask(i);
}
\end{lstlisting}

Optimized implementations compute partial-word masks for the first and last words and full-word masks for interior words.

\subsubsection{14.4 Count bits}\label{count-bits}

\begin{lstlisting}[language=C]
int Bit_count(Bit_T b) {
    assert(b != NULL);
    int count = 0;
    for (int i = 0; i < b->length; i++)
        count += Bit_get(b, i);
    return count;
}
\end{lstlisting}

\subsubsection{14.5 Bitwise set algebra}\label{bitwise-set-algebra}

\begin{lstlisting}[language=C]
Bit_T Bit_union(Bit_T a, Bit_T b) {
    require_same_length(a, b);
    Bit_T r = Bit_new(a->length);
    for (int i = 0; i < a->nwords; i++)
        r->words[i] = a->words[i] | b->words[i];
    mask_unused_tail_bits(r);
    return r;
}

Bit_T Bit_inter(Bit_T a, Bit_T b) {
    require_same_length(a, b);
    Bit_T r = Bit_new(a->length);
    for (int i = 0; i < a->nwords; i++)
        r->words[i] = a->words[i] & b->words[i];
    return r;
}

Bit_T Bit_minus(Bit_T a, Bit_T b) {
    require_same_length(a, b);
    Bit_T r = Bit_new(a->length);
    for (int i = 0; i < a->nwords; i++)
        r->words[i] = a->words[i] & ~b->words[i];
    mask_unused_tail_bits(r);
    return r;
}

Bit_T Bit_diff(Bit_T a, Bit_T b) {
    require_same_length(a, b);
    Bit_T r = Bit_new(a->length);
    for (int i = 0; i < a->nwords; i++)
        r->words[i] = a->words[i] ^ b->words[i];
    mask_unused_tail_bits(r);
    return r;
}
\end{lstlisting}

\subsubsection{14.6 Equality and subset tests}\label{equality-and-subset-tests}

\begin{lstlisting}[language=C]
bool Bit_eq(Bit_T a, Bit_T b) {
    require_same_length(a, b);
    for (int i = 0; i < a->nwords; i++)
        if (a->words[i] != b->words[i])
            return false;
    return true;
}

bool Bit_leq(Bit_T a, Bit_T b) {
    require_same_length(a, b);
    for (int i = 0; i < a->nwords; i++)
        if ((a->words[i] & ~b->words[i]) != 0)
            return false;
    return true;
}

bool Bit_lt(Bit_T a, Bit_T b) {
    return Bit_leq(a, b) && !Bit_eq(a, b);
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{15. Formatting Algorithms}\label{formatting-algorithms}

The \passthrough{\lstinline!Fmt!} module is a small extensible \passthrough{\lstinline!printf!}-like engine. It maps conversion characters to conversion functions.

\subsubsection{15.1 Converter registry}\label{converter-registry}

\begin{lstlisting}[language=C]
typedef void (*Fmt_T)(int code, va_list *args,
                      int put(int c, void *cl), void *cl,
                      unsigned char flags[256], int width, int precision);

Fmt_T converters[256];

Fmt_T Fmt_register(int code, Fmt_T cvt) {
    assert(0 <= code && code <= 255);
    Fmt_T old = converters[code];
    converters[code] = cvt;
    return old;
}
\end{lstlisting}

\subsubsection{15.2 Format-string scanner}\label{format-string-scanner}

\begin{lstlisting}[language=C]
void Fmt_vfmt(int put(int c, void *cl), void *cl,
              const char *fmt, va_list ap) {
    assert(put != NULL && fmt != NULL);

    for (const char *p = fmt; *p != '\0'; p++) {
        if (*p != '%') {
            put(*p, cl);
            continue;
        }

        p++; // skip '%'
        if (*p == '%') {
            put('%', cl);
            continue;
        }

        unsigned char flags[256] = {0};
        int width = DEFAULT_WIDTH;
        int precision = DEFAULT_PRECISION;

        while (is_flag_char(*p))
            flags[(unsigned char)*p++] = 1;

        if (*p == '*') {
            width = va_arg(ap, int);
            p++;
        } else {
            width = scan_unsigned_decimal(&p, DEFAULT_WIDTH);
        }

        if (*p == '.') {
            p++;
            if (*p == '*') {
                precision = va_arg(ap, int);
                p++;
            } else {
                precision = scan_unsigned_decimal(&p, 0);
            }
        }

        int code = (unsigned char)*p;
        Fmt_T cvt = converters[code];
        assert(cvt != NULL);
        cvt(code, &ap, put, cl, flags, width, precision);
    }
}
\end{lstlisting}

\subsubsection{15.3 Emit string with padding and precision}\label{emit-string-with-padding-and-precision}

\begin{lstlisting}[language=C]
void Fmt_puts(const char *s, int len, int put(int, void *), void *cl,
              unsigned char flags[256], int width, int precision) {
    assert(s != NULL && len >= 0 && flags != NULL);

    if (precision >= 0 && precision < len)
        len = precision;

    int pad = max(0, width - len);
    if (!flags['-'])
        repeat_put(' ', pad, put, cl);

    for (int i = 0; i < len; i++)
        put(s[i], cl);

    if (flags['-'])
        repeat_put(' ', pad, put, cl);
}
\end{lstlisting}

\subsubsection{15.4 Emit number with sign, zero padding, and width}\label{emit-number-with-sign-zero-padding-and-width}

\begin{lstlisting}[language=C]
void Fmt_putd(const char *digits, int len, int put(int, void *), void *cl,
              unsigned char flags[256], int width, int precision) {
    assert(digits != NULL && flags != NULL);

    char sign = extract_optional_sign(&digits, &len);
    int zeroes = max(0, precision - len);
    int total = len + zeroes + (sign != 0);
    int blanks = max(0, width - total);

    if (!flags['-'] && !flags['0']) repeat_put(' ', blanks, put, cl);
    if (sign) put(sign, cl);
    if (!flags['-'] && flags['0']) repeat_put('0', blanks, put, cl);
    repeat_put('0', zeroes, put, cl);
    for (int i = 0; i < len; i++) put(digits[i], cl);
    if (flags['-']) repeat_put(' ', blanks, put, cl);
}
\end{lstlisting}

\subsubsection{15.5 Buffer writer with overflow detection}\label{buffer-writer-with-overflow-detection}

\begin{lstlisting}[language=C]
typedef struct BufferOut {
    char *buf;
    int size;
    int len;
} BufferOut;

int buffer_put(int c, void *cl) {
    BufferOut *b = cl;
    if (b->len >= b->size - 1)
        raise(Fmt_Overflow, current_file(), current_line());
    b->buf[b->len++] = (char)c;
    return c;
}

int Fmt_vsfmt(char *buf, int size, const char *fmt, va_list ap) {
    assert(buf != NULL && size > 0 && fmt != NULL);
    BufferOut b = { buf, size, 0 };
    Fmt_vfmt(buffer_put, &b, fmt, ap);
    buf[b.len] = '\0';
    return b.len;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{16. Low-Level String Algorithms (\texttt{Str})}{16. Low-Level String Algorithms (Str)}}\label{low-level-string-algorithms-str}

\passthrough{\lstinline!Str!} provides substring operations over ordinary null-terminated C strings, with normalized inclusive index ranges.

\subsubsection{16.1 Position normalization}\label{position-normalization}

\begin{lstlisting}[language=C]
int Str_pos(const char *s, int i) {
    assert(s != NULL);
    int n = strlen(s);

    if (i > 0) return i - 1;       // 1-based positive index
    if (i < 0) return n + i;       // negative index from end
    return n;                      // convention for 0
}

Range normalize_range(const char *s, int i, int j) {
    int lo = Str_pos(s, i);
    int hi = Str_pos(s, j);
    assert(0 <= lo && lo <= strlen(s));
    assert(0 <= hi && hi <= strlen(s));
    if (lo > hi) return empty_range_at(lo);
    return Range{lo, hi};
}
\end{lstlisting}

\subsubsection{16.2 Substring and duplication}\label{substring-and-duplication}

\begin{lstlisting}[language=C]
char *Str_sub(const char *s, int i, int j) {
    assert(s != NULL);
    Range r = normalize_range(s, i, j);
    int n = range_length(r);

    char *out = alloc(n + 1);
    copy(out, s + r.lo, n);
    out[n] = '\0';
    return out;
}

char *Str_dup(const char *s, int i, int j, int n) {
    assert(n >= 0);
    char *piece = Str_sub(s, i, j);
    int len = strlen(piece);
    char *out = alloc(n * len + 1);

    char *p = out;
    for (int k = 0; k < n; k++) {
        copy(p, piece, len);
        p += len;
    }
    *p = '\0';
    free(piece);
    return out;
}
\end{lstlisting}

\subsubsection{16.3 Character and set search}\label{character-and-set-search}

\begin{lstlisting}[language=C]
int Str_chr(const char *s, int i, int j, int c) {
    Range r = normalize_range(s, i, j);
    for (int k = r.lo; k <= r.hi; k++)
        if ((unsigned char)s[k] == c)
            return external_index(k);
    return 0;
}

int Str_rchr(const char *s, int i, int j, int c) {
    Range r = normalize_range(s, i, j);
    for (int k = r.hi; k >= r.lo; k--)
        if ((unsigned char)s[k] == c)
            return external_index(k);
    return 0;
}

int Str_upto(const char *s, int i, int j, const char *set) {
    Range r = normalize_range(s, i, j);
    for (int k = r.lo; k <= r.hi; k++)
        if (char_in_set(s[k], set))
            return external_index(k);
    return 0;
}

int Str_many(const char *s, int i, int j, const char *set) {
    Range r = normalize_range(s, i, j);
    int k = r.lo;
    while (k <= r.hi && char_in_set(s[k], set))
        k++;
    return external_index(k - 1);
}
\end{lstlisting}

\subsubsection{16.4 Match and find substring}\label{match-and-find-substring}

\begin{lstlisting}[language=C]
bool Str_match(const char *s, int i, int j, const char *pattern) {
    Range r = normalize_range(s, i, j);
    int n = strlen(pattern);
    return r.lo + n - 1 <= r.hi && bytes_equal(s + r.lo, pattern, n);
}

int Str_find(const char *s, int i, int j, const char *pattern) {
    Range r = normalize_range(s, i, j);
    int n = strlen(pattern);

    for (int k = r.lo; k + n - 1 <= r.hi; k++)
        if (bytes_equal(s + k, pattern, n))
            return external_index(k);
    return 0;
}

int Str_rfind(const char *s, int i, int j, const char *pattern) {
    Range r = normalize_range(s, i, j);
    int n = strlen(pattern);

    for (int k = r.hi - n + 1; k >= r.lo; k--)
        if (bytes_equal(s + k, pattern, n))
            return external_index(k);
    return 0;
}
\end{lstlisting}

\subsubsection{16.5 Concatenate, reverse, map}\label{concatenate-reverse-map}

\begin{lstlisting}[language=C]
char *Str_cat(const char *s1, int i1, int j1,
              const char *s2, int i2, int j2) {
    char *a = Str_sub(s1, i1, j1);
    char *b = Str_sub(s2, i2, j2);
    char *out = alloc(strlen(a) + strlen(b) + 1);
    copy_string(out, a);
    append_string(out, b);
    free(a); free(b);
    return out;
}

char *Str_reverse(const char *s, int i, int j) {
    Range r = normalize_range(s, i, j);
    int n = range_length(r);
    char *out = alloc(n + 1);
    for (int k = 0; k < n; k++)
        out[k] = s[r.hi - k];
    out[n] = '\0';
    return out;
}

char *Str_map(const char *s, int i, int j, const char *from, const char *to) {
    Range r = normalize_range(s, i, j);
    int n = range_length(r);
    char *out = alloc(n + 1);

    for (int k = 0; k < n; k++) {
        char ch = s[r.lo + k];
        int p = index_in_string(from, ch);
        out[k] = (p >= 0) ? to[p] : ch;
    }
    out[n] = '\0';
    return out;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{17. High-Level Text Algorithms (\texttt{Text})}{17. High-Level Text Algorithms (Text)}}\label{high-level-text-algorithms-text}

\passthrough{\lstinline!Text\_T!} values carry length explicitly, so they can represent substrings without relying on null termination.

\subsubsection{17.1 Representation and creation}\label{representation-and-creation}

\begin{lstlisting}[language=C]
typedef struct Text {
    int len;
    const char *str;
} Text_T;

Text_T Text_box(const char *s, int len) {
    assert(s != NULL);
    assert(len >= 0);
    return Text_T{ len, s };
}

Text_T Text_put(const char *s) {
    assert(s != NULL);
    int n = strlen(s);
    char *copy = alloc(n + 1);
    copy_string(copy, s);
    return Text_T{ n, copy };
}
\end{lstlisting}

\subsubsection{17.2 Text concatenation and substring}\label{text-concatenation-and-substring}

\begin{lstlisting}[language=C]
Text_T Text_cat(Text_T a, Text_T b) {
    char *s = alloc(a.len + b.len + 1);
    copy(s, a.str, a.len);
    copy(s + a.len, b.str, b.len);
    s[a.len + b.len] = '\0';
    return Text_T{ a.len + b.len, s };
}

Text_T Text_sub(Text_T t, int i, int j) {
    Range r = normalize_text_range(t, i, j);
    return Text_T{ range_length(r), t.str + r.lo };
}
\end{lstlisting}

\subsubsection{17.3 Text copy-out}\label{text-copy-out}

\begin{lstlisting}[language=C]
char *Text_get(char *dst, int size, Text_T t) {
    assert(size >= 0);
    if (dst == NULL)
        dst = alloc(t.len + 1);
    else
        assert(size > t.len);

    copy(dst, t.str, t.len);
    dst[t.len] = '\0';
    return dst;
}
\end{lstlisting}

\subsubsection{17.4 Search/match operations}\label{searchmatch-operations}

\begin{lstlisting}[language=C]
bool Text_match(Text_T s, int i, int j, Text_T pat) {
    Range r = normalize_text_range(s, i, j);
    return r.lo + pat.len - 1 <= r.hi && bytes_equal(s.str + r.lo, pat.str, pat.len);
}

int Text_find(Text_T s, int i, int j, Text_T pat) {
    Range r = normalize_text_range(s, i, j);
    for (int k = r.lo; k + pat.len - 1 <= r.hi; k++)
        if (bytes_equal(s.str + k, pat.str, pat.len))
            return external_index(k);
    return 0;
}
\end{lstlisting}

\subsubsection{17.5 Save and restore allocated text pool}\label{save-and-restore-allocated-text-pool}

The high-level string implementation uses a storage discipline that can save a point and later release all text allocated after that point.

\begin{lstlisting}[language=C]
typedef struct Text_save {
    Arena_T arena_snapshot;
    AllocationMarker marker;
} *Text_save_T;

Text_save_T Text_save(void) {
    Text_save_T s = alloc(sizeof *s);
    s->marker = current_text_allocation_marker();
    return s;
}

void Text_restore(Text_save_T *sp) {
    assert(sp != NULL && *sp != NULL);
    release_text_allocations_after((*sp)->marker);
    free(*sp);
    *sp = NULL;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{18. Extended Precision Arithmetic (\texttt{XP})}{18. Extended Precision Arithmetic (XP)}}\label{extended-precision-arithmetic-xp}

\passthrough{\lstinline!XP!} operates on unsigned little-endian byte vectors. Each digit is normally one byte, so carries and borrows propagate base 256.

\subsubsection{18.1 Compare and length}\label{compare-and-length}

\begin{lstlisting}[language=C]
int XP_length(int n, unsigned char x[]) {
    while (n > 0 && x[n - 1] == 0)
        n--;
    return n;
}

int XP_cmp(int n, unsigned char x[], unsigned char y[]) {
    for (int i = n - 1; i >= 0; i--) {
        if (x[i] < y[i]) return -1;
        if (x[i] > y[i]) return 1;
    }
    return 0;
}
\end{lstlisting}

\subsubsection{18.2 Addition and subtraction}\label{addition-and-subtraction}

\begin{lstlisting}[language=C]
int XP_add(int n, byte z[], byte x[], byte y[], int carry) {
    for (int i = 0; i < n; i++) {
        int sum = x[i] + y[i] + carry;
        z[i] = sum & 0xFF;
        carry = sum >> 8;
    }
    return carry;
}

int XP_sub(int n, byte z[], byte x[], byte y[], int borrow) {
    for (int i = 0; i < n; i++) {
        int diff = x[i] - y[i] - borrow;
        if (diff < 0) {
            diff += 256;
            borrow = 1;
        } else {
            borrow = 0;
        }
        z[i] = diff;
    }
    return borrow;
}
\end{lstlisting}

\subsubsection{18.3 Single-digit operations}\label{single-digit-operations}

\begin{lstlisting}[language=C]
int XP_sum(int n, byte z[], byte x[], int y) {
    int carry = y;
    for (int i = 0; i < n; i++) {
        int sum = x[i] + carry;
        z[i] = sum & 0xFF;
        carry = sum >> 8;
    }
    return carry;
}

int XP_diff(int n, byte z[], byte x[], int y) {
    int borrow = y;
    for (int i = 0; i < n; i++) {
        int diff = x[i] - borrow;
        if (diff < 0) { diff += 256; borrow = 1; }
        else          { borrow = 0; }
        z[i] = diff;
    }
    return borrow;
}

int XP_product(int n, byte z[], byte x[], int y) {
    int carry = 0;
    for (int i = 0; i < n; i++) {
        int prod = x[i] * y + carry;
        z[i] = prod & 0xFF;
        carry = prod >> 8;
    }
    return carry;
}

int XP_quotient(int n, byte z[], byte x[], int y) {
    int rem = 0;
    for (int i = n - 1; i >= 0; i--) {
        int num = rem * 256 + x[i];
        z[i] = num / y;
        rem = num % y;
    }
    return rem;
}
\end{lstlisting}

\subsubsection{18.4 Multiplication}\label{multiplication}

\begin{lstlisting}[language=C]
int XP_mul(byte z[], int n, byte x[], int m, byte y[]) {
    zero(z, n + m);

    for (int i = 0; i < n; i++) {
        int carry = 0;
        for (int j = 0; j < m; j++) {
            int t = z[i + j] + x[i] * y[j] + carry;
            z[i + j] = t & 0xFF;
            carry = t >> 8;
        }
        z[i + m] += carry;
    }
    return overflow_carry_if_any(z, n + m);
}
\end{lstlisting}

\subsubsection{18.5 Division by multi-byte divisor}\label{division-by-multi-byte-divisor}

\begin{lstlisting}[language=C]
int XP_div(int n, byte q[], byte x[], int m, byte y[], byte r[], byte tmp[]) {
    assert(m > 0);
    assert(!is_zero(y, m));

    zero(q, n);
    copy(r, x, n);

    // Long division over base-256 digits. The implementation normalizes the
    // divisor, estimates quotient digits, corrects overestimates, subtracts,
    // and unnormalizes the remainder.
    normalize(y, r, &shift);
    for (int k = n - m; k >= 0; k--) {
        int qhat = estimate_quotient_digit(r, y, k, m);
        multiply_shifted(tmp, y, m, qhat, k);
        while (XP_cmp(n, tmp, r) > 0) {
            qhat--;
            subtract_shifted(tmp, y, m, k);
        }
        XP_sub(n, r, r, tmp, 0);
        q[k] = qhat;
    }
    unnormalize(r, shift);
    return 0;
}
\end{lstlisting}

\subsubsection{18.6 Bit shifts}\label{bit-shifts}

\begin{lstlisting}[language=C]
void XP_lshift(int n, byte z[], int m, byte x[], int s, int fill) {
    int byte_shift = s / 8;
    int bit_shift = s % 8;
    zero(z, n);

    for (int i = m - 1; i >= 0; i--) {
        int j = i + byte_shift;
        if (j < n)
            z[j] |= x[i] << bit_shift;
        if (bit_shift != 0 && j + 1 < n)
            z[j + 1] |= x[i] >> (8 - bit_shift);
    }
    apply_fill_to_low_bits_if_requested(z, s, fill);
}

void XP_rshift(int n, byte z[], int m, byte x[], int s, int fill) {
    int byte_shift = s / 8;
    int bit_shift = s % 8;
    zero(z, n);

    for (int i = 0; i < m; i++) {
        int j = i - byte_shift;
        if (j >= 0)
            z[j] |= x[i] >> bit_shift;
        if (bit_shift != 0 && j - 1 >= 0)
            z[j - 1] |= x[i] << (8 - bit_shift);
    }
    apply_fill_to_high_bits_if_requested(z, n, fill);
}
\end{lstlisting}

\subsubsection{18.7 String conversion}\label{string-conversion}

\begin{lstlisting}[language=C]
int XP_fromstr(int n, byte z[], const char *s, int base, char **end) {
    zero(z, n);
    const char *p = s;
    int any = 0;

    while (digit_value(*p) is valid in base) {
        int d = digit_value(*p++);
        int carry = XP_product(n, z, z, base);
        carry += XP_sum(n, z, z, d);
        if (carry != 0) break;
        any = 1;
    }
    if (end != NULL) *end = (char *)(any ? p : s);
    return any;
}

char *XP_tostr(char *str, int size, int base, int n, byte x[]) {
    byte tmp[MAX_N];
    copy(tmp, x, n);
    char reversed[MAX_OUTPUT];
    int k = 0;

    do {
        int rem = XP_quotient(n, tmp, tmp, base);
        reversed[k++] = digit_char(rem);
    } while (!is_zero(tmp, n));

    assert(k + 1 <= size);
    reverse_copy(str, reversed, k);
    str[k] = '\0';
    return str;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{19. Arbitrary Precision Arithmetic (\texttt{AP})}{19. Arbitrary Precision Arithmetic (AP)}}\label{arbitrary-precision-arithmetic-ap}

\passthrough{\lstinline!AP!} wraps XP digit vectors with sign and dynamic length.

\subsubsection{19.1 Representation}\label{representation-6}

\begin{lstlisting}[language=C]
typedef struct AP {
    int sign;       // -1, 0, +1
    int ndigits;
    byte *digits;   // magnitude, little-endian base 256
} *AP_T;
\end{lstlisting}

\subsubsection{19.2 Normalize signed integer}\label{normalize-signed-integer}

\begin{lstlisting}[language=C]
void AP_normalize(AP_T x) {
    x->ndigits = XP_length(x->ndigits, x->digits);
    if (x->ndigits == 0)
        x->sign = 0;
}
\end{lstlisting}

\subsubsection{19.3 Create from long}\label{create-from-long}

\begin{lstlisting}[language=C]
AP_T AP_new(long n) {
    AP_T z = alloc(sizeof *z);
    z->sign = (n > 0) - (n < 0);
    z->digits = allocate_enough_for_long();

    unsigned long mag = absolute_value_as_unsigned(n);
    z->ndigits = fill_digits_from_unsigned(z->digits, mag);
    AP_normalize(z);
    return z;
}
\end{lstlisting}

\subsubsection{19.4 Signed addition and subtraction}\label{signed-addition-and-subtraction}

\begin{lstlisting}[language=C]
AP_T AP_add(AP_T x, AP_T y) {
    if (x->sign == 0) return AP_copy(y);
    if (y->sign == 0) return AP_copy(x);

    AP_T z = allocate_result(max(x->ndigits, y->ndigits) + 1);

    if (x->sign == y->sign) {
        z->sign = x->sign;
        XP_add(z->capacity, z->digits, x->digits, y->digits, 0);
    } else {
        int cmp = compare_magnitudes(x, y);
        if (cmp == 0) return AP_new(0);
        if (cmp > 0) {
            z->sign = x->sign;
            XP_sub(z->capacity, z->digits, x->digits, y->digits, 0);
        } else {
            z->sign = y->sign;
            XP_sub(z->capacity, z->digits, y->digits, x->digits, 0);
        }
    }

    AP_normalize(z);
    return z;
}

AP_T AP_sub(AP_T x, AP_T y) {
    AP_T neg_y = AP_neg(y);
    AP_T z = AP_add(x, neg_y);
    AP_free(&neg_y);
    return z;
}
\end{lstlisting}

\subsubsection{19.5 Multiplication}\label{multiplication-1}

\begin{lstlisting}[language=C]
AP_T AP_mul(AP_T x, AP_T y) {
    if (x->sign == 0 || y->sign == 0)
        return AP_new(0);

    AP_T z = allocate_result(x->ndigits + y->ndigits);
    z->sign = x->sign * y->sign;
    XP_mul(z->digits, x->ndigits, x->digits, y->ndigits, y->digits);
    AP_normalize(z);
    return z;
}
\end{lstlisting}

\subsubsection{19.6 Division and modulo}\label{division-and-modulo}

\begin{lstlisting}[language=C]
Pair AP_divmod(AP_T x, AP_T y) {
    assert(y->sign != 0);

    AP_T q = allocate_result(max(1, x->ndigits - y->ndigits + 1));
    AP_T r = allocate_result(y->ndigits);

    XP_div(x->ndigits, q->digits,
           x->digits,
           y->ndigits, y->digits,
           r->digits, temp_buffer());

    q->sign = (is_zero(q)) ? 0 : x->sign * y->sign;
    r->sign = (is_zero(r)) ? 0 : x->sign;

    AP_normalize(q);
    AP_normalize(r);

    // Adjust to floor-division/mod semantics if signs differ and remainder nonzero.
    if (r->sign != 0 && x->sign != y->sign) {
        q = AP_subi_in_place(q, 1);
        r = AP_add_in_place(r, y);
    }

    return Pair{q, r};
}

AP_T AP_div(AP_T x, AP_T y) { return AP_divmod(x, y).quotient; }
AP_T AP_mod(AP_T x, AP_T y) { return AP_divmod(x, y).remainder; }
\end{lstlisting}

\subsubsection{19.7 Exponentiation with optional modulus}\label{exponentiation-with-optional-modulus}

\begin{lstlisting}[language=C]
AP_T AP_pow(AP_T x, AP_T y, AP_T modulus) {
    assert(y->sign >= 0);
    if (modulus != NULL)
        assert(AP_cmpi(modulus, 2) >= 0);

    AP_T result = AP_new(1);
    AP_T base = (modulus != NULL) ? AP_mod(x, modulus) : AP_copy(x);
    AP_T exp = AP_copy(y);

    while (AP_cmpi(exp, 0) > 0) {
        if (least_significant_bit(exp) == 1) {
            result = AP_mul(result, base);
            if (modulus != NULL) result = AP_mod(result, modulus);
        }
        exp = AP_rshift(exp, 1);
        if (AP_cmpi(exp, 0) > 0) {
            base = AP_mul(base, base);
            if (modulus != NULL) base = AP_mod(base, modulus);
        }
    }
    return result;
}
\end{lstlisting}

\subsubsection{19.8 Parse and format}\label{parse-and-format}

\begin{lstlisting}[language=C]
AP_T AP_fromstr(const char *s, int base, char **end) {
    assert(s != NULL);
    assert(2 <= base && base <= 36);

    skip_spaces(&s);
    int sign = scan_optional_sign(&s);
    AP_T z = AP_new(0);
    bool any = false;

    while (digit_value(*s) is valid in base) {
        z = AP_muli(z, base);
        z = AP_addi(z, digit_value(*s));
        s++;
        any = true;
    }

    if (!any) {
        if (end) *end = original_s;
        AP_free(&z);
        return NULL;
    }
    z->sign *= sign;
    if (end) *end = (char *)s;
    return z;
}

char *AP_tostr(char *str, int size, int base, AP_T x) {
    assert(2 <= base && base <= 36);

    AP_T tmp = AP_abs(x);
    char reversed[MAX_OR_DYNAMIC];
    int k = 0;

    do {
        long rem = AP_modi(tmp, base);
        reversed[k++] = digit_char(rem);
        tmp = AP_divi(tmp, base);
    } while (AP_cmpi(tmp, 0) != 0);

    if (x->sign < 0)
        reversed[k++] = '-';

    if (str == NULL)
        str = alloc(k + 1);
    else
        assert(size > k);

    reverse_copy(str, reversed, k);
    str[k] = '\0';
    return str;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{\texorpdfstring{20. Multiple Precision Arithmetic (\texttt{MP})}{20. Multiple Precision Arithmetic (MP)}}\label{multiple-precision-arithmetic-mp}

\passthrough{\lstinline!MP!} exposes fixed-width multiple precision arithmetic over client-supplied digit arrays. It provides unsigned operations and signed two's-complement-style operations.

\subsubsection{20.1 Fixed-width conversion}\label{fixed-width-conversion}

\begin{lstlisting}[language=C]
MP_T MP_fromintu(int n, MP_T z, unsigned long u) {
    XP_fromint(n, z, u);
    return z;
}

unsigned long MP_tointu(int n, MP_T x) {
    return XP_toint(n, x);
}

MP_T MP_cvt(int n, MP_T z, MP_T x) {
    // signed conversion from source width to destination width
    int fill = sign_bit_of(x) ? 0xFF : 0x00;
    copy_or_sign_extend(z, n, x, source_length(x), fill);
    if (value_changed_by_truncation(z, x))
        raise(MP_Overflow, current_file(), current_line());
    return z;
}
\end{lstlisting}

\subsubsection{20.2 Unsigned add/subtract and signed add/subtract}\label{unsigned-addsubtract-and-signed-addsubtract}

\begin{lstlisting}[language=C]
MP_T MP_addu(int n, MP_T z, MP_T x, MP_T y) {
    int carry = XP_add(n, z, x, y, 0);
    if (carry != 0)
        raise(MP_Overflow, current_file(), current_line());
    return z;
}

MP_T MP_subu(int n, MP_T z, MP_T x, MP_T y) {
    int borrow = XP_sub(n, z, x, y, 0);
    if (borrow != 0)
        raise(MP_Overflow, current_file(), current_line());
    return z;
}

MP_T MP_add(int n, MP_T z, MP_T x, MP_T y) {
    int sx = sign_bit_of(x), sy = sign_bit_of(y);
    XP_add(n, z, x, y, 0);
    int sz = sign_bit_of(z);
    if (sx == sy && sz != sx)
        raise(MP_Overflow, current_file(), current_line());
    return z;
}

MP_T MP_sub(int n, MP_T z, MP_T x, MP_T y) {
    int sx = sign_bit_of(x), sy = sign_bit_of(y);
    XP_sub(n, z, x, y, 0);
    int sz = sign_bit_of(z);
    if (sx != sy && sz != sx)
        raise(MP_Overflow, current_file(), current_line());
    return z;
}
\end{lstlisting}

\subsubsection{20.3 Bitwise operations}\label{bitwise-operations}

\begin{lstlisting}[language=C]
MP_T MP_and(int n, MP_T z, MP_T x, MP_T y) {
    for (int i = 0; i < n; i++) z[i] = x[i] & y[i];
    return z;
}

MP_T MP_or(int n, MP_T z, MP_T x, MP_T y) {
    for (int i = 0; i < n; i++) z[i] = x[i] | y[i];
    return z;
}

MP_T MP_xor(int n, MP_T z, MP_T x, MP_T y) {
    for (int i = 0; i < n; i++) z[i] = x[i] ^ y[i];
    return z;
}

MP_T MP_not(int n, MP_T z, MP_T x) {
    for (int i = 0; i < n; i++) z[i] = ~x[i];
    return z;
}
\end{lstlisting}

\subsubsection{20.4 Shifts}\label{shifts}

\begin{lstlisting}[language=C]
MP_T MP_lshift(int n, MP_T z, MP_T x, int s) {
    assert(s >= 0);
    XP_lshift(n, z, n, x, s, 0);
    return z;
}

MP_T MP_rshiftu(int n, MP_T z, MP_T x, int s) {
    assert(s >= 0);
    XP_rshift(n, z, n, x, s, 0);
    return z;
}

MP_T MP_rshift(int n, MP_T z, MP_T x, int s) {
    assert(s >= 0);
    int fill = sign_bit_of(x) ? 1 : 0;
    XP_rshift(n, z, n, x, s, fill);
    return z;
}
\end{lstlisting}

\subsubsection{20.5 Multiplication and division}\label{multiplication-and-division}

\begin{lstlisting}[language=C]
MP_T MP_mulu(int n, MP_T z, MP_T x, MP_T y) {
    byte tmp[2 * n];
    XP_mul(tmp, n, x, n, y);
    if (high_half_nonzero(tmp, n))
        raise(MP_Overflow, current_file(), current_line());
    copy(z, tmp, n);
    return z;
}

MP_T MP_divu(int n, MP_T z, MP_T x, MP_T y) {
    if (is_zero(y, n))
        raise(MP_DivideByZero, current_file(), current_line());
    byte r[n], tmp[temporary_division_size(n)];
    XP_div(n, z, x, n, y, r, tmp);
    return z;
}

MP_T MP_div(int n, MP_T z, MP_T x, MP_T y) {
    if (is_zero(y, n))
        raise(MP_DivideByZero, current_file(), current_line());

    int sx = sign_bit_of(x), sy = sign_bit_of(y);
    byte ax[n], ay[n];
    abs_twos_complement(ax, x, n);
    abs_twos_complement(ay, y, n);

    MP_divu(n, z, ax, ay);
    if (sx != sy)
        twos_complement_in_place(z, n);
    return z;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{21. Thread, Semaphore, and Channel Algorithms}\label{thread-semaphore-and-channel-algorithms}

The thread subsystem is a user-level cooperative/preemptive thread package with ready queues, blocking queues, semaphores, and synchronous channels.

\subsubsection{21.1 Thread representation}\label{thread-representation}

\begin{lstlisting}[language=C]
typedef enum { RUNNING, READY, BLOCKED, DEAD } ThreadState;

typedef struct Thread {
    Context ctx;
    ThreadState state;
    int exit_code;
    int alerted;
    Queue joiners;
    Except_Frame *exception_stack;
    void *stack;
    Thread *next;
} *Thread_T;

Queue readyq;
Thread_T current;
int nthreads;
\end{lstlisting}

\subsubsection{21.2 Initialize thread system}\label{initialize-thread-system}

\begin{lstlisting}[language=C]
int Thread_init(int preempt, ...) {
    if (already_initialized)
        return 0;

    current = allocate_main_thread_record();
    current->state = RUNNING;
    readyq = Queue_new();
    nthreads = 1;

    if (preempt)
        install_timer_interrupt_for_periodic_pause();

    already_initialized = true;
    return 1;
}
\end{lstlisting}

\subsubsection{21.3 Create a thread}\label{create-a-thread}

\begin{lstlisting}[language=C]
Thread_T Thread_new(int apply(void *), void *args, int nbytes, ...) {
    assert(apply != NULL);
    assert(nbytes >= 0);

    Thread_T t = alloc(sizeof *t);
    t->stack = allocate_thread_stack(optional_stack_size());
    if (t->stack == NULL)
        raise(Thread_Failed, current_file(), current_line());

    void *argcopy = NULL;
    if (nbytes > 0) {
        argcopy = alloc(nbytes);
        copy(argcopy, args, nbytes);
    } else {
        argcopy = args;
    }

    initialize_context_to_start_at(t->ctx, thread_start_trampoline,
                                   apply, argcopy, t->stack);
    t->state = READY;
    t->alerted = 0;
    t->joiners = Queue_new();
    t->exception_stack = NULL;

    enqueue(readyq, t);
    nthreads++;
    return t;
}

void thread_start_trampoline(int apply(void *), void *arg) {
    int code = apply(arg);
    Thread_exit(code);
}
\end{lstlisting}

\subsubsection{21.4 Scheduler and pause}\label{scheduler-and-pause}

\begin{lstlisting}[language=C]
void Thread_pause(void) {
    disable_interrupts();

    Thread_T old = current;
    if (old->state == RUNNING) {
        old->state = READY;
        enqueue(readyq, old);
    }

    Thread_T next = dequeue(readyq);
    if (next == NULL)
        terminate_process_or_wait_for_event();

    next->state = RUNNING;
    current = next;
    context_switch(&old->ctx, &next->ctx);

    enable_interrupts();
}
\end{lstlisting}

\subsubsection{21.5 Blocking and releasing}\label{blocking-and-releasing}

\begin{lstlisting}[language=C]
void block_current_on(Queue q) {
    current->state = BLOCKED;
    enqueue(q, current);
    Thread_pause();
}

void release_one(Queue q) {
    Thread_T t = dequeue(q);
    if (t != NULL) {
        t->state = READY;
        enqueue(readyq, t);
    }
}
\end{lstlisting}

\subsubsection{21.6 Alert, join, and exit}\label{alert-join-and-exit}

\begin{lstlisting}[language=C]
void Thread_alert(Thread_T t) {
    assert(t != NULL);
    t->alerted = 1;
    if (t->state == BLOCKED)
        move_from_blocked_queue_to_readyq(t);
}

int Thread_join(Thread_T t) {
    assert(t != NULL);
    if (current->alerted)
        raise(Thread_Alerted, current_file(), current_line());

    while (t->state != DEAD)
        block_current_on(t->joiners);

    return t->exit_code;
}

void Thread_exit(int code) {
    disable_interrupts();
    current->exit_code = code;
    current->state = DEAD;

    while (!Queue_empty(current->joiners))
        release_one(current->joiners);

    nthreads--;
    Thread_pause();
    unreachable();
}
\end{lstlisting}

\subsubsection{21.7 Semaphore algorithms}\label{semaphore-algorithms}

\begin{lstlisting}[language=C]
typedef struct Sem {
    int count;
    Queue waiting;
} Sem_T;

void Sem_init(Sem_T *s, int count) {
    assert(s != NULL);
    assert(count >= 0);
    s->count = count;
    s->waiting = Queue_new();
}

void Sem_wait(Sem_T *s) {
    assert(s != NULL);
    disable_interrupts();

    if (current->alerted) {
        enable_interrupts();
        raise(Thread_Alerted, current_file(), current_line());
    }

    while (s->count == 0)
        block_current_on(s->waiting);

    s->count--;
    enable_interrupts();
}

void Sem_signal(Sem_T *s) {
    assert(s != NULL);
    disable_interrupts();
    s->count++;
    release_one(s->waiting);
    enable_interrupts();
}
\end{lstlisting}

\subsubsection{21.8 Channel rendezvous algorithms}\label{channel-rendezvous-algorithms}

Channels synchronize exactly one sender with one receiver and copy bytes directly from sender buffer to receiver buffer.

\begin{lstlisting}[language=C]
typedef struct ChanWaiter {
    Thread_T thread;
    void *ptr;
    int size;
    int transferred;
} ChanWaiter;

typedef struct Chan {
    Queue senders;
    Queue receivers;
} *Chan_T;

Chan_T Chan_new(void) {
    Chan_T c = alloc(sizeof *c);
    c->senders = Queue_new();
    c->receivers = Queue_new();
    return c;
}
\end{lstlisting}

\begin{lstlisting}[language=C]
int Chan_send(Chan_T c, const void *ptr, int size) {
    assert(c != NULL && ptr != NULL && size >= 0);
    disable_interrupts();

    ChanWaiter *r = dequeue(c->receivers);
    if (r != NULL) {
        int n = min(size, r->size);
        copy(r->ptr, ptr, n);
        r->transferred = n;
        release_thread(r->thread);
        enable_interrupts();
        return n;
    }

    ChanWaiter self = { current, (void *)ptr, size, 0 };
    enqueue(c->senders, &self);
    block_current_on(c->senders);

    if (current->alerted) {
        enable_interrupts();
        raise(Thread_Alerted, current_file(), current_line());
    }

    enable_interrupts();
    return self.transferred;
}
\end{lstlisting}

\begin{lstlisting}[language=C]
int Chan_receive(Chan_T c, void *ptr, int size) {
    assert(c != NULL && ptr != NULL && size >= 0);
    disable_interrupts();

    ChanWaiter *s = dequeue(c->senders);
    if (s != NULL) {
        int n = min(size, s->size);
        copy(ptr, s->ptr, n);
        s->transferred = n;
        release_thread(s->thread);
        enable_interrupts();
        return n;
    }

    ChanWaiter self = { current, ptr, size, 0 };
    enqueue(c->receivers, &self);
    block_current_on(c->receivers);

    if (current->alerted) {
        enable_interrupts();
        raise(Thread_Alerted, current_file(), current_line());
    }

    enable_interrupts();
    return self.transferred;
}
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{22. Cross-Cutting Design Algorithms}\label{cross-cutting-design-algorithms}

\subsubsection{22.1 Opaque pointer discipline}\label{opaque-pointer-discipline}

\begin{lstlisting}[language=C]
// Interface file
struct Hidden;
typedef struct Hidden *T;

// Implementation file
struct Hidden {
    representation_fields;
};
\end{lstlisting}

\textbf{Algorithmic effect:} clients can pass handles but cannot inspect or depend on representation fields.

\subsubsection{22.2 Checked runtime error discipline}\label{checked-runtime-error-discipline}

\begin{lstlisting}[language=C]
void public_function(T object, int index) {
    assert(object != NULL);
    assert(0 <= index && index < object->length);
    implementation_body();
}
\end{lstlisting}

\subsubsection{22.3 Allocation macro discipline}\label{allocation-macro-discipline}

\begin{lstlisting}[language=C]
#define NEW(p)  ((p) = ALLOC(sizeof *(p)))
#define NEW0(p) ((p) = CALLOC(1, sizeof *(p)))
#define FREE(p) (Mem_free((p), file,line), (p) = NULL)
\end{lstlisting}

\textbf{Algorithmic effect:} size follows the pointed-to type rather than a manually repeated type name; freed variables are nulled to reduce dangling-pointer errors.

\subsubsection{22.4 Map callback pattern}\label{map-callback-pattern}

\begin{lstlisting}[language=C]
void ADT_map(T object, void apply(Element *e, void *cl), void *cl) {
    assert(object != NULL && apply != NULL);
    for each element slot in object
        apply(&slot, cl);
}
\end{lstlisting}

\textbf{Algorithmic effect:} the ADT owns traversal order and representation; clients supply behavior.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{23. Reimplementation Checklist}\label{reimplementation-checklist}

Use this checklist when converting the pseudocode into a real C library:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Preserve every interface contract separately from the implementation.
\item
  Decide which runtime errors are checked and which are unchecked.
\item
  Use opaque pointers for ADTs unless representation exposure is intentional.
\item
  Keep allocation ownership explicit in every function.
\item
  For hash-based structures, require compatible \passthrough{\lstinline!cmp!} and \passthrough{\lstinline!hash!} functions for binary operations.
\item
  For circular buffers and rings, centralize index normalization.
\item
  For bit vectors, mask unused bits in the final word after complementing or union-like operations.
\item
  For numeric code, separate magnitude algorithms from signed semantics.
\item
  For thread code, keep scheduling, interrupt masking, and queue transitions atomic.
\item
  Keep source-coordinate propagation for allocation and assertion failures.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{24. Compressed Algorithm Map by Interface Function}\label{compressed-algorithm-map-by-interface-function}

This section maps the public surface to the algorithmic pattern used.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Interface
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Functions
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Algorithmic pattern
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\passthrough{\lstinline!Arith!} & \passthrough{\lstinline!max!}, \passthrough{\lstinline!min!}, \passthrough{\lstinline!div!}, \passthrough{\lstinline!mod!}, \passthrough{\lstinline!floor!}, \passthrough{\lstinline!ceiling!} & branch selection; corrected floor division \\
\passthrough{\lstinline!Stack!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!empty!}, \passthrough{\lstinline!push!}, \passthrough{\lstinline!pop!}, \passthrough{\lstinline!free!} & singly linked LIFO \\
\passthrough{\lstinline!Atom!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!string!}, \passthrough{\lstinline!int!}, \passthrough{\lstinline!length!} & hash-consing; byte hashing; table scan for length \\
\passthrough{\lstinline!Except!} & \passthrough{\lstinline!raise!}, \passthrough{\lstinline!RAISE!}, \passthrough{\lstinline!RERAISE!}, \passthrough{\lstinline!TRY!}, \passthrough{\lstinline!EXCEPT!}, \passthrough{\lstinline!FINALLY!}, \passthrough{\lstinline!RETURN!} & frame stack plus nonlocal jump \\
\passthrough{\lstinline!Mem!} & \passthrough{\lstinline!alloc!}, \passthrough{\lstinline!calloc!}, \passthrough{\lstinline!resize!}, \passthrough{\lstinline!free!} & checked allocation wrapper or debug descriptor allocator \\
\passthrough{\lstinline!Arena!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!alloc!}, \passthrough{\lstinline!calloc!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!dispose!} & bump allocation in chunks; whole-arena release \\
\passthrough{\lstinline!List!} & \passthrough{\lstinline!append!}, \passthrough{\lstinline!copy!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!}, \passthrough{\lstinline!list!}, \passthrough{\lstinline!map!}, \passthrough{\lstinline!pop!}, \passthrough{\lstinline!push!}, \passthrough{\lstinline!reverse!}, \passthrough{\lstinline!toArray!} & singly linked traversal and node manipulation \\
\passthrough{\lstinline!Table!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!get!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!remove!}, \passthrough{\lstinline!map!}, \passthrough{\lstinline!toArray!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!} & separate chaining hash map \\
\passthrough{\lstinline!Set!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!member!}, \passthrough{\lstinline!remove!}, \passthrough{\lstinline!union!}, \passthrough{\lstinline!inter!}, \passthrough{\lstinline!minus!}, \passthrough{\lstinline!diff!}, \passthrough{\lstinline!map!}, \passthrough{\lstinline!toArray!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!} & separate chaining hash set plus set algebra \\
\passthrough{\lstinline!Array!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!get!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!resize!}, \passthrough{\lstinline!copy!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!}, \passthrough{\lstinline!size!} & contiguous fixed-size slots \\
\passthrough{\lstinline!Seq!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!addhi!}, \passthrough{\lstinline!addlo!}, \passthrough{\lstinline!remhi!}, \passthrough{\lstinline!remlo!}, \passthrough{\lstinline!get!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!} & circular dynamic buffer deque \\
\passthrough{\lstinline!Ring!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!ring!}, \passthrough{\lstinline!add!}, \passthrough{\lstinline!addhi!}, \passthrough{\lstinline!addlo!}, \passthrough{\lstinline!remove!}, \passthrough{\lstinline!remhi!}, \passthrough{\lstinline!remlo!}, \passthrough{\lstinline!get!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!rotate!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!} & circular doubly linked list \\
\passthrough{\lstinline!Bit!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!get!}, \passthrough{\lstinline!set!}, \passthrough{\lstinline!clear!}, \passthrough{\lstinline!not!}, \passthrough{\lstinline!count!}, \passthrough{\lstinline!union!}, \passthrough{\lstinline!inter!}, \passthrough{\lstinline!minus!}, \passthrough{\lstinline!diff!}, \passthrough{\lstinline!eq!}, \passthrough{\lstinline!leq!}, \passthrough{\lstinline!lt!}, \passthrough{\lstinline!map!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!length!} & packed word bit manipulation \\
\passthrough{\lstinline!Fmt!} & \passthrough{\lstinline!fmt!}, \passthrough{\lstinline!vfmt!}, \passthrough{\lstinline!sfmt!}, \passthrough{\lstinline!vsfmt!}, \passthrough{\lstinline!print!}, \passthrough{\lstinline!fprint!}, \passthrough{\lstinline!string!}, \passthrough{\lstinline!vstring!}, \passthrough{\lstinline!register!}, \passthrough{\lstinline!putd!}, \passthrough{\lstinline!puts!} & format parser plus converter dispatch \\
\passthrough{\lstinline!Str!} & \passthrough{\lstinline!sub!}, \passthrough{\lstinline!dup!}, \passthrough{\lstinline!cat!}, \passthrough{\lstinline!catv!}, \passthrough{\lstinline!reverse!}, \passthrough{\lstinline!map!}, \passthrough{\lstinline!find!}, \passthrough{\lstinline!rfind!}, \passthrough{\lstinline!match!}, \passthrough{\lstinline!rmatch!}, \passthrough{\lstinline!chr!}, \passthrough{\lstinline!rchr!}, \passthrough{\lstinline!upto!}, \passthrough{\lstinline!rupto!}, \passthrough{\lstinline!many!}, \passthrough{\lstinline!rmany!}, \passthrough{\lstinline!any!}, \passthrough{\lstinline!len!}, \passthrough{\lstinline!pos!}, \passthrough{\lstinline!fmt!} & normalized range operations over C strings \\
\passthrough{\lstinline!Text!} & \passthrough{\lstinline!box!}, \passthrough{\lstinline!put!}, \passthrough{\lstinline!get!}, \passthrough{\lstinline!sub!}, \passthrough{\lstinline!cat!}, \passthrough{\lstinline!dup!}, \passthrough{\lstinline!reverse!}, \passthrough{\lstinline!map!}, \passthrough{\lstinline!find!}, \passthrough{\lstinline!rfind!}, \passthrough{\lstinline!match!}, \passthrough{\lstinline!rmatch!}, \passthrough{\lstinline!chr!}, \passthrough{\lstinline!rchr!}, \passthrough{\lstinline!upto!}, \passthrough{\lstinline!rupto!}, \passthrough{\lstinline!many!}, \passthrough{\lstinline!rmany!}, \passthrough{\lstinline!any!}, \passthrough{\lstinline!cmp!}, \passthrough{\lstinline!save!}, \passthrough{\lstinline!restore!}, \passthrough{\lstinline!fmt!} & length-bearing string slices and allocated text pool \\
\passthrough{\lstinline!XP!} & \passthrough{\lstinline!add!}, \passthrough{\lstinline!sub!}, \passthrough{\lstinline!mul!}, \passthrough{\lstinline!div!}, \passthrough{\lstinline!sum!}, \passthrough{\lstinline!diff!}, \passthrough{\lstinline!product!}, \passthrough{\lstinline!quotient!}, \passthrough{\lstinline!cmp!}, \passthrough{\lstinline!neg!}, \passthrough{\lstinline!lshift!}, \passthrough{\lstinline!rshift!}, \passthrough{\lstinline!fromint!}, \passthrough{\lstinline!toint!}, \passthrough{\lstinline!fromstr!}, \passthrough{\lstinline!tostr!}, \passthrough{\lstinline!length!} & unsigned base-256 digit-vector arithmetic \\
\passthrough{\lstinline!AP!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!free!}, \passthrough{\lstinline!add!}, \passthrough{\lstinline!sub!}, \passthrough{\lstinline!mul!}, \passthrough{\lstinline!div!}, \passthrough{\lstinline!mod!}, \passthrough{\lstinline!neg!}, \passthrough{\lstinline!pow!}, \passthrough{\lstinline!cmp!}, \passthrough{\lstinline!lshift!}, \passthrough{\lstinline!rshift!}, \passthrough{\lstinline!fromstr!}, \passthrough{\lstinline!tostr!}, \passthrough{\lstinline!toint!}, \passthrough{\lstinline!fmt!} & signed dynamic big integers layered on XP \\
\passthrough{\lstinline!MP!} & arithmetic, bitwise, shifts, comparison, conversion, parsing, formatting & fixed-width multi-precision arithmetic with overflow checks \\
\passthrough{\lstinline!Thread!} & \passthrough{\lstinline!init!}, \passthrough{\lstinline!new!}, \passthrough{\lstinline!self!}, \passthrough{\lstinline!pause!}, \passthrough{\lstinline!join!}, \passthrough{\lstinline!alert!}, \passthrough{\lstinline!exit!} & user-level scheduling and context switching \\
\passthrough{\lstinline!Sem!} & \passthrough{\lstinline!init!}, \passthrough{\lstinline!new!}, \passthrough{\lstinline!wait!}, \passthrough{\lstinline!signal!} & counting semaphore over thread wait queues \\
\passthrough{\lstinline!Chan!} & \passthrough{\lstinline!new!}, \passthrough{\lstinline!send!}, \passthrough{\lstinline!receive!} & synchronous rendezvous and byte-copy transfer \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{25. Key Takeaways}\label{key-takeaways}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The book's algorithms are best understood as \textbf{interface-enforced data structure algorithms}, not isolated textbook routines.
\item
  Many modules use simple structures---linked lists, arrays, hash tables, circular buffers---but combine them with strict contracts.
\item
  The most reusable pattern is the separation of \textbf{public ADT handle}, \textbf{private representation}, \textbf{checked runtime errors}, and \textbf{client-owned callback functions}.
\item
  The numeric chapters form a layered architecture: \passthrough{\lstinline!XP!} handles unsigned digit vectors; \passthrough{\lstinline!AP!} adds dynamic allocation and sign; \passthrough{\lstinline!MP!} uses fixed-width buffers and checked overflow.
\item
  The systems chapters---exceptions, memory, arenas, and threads---show how C can emulate higher-level runtime services through disciplined interfaces.
\end{enumerate}


\end{document}
