
\documentclass[11pt]{article}

\usepackage[letterpaper,margin=0.72in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{bookmark}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{calc}
\usepackage{amsmath,amssymb}
\usepackage{textcomp}
\usepackage{upquote}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{etoolbox}

\definecolor{OOPBlue}{HTML}{1F4E79}
\definecolor{OOPLightBlue}{HTML}{EAF3F8}
\definecolor{OOPGold}{HTML}{B8860B}
\definecolor{OOPGoldLight}{HTML}{FFF8E6}
\definecolor{OOPGreen}{HTML}{2E6F40}
\definecolor{OOPPurple}{HTML}{5B4B8A}
\definecolor{OOPRule}{HTML}{B7C9D6}
\definecolor{OOPGrey}{HTML}{F7F7F7}
\definecolor{OOPDarkGrey}{HTML}{444444}

\hypersetup{
  colorlinks=true,
  linkcolor=OOPBlue,
  urlcolor=OOPBlue,
  citecolor=OOPBlue,
  pdftitle={Object-Oriented Programming with ANSI-C Algorithms as C-like Pseudocode},
  pdfauthor={Jordan Suber / ChatGPT},
  pdfsubject={Pseudocode handout based on Axel-Tobias Schreiner's Object-Oriented Programming with ANSI-C},
  pdfkeywords={ANSI C, object-oriented programming, pseudocode, algorithms, dynamic dispatch, metaclasses, ooc, callbacks, persistence, exceptions}
}

\pagestyle{fancy}
\fancyhf{}
\lhead{\textit{Object-Oriented Programming with ANSI-C} Pseudocode}
\rhead{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\headheight}{14pt}

\titleformat{\section}{\Large\bfseries\color{OOPBlue}}{}{0pt}{}
\titleformat{\subsection}{\large\bfseries\color{OOPBlue}}{}{0pt}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{OOPDarkGrey}}{}{0pt}{}
\titlespacing*{\section}{0pt}{1.35em}{0.5em}
\titlespacing*{\subsection}{0pt}{0.95em}{0.35em}
\titlespacing*{\subsubsection}{0pt}{0.75em}{0.25em}

\setcounter{secnumdepth}{0}
\setcounter{tocdepth}{2}
\setlist[itemize]{leftmargin=1.4em,itemsep=0.2em,topsep=0.2em}
\setlist[enumerate]{leftmargin=1.6em,itemsep=0.25em,topsep=0.25em}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.52em}
\setlength{\LTpre}{0.35em}
\setlength{\LTpost}{0.35em}
\emergencystretch=4em
\sloppy

\lstdefinestyle{oopcode}{
  language=C,
  basicstyle=\ttfamily\scriptsize,
  keywordstyle=\bfseries\color{OOPBlue},
  commentstyle=\itshape\color{OOPDarkGrey},
  stringstyle=\color{OOPGreen},
  numbers=left,
  numberstyle=\tiny\color{OOPDarkGrey},
  stepnumber=1,
  numbersep=0.7em,
  breaklines=true,
  breakatwhitespace=false,
  columns=fullflexible,
  keepspaces=true,
  tabsize=4,
  showstringspaces=false,
  frame=single,
  framerule=0.3pt,
  rulecolor=\color{OOPRule},
  backgroundcolor=\color{OOPGrey},
  xleftmargin=1.0em,
  framexleftmargin=0.8em,
  framexrightmargin=0.35em,
  framesep=0.42em,
  aboveskip=0.55em,
  belowskip=0.55em,
  morekeywords={function,procedure,method,class,type,struct,record,extends,implements,selector,metaclass,constructor,destructor,override,dispatch,super,self,NULL,true,false,bool,size_t,uintptr_t,va_list,va_arg,va_start,va_end,assert,raise,throw,catch,try,finally,repeat,until,foreach,continue,break,new,delete,calloc,malloc,realloc,free,memcpy,memmove,strcmp,strcpy,strlen,strtod,strspn,strncpy,setjmp,longjmp,scan,screen,install,exec,process,cast,isA,isOf,classOf,sizeOf,super,ctor,dtor,puto,geto,retrieve,respondsTo,forward,gate,wire,signal,wait,send,receive,enqueue,dequeue,push,pop,clone,differ,contains,drop,find,add}
}
\lstset{style=oopcode}

\newtcolorbox{scopebox}{
  enhanced,
  colback=OOPLightBlue,
  colframe=OOPBlue,
  boxrule=0.5pt,
  arc=2mm,
  left=1em,
  right=1em,
  top=0.7em,
  bottom=0.7em
}

\newtcolorbox{methodbox}{
  enhanced,
  colback=OOPGrey,
  colframe=OOPRule,
  boxrule=0.4pt,
  arc=2mm,
  left=0.9em,
  right=0.9em,
  top=0.6em,
  bottom=0.6em
}

\newtcolorbox{warningbox}{
  enhanced,
  colback=OOPGoldLight,
  colframe=OOPGold,
  boxrule=0.45pt,
  arc=2mm,
  left=0.9em,
  right=0.9em,
  top=0.6em,
  bottom=0.6em
}

% Pandoc compatibility helpers
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\providecommand{\passthrough}[1]{#1}
\AtBeginEnvironment{longtable}{\small}
\renewcommand{\arraystretch}{1.14}

\begin{document}
\hypersetup{pageanchor=false}
\pagenumbering{gobble}

\begin{titlepage}
  \centering
  \vspace*{1.4cm}
  {\Huge\bfseries\color{OOPBlue}Object-Oriented Programming with ANSI-C\par}
  \vspace{0.35cm}
  {\LARGE Algorithms as C-like Pseudocode\par}
  \vspace{0.25cm}
  {\large Based on Axel-Tobias Schreiner, \textit{Object-Oriented Programming with ANSI-C}\par}
  \vspace{0.95cm}
  \begin{scopebox}
  This handout reorganizes the book's ANSI-C examples into implementation-oriented pseudocode. It is designed for studying how object-oriented mechanisms can be built directly from C structures, function pointers, descriptors, selectors, metaclasses, generated boilerplate, callbacks, persistence, exceptions, and message forwarding.
  \end{scopebox}
  \vspace{0.75cm}
  \begin{methodbox}
  \textbf{Scope:} systematic extraction from the uploaded book into independent C-like pseudocode.\\
  \textbf{Primary mechanisms:} abstract data types, descriptor-based allocation, generic constructors/destructors, dynamic linkage, inheritance by structure prefixing, symbol tables, metaclass-driven method tables, the \texttt{ooc} preprocessor, defensive runtime type checks, static construction, delegates, class methods, persistence, exception handling, and GUI message forwarding.\\
  \textbf{Boundary:} routines are rewritten as study-grade pseudocode rather than copied source listings.
  \end{methodbox}
  \vfill
  {\large Prepared May 29, 2026\par}
\end{titlepage}

\hypersetup{pageanchor=true}
\pagenumbering{roman}
\tableofcontents
\clearpage
\pagenumbering{arabic}

\begin{warningbox}
\textbf{Implementation note.} The original text is a practical ANSI-C construction of object-oriented mechanisms. This handout keeps exported names where they clarify orientation, but condenses routine bodies into pseudocode. Treat the snippets as architectural algorithms, not drop-in source code.
\end{warningbox}

\textbf{Source analyzed:} Axel-Tobias Schreiner, \emph{Object-Oriented
Programming with ANSI-C} (1993), uploaded PDF.

\subsection{Scope and reconstruction
rules}\label{scope-and-reconstruction-rules}

This document extracts the algorithmic content from the book's C
examples and rewrites it as \textbf{independent C-like pseudocode}. It
is not a source-code clone. Names such as \passthrough{\lstinline!new!},
\passthrough{\lstinline!delete!}, \passthrough{\lstinline!Class!},
\passthrough{\lstinline!Object!}, \passthrough{\lstinline!ctor!},
\passthrough{\lstinline!dtor!}, \passthrough{\lstinline!cast!},
\passthrough{\lstinline!puto!}, \passthrough{\lstinline!geto!},
\passthrough{\lstinline!retrieve!},
\passthrough{\lstinline!respondsTo!}, \passthrough{\lstinline!gate!},
and \passthrough{\lstinline!wire!} are retained where they are part of
the book's architectural vocabulary.

The book is not a traditional algorithms textbook. Its ``algorithms''
are implementation patterns for building object-oriented mechanisms in
ANSI C: abstract data types, constructors/destructors, dynamic dispatch,
inheritance by structure prefixing, metaclasses, defensive type checks,
preprocessing, callbacks, persistence, exceptions, and message
forwarding.

\subsection{High-level algorithm
inventory}\label{high-level-algorithm-inventory}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Area
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Algorithmic material extracted
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Abstract data types & Fixed-heap set, dynamically allocated bag,
membership, add/drop/find, descriptor-based allocation \\
Dynamic linkage & \passthrough{\lstinline!Class!} descriptor tables,
constructors/destructors, selectors, polymorphic
\passthrough{\lstinline!differ!}, string methods, atom interning \\
Expression processing & Scanner, recursive-descent recognizer,
expression-tree construction, tree destruction, postfix output,
evaluation, infix output \\
Inheritance & Prefix-based subclass layout, up-casting, constructor
chaining, inherited static methods, overwritten dynamic methods \\
Symbol table & Sorted dynamic array with binary search/insertion,
\passthrough{\lstinline!screen!}, \passthrough{\lstinline!install!},
variable/function/constant nodes \\
Metaclass system & \passthrough{\lstinline!Object!} and
\passthrough{\lstinline!Class!} roots, metaclass constructors, binary
inheritance of method tables, selectors and superclass selectors \\
\passthrough{\lstinline!ooc!} preprocessor & Class-description parsing,
interface/representation/source generation, report expansion,
method-header generation \\
Dynamic checking & \passthrough{\lstinline!isA!},
\passthrough{\lstinline!isOf!}, \passthrough{\lstinline!cast!},
magic-number checks, defensive import rules \\
Static construction & Initializer-list generation by
\passthrough{\lstinline!munch!}, class-description functions, automatic
class initialization \\
Delegates and callbacks & \passthrough{\lstinline!respondsTo!}, callback
discovery by tags, \passthrough{\lstinline!Filter!} framework, option
parsing, line/file callbacks, sort delegate \\
Class methods & Class-dispatched \passthrough{\lstinline!new!},
class-specific allocation, node chain tracking, reclaiming leaked AST
nodes \\
Persistence & \passthrough{\lstinline!puto!},
\passthrough{\lstinline!geto!}, \passthrough{\lstinline!retrieve!},
class-name lookup, symbol attachment, shallow transfer by
\passthrough{\lstinline!move!} \\
Exceptions & Exception stack, \passthrough{\lstinline!catch!},
\passthrough{\lstinline!cause!}, nested handler cleanup \\
Message forwarding and GUI & Fallback selector dispatch,
\passthrough{\lstinline!forward!}, event propagation by
\passthrough{\lstinline!gate!}, wiring by
\passthrough{\lstinline!wire!}, multiplexer, calculator FSM, curses/Xt
wrappers \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{1. Abstract Data Types and Information
Hiding}\label{abstract-data-types-and-information-hiding}

\subsection{1.1 Descriptor-based
allocation}\label{descriptor-based-allocation}

The first design hides representation. Clients receive only descriptors
and generic pointers.

\begin{lstlisting}[language=C]
function NEW_FROM_SIZE_DESCRIPTOR(typeDescriptor):
    size = *(size_t*)typeDescriptor
    object = calloc(1, size)
    assert(object != NULL)
    return object

function DELETE_GENERIC(object):
    free(object)       // Free accepts NULL.
\end{lstlisting}

\subsection{1.2 Fixed heap allocator for tiny set/object
experiment}\label{fixed-heap-allocator-for-tiny-setobject-experiment}

The initial set implementation represents every set or object as a slot
in a fixed integer heap.

\begin{lstlisting}[language=C]
constant MANY = configured_capacity
global int heap[MANY] = {0}

function NEW_FIXED_HEAP(typeDescriptor):
    // typeDescriptor is intentionally ignored in this experiment.
    for p = &heap[1] to &heap[MANY - 1]:
        if *p == 0:
            *p = MANY          // allocated but not in any set
            return p

    assert(false)              // heap exhausted

function DELETE_FIXED_HEAP(item):
    if item == NULL:
        return

    assert(item points inside heap[1..MANY-1])
    *item = 0                  // mark free
\end{lstlisting}

\subsection{1.3 Fixed heap set: add, find, contains,
drop}\label{fixed-heap-set-add-find-contains-drop}

A set is represented indirectly: each object slot stores the index of
the set that contains it.

\begin{lstlisting}[language=C]
function SET_ADD(set, element):
    assert(set is pointer into heap)
    assert(*set == MANY)       // a set cannot itself be inside another set
    assert(element is pointer into heap)

    if *element == MANY:
        *element = index_of(set, heap)
    else:
        assert(*element == index_of(set, heap))

    return element

function SET_FIND(set, element):
    assert(set is pointer into heap)
    assert(*set == MANY)
    assert(element is pointer into heap)
    assert(*element != 0)      // allocated

    if *element == index_of(set, heap):
        return element
    return NULL

function SET_CONTAINS(set, element):
    return SET_FIND(set, element) != NULL

function SET_DROP(set, element):
    found = SET_FIND(set, element)
    if found != NULL:
        *found = MANY          // allocated but not in a set
    return found

function OBJECT_DIFFER(a, b):
    return a != b
\end{lstlisting}

\subsection{1.4 Dynamically allocated
bag}\label{dynamically-allocated-bag}

The second implementation changes the hidden representation without
changing the public interface. A set tracks total membership count. An
object tracks its owning set and reference count.

\begin{lstlisting}[language=C]
struct SetRep {
    unsigned count
}

struct ObjectRep {
    unsigned count
    SetRep* in
}

function BAG_ADD(set, element):
    assert(set != NULL)
    assert(element != NULL)

    if element->in == NULL:
        element->in = set
    else:
        assert(element->in == set)

    element->count += 1
    set->count += 1
    return element

function BAG_FIND(set, element):
    assert(set != NULL)
    assert(element != NULL)

    if element->in == set:
        return element
    return NULL

function BAG_DROP(set, element):
    element = BAG_FIND(set, element)
    if element != NULL:
        element->count -= 1
        if element->count == 0:
            element->in = NULL
        set->count -= 1

    return element

function BAG_COUNT(set):
    assert(set != NULL)
    return set->count
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{2. Dynamic Linkage, Generic Functions, and
Polymorphism}\label{dynamic-linkage-generic-functions-and-polymorphism}

\subsection{2.1 Object layout invariant}\label{object-layout-invariant}

Every dynamically dispatched object begins with a pointer to a class
descriptor.

\begin{lstlisting}[language=C]
struct ClassDescriptor {
    size_t size
    function ctor(self, va_list*) -> void*
    function dtor(self) -> void*
    function clone(self) -> void*
    function differ(self, other) -> int
}

struct AnyObject {
    ClassDescriptor* class       // must be first
    ...
}
\end{lstlisting}

\subsection{2.2 Generic constructor
dispatcher}\label{generic-constructor-dispatcher}

\begin{lstlisting}[language=C]
function new(classDescriptor, ...):
    assert(classDescriptor != NULL)
    object = calloc(1, classDescriptor->size)
    assert(object != NULL)

    object->class = classDescriptor

    if classDescriptor->ctor != NULL:
        ap = start_varargs_after(classDescriptor)
        object = classDescriptor->ctor(object, &ap)
        end_varargs(ap)

    return object
\end{lstlisting}

\subsection{2.3 Generic destructor
dispatcher}\label{generic-destructor-dispatcher}

\begin{lstlisting}[language=C]
function delete(object):
    if object == NULL:
        return

    class = object->class
    assert(class != NULL)

    if class->dtor != NULL:
        object = class->dtor(object)

    free(object)        // destructor may return NULL to suppress free
\end{lstlisting}

\subsection{2.4 Generic selector for dynamic method
dispatch}\label{generic-selector-for-dynamic-method-dispatch}

Selectors are stable public functions. They retrieve the class
descriptor from the receiver and call the class-specific method.

\begin{lstlisting}[language=C]
function differ(self, other):
    assert(self != NULL)
    class = self->class
    assert(class != NULL)
    assert(class->differ != NULL)

    return class->differ(self, other)

function sizeOf(self):
    assert(self != NULL)
    class = self->class
    assert(class != NULL)
    return class->size
\end{lstlisting}

\subsection{2.5 String constructor, destructor, clone,
comparison}\label{string-constructor-destructor-clone-comparison}

\begin{lstlisting}[language=C]
struct StringRep {
    ClassDescriptor* class
    char* text
}

function STRING_CTOR(self, ap):
    text = va_arg(ap, char*)
    self->text = malloc(strlen(text) + 1)
    assert(self->text != NULL)
    strcpy(self->text, text)
    return self

function STRING_DTOR(self):
    free(self->text)
    self->text = NULL
    return self

function STRING_CLONE(self):
    return new(String, self->text)

function STRING_DIFFER(self, other):
    if self == other:
        return false

    if other == NULL || other->class != String:
        return true

    return strcmp(self->text, other->text)
\end{lstlisting}

\subsection{2.6 Atom interning with a circular
list}\label{atom-interning-with-a-circular-list}

Atoms are unique string objects. If an identical atom exists,
construction returns the existing object and discards the newly
allocated one.

\begin{lstlisting}[language=C]
global StringRep* atomRing = NULL

function ATOM_CLONE(self):
    self->count += 1
    return self

function ATOM_CTOR(self, ap):
    text = va_arg(ap, char*)

    if atomRing != NULL:
        p = atomRing
        do:
            if strcmp(p->text, text) == 0:
                p->count += 1
                free(self)             // discard newly allocated shell
                return p
            p = p->next
        while p != atomRing
    else:
        atomRing = self

    self->text = duplicate(text)
    self->count = 1

    // insert self into circular ring after atomRing
    self->next = atomRing->next
    atomRing->next = self
    return self

function ATOM_DTOR(self):
    self->count -= 1
    if self->count > 0:
        return NULL                    // suppress free by delete()

    assert(atomRing != NULL)

    if atomRing == self:
        atomRing = self->next

    if atomRing == self:
        atomRing = NULL
    else:
        p = atomRing
        while p->next != self:
            p = p->next
            assert(p != atomRing)
        p->next = self->next

    free(self->text)
    return self
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{3. Arithmetic Expressions: Scanner, Parser, and Expression
Trees}\label{arithmetic-expressions-scanner-parser-and-expression-trees}

\subsection{3.1 Main input loop with nonlocal error
recovery}\label{main-input-loop-with-nonlocal-error-recovery}

\begin{lstlisting}[language=C]
global token
global jump_buffer onError

function error(format, ...):
    print_formatted_error(format, ...)
    longjmp(onError, 1)

function main():
    errors = 0

    if setjmp(onError) != 0:
        errors += 1

    while line = read_line(stdin):
        if scan(line) != END:
            expr = parse_statement_or_sum()
            if token != END:
                error("trailing input")
            process(expr)
            delete(expr)

    return errors > 0
\end{lstlisting}

\subsection{3.2 Scanner for numbers and single-character
tokens}\label{scanner-for-numbers-and-single-character-tokens}

\begin{lstlisting}[language=C]
global cursor
global number
global token

function scan(lineOrNull):
    if lineOrNull != NULL:
        cursor = lineOrNull

    while isspace(*cursor):
        cursor += 1

    if isdigit(*cursor) || *cursor == '.':
        errno = 0
        number, cursor = parse_double(cursor)
        if errno == RANGE_ERROR:
            error("bad numeric value")
        token = NUMBER
    else if *cursor != '\0':
        token = *cursor
        cursor += 1
    else:
        token = END

    return token
\end{lstlisting}

\subsection{3.3 Recursive descent
grammar}\label{recursive-descent-grammar}

A representative grammar:

\begin{lstlisting}
sum     := product { ('+' | '-') product }*
product := factor  { ('*' | '/') factor }*
factor  := '+' factor
         | '-' factor
         | NUMBER
         | '(' sum ')'
\end{lstlisting}

\subsection{3.4 Parser constructs expression nodes while recognizing
input}\label{parser-constructs-expression-nodes-while-recognizing-input}

\begin{lstlisting}[language=C]
function parse_sum():
    result = parse_product()

    while token == '+' || token == '-':
        if token == '+':
            nodeType = Add
        else:
            nodeType = Sub

        scan(NULL)
        rhs = parse_product()
        result = new(nodeType, result, rhs)

    return result

function parse_product():
    result = parse_factor()

    while token == '*' || token == '/':
        if token == '*':
            nodeType = Mult
        else:
            nodeType = Div

        scan(NULL)
        rhs = parse_factor()
        result = new(nodeType, result, rhs)

    return result

function parse_factor():
    switch token:
        case '+':
            scan(NULL)
            return parse_factor()

        case '-':
            scan(NULL)
            return new(Minus, parse_factor())

        case NUMBER:
            value = number
            scan(NULL)
            return new(Value, value)

        case '(':
            scan(NULL)
            result = parse_sum()
            if token != ')':
                error("expected ')'")
            scan(NULL)
            return result

        default:
            error("bad factor")
\end{lstlisting}

\subsection{3.5 Expression-node factory by type
descriptor}\label{expression-node-factory-by-type-descriptor}

\begin{lstlisting}[language=C]
struct ExprType {
    char* name
    function make(ap) -> Node*
    function exec(node, ...) -> any
    function destroy(node)
}

function expr_new(type, ...):
    assert(type != NULL)
    assert(type->make != NULL)

    ap = start_varargs_after(type)
    node = type->make(ap)
    node->type = type
    end_varargs(ap)
    return node
\end{lstlisting}

\subsection{3.6 Binary-node constructor and
destructor}\label{binary-node-constructor-and-destructor}

\begin{lstlisting}[language=C]
struct BinaryNode {
    ExprType* type
    Node* left
    Node* right
}

function MAKE_BINARY(ap):
    node = malloc(sizeof(BinaryNode))
    assert(node != NULL)
    node->left = va_arg(ap, Node*)
    node->right = va_arg(ap, Node*)
    return node

function FREE_BINARY(node):
    delete(node->left)
    delete(node->right)
    free(node)
\end{lstlisting}

\subsection{3.7 Postfix traversal}\label{postfix-traversal}

\begin{lstlisting}[language=C]
function EXEC_POSTFIX(node):
    type = node->type
    assert(type != NULL)
    return type->exec(node)

function POSTFIX_BINARY(node):
    EXEC_POSTFIX(node->left)
    EXEC_POSTFIX(node->right)
    print(" ", node->type->name)

function POSTFIX_VALUE(node):
    print(" ", node->value)

function PROCESS_POSTFIX(root):
    print("\t")
    EXEC_POSTFIX(root)
    print("\n")
\end{lstlisting}

\subsection{3.8 Arithmetic evaluation}\label{arithmetic-evaluation}

\begin{lstlisting}[language=C]
function EVAL(node):
    type = node->type
    assert(type != NULL && type->exec != NULL)
    return type->exec(node)

function EVAL_VALUE(node):
    return node->value

function EVAL_ADD(node):
    return EVAL(node->left) + EVAL(node->right)

function EVAL_SUB(node):
    return EVAL(node->left) - EVAL(node->right)

function EVAL_MULT(node):
    return EVAL(node->left) * EVAL(node->right)

function EVAL_DIV(node):
    rhs = EVAL(node->right)
    assert(rhs != 0)
    return EVAL(node->left) / rhs

function EVAL_MINUS(node):
    return -EVAL(node->down)
\end{lstlisting}

\subsection{3.9 Minimal-parentheses infix
output}\label{minimal-parentheses-infix-output}

Each operator carries a precedence rank and a flag that says whether the
right operand needs parentheses at equal rank.

\begin{lstlisting}[language=C]
function PRINT_INFIX(node, parentRank, parentRequiresParensAtEqual):
    type = node->type

    needParens =
        type->rank < parentRank ||
        (parentRequiresParensAtEqual && type->rank == parentRank)

    if needParens:
        print("(")

    if node is binary:
        PRINT_INFIX(node->left, type->rank, false)
        print(" ", type->name, " ")
        PRINT_INFIX(node->right, type->rank, type->rightAssociativeNeedsParens)
    else if node is unary:
        print(type->name)
        PRINT_INFIX(node->down, type->rank, true)
    else if node is value:
        print(node->value)

    if needParens:
        print(")")
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{4. Inheritance by Structure
Prefixing}\label{inheritance-by-structure-prefixing}

\subsection{4.1 Layout rule}\label{layout-rule}

A subclass object begins with an entire superclass object
representation.

\begin{lstlisting}[language=C]
struct Point {
    ClassDescriptor* class
    int x
    int y
}

struct Circle {
    Point superPart     // must be first
    int radius
}
\end{lstlisting}

An up-cast from \passthrough{\lstinline!Circle*!} to
\passthrough{\lstinline!Point*!} is valid because
\passthrough{\lstinline!Circle!} begins with
\passthrough{\lstinline!Point!}.

\subsection{4.2 Point constructor, move,
draw}\label{point-constructor-move-draw}

\begin{lstlisting}[language=C]
function POINT_CTOR(self, ap):
    self->x = va_arg(ap, int)
    self->y = va_arg(ap, int)
    return self

function POINT_MOVE(point, dx, dy):
    point->x += dx
    point->y += dy

function POINT_DRAW(point):
    print_point(point->x, point->y)
\end{lstlisting}

\subsection{4.3 Circle constructor chains to Point
constructor}\label{circle-constructor-chains-to-point-constructor}

\begin{lstlisting}[language=C]
function CIRCLE_CTOR(self, ap):
    // Initialize superclass prefix first.
    self = super_ctor(Circle, self, ap)

    // Initialize subclass extension.
    self->radius = va_arg(ap, int)
    return self

function CIRCLE_DRAW(circle):
    print_circle(x(circle), y(circle), circle->radius)
\end{lstlisting}

\subsection{4.4 Static versus dynamic
inheritance}\label{static-versus-dynamic-inheritance}

\begin{lstlisting}[language=C]
// Static method inherited unchanged:
POINT_MOVE((Point*)circle, dx, dy)

// Dynamic method selected by actual class descriptor:
draw((Point*)circle)       // dispatches to CIRCLE_DRAW if circle->class == Circle
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{5. Symbol Table and Extended
Calculator}\label{symbol-table-and-extended-calculator}

\subsection{5.1 Identifier scanning}\label{identifier-scanning}

\begin{lstlisting}[language=C]
function scan(lineOrNull):
    ...
    if isalpha(*cursor) || *cursor == '_':
        name = read_while(cursor, is_alnum_or_underscore)
        cursor += length(name)
        token = screen(name)
        return token
    ...
\end{lstlisting}

\subsection{5.2 Sorted dynamic-array binary search with
insertion}\label{sorted-dynamic-array-binary-search-with-insertion}

The \passthrough{\lstinline!binary!} routine is a hybrid of binary
search and ordered insertion.

\begin{lstlisting}[language=C]
function BINARY_SEARCH_OR_INSERT(key, base, countRef, width, compare):
    count = *countRef
    low = 0
    high = count - 1

    while low <= high:
        mid = low + (high - low) / 2
        c = compare(key, element_at(base, mid, width))

        if c < 0:
            high = mid - 1
        else if c > 0:
            low = mid + 1
        else:
            return address_of_element(base, mid, width)

    // Not found: low is the insertion point.
    shift_right(base, low, count - low, width)
    *countRef += 1
    copy_element(address_of_element(base, low, width), key, width)
    return address_of_element(base, low, width)
\end{lstlisting}

\subsection{5.3 Symbol table storage
growth}\label{symbol-table-storage-growth}

\begin{lstlisting}[language=C]
function SYMBOL_SEARCH(table, nameRef):
    if table.used >= table.capacity:
        if table.capacity == 0:
            table.capacity = INITIAL_SYMBOL_CAPACITY
            table.entries = malloc(table.capacity * sizeof(Symbol*))
        else:
            table.capacity *= 2
            table.entries = realloc(table.entries,
                                    table.capacity * sizeof(Symbol*))
        assert(table.entries != NULL)

    return BINARY_SEARCH_OR_INSERT(nameRef,
                                   table.entries,
                                   &table.used,
                                   sizeof(Symbol*),
                                   compare_name_to_symbol)
\end{lstlisting}

\subsection{5.4 Install predefined
symbols}\label{install-predefined-symbols}

\begin{lstlisting}[language=C]
function INSTALL(symbol):
    name = symbol->name
    slot = SYMBOL_SEARCH(symbolTable, &name)

    if *slot != name:
        error("symbol already installed")

    *slot = symbol
\end{lstlisting}

\subsection{5.5 Screen an input
identifier}\label{screen-an-input-identifier}

\begin{lstlisting}[language=C]
function SCREEN(name):
    slot = SYMBOL_SEARCH(symbolTable, &name)

    if *slot == name:
        // The name string was inserted as a placeholder.
        *slot = new(Var, name)

    global symbol = *slot
    return symbol->token
\end{lstlisting}

\subsection{5.6 Assignment statement
recognition}\label{assignment-statement-recognition}

\begin{lstlisting}[language=C]
function parse_statement():
    if token == LET:
        scan(NULL)
        if token != VAR:
            error("bad assignment")

        variable = symbol

        scan(NULL)
        if token != '=':
            error("expected '='")

        scan(NULL)
        return new(Assign, variable, parse_sum())

    return parse_sum()
\end{lstlisting}

\subsection{5.7 Variable execution and
assignment}\label{variable-execution-and-assignment}

\begin{lstlisting}[language=C]
function EXEC_VAR(varSymbol):
    return varSymbol->value

function EXEC_ASSIGN(assignNode):
    value = EVAL(assignNode->right)
    assignNode->leftSymbol->value = value
    return value
\end{lstlisting}

\subsection{5.8 Mathematical function call
node}\label{mathematical-function-call-node}

\begin{lstlisting}[language=C]
function INIT_MATH_SYMBOLS():
    for each entry in mathFunctionTable:
        install(new_static_math_symbol(entry.name, MATH, entry.functionPointer))

function parse_factor_for_math():
    if token == MATH:
        functionSymbol = symbol
        scan(NULL)
        require('(')
        scan(NULL)
        arg = parse_sum()
        require(')')
        scan(NULL)
        return new(MathNode, functionSymbol, arg)

function EXEC_MATH(node):
    argValue = EVAL(node->right)
    errno = 0
    result = node->functionSymbol->function(argValue)

    if errno != 0:
        error("math function failed")

    return result
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{6. Class Hierarchy and
Metaclasses}\label{class-hierarchy-and-metaclasses}

\subsection{6.1 Root object and class
descriptors}\label{root-object-and-class-descriptors}

\begin{lstlisting}[language=C]
struct Object {
    Class* class
}

struct Class {
    Object superObject       // Class is itself an Object.
    char* name
    Class* super
    size_t size

    Method ctor
    Method dtor
    Method differ
    Method puto
}
\end{lstlisting}

\subsection{\texorpdfstring{6.2 \texttt{classOf}, \texttt{sizeOf},
\texttt{super}}{6.2 classOf, sizeOf, super}}\label{classof-sizeof-super}

\begin{lstlisting}[language=C]
function classOf(object):
    assert(object != NULL)
    assert(object->class != NULL)
    return object->class

function sizeOf(object):
    class = classOf(object)
    return class->size

function super(classObject):
    assert(classObject != NULL)
    assert(classObject->super != NULL)
    return classObject->super
\end{lstlisting}

\subsection{6.3 Object creation through class
descriptors}\label{object-creation-through-class-descriptors}

\begin{lstlisting}[language=C]
function object_new(class, ...):
    assert(class != NULL)
    assert(class->size > 0)

    object = calloc(1, class->size)
    assert(object != NULL)
    object->class = class

    ap = start_varargs_after(class)
    object = ctor(object, &ap)
    end_varargs(ap)

    return object
\end{lstlisting}

\subsection{6.4 Object deletion through destructor
selector}\label{object-deletion-through-destructor-selector}

\begin{lstlisting}[language=C]
function object_delete(object):
    if object == NULL:
        return

    object = dtor(object)
    free(object)
\end{lstlisting}

\subsection{6.5 Generic constructor
selector}\label{generic-constructor-selector}

\begin{lstlisting}[language=C]
function ctor(self, ap):
    class = classOf(self)
    assert(class->ctor != NULL)
    return class->ctor(self, ap)

function dtor(self):
    class = classOf(self)
    assert(class->dtor != NULL)
    return class->dtor(self)
\end{lstlisting}

\subsection{6.6 Class constructor implements binary
inheritance}\label{class-constructor-implements-binary-inheritance}

The metaclass constructor creates class descriptors. It copies inherited
method slots from the superclass descriptor, then overwrites only the
explicitly supplied methods.

\begin{lstlisting}[language=C]
function CLASS_CTOR(self, ap):
    self->name = va_arg(ap, char*)
    self->super = va_arg(ap, Class*)
    self->size = va_arg(ap, size_t)

    assert(self->super != NULL)

    // Copy inherited method table region from superclass.
    offset = offsetof(Class, ctor)
    byteCount = sizeOf(self->super) - offset
    memcpy((char*)self + offset,
           (char*)self->super + offset,
           byteCount)

    // Apply selector/method overrides.
    while true:
        selector = va_arg(ap, Method)
        if selector == NULL:
            break

        method = va_arg(ap, Method)

        if selector == ctor:
            self->ctor = method
        else if selector == dtor:
            self->dtor = method
        else if selector == differ:
            self->differ = method
        else if selector == puto:
            self->puto = method
        else:
            ignore_unknown_selector_for_this_metaclass()

    return self
\end{lstlisting}

\subsection{6.7 Superclass selector}\label{superclass-selector}

A method that wants to call the method it overrides must name the class
where the current method is defined, not
\passthrough{\lstinline!classOf(self)!}.

\begin{lstlisting}[language=C]
function super_ctor(currentClass, self, ap):
    superclass = super(currentClass)
    assert(superclass->ctor != NULL)
    return superclass->ctor(self, ap)

function super_dtor(currentClass, self):
    superclass = super(currentClass)
    assert(superclass->dtor != NULL)
    return superclass->dtor(self)
\end{lstlisting}

\subsection{\texorpdfstring{6.8 Metaclass extension for
\texttt{draw}}{6.8 Metaclass extension for draw}}\label{metaclass-extension-for-draw}

\begin{lstlisting}[language=C]
struct PointClass extends Class {
    Method draw
}

function POINT_CLASS_CTOR(self, ap):
    self = super_ctor(PointClass, self, ap)

    // Shared argument list: process method overrides for new slots.
    while true:
        selector = next_selector_without_consuming_super_processed_entries(ap)
        if selector == NULL:
            break

        method = next_method(ap)
        if selector == draw:
            self->draw = method

    return self

function draw(self):
    class = classOf(self)
    assert(class is PointClass or subclass of PointClass)
    assert(class->draw != NULL)
    class->draw(self)
\end{lstlisting}

\subsection{6.9 Class initialization
function}\label{class-initialization-function}

\begin{lstlisting}[language=C]
global PointClassDescription = NULL
global PointDescription = NULL

function initPoint():
    if PointClassDescription == NULL:
        PointClassDescription =
            new(Class, "PointClass",
                Class, sizeof(PointClass),
                ctor, POINT_CLASS_CTOR,
                NULL)

    if PointDescription == NULL:
        PointDescription =
            new(PointClassDescription, "Point",
                Object, sizeof(Point),
                ctor, POINT_CTOR,
                draw, POINT_DRAW,
                NULL)
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{\texorpdfstring{7. The \texttt{ooc} Preprocessor and Generated
Boilerplate}{7. The ooc Preprocessor and Generated Boilerplate}}\label{the-ooc-preprocessor-and-generated-boilerplate}

\subsection{7.1 Class-description model}\label{class-description-model}

\passthrough{\lstinline!ooc!} reads compact class descriptions and
generates C declarations and dispatch boilerplate.

\begin{lstlisting}[language=C]
record ClassDescription:
    className
    superName
    metaName
    metaSuperName
    objectFields[]
    staticMethods[]
    dynamicMethods[]
    classMethods[]
    protectedText[]
    publicText[]
\end{lstlisting}

\subsection{7.2 Recursive loading of superclass
descriptions}\label{recursive-loading-of-superclass-descriptions}

\begin{lstlisting}[language=C]
function LOAD_DESCRIPTION(className):
    if database.contains(className):
        return database[className]

    desc = parse_file(className + ".d")

    if desc.superName exists:
        desc.super = LOAD_DESCRIPTION(desc.superName)

    if desc.definesNewMetaclass:
        desc.metaSuper = desc.super.metaName
    else:
        desc.metaName = desc.super.metaName

    database[className] = desc
    return desc
\end{lstlisting}

\subsection{7.3 Generate interface file}\label{generate-interface-file}

\begin{lstlisting}[language=C]
function GENERATE_INTERFACE(desc):
    emit_include_guard(desc.className + "_h")

    emit(desc.publicText)
    if desc.superName exists:
        emit('#include "' + desc.superName + '.h"')

    emit("extern const void* " + desc.className + ";")

    for method in desc.staticMethods:
        emit_public_method_declaration(method)

    if desc.definesNewMetaclass:
        emit("extern const void* " + desc.metaName + ";")
        for method in desc.dynamicMethods:
            emit_selector_declaration(method)

    emit_end_guard()
\end{lstlisting}

\subsection{7.4 Generate representation
file}\label{generate-representation-file}

\begin{lstlisting}[language=C]
function GENERATE_REPRESENTATION(desc):
    emit_include_guard(desc.className + "_r")

    if desc.superName exists:
        emit('#include "' + desc.superName + '.r"')

    emit("struct " + desc.className + " {")
    emit("    const struct " + desc.superName + " _;")
    for field in desc.objectFields:
        emit("    " + field.declaration + ";")
    emit("};")

    emit(desc.protectedText)

    if desc.definesNewMetaclass:
        emit("struct " + desc.metaName + " {")
        emit("    const struct " + desc.metaSuperName + " _;")
        for method in desc.dynamicMethods:
            emit("    MethodSlot " + method.name + ";")
        emit("};")

        for method in desc.dynamicMethods:
            emit_super_selector_declaration(method)

    emit_end_guard()
\end{lstlisting}

\subsection{7.5 Preprocess implementation
file}\label{preprocess-implementation-file}

\begin{lstlisting}[language=C]
function PREPROCESS_IMPLEMENTATION(desc, implementationTemplate):
    for directive in implementationTemplate:
        switch directive.kind:
            case METHOD_HEADER:
                method = find_method(desc, directive.methodName)
                emit_generated_method_header(desc, method)
            case CASTS:
                emit_parameter_imports_and_casts(currentMethod)
            case INIT:
                emit_selectors_if_needed(desc)
                emit_super_selectors_if_needed(desc)
                emit_metaclass_constructor_if_needed(desc)
                emit_class_initialization(desc)
            default:
                emit_raw_text(directive.text)
\end{lstlisting}

\subsection{7.6 Report-expansion engine}\label{report-expansion-engine}

The report language provides loops over methods and parameters,
conditionals, and substitution.

\begin{lstlisting}[language=C]
function EXPAND_REPORT(report, context):
    output = ""

    for token in tokenize_report(report):
        if token is literal:
            output += token.text

        else if token is substitution:
            output += context.lookup(token.name)

        else if token is conditional:
            if evaluate_condition(token.condition, context):
                output += EXPAND_REPORT(token.body, context)

        else if token is loop:
            for item in context.sequence(token.sequenceName):
                child = context.with_current(item)
                output += EXPAND_REPORT(token.body, child)

        else if token is reportCall:
            output += EXPAND_REPORT(load_report(token.name), context)

    return normalize_whitespace(output)
\end{lstlisting}

\subsection{7.7 Dynamic ring-buffer
list}\label{dynamic-ring-buffer-list}

\begin{lstlisting}[language=C]
struct List {
    Object super
    Object** buf
    unsigned dim
    unsigned count
    unsigned begin
    unsigned end
}

function LIST_CTOR(self, ap):
    initialDim = va_arg(ap, unsigned)
    if initialDim == 0:
        initialDim = MINIMUM_DIMENSION

    self->dim = initialDim
    self->buf = malloc(self->dim * sizeof(Object*))
    assert(self->buf != NULL)
    return self

function LIST_EXTEND_FOR_ONE_MORE(self):
    if self->count < self->dim:
        self->count += 1
        return

    self->buf = realloc(self->buf, 2 * self->dim * sizeof(Object*))
    assert(self->buf != NULL)

    if self->begin != 0 && self->begin != self->dim:
        // The logical ring wraps; move the front segment into the new half.
        move_count = self->dim - self->begin
        copy(self->buf + self->dim + self->begin,
             self->buf + self->begin,
             move_count)
        self->begin += self->dim
    else:
        self->begin = 0

    self->dim *= 2
    self->count += 1

function LIST_ADD_FIRST(self, element):
    if self->count == 0:
        self->begin = 0
        self->end = 1
        self->count = 1
        self->buf[0] = element
        return element

    LIST_EXTEND_FOR_ONE_MORE(self)

    if self->begin == 0:
        self->begin = self->dim

    self->begin -= 1
    self->buf[self->begin] = element
    return element

function LIST_ADD_LAST(self, element):
    if self->count == 0:
        self->begin = 0
        self->end = 1
        self->count = 1
        self->buf[0] = element
        return element

    LIST_EXTEND_FOR_ONE_MORE(self)

    if self->end >= self->dim:
        self->end = 0

    self->buf[self->end] = element
    self->end += 1
    return element

function LIST_LOOK_AT(self, n):
    if n >= self->count:
        return NULL
    return self->buf[(self->begin + n) % self->dim]

function LIST_TAKE_FIRST(self):
    if self->count == 0:
        return NULL

    self->count -= 1

    if self->begin >= self->dim:
        self->begin = 0

    result = self->buf[self->begin]
    self->begin += 1
    return result

function LIST_TAKE_LAST(self):
    if self->count == 0:
        return NULL

    self->count -= 1

    if self->end == 0:
        self->end = self->dim

    self->end -= 1
    return self->buf[self->end]
\end{lstlisting}

\subsection{7.8 Stack and queue
specializations}\label{stack-and-queue-specializations}

\begin{lstlisting}[language=C]
function STACK_ADD(self, element):
    return LIST_ADD_LAST(self, element)

function STACK_TAKE(self):
    return LIST_TAKE_LAST(self)

function QUEUE_ADD(self, element):
    return LIST_ADD_LAST(self, element)

function QUEUE_TAKE(self):
    return LIST_TAKE_FIRST(self)
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{8. Dynamic Type Checking and Defensive
Programming}\label{dynamic-type-checking-and-defensive-programming}

\subsection{8.1 Basic type predicates}\label{basic-type-predicates}

\begin{lstlisting}[language=C]
function isA(self, class):
    return self != NULL && classOf(self) == class

function isOf(self, class):
    if self == NULL:
        return false

    myClass = classOf(self)

    if class == Object:
        return true

    while myClass != class:
        if myClass == Object:
            return false
        myClass = super(myClass)

    return true
\end{lstlisting}

\subsection{8.2 Safe cast}\label{safe-cast}

\begin{lstlisting}[language=C]
function cast(expectedClass, self):
    assert(self != NULL)

    // First verify that self looks like an object.
    assert(self->magic == MAGIC)

    myClass = self->class
    assert(myClass != NULL)
    assert(myClass->magic == MAGIC)

    if expectedClass != Object:
        assert(expectedClass != NULL)
        assert(expectedClass->magic == MAGIC)

        while myClass != expectedClass:
            assert(myClass != Object)
            assert(myClass is a valid Class descriptor)
            myClass = myClass->super

    return self
\end{lstlisting}

\subsection{8.3 Method import rules}\label{method-import-rules}

\begin{lstlisting}[language=C]
// Static method:
function STATIC_METHOD(self, objectArg):
    self = cast(CurrentClass, self)
    cast(ExpectedClass, objectArg)
    ...

// Selector:
function SELECTOR(self, objectArg):
    class = cast(ExpectedMetaclass, classOf(self))
    cast(ExpectedClass, objectArg)
    assert(class->method != NULL)
    return class->method(self, objectArg)

// Dynamically linked method:
function CLASS_METHOD(self, objectArg):
    self = cast(CurrentClass, self)
    ...
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{9. Static Construction and
Self-Organization}\label{static-construction-and-self-organization}

\subsection{9.1 Initializer-list
execution}\label{initializer-list-execution}

\begin{lstlisting}[language=C]
global function initializers[] = {
    NULL
}

function RUN_INITIALIZERS():
    for i = 0; initializers[i] != NULL; ++i:
        initializers[i]()
\end{lstlisting}

\subsection{\texorpdfstring{9.2 \texttt{munch}: generate initializer
list from symbol
table}{9.2 munch: generate initializer list from symbol table}}\label{munch-generate-initializer-list-from-symbol-table}

\begin{lstlisting}[language=C]
function MUNCH_SYMBOL_LIST(nmLines):
    names = empty_set()

    for line in nmLines:
        fields = split(line)
        if count(fields) != 3:
            continue

        address = fields[0]
        kind = fields[1]
        name = fields[2]

        if kind != "T":
            continue
        if not is_hex_address(address):
            continue
        if not matches(name, optional_underscore + "init" + CapitalizedIdentifier):
            continue

        name = remove_leading_underscore_if_present(name)
        names.add(name)

    emit_extern_declarations(names)
    emit("void (*initializers[])(void) = {")
    for name in names:
        emit("    " + name + ",")
    emit("    0")
    emit("};")
\end{lstlisting}

\subsection{9.3 Class description as a
function}\label{class-description-as-a-function}

Instead of exposing \passthrough{\lstinline!Point!} as a raw pointer,
expose \passthrough{\lstinline!Point()!} as a lazy initialization
function.

\begin{lstlisting}[language=C]
static void* _Point = NULL

function Point():
    if _Point != NULL:
        return _Point

    _Point = new(PointClass(),
                 "Point",
                 Object(),
                 sizeof(Point),
                 ctor, POINT_CTOR,
                 draw, POINT_DRAW,
                 NULL)

    return _Point
\end{lstlisting}

\subsection{9.4 Generate class list rather than initializer
list}\label{generate-class-list-rather-than-initializer-list}

\begin{lstlisting}[language=C]
function MUNCH_CLASS_FUNCTIONS(nmLines):
    textSymbols = set()
    staticDataSymbols = set()
    systemVDataSymbols = set()

    for line in nmLines:
        address, kind, name = parse_nm_line(line)
        if not is_hex_address(address):
            continue

        if kind in {"b", "s"}:
            staticDataSymbols.add(name)
        else if kind == "d":
            systemVDataSymbols.add(name)
        else if kind == "T":
            textSymbols.add(name)

    classNames = []
    for name in textSymbols:
        if "_" + name in staticDataSymbols:
            classNames.append(name)
        else if "_" + name in systemVDataSymbols:
            classNames.append(name)

    shellsort(classNames)

    emit_class_function_array(classNames)
\end{lstlisting}

\subsection{\texorpdfstring{9.5 Shellsort used by
\texttt{munch}}{9.5 Shellsort used by munch}}\label{shellsort-used-by-munch}

\begin{lstlisting}[language=C]
function SHELLSORT(names):
    gap = length(names) / 2
    while gap > 0:
        for i = gap to length(names) - 1:
            j = i - gap
            while j >= 0 && names[j] > names[j + gap]:
                swap(names[j], names[j + gap])
                j -= gap
        gap = gap / 2
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{10. Delegates, Callback Functions, and Application
Frameworks}\label{delegates-callback-functions-and-application-frameworks}

\subsection{\texorpdfstring{10.1 \texttt{respondsTo}: search method
table by
tag}{10.1 respondsTo: search method table by tag}}\label{respondsto-search-method-table-by-tag}

Each method table entry contains a tag, selector, and method pointer.

\begin{lstlisting}[language=C]
struct MethodSlot {
    char* tag
    Method selector
    Method method
}

function respondsTo(self, tag):
    if tag == NULL || tag[0] == '\0':
        return NULL

    class = classOf(self)
    methodStart = address_of(class->ctor)
    methodCount = (sizeOf(class) - offsetof(Class, ctor)) / sizeof(MethodSlot)

    p = methodStart
    for i = 0 to methodCount - 1:
        if p->tag != NULL && strcmp(p->tag, tag) == 0:
            if p->method != NULL:
                return p->selector
            return NULL
        p += 1

    return NULL
\end{lstlisting}

\subsection{10.2 Filter constructor discovers delegate
callbacks}\label{filter-constructor-discovers-delegate-callbacks}

\begin{lstlisting}[language=C]
function FILTER_CTOR(self, ap):
    self = super_ctor(Filter, self, ap)
    self->delegate = va_arg(ap, Object*)

    self->flag = respondsTo(self->delegate, "flag")
    self->name = respondsTo(self->delegate, "name")
    self->file = respondsTo(self->delegate, "file")
    self->line = respondsTo(self->delegate, "line")
    self->wrap = respondsTo(self->delegate, "wrap")
    self->quit = respondsTo(self->delegate, "quit")

    return self
\end{lstlisting}

\subsection{10.3 Filter main loop: parse options and
files}\label{filter-main-loop-parse-options-and-files}

\begin{lstlisting}[language=C]
function FILTER_MAIN_LOOP(self, argv):
    self->progname = argv[0]
    argv = argv + 1

    while argv[0] != NULL && starts_with(argv[0], "-"):
        arg = argv[0]

        if arg == "-":
            break                   // standard input as filename
        if arg == "--":
            argv += 1
            break                   // end options

        // Bundled single-character flags.
        for each flagChar in arg[1..]:
            if self->flag == NULL:
                print_error("no flags allowed")
                return 1

            self->argv = argv
            self->flag(self->delegate, self, flagChar)
            argv = self->argv       // callback may consume an option argument

        argv += 1

    if argv[0] != NULL:
        do:
            result = doit(self, argv[0])
            if result != 0:
                break
            argv += 1
        while argv[0] != NULL
    else:
        result = doit(self, NULL)

    if self->quit != NULL:
        result = self->quit(self->delegate, self)

    return result
\end{lstlisting}

\subsection{10.4 Option value retrieval}\label{option-value-retrieval}

\begin{lstlisting}[language=C]
function FILTER_ARGVAL(self):
    assert(self->argv != NULL)
    assert(self->argv[0] != NULL)

    current = self->argv[0]

    if current[1] != '\0':
        // -fVALUE
        result = &current[1]
        advance_current_argument_to_end(self->argv)
    else if self->argv[1] != NULL:
        // -f VALUE
        self->argv += 1
        result = self->argv[0]
        advance_current_argument_to_end(self->argv)
    else:
        result = NULL

    return result
\end{lstlisting}

\subsection{10.5 Process one file or standard
input}\label{process-one-file-or-standard-input}

\begin{lstlisting}[language=C]
function FILTER_DOIT(self, arg):
    if self->name != NULL:
        return self->name(self->delegate, self, arg)

    if arg == NULL || strcmp(arg, "-") == 0:
        fp = stdin
        clearerr(fp)
    else if arg[0] == '\0':
        print_error("null filename")
        return 1
    else:
        fp = fopen(arg, "r")
        if fp == NULL:
            perror(arg)
            return 1

    result = 0

    if self->file != NULL:
        result = self->file(self->delegate, self, arg, fp)
    else:
        ensure_dynamic_line_buffer(self)

        while fgets_dynamic(self->buf, self->blen, fp):
            if self->line != NULL:
                result = self->line(self->delegate, self, arg, self->buf)
                if result != 0:
                    break

        if self->wrap != NULL:
            result = self->wrap(self->delegate, self, arg)

    if fp != stdin:
        fclose(fp)

    fflush(stdout)
    if ferror(stdout):
        print_error("output error")
        result = 1

    return result
\end{lstlisting}

\subsection{10.6 Word-count delegate
callbacks}\label{word-count-delegate-callbacks}

\begin{lstlisting}[language=C]
function WC_LINE(self, filter, filename, buffer):
    self->lines += 1
    self->chars += strlen(buffer)
    return 0

function WC_WRAP(self, filter, filename):
    print_file_counts_if_needed(filename, self->lines, self->chars)

    self->allLines += self->lines
    self->allChars += self->chars
    self->lines = 0
    self->chars = 0
    self->files += 1
    return 0

function WC_QUIT(self, filter):
    if self->files > 1:
        print_total_counts(self->allLines, self->allChars, self->files)
    return 0
\end{lstlisting}

\subsection{10.7 Sort delegate}\label{sort-delegate}

\begin{lstlisting}[language=C]
function SORT_FLAG(self, filter, flag):
    if flag == 'r':
        self->reverse = true
    else:
        print_usage(progname(filter))
        exit(1)

function SORT_LINE(self, filter, filename, buffer):
    addLast(self, new(String, buffer))
    return 0

function SORT_QUIT(self, filter):
    if count(self) == 0:
        return 0

    sort(self)

    while count(self) > 0:
        if self->reverse:
            puto(takeLast(self), stdout)
        else:
            puto(takeFirst(self), stdout)

    return 0
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{11. Class Methods and Memory-Leak
Reclamation}\label{class-methods-and-memory-leak-reclamation}

\subsection{11.1 New as a class method}\label{new-as-a-class-method}

\passthrough{\lstinline!new!} becomes dynamically dispatched on a class
description.

\begin{lstlisting}[language=C]
function Object_new(classDescription, ap):
    object = allocate(classDescription)
    return ctor(object, ap)

function allocate(classDescription):
    assert(classDescription->size > 0)

    object = calloc(1, classDescription->size)
    assert(object != NULL)
    object->magic = MAGIC
    object->class = classDescription
    return object

function Object_delete(self):
    free(dtor(self))
\end{lstlisting}

\subsection{11.2 Class method selector}\label{class-method-selector}

\begin{lstlisting}[language=C]
function new(classDescription, ...):
    class = cast(Class, classDescription)
    ap = start_varargs_after(classDescription)

    if class->new.method != NULL:
        result = class->new.method(classDescription, &ap)
    else:
        forward(classDescription, &result, new, "new", classDescription, &ap)

    end_varargs(ap)
    return result
\end{lstlisting}

\subsection{11.3 Node class tracks all expression
nodes}\label{node-class-tracks-all-expression-nodes}

\begin{lstlisting}[language=C]
global Node* nodes = NULL

function NODE_NEW(classDescription, ap):
    result = super_new(Node, classDescription, ap)
    result = cast(Node, result)

    result->next = nodes
    nodes = result

    return result
\end{lstlisting}

\subsection{11.4 Sunder a node from the
chain}\label{sunder-a-node-from-the-chain}

\begin{lstlisting}[language=C]
function NODE_SUNDER(self):
    self = cast(Node, self)

    if nodes == self:
        nodes = self->next

    else if nodes != NULL:
        p = nodes
        while p->next != NULL && p->next != self:
            p = p->next

        if p->next == self:
            p->next = self->next

    self->next = NULL
\end{lstlisting}

\subsection{11.5 Node deletion removes node from recovery
chain}\label{node-deletion-removes-node-from-recovery-chain}

\begin{lstlisting}[language=C]
function NODE_DELETE(self):
    self = cast(Node, self)
    NODE_SUNDER(self)
    super_delete(Node, self)
\end{lstlisting}

\subsection{11.6 Reclaim all unreleased nodes after a parse
error}\label{reclaim-all-unreleased-nodes-after-a-parse-error}

\begin{lstlisting}[language=C]
function NODE_RECLAIM(classDescription, how):
    cast(NodeClass, classDescription)

    while nodes != NULL:
        how(nodes)
\end{lstlisting}

Typical use:

\begin{lstlisting}[language=C]
if parse_error_was_caught:
    reclaim(Node, delete)
\end{lstlisting}

\subsection{11.7 Preserve a successfully stored function
body}\label{preserve-a-successfully-stored-function-body}

If a parsed expression is stored in a symbol table as a function
definition, do not let the next error reclaim it.

\begin{lstlisting}[language=C]
function SET_FUNCTION(symbol, expressionTree):
    if symbol->fun != NULL:
        delete(symbol->fun)

    symbol->fun = expressionTree
    reclaim(Node, sunder)       // detach kept tree nodes from cleanup chain
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{12. Persistent Objects}\label{persistent-objects}

\subsection{\texorpdfstring{12.1 Cascading output with
\texttt{puto}}{12.1 Cascading output with puto}}\label{cascading-output-with-puto}

Each class writes only its own contribution after calling the superclass
method.

\begin{lstlisting}[language=C]
function OBJECT_PUTO(self, fp):
    class = classOf(self)
    return fprintf(fp, "%s at %p\n", class->name, self)

function SYMBOL_PUTO(self, fp):
    n = super_puto(Symbol, self, fp)
    n += fprintf(fp, "\tname %s\n", self->name)
    n += fprintf(fp, "\tlex %d\n", self->lex)
    return n

function VAR_PUTO(self, fp):
    n = super_puto(Var, self, fp)
    n += fprintf(fp, "\tvalue %g\n", self->value)
    return n
\end{lstlisting}

\subsection{12.2 Preorder persistence for expression
trees}\label{preorder-persistence-for-expression-trees}

\begin{lstlisting}[language=C]
function BINARY_PUTO(self, fp):
    n = super_puto(Binary, self, fp)
    n += puto(self->left, fp)
    n += puto(self->right, fp)
    return n

function REF_PUTO(self, fp):
    n = super_puto(Ref, self, fp)

    // Write only a symbolic reference, not the full referenced symbol.
    n += putsymbol(self->leftSymbol, fp)

    n += puto(self->right, fp)
    return n
\end{lstlisting}

\subsection{\texorpdfstring{12.3 Cascading input with
\texttt{geto}}{12.3 Cascading input with geto}}\label{cascading-input-with-geto}

\passthrough{\lstinline!retrieve!} allocates the object first, then
\passthrough{\lstinline!geto!} fills it.

\begin{lstlisting}[language=C]
function OBJECT_GETO(self, fp):
    if fscanf(fp, " at %p\n", &dummyAddress) != 1:
        assert(false)
    return self

function SYMBOL_GETO(self, fp):
    self = super_geto(Symbol, self, fp)

    if fscanf(fp, "\tname %s\n\tlex %d\n", buffer, &self->lex) != 2:
        assert(false)

    self->name = duplicate(buffer)
    return self

function VAR_GETO(self, fp):
    self = super_geto(Var, self, fp)

    if fscanf(fp, "\tvalue %lg\n", &self->value) != 1:
        assert(false)

    return self
\end{lstlisting}

\subsection{12.4 Retrieve object from class
name}\label{retrieve-object-from-class-name}

\begin{lstlisting}[language=C]
struct ClassListEntry {
    char* name
    Class* class
}

static ClassListEntry* classList = NULL
static int classCount = -1

function BUILD_CLASS_LIST_ONCE():
    if classCount >= 0:
        return

    classCount = count_class_functions(classes)

    if classCount == 0:
        classList = NULL
        return

    classList = malloc(classCount * sizeof(ClassListEntry))
    assert(classList != NULL)

    for i = 0 to classCount - 1:
        classList[i].class = classes[i]()
        classList[i].name = nameOf(classList[i].class)

function retrieve(fp):
    BUILD_CLASS_LIST_ONCE()

    if classCount == 0:
        print_error("no classes known")
        return NULL

    if fp == NULL || feof(fp):
        return NULL

    if fscanf(fp, "%s", className) != 1:
        return NULL

    key.name = className
    entry = bsearch(key, classList, classCount, compare_class_names)

    if entry == NULL:
        print_error(className + ": cannot retrieve")
        return NULL

    object = allocate(entry->class)
    return geto(object, fp)
\end{lstlisting}

\subsection{12.5 Attach symbolic references to current symbol
table}\label{attach-symbolic-references-to-current-symbol-table}

\begin{lstlisting}[language=C]
function getsymbol(expectedClass, fp):
    if fscanf(fp, "\tname %s\n\tlex %d\n", nameBuffer, &token) != 2:
        assert(false)

    result = screen(symbolTable, nameBuffer, UNDEF)

    if lex(result) == UNDEF:
        install(symbolTable, result = new(expectedClass, name(result), token))

    else if lex(result) != token:
        fclose(fp)
        error("%s: need %s, got %s",
              nameBuffer,
              nameOf(expectedClass),
              nameOf(classOf(result)))

    return result

function GLOBAL_GETO(self, fp):
    self = super_geto(Global, self, fp)
    self->down = getsymbol(Var, fp)
    return self

function PARM_GETO(self, fp):
    self = super_geto(Parm, self, fp)
    self->down = getsymbol(Fun, fp)
    return self
\end{lstlisting}

\subsection{12.6 Load symbol definitions
safely}\label{load-symbol-definitions-safely}

\begin{lstlisting}[language=C]
function SYMTAB_LOAD(self, filename):
    fp = fopen(filename, "r")
    if fp == NULL:
        return EOF

    result = 1

    while in = retrieve(fp):
        name = symbol_name(in)
        slot = search(self, &name)

        if *slot == name:
            // New name: insert retrieved symbol.
            *slot = in

        else if isA(*slot, Symbol):
            // Undefined placeholder: replace it.
            oldName = name(*slot)
            delete(*slot)
            free(oldName)
            *slot = in

        else if !respondsTo(*slot, "move"):
            // Existing object is protected; discard input.
            inName = name(in)
            delete(in)
            free(inName)
            continue

        else:
            // Existing object must remain stable because expression trees may point to it.
            move(*slot, in)
            delete(in)
            free(name)

    fclose(fp)
    return result
\end{lstlisting}

\subsection{\texorpdfstring{12.7 Shallow transfer by
\texttt{move}}{12.7 Shallow transfer by move}}\label{shallow-transfer-by-move}

\begin{lstlisting}[language=C]
function VAR_MOVE(self, from):
    self->value = from->value

function CONST_MOVE(self, from):
    // Intentionally untagged: constants should not respond to "move".
    do_nothing()

function FUN_MOVE(self, from):
    self->fun = from->fun
    from->fun = NULL        // prevent source deletion from deleting transferred tree
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{13. Exceptions and Disciplined Error
Recovery}\label{exceptions-and-disciplined-error-recovery}

\subsection{13.1 Exception object stack}\label{exception-object-stack}

\begin{lstlisting}[language=C]
struct Exception {
    Object super
    int armed
    jmp_buf label
}

global List* exceptionStack = NULL

function EXCEPTION_CTOR(self, ap):
    self = super_ctor(Exception, self, ap)

    if exceptionStack == NULL:
        exceptionStack = new(List, 10)

    addLast(exceptionStack, self)
    return self

function EXCEPTION_DTOR(self):
    assert(exceptionStack != NULL)

    top = takeLast(exceptionStack)
    assert(top == self)

    return super_dtor(Exception, self)
\end{lstlisting}

\subsection{13.2 Arm an exception object}\label{arm-an-exception-object}

The macro \passthrough{\lstinline!catch(e)!} expands to
\passthrough{\lstinline!setjmp(catchException(e))!}.

\begin{lstlisting}[language=C]
function catchException(self):
    self = cast(Exception, self)
    self->armed = true
    return self->label
\end{lstlisting}

\subsection{13.3 Cause an exception}\label{cause-an-exception}

\begin{lstlisting}[language=C]
function cause(number):
    if number == 0:
        return

    if exceptionStack == NULL:
        return

    cnt = count(exceptionStack)
    if cnt == 0:
        return

    top = lookAt(exceptionStack, cnt - 1)
    e = cast(Exception, top)

    if e->armed:
        longjmp(e->label, number)

    // If no armed handler exists, return to caller.
\end{lstlisting}

\subsection{13.4 Main loop recovery}\label{main-loop-recovery}

\begin{lstlisting}[language=C]
function main():
    errors = 0
    retry = new(Exception)

    if catch(retry) != 0:
        errors += 1
        reclaim(Node, delete)

    while line = read_line(stdin):
        parse_and_execute(line)

    delete(retry)
    return errors > 0
\end{lstlisting}

\subsection{13.5 Nested cleanup handler}\label{nested-cleanup-handler}

\begin{lstlisting}[language=C]
function SYMTAB_LOAD_WITH_EXCEPTION(self, filename):
    fp = fopen(filename, "r")
    if fp == NULL:
        return EOF

    cleanup = new(Exception)

    if catch(cleanup) != 0:
        fclose(fp)
        delete(cleanup)
        cause(1)             // propagate to outer handler
        assert(false)

    while object = retrieve(fp):
        merge_loaded_object(self, object)

    fclose(fp)
    delete(cleanup)
    return result
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{14. Message Forwarding and GUI
Calculator}\label{message-forwarding-and-gui-calculator}

\subsection{14.1 Forwarding selector}\label{forwarding-selector}

A selector first tries normal dispatch. If the receiver does not
implement the method, it calls \passthrough{\lstinline!forward!}.

\begin{lstlisting}[language=C]
function add(self, element):
    result = NULL
    class = classOf(self)

    if isOf(class, ListClass) && class->add.method != NULL:
        cast(Object, element)
        result = class->add.method(self, element)
    else:
        forward(self, &result, add, "add", self, element)

    return result
\end{lstlisting}

\subsection{14.2 Default forward method}\label{default-forward-method}

\begin{lstlisting}[language=C]
function OBJECT_FORWARD(self, resultArea, selector, name, ...):
    print_error("%s at %p does not answer %s",
                nameOf(classOf(self)), self, name)
    assert(false)
\end{lstlisting}

\subsection{14.3 Forwarding-aware selector
generation}\label{forwarding-aware-selector-generation}

\begin{lstlisting}[language=C]
function GENERATED_SELECTOR(self, args...):
    declare result if method returns nonvoid
    class = classOf(self)

    if class_is_compatible_with_method_metaclass(class) &&
       class->methodSlot.method != NULL:

        check_nonself_object_args(args)
        result = class->methodSlot.method(self, args...)

    else:
        if selector_being_generated_is_forward:
            assert(false)
        else:
            forward(self,
                    address_of_result_or_NULL,
                    selector_function_pointer,
                    selector_name,
                    self,
                    args...)

    return result if needed
\end{lstlisting}

\subsection{\texorpdfstring{14.4 Information-flow base class
\texttt{Ic}}{14.4 Information-flow base class Ic}}\label{information-flow-base-class-ic}

\begin{lstlisting}[language=C]
enum React { reject, accept }

struct Ic {
    Object super
    Object* out
}

function IC_WIRE(to, self):
    self = cast(Ic, self)
    self->out = to

function IC_GATE(self, item):
    self = cast(Ic, self)

    if self->out != NULL:
        return gate(self->out, item)

    return reject
\end{lstlisting}

\subsection{14.5 Multiplexer}\label{multiplexer}

\begin{lstlisting}[language=C]
struct Mux extends Ic {
    List* list
}

function MUX_WIRE(to, self):
    addLast(self->list, to)

function MUX_GATE(self, item):
    result = reject
    n = count(self->list)

    for i = 0 to n - 1:
        result = gate(lookAt(self->list, i), item)
        if result == accept:
            break

    return result
\end{lstlisting}

\subsection{14.6 Event object}\label{event-object}

\begin{lstlisting}[language=C]
struct Event {
    Object super
    int kind
    void* data
}

function EVENT_CTOR(self, ap):
    self = super_ctor(Event, self, ap)
    self->kind = va_arg(ap, int)
    self->data = va_arg(ap, void*)
    return self
\end{lstlisting}

\subsection{14.7 Button filter}\label{button-filter}

\begin{lstlisting}[language=C]
function BUTTON_CTOR(self, ap):
    self = super_ctor(Button, self, ap)
    self->text = va_arg(ap, char*)
    return self

function BUTTON_GATE(self, item):
    if item != NULL && kind(item) == TEXT_EVENT:
        if strcmp(data(item), self->text) != 0:
            return reject

    // NULL or matching text event means "press this button".
    return super_gate(Button, self, self->text)
\end{lstlisting}

\subsection{14.8 Line output sink}\label{line-output-sink}

\begin{lstlisting}[language=C]
function LINEOUT_GATE(self, item):
    assert(item != NULL)
    puts(item)
    return accept
\end{lstlisting}

\subsection{14.9 Calculator finite state
machine}\label{calculator-finite-state-machine}

The calculator is event-driven. It reacts to every button event.

\begin{lstlisting}[language=C]
struct Calc {
    Ic super
    int value[2]
    int op
    int state
}

function CALC_GATE(self, item):
    text = item_as_text(item)
    ch = text[0]

    if isdigit(ch):
        self->value[self->state] *= 10
        self->value[self->state] += digit_value(ch)
        return super_gate(Calc, self, format(self->value[self->state]))

    switch ch:
        case '+':
        case '-':
        case '*':
        case '/':
            if self->state == 1:
                apply_pending_operator(self)
            self->op = ch
            self->state = 1
            self->value[1] = 0
            return accept

        case '=':
            if self->state == 1:
                apply_pending_operator(self)
                self->state = 0
            return super_gate(Calc, self, format(self->value[0]))

        case 'C':
            self->value[self->state] = 0
            return super_gate(Calc, self, format(self->value[self->state]))

        case 'Q':
            exit(0)

        default:
            return reject

function apply_pending_operator(self):
    switch self->op:
        case '+': self->value[0] += self->value[1]; break
        case '-': self->value[0] -= self->value[1]; break
        case '*': self->value[0] *= self->value[1]; break
        case '/': self->value[0] /= self->value[1]; break
\end{lstlisting}

\subsection{14.10 Character-input test
harness}\label{character-input-test-harness}

\begin{lstlisting}[language=C]
function RUN_TEXT_CALCULATOR():
    calc = new(Calc)
    lineOut = new(LineOut)
    mux = new(Mux)

    wire(lineOut, calc)

    for label in ["C","Q","0","1","2","3","4","5","6","7","8","9","+","-","*","/","="]:
        button = new(Button, label)
        wire(calc, button)
        wire(button, mux)

    while ch = getchar():
        if isspace(ch):
            continue

        buffer[0] = ch
        buffer[1] = '\0'

        event = new(Event, TEXT_EVENT, buffer)
        gate(mux, event)
        delete(event)
\end{lstlisting}

\subsection{14.11 Curses wrapper
behavior}\label{curses-wrapper-behavior}

\begin{lstlisting}[language=C]
function CRT_GATE(self, ignored):
    while true:
        ch = get_character_from_terminal()

        if ch == CONTROL_D:
            return reject

        if ch is cursor_control:
            update_cursor_position(self, ch)
            continue

        if ch == ENTER:
            coords = current_cursor_coordinates(self)
            event = new(Event, POSITION_EVENT, coords)
        else:
            text = one_character_string(ch)
            event = new(Event, TEXT_EVENT, text)

        super_gate(Crt, self, event)
        delete(event)
\end{lstlisting}

\subsection{14.12 Curses button wrapper}\label{curses-button-wrapper}

\begin{lstlisting}[language=C]
function CBUTTON_CTOR(self, ap):
    self = super_ctor(CButton, self, ap)

    text = va_arg(ap, char*)
    self->button = new(Button, text)
    self->y = va_arg(ap, int)
    self->x = va_arg(ap, int)

    makeWindow(self, height = 3, width = strlen(text) + 4, self->y, self->x)
    addStr(self, row = 1, col = 2, text)
    draw_box(self)
    return self

function CBUTTON_WIRE(to, self):
    wire(to, self->button)

function CBUTTON_GATE(self, item):
    if kind(item) == POSITION_EVENT:
        pos = data(item)

        if point_inside_box(pos, self->x, self->y, cols(self), rows(self)):
            return gate(self->button, NULL)

        return reject

    return gate(self->button, item)
\end{lstlisting}

\subsection{14.13 X toolkit output
forwarding}\label{x-toolkit-output-forwarding}

\passthrough{\lstinline!XLineOut!} is not an
\passthrough{\lstinline!Ic!}, so it receives forwarded
\passthrough{\lstinline!gate!} calls.

\begin{lstlisting}[language=C]
function XLINEOUT_FORWARD(self, resultArea, selector, name, ap):
    if selector == gate:
        discard_original_self_from(ap)
        text = va_arg(ap, void*)
        setLabel(self, text)
        *(React*)resultArea = accept
    else:
        super_forward(XLineOut, self, resultArea, selector, name, ap)
\end{lstlisting}

\subsection{14.14 X toolkit button
forwarding}\label{x-toolkit-button-forwarding}

\passthrough{\lstinline!XButton!} owns a regular
\passthrough{\lstinline!Button!}; forwarded
\passthrough{\lstinline!wire!} calls are sent to the owned button.

\begin{lstlisting}[language=C]
function XBUTTON_CTOR(self, ap):
    self = super_ctor(XButton, self, ap)

    parent = already_consumed_by_super()
    name = already_consumed_by_super()
    text = va_arg(ap, char*)

    self->button = new(Button, text)
    setLabel(self, text)
    addCallback(self, TELL_CALLBACK, self->button)
    return self

function TELL_CALLBACK(widget, clientData, callData):
    gate(clientData, NULL)

function XBUTTON_FORWARD(self, resultArea, selector, name, ap):
    if selector == wire:
        to = va_arg(ap, void*)
        wire(to, self->button)
    else:
        super_forward(XButton, self, resultArea, selector, name, ap)
\end{lstlisting}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Appendix A. Practical Algorithm Map by Function
Name}\label{appendix-a.-practical-algorithm-map-by-function-name}

\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
Function / mechanism & Pseudocode location \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\passthrough{\lstinline!new!}, \passthrough{\lstinline!delete!} &
Sections 1.1, 2.2, 2.3, 6.3, 11.1 \\
\passthrough{\lstinline!add!}, \passthrough{\lstinline!find!},
\passthrough{\lstinline!drop!}, \passthrough{\lstinline!contains!} &
Sections 1.3, 1.4 \\
\passthrough{\lstinline!ctor!}, \passthrough{\lstinline!dtor!} &
Sections 2.2, 2.3, 6.5 \\
\passthrough{\lstinline!clone!}, \passthrough{\lstinline!differ!},
\passthrough{\lstinline!sizeOf!} & Sections 2.4, 2.5 \\
Atom ring & Section 2.6 \\
\passthrough{\lstinline!scan!}, \passthrough{\lstinline!sum!},
\passthrough{\lstinline!product!}, \passthrough{\lstinline!factor!} &
Sections 3.2-3.4 \\
\passthrough{\lstinline!exec!}, \passthrough{\lstinline!process!},
postfix/eval/infix & Sections 3.7-3.9 \\
\passthrough{\lstinline!move!}, \passthrough{\lstinline!draw!},
superclass selectors & Sections 4.2-4.4, 6.7 \\
\passthrough{\lstinline!binary!}, \passthrough{\lstinline!screen!},
\passthrough{\lstinline!install!} & Sections 5.2-5.5 \\
\passthrough{\lstinline!Class\_ctor!} & Section 6.6 \\
\passthrough{\lstinline!PointClass\_ctor!},
\passthrough{\lstinline!initPoint!} & Sections 6.8-6.9 \\
\passthrough{\lstinline!ooc!} parse/generate/report & Sections
7.1-7.6 \\
\passthrough{\lstinline!List!} ring buffer & Section 7.7 \\
\passthrough{\lstinline!isA!}, \passthrough{\lstinline!isOf!},
\passthrough{\lstinline!cast!} & Section 8 \\
\passthrough{\lstinline!munch!}, \passthrough{\lstinline!Point()!} lazy
class function & Section 9 \\
\passthrough{\lstinline!respondsTo!},
\passthrough{\lstinline!mainLoop!}, \passthrough{\lstinline!doit!},
\passthrough{\lstinline!argval!} & Section 10 \\
\passthrough{\lstinline!Node\_new!}, \passthrough{\lstinline!sunder!},
\passthrough{\lstinline!reclaim!} & Section 11 \\
\passthrough{\lstinline!puto!}, \passthrough{\lstinline!geto!},
\passthrough{\lstinline!retrieve!}, \passthrough{\lstinline!getsymbol!},
\passthrough{\lstinline!move!} & Section 12 \\
\passthrough{\lstinline!catch!}, \passthrough{\lstinline!cause!},
exception stack & Section 13 \\
\passthrough{\lstinline!forward!}, \passthrough{\lstinline!gate!},
\passthrough{\lstinline!wire!}, GUI wrappers & Section 14 \\
\end{longtable}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Appendix B. Design implications for modern C
work}\label{appendix-b.-design-implications-for-modern-c-work}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Opaque headers remain the central discipline.} The public
  interface exposes descriptors and functions; representation files
  expose only what subclass implementations need.
\item
  \textbf{The object header must be stable.} Dynamic dispatch depends on
  the first field being a class pointer.
\item
  \textbf{Inheritance is structure-prefix compatibility.} It is not
  compiler-supported inheritance; it is a layout invariant.
\item
  \textbf{Dynamic dispatch is ordinary table lookup.} Selectors are
  small, auditable dispatch algorithms.
\item
  \textbf{Metaclasses are method-table constructors.} Their main
  algorithmic job is to copy inherited slots and overwrite explicitly
  supplied slots.
\item
  \textbf{Generated boilerplate is a correctness tool.}
  \passthrough{\lstinline!ooc!} reduces dispatch-maintenance errors by
  centralizing selector, superclass selector, import/cast, and
  initialization generation.
\item
  \textbf{\passthrough{\lstinline!cast!} is the runtime contract
  checker.} It prevents arbitrary \passthrough{\lstinline!void*!} values
  from silently corrupting object-method assumptions.
\item
  \textbf{Class methods are useful when allocation policy is
  class-specific.} The \passthrough{\lstinline!Node!} example uses
  class-dispatched \passthrough{\lstinline!new!} to track all nodes for
  cleanup.
\item
  \textbf{Persistence requires class-name-to-descriptor lookup.} The
  program must already contain the code for the persisted class before
  \passthrough{\lstinline!retrieve!} can reconstruct objects.
\item
  \textbf{Message forwarding substitutes for some multiple-inheritance
  use cases.} Wrapper objects can behave as if they understand methods
  from another hierarchy by forwarding unrecognized selectors.
\end{enumerate}


\end{document}
