\documentclass[11pt]{article}

\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{tcolorbox}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}

\definecolor{titleblue}{HTML}{1F4E79}
\definecolor{sectionblue}{HTML}{245B8A}
\definecolor{lightblue}{HTML}{EAF3FA}
\definecolor{lightgray}{HTML}{F6F8FA}
\definecolor{codeframe}{HTML}{D0D7DE}
\definecolor{darkgray}{HTML}{333333}
\definecolor{green}{HTML}{116329}
\definecolor{red}{HTML}{A40E26}

\hypersetup{
  colorlinks=true,
  linkcolor=titleblue,
  urlcolor=titleblue,
  citecolor=titleblue,
  pdftitle={Point vs. Vertex in C: Use Composition, Not Inheritance},
  pdfauthor={Jordan Suber}
}

\setlength{\headheight}{14pt}
\pagestyle{fancy}
\fancyhf{}
\lhead{Point vs. Vertex in C}
\rhead{Composition over Inheritance}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\titleformat{\section}{\Large\bfseries\color{sectionblue}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries\color{darkgray}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{darkgray}}{}{0em}{}

\setlist[itemize]{leftmargin=*,topsep=3pt,itemsep=2pt}
\setlist[enumerate]{leftmargin=*,topsep=3pt,itemsep=2pt}
\renewcommand{\arraystretch}{1.2}

\lstdefinestyle{cstyle}{
  language=C,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{titleblue}\bfseries,
  commentstyle=\color{green},
  stringstyle=\color{red},
  numbers=left,
  numberstyle=\tiny\color{darkgray},
  stepnumber=1,
  numbersep=8pt,
  showstringspaces=false,
  breaklines=true,
  breakatwhitespace=false,
  frame=single,
  rulecolor=\color{codeframe},
  backgroundcolor=\color{lightgray},
  tabsize=4,
  columns=fullflexible,
  keepspaces=true,
  captionpos=b
}
\lstset{style=cstyle}

\newtcolorbox{decisionbox}{
  colback=lightblue,
  colframe=titleblue,
  boxrule=0.6pt,
  arc=2pt,
  left=8pt,
  right=8pt,
  top=6pt,
  bottom=6pt
}

\newtcolorbox{warningbox}{
  colback=red!5,
  colframe=red,
  boxrule=0.6pt,
  arc=2pt,
  left=8pt,
  right=8pt,
  top=6pt,
  bottom=6pt
}

\newtcolorbox{checkbox}{
  colback=green!5,
  colframe=green,
  boxrule=0.6pt,
  arc=2pt,
  left=8pt,
  right=8pt,
  top=6pt,
  bottom=6pt
}

\begin{document}

\begin{center}
  {\Huge\bfseries\color{titleblue} Point vs. Vertex in C}\\[4pt]
  {\Large Use Composition, Not Inheritance}\\[8pt]
  {\normalsize A design handout for computational geometry, graph algorithms, and mesh data structures}
\end{center}

\vspace{0.5em}

\begin{decisionbox}
\textbf{Core design rule:} A \texttt{Point} models geometry. A \texttt{Vertex} models topology. Do not put them in an inheritance hierarchy. Use separate data types for pure geometry and pure graph algorithms. Use composition when a topological vertex needs a geometric position.
\end{decisionbox}

\section{Purpose}

This handout updates the earlier \texttt{Point} versus \texttt{Vertex} guidance with C-style examples aligned to two algorithm families:

\begin{itemize}
  \item \textbf{Computational geometry style:} points are coordinate value types used by predicates such as orientation, sidedness, equality, intersection, and convex hull logic.
  \item \textbf{Stanford GraphBase style:} vertices and arcs form adjacency-list graph structures used by traversal and graph algorithms.
\end{itemize}

The goal is to preserve a clean separation between \textbf{geometry} and \textbf{topology} while still supporting hybrid programs such as pathfinding visualizers, mesh kernels, graph drawing tools, GIS systems, and computational-geometry applications.

\section{The Conceptual Boundary}

\begin{center}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth} >{\raggedright\arraybackslash}X >{\raggedright\arraybackslash}X}
\toprule
\textbf{Type} & \textbf{Primary Concern} & \textbf{Typical Fields} \\
\midrule
\texttt{Point2}, \texttt{Point3} & Geometry: where something is & \texttt{x}, \texttt{y}, \texttt{z}; no adjacency list \\
\texttt{Vertex} & Graph topology: how nodes connect & \texttt{arcs}, \texttt{name}, labels, marks, traversal state \\
\texttt{Arc} / \texttt{Edge} & Connectivity between vertices & destination pointer, next pointer, weight, incident faces \\
\texttt{Face} & Mesh topology and surface structure & vertices, edges, visibility flag, orientation metadata \\
\bottomrule
\end{tabularx}
\end{center}

\begin{warningbox}
\textbf{Do not model this as inheritance.} In C, inheritance is not native, but the same design mistake appears when one structure embeds another to pretend it is a subtype. Avoid designs that imply \texttt{Point IS-A Vertex} or \texttt{Vertex IS-A Point}.
\end{warningbox}

\section{Option 1: Keep Geometry and Topology Separate}

Use separate types when your program has pure geometry algorithms and pure graph algorithms that should not depend on each other.

\subsection{Geometry-Only Point Types}

A point is coordinate data. It should not know about neighbors, arcs, incident edges, traversal marks, or graph labels.

\begin{lstlisting}[caption={Coordinate value types for 2D and 3D geometry}]
typedef struct Point2 {
    double x;
    double y;
} Point2;

typedef struct Point3 {
    double x;
    double y;
    double z;
} Point3;
\end{lstlisting}

Geometry functions should operate directly on points:

\begin{lstlisting}[caption={A 2D orientation predicate based on twice-signed area}]
double area2(Point2 a, Point2 b, Point2 c) {
    return (b.x - a.x) * (c.y - a.y)
         - (c.x - a.x) * (b.y - a.y);
}

int left(Point2 a, Point2 b, Point2 c) {
    return area2(a, b, c) > 0.0;
}
\end{lstlisting}

This design keeps geometric predicates compact, reusable, and independent of graph storage.

\subsection{Topology-Only Graph Types}

A graph vertex represents connectivity. It does not require geometric coordinates.

\begin{lstlisting}[caption={Graph adjacency-list structures in C style}]
typedef struct Arc Arc;
typedef struct Vertex Vertex;

struct Arc {
    Vertex *tip;      /* destination vertex */
    Arc    *next;     /* next outgoing arc from same source */
    long    len;      /* arc length, cost, or weight */
};

struct Vertex {
    Arc  *arcs;       /* outgoing adjacency list */
    char *name;       /* symbolic graph name */
};
\end{lstlisting}

Graph traversal should operate on adjacency, not coordinates:

\begin{lstlisting}[caption={Traversing outgoing neighbors through an adjacency list}]
void visit_neighbors(Vertex *u) {
    for (Arc *a = u->arcs; a != NULL; a = a->next) {
        Vertex *v = a->tip;
        /* process neighbor v */
    }
}
\end{lstlisting}

This model supports graph algorithms where vertices may represent servers, tasks, users, cities, modules, applications, dependency nodes, or abstract states.

\section{Option 2: Use Composition for Geometric Graphs}

Use composition when a graph vertex must also have a geometric location. This is common in graph drawing, GIS, games, pathfinding, computational geometry, and visualization systems.

\begin{lstlisting}[caption={A geometric graph vertex composed from topology plus a point}]
typedef struct GeoVertex GeoVertex;
typedef struct GeoArc GeoArc;

typedef struct Point2 {
    double x;
    double y;
} Point2;

struct GeoArc {
    GeoVertex *tip;
    GeoArc    *next;
    double     weight;
};

struct GeoVertex {
    char    *name;
    Point2   position;   /* GeoVertex HAS-A Point2 */
    GeoArc  *arcs;       /* graph topology */
    int      mark;       /* traversal or algorithm state */
};
\end{lstlisting}

\begin{decisionbox}
\textbf{Correct relationship:}
\begin{center}
\texttt{GeoVertex HAS-A Point2} \\[-1pt]
\texttt{GeoVertex IS-NOT-A Point2} \\[-1pt]
\texttt{Point2 IS-NOT-A GeoVertex}
\end{center}
\end{decisionbox}

The geometry side can extract the position field:

\begin{lstlisting}[caption={Applying a geometric predicate to positioned vertices}]
int is_left_of_edge(GeoVertex *a, GeoVertex *b, GeoVertex *q) {
    return left(a->position, b->position, q->position);
}
\end{lstlisting}

The graph side can still traverse neighbors:

\begin{lstlisting}[caption={Traversing neighbors in a geometric graph}]
void visit_geo_neighbors(GeoVertex *u) {
    for (GeoArc *a = u->arcs; a != NULL; a = a->next) {
        GeoVertex *v = a->tip;
        /* process neighbor v */
    }
}
\end{lstlisting}

This design allows graph algorithms and geometry algorithms to share data without collapsing into one overloaded type.

\section{Option 3: Use Mesh-Oriented Composition for 3D Geometry}

For 3D computational geometry, a vertex often owns a point but also participates in edges and faces. The point stores coordinates; the vertex stores mesh membership and algorithm metadata.

\begin{lstlisting}[caption={3D mesh structures using composition}]
typedef struct Point3 {
    double x;
    double y;
    double z;
} Point3;

typedef struct Vertex3 Vertex3;
typedef struct Edge3   Edge3;
typedef struct Face3   Face3;

struct Vertex3 {
    Point3 p;          /* Vertex3 HAS-A Point3 */
    int    onHull;
    int    markDelete;
};

struct Edge3 {
    Vertex3 *a;
    Vertex3 *b;
    Face3   *adj[2];
    Face3   *newFace;
    int      markDelete;
};

struct Face3 {
    Vertex3 *v[3];
    Edge3   *e[3];
    int      visible;
};
\end{lstlisting}

This is not inheritance. \texttt{Vertex3} is a topological mesh element that owns a \texttt{Point3}. A \texttt{Point3} remains a plain coordinate value.

\subsection{Example: Face Visibility Predicate}

A visibility predicate should extract points from vertices before applying geometric operations.

\begin{lstlisting}[caption={Point-vector operations for a 3D visibility test}]
double dot3(Point3 a, Point3 b) {
    return a.x*b.x + a.y*b.y + a.z*b.z;
}

Point3 sub3(Point3 a, Point3 b) {
    Point3 r = { a.x - b.x, a.y - b.y, a.z - b.z };
    return r;
}

Point3 cross3(Point3 a, Point3 b) {
    Point3 r = {
        a.y*b.z - a.z*b.y,
        a.z*b.x - a.x*b.z,
        a.x*b.y - a.y*b.x
    };
    return r;
}

double volume6(Point3 a, Point3 b, Point3 c, Point3 d) {
    return dot3(cross3(sub3(b, a), sub3(c, a)), sub3(d, a));
}

int face_visible_from_point(Face3 *f, Point3 p) {
    Point3 a = f->v[0]->p;
    Point3 b = f->v[1]->p;
    Point3 c = f->v[2]->p;

    return volume6(a, b, c, p) > 0.0;
}
\end{lstlisting}

The algorithm works because it respects the boundary between mesh topology and coordinate geometry.

\section{Why Avoid Inheritance?}

\subsection{1. It Violates the Domain Model}

A raw coordinate point does not naturally have outgoing arcs, incident edges, labels, traversal marks, or graph metadata.

A graph vertex does not always have coordinates. In a pure graph, a vertex may represent a word, task, server, city, dependency, application, security control, or abstract state.

\subsection{2. It Breaks Algorithm Boundaries}

A geometric predicate should not need graph fields:

\begin{lstlisting}[caption={Geometry should remain geometry}]
double area2(Point2 a, Point2 b, Point2 c);
\end{lstlisting}

A graph traversal should not need coordinate fields:

\begin{lstlisting}[caption={Graph traversal should remain topology-driven}]
for (Arc *a = u->arcs; a != NULL; a = a->next) {
    Vertex *v = a->tip;
}
\end{lstlisting}

Mixing these concerns makes both algorithm families harder to reuse.

\subsection{3. It Causes Unnecessary Memory Coupling}

If every \texttt{Vertex} carries coordinate fields, pure graph programs pay for unused geometry data.

If every \texttt{Point} carries graph fields, pure geometry programs pay for unused adjacency and traversal data.

\subsection{4. Composition Handles Hybrid Cases Cleanly}

When topology needs geometry, embed or reference the point as a field:

\begin{lstlisting}[caption={The preferred pattern for a geometric graph node}]
struct GeoVertex {
    Point2 position;
    ArcList adjacency;
};
\end{lstlisting}

For mesh kernels, use the same idea:

\begin{lstlisting}[caption={The preferred pattern for a 3D mesh vertex}]
struct Vertex3 {
    Point3 p;
    Edge3 *incident_edges;
    Face3 *incident_faces;
};
\end{lstlisting}

This gives you a clean \textbf{has-a} relationship without pretending that one concept is a subtype of the other.

\section{Decision Table}

\begin{center}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.25\textwidth} >{\raggedright\arraybackslash}p{0.30\textwidth} >{\raggedright\arraybackslash}X}
\toprule
\textbf{Use Case} & \textbf{Recommended Type Design} & \textbf{Reason} \\
\midrule
Pure computational geometry & \texttt{Point2}, \texttt{Point3}, \texttt{Segment2}, \texttt{Polygon} & Algorithms depend on coordinates and predicates, not graph adjacency. \\
Pure graph analysis & \texttt{Graph}, \texttt{Vertex}, \texttt{Arc} & Algorithms depend on connectivity, weights, labels, and traversal state. \\
Pathfinding on a map & \texttt{GeoVertex} with \texttt{Point2 position} and adjacency list & The vertex has both a spatial location and neighbors. \\
Graph drawing & \texttt{Vertex} plus optional layout position & Layout is geometry attached to topology; it should be replaceable. \\
3D mesh processing & \texttt{Vertex3} has \texttt{Point3}; edges/faces reference vertices & Mesh topology and coordinate geometry are related but not identical. \\
Game engine navigation mesh & Vertices own/refer to points; faces/edges store topology & Geometry supports spatial queries; topology supports traversal. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Implementation Checklist}

\begin{checkbox}
Use this checklist when designing a C data model for geometry, graphs, or meshes.
\begin{enumerate}
  \item Define \texttt{Point2} and \texttt{Point3} as coordinate value types.
  \item Keep geometric predicates independent of graph fields.
  \item Define graph \texttt{Vertex} and \texttt{Arc} structures around adjacency and traversal needs.
  \item Add \texttt{Point2 position} or \texttt{Point3 p} only when the topological object needs physical location.
  \item Avoid pretending that embedding means subtype inheritance.
  \item Keep algorithm-specific metadata localized to the structure that actually needs it.
  \item Write conversion or adapter functions when a graph algorithm needs positions for cost or heuristic calculations.
\end{enumerate}
\end{checkbox}

\section{Final Rule}

Use \texttt{Point2} and \texttt{Point3} for geometric coordinates and predicates.

Use \texttt{Vertex}, \texttt{Arc}, \texttt{Edge}, and \texttt{Face} for graph or mesh topology.

Use composition when topology needs geometry:

\begin{decisionbox}
\begin{center}
\texttt{Vertex HAS-A Point} \\[2pt]
\texttt{Point is not a Vertex} \\[2pt]
\texttt{Vertex is not necessarily a Point}
\end{center}
\end{decisionbox}

\section*{Referenced Input Handouts}

This handout was prepared from the design guidance developed in the conversation and aligned with the uploaded source handouts:

\begin{itemize}
  \item \texttt{stanford-graphbase-algorithms-pseudocode-handout.tex}
  \item \texttt{computational-geometry-in-c-algorithms-pseudocode-handout.tex}
\end{itemize}

\end{document}
