\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=0.75in]{geometry}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{verbatim}

\definecolor{HeadingBlue}{HTML}{17365D}
\definecolor{RuleBlue}{HTML}{2F5597}
\definecolor{LightBlue}{HTML}{EAF2F8}
\definecolor{SoftGray}{HTML}{F4F6F7}
\definecolor{DarkGray}{HTML}{333333}

\hypersetup{
  colorlinks=true,
  linkcolor=HeadingBlue,
  urlcolor=RuleBlue,
  pdftitle={DongTai IAST Handout},
  pdfauthor={Prepared with ChatGPT},
  pdfsubject={Open-source Interactive Application Security Testing overview},
  pdfkeywords={DongTai, IAST, AppSec, DevSecOps, Java, Python, Docker Compose, Kubernetes}
}

\setlength{\headheight}{14pt}
\pagestyle{fancy}
\fancyhf{}
\lhead{DongTai IAST Handout}
\rhead{AppSec Reference}
\cfoot{Page \thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.2pt}

\titleformat{\section}{\Large\bfseries\color{HeadingBlue}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries\color{HeadingBlue}}{}{0em}{}
\titleformat{\subsubsection}{\bfseries\color{DarkGray}}{}{0em}{}

\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\setlist[itemize]{topsep=3pt,itemsep=2pt,leftmargin=1.2em}
\setlist[enumerate]{topsep=3pt,itemsep=2pt,leftmargin=1.4em}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\emergencystretch=2em

\newcommand{\keyterm}[1]{\textbf{\color{HeadingBlue}#1}}
\newcommand{\risk}[1]{\textbf{\color{DarkGray}#1}}

\begin{document}

\begin{center}
  {\Huge\bfseries\color{HeadingBlue} DongTai IAST}\\[4pt]
  {\Large Open-Source Runtime Security Testing Handout}\\[8pt]
  {\normalsize Prepared for AppSec, DevSecOps, and engineering implementation planning}\\[4pt]
  {\normalsize May 6, 2026}
\end{center}

\vspace{0.3em}
\colorbox{LightBlue}{%
\begin{minipage}{0.97\linewidth}
\textbf{Purpose.} This handout summarizes DongTai IAST as an open-source Interactive Application Security Testing capability for instrumented application testing. It explains what DongTai does, where it fits in a DevSecOps program, how its major components work together, and how to pilot it safely in a CI/CD or pre-release testing workflow.
\end{minipage}}

\section{Executive Summary}

\keyterm{DongTai IAST} is an open-source Interactive Application Security Testing tool that observes application behavior at runtime through passive instrumentation. Its value is strongest during functional, integration, QA, and security regression testing, where the application is actively exercised and the IAST agent can observe request handling, method invocation, data flow, third-party component usage, and security-sensitive execution paths.

For an enterprise AppSec program, DongTai should be treated as a \keyterm{runtime evidence source}, not a replacement for SAST, SCA, DAST, threat modeling, or manual security review. The practical goal is to add exploitability and runtime-path context to the software security toolchain.

\begin{center}
\begin{tabularx}{0.96\linewidth}{L{0.26\linewidth}Y}
\toprule
\textbf{Question} & \textbf{Answer} \\
\midrule
What does it test? & Running applications and observed execution paths during test activity. \\
Where does it fit? & Test-stage security validation, security regression testing, and release-readiness review. \\
What does it complement? & SAST, SCA, DAST, secret scanning, dependency scanning, and manual AppSec review. \\
What makes results useful? & High-quality functional coverage, representative test data, correct agent configuration, and disciplined triage. \\
Primary caution & IAST can only observe code paths that are executed while the agent is attached. \\
\bottomrule
\end{tabularx}
\end{center}

\section{What DongTai Provides}

DongTai provides runtime application security visibility by attaching agents to applications and reporting observed behavior to the DongTai server platform. The uploaded repository snapshot positions DongTai as particularly suitable for the testing phase of a development pipeline and describes DevSecOps use cases such as automated vulnerability detection, third-party component review, and pre-release security testing.

\subsection{Core Capabilities}

\begin{itemize}
  \item Runtime vulnerability detection through application instrumentation.
  \item Request-chain and method-call observation during application execution.
  \item Vulnerability analysis using reported runtime data and taint-tracking logic.
  \item Third-party component visibility and component-risk analysis.
  \item Web-based review of vulnerabilities, affected projects, agents, and supporting evidence.
  \item Localized deployment options using Docker Compose or Kubernetes.
  \item Support for agent-based testing workflows, with the supplied context identifying Java and Python agent usage.
\end{itemize}

\section{Where DongTai Fits in the AppSec Toolchain}

\begin{center}
\begin{tabularx}{\linewidth}{L{0.18\linewidth}L{0.31\linewidth}Y}
\toprule
\textbf{Control} & \textbf{Primary Function} & \textbf{DongTai Relationship} \\
\midrule
SAST & Finds source-level or bytecode-level weaknesses before runtime. & DongTai complements SAST by adding runtime evidence for executed paths. \\
SCA & Identifies vulnerable dependencies and license/component risk. & DongTai can contribute component visibility observed during runtime testing. \\
DAST & Tests externally exposed application behavior from the outside. & DongTai complements DAST by showing internal call paths and runtime context. \\
IAST & Observes application behavior from inside the runtime during tests. & DongTai belongs in this layer. \\
CI/CD Policy & Enforces release criteria and records evidence. & DongTai findings can inform gates, advisories, or manual review checkpoints. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Recommended Control Positioning}

Use DongTai as a \keyterm{test-stage runtime security control}. It is best suited for environments where the application can be deployed, instrumented, and exercised under realistic test scenarios before release.

Recommended initial posture:
\begin{enumerate}
  \item Run DongTai in a controlled non-production environment.
  \item Attach the agent only to test workloads or designated pilot applications.
  \item Run functional and integration test suites while the agent is active.
  \item Review high-severity findings manually before using results as hard release gates.
  \item Track validated findings in the organization's vulnerability management or ticketing workflow.
\end{enumerate}

\section{Reference Architecture}

The repository describes DongTai as a multi-service platform. The following table translates those components into an AppSec implementation view.

\begin{center}
\begin{tabularx}{\linewidth}{L{0.24\linewidth}Y}
\toprule
\textbf{Component} & \textbf{Operational Role} \\
\midrule
\texttt{DongTai-web} & User-facing product interface for reviewing projects, agents, vulnerabilities, and security results. \\
\texttt{dongtai\_web} & API layer for web interactions, user actions, projects, vulnerabilities, and management operations. \\
\texttt{dongtai\_protocol} & Agent communication layer. It handles registration, heartbeat, method data, third-party component data, error logs, hook strategy delivery, and probe-control commands. \\
\texttt{dongtai\_engine} & Detection and processing layer. It analyzes reported runtime data and uses taint-tracking logic to determine whether requests contain vulnerabilities. \\
\texttt{agent} & Runtime probe installed with the target application. It collects runtime data and reports it to DongTai services. \\
\texttt{deploy} & Deployment assets for Docker Compose and Kubernetes installation patterns. \\
\texttt{DongTai-Base-Image} & Runtime dependency layer that includes foundational services such as MySQL and Redis. \\
\texttt{DongTai-Plugin-IDEA} & IntelliJ IDEA integration path for Java probe usage and local developer testing. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Standard Usage Workflow}

\subsection{1. Deploy the DongTai Platform}

For a local pilot, use Docker Compose when the objective is fast evaluation. Use Kubernetes when the objective is a more production-like shared test service.

Repository-provided Docker Compose startup pattern:

\begin{verbatim}
git clone git@github.com:HXSecurity/DongTai.git
cd DongTai
chmod u+x build_with_docker_compose.sh
./build_with_docker_compose.sh
\end{verbatim}

The deployment assets include components for the web interface, server/API layer, engine workers, MySQL, Redis, and supporting configuration.

\subsection{2. Attach the Agent to the Application}

DongTai depends on agent instrumentation. The agent must be connected to the running application so that runtime behavior can be reported back to DongTai.

\begin{center}
\begin{tabularx}{\linewidth}{L{0.21\linewidth}Y}
\toprule
\textbf{Runtime} & \textbf{Instrumentation Pattern} \\
\midrule
Java & Attach the Java agent to the JVM with the \texttt{-javaagent} option and configure it to communicate with the DongTai service. \\
Python & Install and configure the Python agent in the application runtime according to the project guidance, then execute tests while the agent is active. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{3. Exercise the Application}

IAST findings depend heavily on exercised code paths. Run tests that cover authentication, authorization, input validation, API routes, database interactions, file handling, outbound service calls, and security-sensitive business workflows.

\subsection{4. Review Findings}

Review DongTai results in the web interface. At minimum, triage should capture:

\begin{itemize}
  \item affected application, project, and version;
  \item vulnerability type and severity;
  \item request path or call chain evidence;
  \item affected component or dependency, if applicable;
  \item proof that the path was exercised during testing;
  \item remediation owner and target date;
  \item final disposition: true positive, false positive, duplicate, accepted risk, or remediated.
\end{itemize}

\section{CI/CD Integration Pattern}

A practical CI/CD pattern is to run DongTai during a test-stage job rather than during source-only build validation.

\begin{enumerate}
  \item Build the application artifact or container image.
  \item Deploy the application to an ephemeral or shared test environment.
  \item Start the application with the DongTai agent attached.
  \item Run functional, integration, and security regression tests.
  \item Pull or review DongTai findings after tests complete.
  \item Fail, warn, or route for manual review based on the pilot policy.
  \item Create or update remediation tickets for validated findings.
\end{enumerate}

\subsection{Suggested Gate Maturity Model}

\begin{center}
\begin{tabularx}{\linewidth}{L{0.18\linewidth}L{0.22\linewidth}Y}
\toprule
\textbf{Phase} & \textbf{Gate Mode} & \textbf{Recommended Use} \\
\midrule
Pilot & Advisory only & Collect findings, validate signal quality, tune instrumentation, and identify noisy patterns. \\
Stabilization & Manual review required & Route high-risk findings to AppSec before release but do not automatically block all releases. \\
Controlled gating & Conditional block & Block releases only for validated high-confidence, high-risk categories with clear remediation expectations. \\
Operationalized & Risk-based policy & Integrate with formal vulnerability SLAs, exception handling, evidence retention, and release governance. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Triage Checklist}

Use the following checklist before treating a DongTai finding as release-blocking evidence.

\begin{itemize}
  \item \risk{Coverage confirmed:} the vulnerable route or workflow was actually executed during testing.
  \item \risk{Environment confirmed:} the finding came from the intended application, branch, build, and test environment.
  \item \risk{Agent confirmed:} the correct agent version and configuration were active during the test run.
  \item \risk{Path evidence reviewed:} the request chain, method path, source, sink, and taint path are understandable.
  \item \risk{Severity validated:} severity aligns with business impact, exploitability, exposure, and compensating controls.
  \item \risk{Duplicate checked:} the issue is not already tracked through SAST, SCA, DAST, or prior IAST results.
  \item \risk{Owner assigned:} remediation has a named engineering owner or service team.
  \item \risk{Disposition recorded:} the final decision and evidence are stored in the vulnerability management workflow.
\end{itemize}

\section{Strengths and Limitations}

\begin{center}
\begin{tabularx}{\linewidth}{Y Y}
\toprule
\textbf{Strengths} & \textbf{Limitations} \\
\midrule
Provides runtime evidence tied to real application execution. & Cannot detect paths that are not executed during testing. \\
Can help reduce uncertainty by showing request chains and method-level context. & Requires correct agent setup and compatible runtime conditions. \\
Fits naturally into functional, integration, QA, and regression testing stages. & Findings still require AppSec triage before strict release enforcement. \\
Complements SAST, SCA, and DAST rather than duplicating their exact roles. & Test environments must be managed carefully to avoid noisy, stale, or misattributed results. \\
Can support DevSecOps workflows and pre-release security validation. & Production deployment requires additional governance and performance review. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Implementation Guardrails}

\subsection{Environment Guardrails}

\begin{itemize}
  \item Start in non-production environments.
  \item Use a known application version and a repeatable test suite.
  \item Separate pilot data from production vulnerability metrics until the process is stable.
  \item Ensure application owners understand that IAST requires executed test coverage.
\end{itemize}

\subsection{Security and Governance Guardrails}

\begin{itemize}
  \item Treat DongTai as a source of security evidence, not as the system of record by default.
  \item Avoid hard-blocking releases until severity mapping, false-positive handling, and exception workflows are mature.
  \item Retain enough evidence to support auditability: build identifier, environment, test run, agent configuration, finding details, triage decision, and remediation status.
  \item Integrate validated findings into the organization's existing vulnerability management process.
\end{itemize}

\section{Pilot Success Criteria}

A DongTai pilot should be considered successful when the team can demonstrate the following outcomes:

\begin{enumerate}
  \item DongTai deploys repeatably using Docker Compose or Kubernetes.
  \item At least one Java or Python application is instrumented successfully.
  \item Functional tests generate observable runtime telemetry in DongTai.
  \item The AppSec team can review request chains and vulnerability details in the web interface.
  \item Findings can be dispositioned consistently using a documented triage workflow.
  \item Validated findings can be routed to engineering through the normal remediation process.
  \item The team has enough data to decide whether to continue with advisory-only use, manual review, or conditional gating.
\end{enumerate}

\section{Source References}

This handout was prepared from the provided DongTai summary and uploaded HXSecurity/DongTai repository snapshot. For operational use, validate commands, versions, and agent configuration against the current upstream documentation before implementation.

\begin{itemize}
  \item DongTai repository: \url{https://github.com/HXSecurity/DongTai}
  \item DongTai documentation: \url{https://docs.dongtai.io/}
  \item DongTai Java agent: \url{https://github.com/HXSecurity/DongTai-agent-java}
  \item DongTai Python agent: \url{https://github.com/HXSecurity/DongTai-agent-python}
  \item DongTai discussions: \url{https://github.com/HXSecurity/DongTai/discussions}
\end{itemize}

\end{document}



