\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.6em}
\setlength{\emergencystretch}{3em}
\raggedbottom
\hbadness=10000

\title{Using ISO/IEC/IEEE 42010:2022, TOGAF, and SEI Views \& Beyond Together\\\large A Layered Strategy for Architecture Documentation}
\author{}
\date{May 7, 2026}

\begin{document}
\maketitle
\thispagestyle{empty}

\section*{Executive Summary}
ISO/IEC/IEEE 42010:2022, TOGAF, and SEI's \emph{Views and Beyond} (V\&B) address different parts of the architecture documentation problem. ISO/IEC/IEEE 42010 defines the conceptual model and required content for an architecture description, including stakeholders, concerns, viewpoints, views, models, and correspondence. TOGAF provides an enterprise architecture method and governance structure centered on the Architecture Development Method (ADM). SEI V\&B provides concrete, software-focused documentation guidance organized around views and templates. Used together, they support a layered documentation strategy: ISO 42010 supplies the rules of the architecture description, TOGAF supplies the enterprise process and governance context, and V\&B supplies the software-level documentation patterns and view packages. This combined pattern is a practical synthesis of complementary sources; it is not a single prescribed method from any one authority.\cite{iso42010cm,iso42010ads,togafstd,sei_vb_factsheet,sei_vb_book}

\section*{What Each Contributes}
\begin{table}[h!]
\small
\centering
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth} >{\raggedright\arraybackslash}p{0.28\textwidth} X}
\toprule
\textbf{Source} & \textbf{Primary Contribution} & \textbf{Role in a Combined Documentation Strategy} \\
\midrule
ISO/IEC/IEEE 42010:2022 & Architecture description concepts and requirements & Establishes the common vocabulary and required structure for the architecture description: stakeholders, concerns, viewpoints, views, model kinds, rationale, and correspondence among architectural elements.\cite{iso42010cm,iso42010ads,iso42010gettingstarted} \\
\addlinespace
TOGAF & Enterprise architecture method and governance & Provides the enterprise-level operating model for planning, developing, governing, and evolving architecture work through the ADM and related guidance.\cite{togafstd,togafadm,togafdownloads} \\
\addlinespace
SEI Views \& Beyond & Prescriptive software architecture documentation guidance & Provides practical documentation patterns, view selection guidance, and reusable templates for software architecture description packages that can satisfy software-layer stakeholder concerns.\cite{sei_vb_factsheet,sei_vb_book} \\
\bottomrule
\end{tabularx}
\end{table}

\section*{How the Three Fit Together}
\begin{enumerate}[leftmargin=1.5em]
    \item \textbf{Start with ISO 42010.} Identify the system of interest, stakeholders, and concerns. Define the viewpoints that will frame those concerns and the model kinds each viewpoint permits. Also define correspondence rules so related architectural elements can be traced across views and models.\cite{iso42010cm,iso42010gettingstarted,iso42010faq}
    \item \textbf{Use TOGAF to organize the work.} Place the architecture effort inside an enterprise method. TOGAF supplies a repeatable process model and governance context for architecture work, including architecture development across business, data, application, and technology domains.\cite{togafstd,togafadm}
    \item \textbf{Use SEI V\&B where software detail is needed.} When a stakeholder concern requires software-architecture depth, use V\&B to produce the concrete documentation package: view catalog, primary views, supporting views, interface information, rationale, and view-specific templates.\cite{sei_vb_factsheet,sei_vb_book}
    \item \textbf{Connect enterprise intent to software reality.} Use ISO 42010 correspondence to link business and enterprise concerns captured through TOGAF to the software views documented with V\&B. This is the control point that keeps detailed software architecture aligned with enterprise architecture intent.\cite{iso42010cm,iso42010faq}
    \item \textbf{Govern change over time.} Maintain the architecture description as a governed asset. TOGAF supplies the governance cadence, ISO 42010 supplies structural consistency, and V\&B supplies stable documentation conventions for teams building and maintaining software.\cite{togafstd,iso42010ads,sei_vb_book}
\end{enumerate}

\section*{Recommended Division of Labor}
\begin{table}[h!]
\small
\centering
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.27\textwidth} >{\raggedright\arraybackslash}p{0.23\textwidth} X}
\toprule
\textbf{Question} & \textbf{Primary Guide} & \textbf{Typical Output} \\
\midrule
What must an architecture description contain? & ISO 42010 & Stakeholders, concerns, viewpoints, views, model kinds, rationale, and correspondence definitions.\cite{iso42010cm,iso42010ads} \\
\addlinespace
When should architecture artifacts be created, reviewed, and governed? & TOGAF & ADM-aligned work products, review points, governance checkpoints, and cross-domain architecture outputs.\cite{togafstd,togafadm} \\
\addlinespace
How should software architecture actually be documented for builders, maintainers, and reviewers? & SEI V\&B & Module views, component-and-connector views, allocation views, interface documentation, and view packages using consistent templates.\cite{sei_vb_factsheet,sei_vb_book} \\
\bottomrule
\end{tabularx}
\end{table}

\section*{Practical Application Pattern}
A practical combined approach is to maintain three related layers of documentation:
\begin{enumerate}[leftmargin=1.5em]
    \item \textbf{Enterprise architecture layer:} TOGAF-aligned architecture work products and governance records across business, data, application, and technology domains.\cite{togafstd,togafadm}
    \item \textbf{Architecture description control layer:} An ISO 42010-conformant description that defines stakeholders, concerns, viewpoints, model kinds, and correspondence rules for the system of interest.\cite{iso42010cm,iso42010ads}
    \item \textbf{Software architecture layer:} SEI V\&B view packages for the application or solution, with detailed module, component-and-connector, and allocation views as needed.\cite{sei_vb_factsheet,sei_vb_book}
\end{enumerate}

This structure prevents two common failures: using TOGAF alone as if it were detailed software documentation, and producing software architecture views with no explicit link to stakeholder concerns or enterprise governance. The three together close both gaps.\cite{togafstd,iso42010cm,sei_vb_book}

\section*{Working Rule of Thumb}
Use \textbf{ISO 42010} as the \emph{architecture description standard}, \textbf{TOGAF} as the \emph{enterprise method and governance framework}, and \textbf{SEI V\&B} as the \emph{software architecture documentation method}. If a team treats them this way, the result is usually clearer architecture intent, better traceability from business concern to software design, and more maintainable architecture documentation.\cite{iso42010ads,togafstd,sei_vb_factsheet}

\section*{Implementation Notes}
\begin{itemize}[leftmargin=1.5em]
    \item Do not treat the three as interchangeable. Their scopes are different and complementary.\cite{iso42010ads,togafstd,sei_vb_book}
    \item Define viewpoints before producing large numbers of diagrams. ISO 42010 is explicit that viewpoints frame concerns and specify the model kinds used in views.\cite{iso42010gettingstarted,iso42010cm}
    \item Use TOGAF to decide \emph{when} and \emph{why} architecture artifacts are needed, but use V\&B to decide \emph{how} detailed software architecture views should be written and organized.\cite{togafstd,sei_vb_book}
    \item Keep rationale and correspondence explicit. This is what makes the documentation reviewable and governable instead of becoming a disconnected diagram set.\cite{iso42010cm,iso42010faq,sei_vb_book}
\end{itemize}

\section*{Bottom Line}
The strongest combined strategy is not to blend ISO 42010, TOGAF, and SEI V\&B into a single indistinguishable method. It is to assign each one a precise role. ISO 42010 defines the architecture description discipline, TOGAF frames enterprise architecture work and governance, and SEI V\&B delivers the software architecture documentation practice needed for implementation teams. Together, they create a coherent path from stakeholder concern to governed enterprise architecture to actionable software documentation.\cite{iso42010cm,iso42010ads,togafstd,sei_vb_factsheet,sei_vb_book}

\begin{thebibliography}{9}
\bibitem{iso42010cm}
ISO/IEC/IEEE 42010 Conceptual Model, \emph{iso-architecture.org}. \url{https://www.iso-architecture.org/42010/cm/}. Accessed May 7, 2026.

\bibitem{iso42010ads}
ISO/IEC/IEEE 42010: Architecture Descriptions, \emph{iso-architecture.org}. \url{https://www.iso-architecture.org/ieee-1471/ads/}. Accessed May 7, 2026.

\bibitem{iso42010gettingstarted}
Getting Started with ISO/IEC/IEEE 42010, \emph{iso-architecture.org}. \url{https://www.iso-architecture.org/ieee-1471/getting-started.html}. Accessed May 7, 2026.

\bibitem{iso42010faq}
ISO/IEC/IEEE 42010 Frequently Asked Questions, \emph{iso-architecture.org}. \url{https://www.iso-architecture.org/ieee-1471/faq.html}. Accessed May 7, 2026.

\bibitem{togafstd}
The TOGAF Standard, \emph{The Open Group}. \url{https://www.opengroup.org/togaf}. Accessed May 7, 2026.

\bibitem{togafadm}
Introduction to the Architecture Development Method (ADM), \emph{The Open Group}. \url{https://www.opengroup.org/public/arch/p2/p2_intro.htm}. Accessed May 7, 2026.

\bibitem{togafdownloads}
TOGAF Standard, 10th Edition Downloads, \emph{The Open Group}. \url{https://www.opengroup.org/togaf-standard-10th-edition-downloads}. Accessed May 7, 2026.

\bibitem{sei_vb_factsheet}
Views and Beyond: The SEI Approach for Architecture Documentation, \emph{Software Engineering Institute}. \url{https://www.sei.cmu.edu/library/views-and-beyond-the-sei-approach-for-architecture-documentation/}. Accessed May 7, 2026.

\bibitem{sei_vb_book}
Paul C. Clements, Felix Bachmann, Len Bass, David Garlan, James Ivers, Reed Little, Paulo Merson, Robert Nord, and Judith Stafford, \emph{Documenting Software Architectures: Views and Beyond, 2nd Edition}, Addison-Wesley Professional, 2010. SEI library page: \url{https://www.sei.cmu.edu/library/documenting-software-architectures-views-and-beyond-second-edition/}. Accessed May 7, 2026.
\end{thebibliography}

\end{document}



