\documentclass[11pt,letterpaper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{xurl}
\usepackage{hyperref}

\hypersetup{
  hidelinks,
  pdftitle={File Naming Conventions: A Practical Handout},
  pdfauthor={OpenAI},
  pdfsubject={File naming best practices},
  pdfcreator={LaTeX}
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.6em}
\setlist[itemize]{leftmargin=1.5em,itemsep=0.35em,topsep=0.35em}
\renewcommand{\arraystretch}{1.2}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\best}[1]{\texttt{#1}}

\begin{document}

\begin{center}
  {\LARGE \textbf{File Naming Conventions}}\\[0.35em]
  {\large A Practical Handout for Consistent, Sortable, and System-Safe Filenames}\\[0.75em]
  {\normalsize May 9, 2026}
\end{center}

\vspace{0.5em}

\textbf{Purpose}\newline
Well-designed file names make documents easier to sort, search, share, preserve, and manage over time. Across academic, records-management, and library guidance, the same core principles recur: use a consistent structure, place the most useful sorting term first, prefer ISO-style dates, avoid problematic characters, and apply explicit versioning rather than ambiguous labels such as \code{final} or \code{latest}.\cite{harvard,nara,purdue,illinois}

\textbf{Recommended baseline pattern}\newline
For most professional, academic, and project-based work, a strong default convention is:

\begin{center}
  \fbox{\best{YYYYMMDD\_Project\_Description\_v01.ext}}
\end{center}

This pattern works well because it supports chronological sorting, communicates subject matter quickly, and distinguishes draft history in a predictable way.\cite{harvard,nara,purdue}

\section*{Core Best Practices}

\begin{itemize}
  \item \textbf{Start with an ISO-style date when time matters.} Use \code{YYYYMMDD} or \code{YYYY-MM-DD} so filenames sort chronologically in standard file explorers and operating systems.\cite{harvard,illinois}
  \item \textbf{Use a consistent delimiter.} Choose either underscores or hyphens and apply the choice uniformly. Consistency is more important than the specific delimiter selected.\cite{harvard,purdue}
  \item \textbf{Avoid spaces and special characters.} Prefer letters, numbers, hyphens, and underscores. Avoid characters such as \code{/ \\ : * ? " < > | \# \% \& \{ \}} because they can create compatibility, portability, or command-line handling issues.\cite{nara,purdue}
  \item \textbf{Be descriptive, but keep names concise.} Include enough context to identify the file without opening it, but avoid long, cluttered names. Records-management guidance commonly recommends relatively short names for readability and interoperability.\cite{nara,harvard}
  \item \textbf{Use explicit version numbers.} Prefer \code{v01}, \code{v02}, and \code{v03} over vague endings such as \code{final}, \code{final2}, or \code{latest}. Structured versioning improves traceability and reduces overwrite risk.\cite{purdue,harvard}
  \item \textbf{Put the highest-value sorting term first.} When date is the main retrieval key, place it first. When topic or client is more important, design the convention accordingly and apply it consistently across the collection.\cite{harvard,nara}
\end{itemize}

\section*{Why This Convention Works}

\begin{tabularx}{\textwidth}{>{\bfseries}p{1.7in}X}
\toprule
Benefit & Practical Effect \\
\midrule
Chronological sorting & ISO-style dates cause files to sort naturally by year, month, and day. \\
Fast identification & Short descriptive labels let users recognize contents without opening the file. \\
Version control & Explicit version numbers clarify draft sequence and review history. \\
System compatibility & Safe characters reduce issues across operating systems, cloud platforms, scripts, and archives. \\
Long-term consistency & Repeated patterns make large folders easier to manage and audit over time. \\
\bottomrule
\end{tabularx}

\section*{Examples}

\begin{tabularx}{\textwidth}{>{\bfseries}p{1.25in}X}
\toprule
Good examples & \best{20260509\_ghas\_review\_queue\_v01.pdf}\\
& \best{20260509\_api\_style\_guide\_v02.tex}\\
& \best{20260509\_uml\_classification\_v01.puml}\\
& \best{2026-05-09\_client\_invoice\_v03.xlsx}\\
\midrule
Avoid & \best{final version really final.pdf}\\
& \best{My File (new).docx}\\
& \best{report!!.pdf}\\
& \best{latest\_notes.docx}\\
\bottomrule
\end{tabularx}

\section*{A Simple Naming Checklist}

Before saving a file, verify the following:

\begin{itemize}
  \item The filename follows a documented pattern.
  \item The most important retrieval key appears first.
  \item The date, if used, is in \code{YYYYMMDD} or \code{YYYY-MM-DD} format.
  \item The name uses only safe characters.
  \item The subject label is clear but not overly long.
  \item The version number is explicit and sequential.
  \item The filename can still be understood by someone else six months later.
\end{itemize}

\section*{Implementation Note}

A single universal pattern is not required for every environment. The important point is to adopt a convention that matches retrieval needs and then apply it consistently across a project, team, or repository. For example, archival collections may prioritize date-first naming, while client-service work may prioritize client or project identifiers before the date.\cite{harvard,nara}

{\small\sloppy
\begin{thebibliography}{9}

\bibitem{harvard}
Harvard Medical School Countway Library.
\newblock \emph{File Naming Conventions}.
\newblock Available at: \url{https://datamanagement.hms.harvard.edu/plan-design/file-naming-conventions}

\bibitem{nara}
U.S. National Archives and Records Administration.
\newblock \emph{Best Practices for File Naming}.
\newblock Available at: \url{https://records-express.blogs.archives.gov/2017/08/22/best-practices-for-file-naming/}

\bibitem{purdue}
Purdue University Libraries and School of Information Studies.
\newblock \emph{File Naming Conventions}.
\newblock Available at: \url{https://guides.lib.purdue.edu/c.php?g=353013&p=2378293}

\bibitem{illinois}
Illinois State University.
\newblock \emph{Naming Conventions for Documents}.
\newblock Available at: \url{https://designsystem.illinoisstate.edu/guidelines/naming-conventions-for-documents/}

\end{thebibliography}
}

\end{document}
