\documentclass[11pt]{article}

% ---------- Common, stable packages ----------
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{amsmath}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% ---------- Helpers ----------
\setlist[itemize]{itemsep=2pt, topsep=4pt}
\setlist[enumerate]{itemsep=4pt, topsep=6pt}
\newcommand{\Section}[1]{\vspace{0.6em}\noindent\textbf{\Large #1}\par\vspace{0.25em}}
\newcommand{\Subsection}[1]{\vspace{0.4em}\noindent\textbf{\large #1}\par\vspace{0.2em}}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\begin{document}

\begin{center}
  {\LARGE \textbf{Chicken Tortilla Soup}}\\[4pt]
  \small By Natalya Drozhzhin \quad\textbullet\quad Updated: May 24, 2024 \quad\textbullet\quad 1{,}062 comments\\
  \small Rating: 4.99 from 1883 votes
\end{center}

\Section{Overview}
Comforting, hearty, one-pot chicken tortilla soup loaded with shredded chicken, sweet corn, black beans, and a zesty tomato--chicken broth. Finish with crispy tortilla strips, avocado, cilantro, and lime.

\Section{At a Glance}
\begin{tabularx}{\textwidth}{@{} l L @{}}
\toprule
\textbf{Yield} & 6 servings \\
\textbf{Prep Time} & 15 minutes \\
\textbf{Cook Time} & 30 minutes \\
\textbf{Total Time} & 45 minutes \\
\textbf{Method} & One pot; stove top \\
\bottomrule
\end{tabularx}

\Section{Ingredients}

\Subsection{Chicken Tortilla Soup}
\begin{itemize}
  \item 1 tablespoon olive oil
  \item 1 medium onion, chopped
  \item 3 garlic cloves, minced
  \item 1 jalape\~no pepper, seeded and diced
  \item 1 teaspoon ground cumin
  \item 1 teaspoon chili powder
  \item 1 lb chicken breasts (about 2 medium), whole
  \item 1 (28 oz) can crushed tomatoes
  \item 32 oz chicken broth (preferably low-sodium; bone broth optional)
  \item 1 (14 oz) can black beans, drained and rinsed
  \item 1 (14 oz) can corn, drained and rinsed
  \item $\tfrac{1}{2}$ cup cilantro, chopped, divided (reserve $\tfrac{1}{4}$ of it for garnish)
  \item 1 lime, juiced
  \item 1 teaspoon salt, or to taste
\end{itemize}

\Subsection{Homemade Tortilla Strips}
\begin{itemize}
  \item $\tfrac{1}{4}$ cup olive oil (for frying)
  \item 8 corn tortillas (6-inch), cut into thin strips
\end{itemize}

\Subsection{Toppings (optional but recommended)}
\begin{itemize}
  \item 1 large avocado, diced
  \item Lime wedges, for serving
  \item Extra cilantro, chopped
  \item Sour cream; shredded cheddar; sliced jalape\~no
\end{itemize}

\Section{Equipment}
Large soup pot or Dutch oven (with lid); large skillet or baking sheet (for tortilla strips); paper towels; tongs; ladle.

\Section{Instructions}

\Subsection{Crispy Tortilla Strips}
\textbf{Pan-Fry:}
\begin{enumerate}
  \item Heat $\tfrac{1}{4}$ cup oil in a skillet over medium-high heat. Fry tortilla strips in batches until crisp. Transfer to a paper-towel-lined plate; season lightly with salt. Repeat, adding oil as needed.
\end{enumerate}
\textbf{Bake (alternative):}
\begin{enumerate}
  \item Toss strips with 2 tablespoons oil. Arrange in a single layer on a baking sheet and bake at $350^\circ$F (175$^\circ$C) for about 15 minutes, until crisp.
\end{enumerate}

\Subsection{Chicken Tortilla Soup}
\begin{enumerate}
  \item \textbf{Saut\'e aromatics.} In a large pot over medium-high heat, heat 1 tablespoon oil. Add onion, garlic, and jalape\~no; saut\'e until tender.
  \item \textbf{Build the soup.} Add whole chicken breasts, corn, beans, chili powder, cumin, crushed tomatoes, salt, $\tfrac{1}{4}$ cup chopped cilantro, and chicken broth. Bring to a boil, then reduce to a simmer; cook at least 25 minutes.
  \item \textbf{Shred chicken.} Remove chicken to a bowl and shred with two forks. Return shredded chicken to the pot; simmer 5 minutes more. Stir in lime juice and adjust salt to taste.
  \item \textbf{Serve.} Ladle into bowls and top with tortilla strips, avocado, fresh cilantro, and lime wedges. Add sour cream, cheese, and jalape\~no if desired.
\end{enumerate}

\Section{Tips}
\begin{itemize}
  \item \textbf{Leftover chicken:} Rotisserie or pre-cooked chicken works well; add shredded and simmer briefly.
  \item \textbf{Broth matters:} Low-sodium chicken bone broth adds depth; adjust final salt to taste.
  \item \textbf{Heat control:} Add extra jalape\~no (or a pinch of cayenne) for more spice.
  \item \textbf{Pantry time-savers:} Canned corn and beans keep this weeknight-fast.
  \item \textbf{Avocado:} Dice just before serving to prevent browning.
\end{itemize}

\Section{Make-Ahead}
The broth deepens in flavor as it rests. Make a big batch and enjoy leftovers for lunch during the week.

\Section{Storage}
\begin{itemize}
  \item \textbf{Refrigerate:} Store soup (without toppings) airtight up to 3--4 days.
  \item \textbf{Freeze:} Up to 2 months in an airtight container. Thaw overnight in the fridge.
  \item \textbf{Reheat:} Bring to a gentle boil on the stovetop; add a splash of broth or water if needed.
\end{itemize}

\Section{Nutrition (Approximate, per serving; 6 servings)}
\begin{tabular}{@{}ll@{}}
\toprule
Calories & 511 kcal \\
Carbohydrates & 54 g \\
Protein & 30 g \\
Fat & 22 g \\
\bottomrule
\end{tabular}

\vfill
\begin{center}
  \footnotesize
  \textit{Serving suggestion: Pair with soft dinner rolls or extra lime wedges for brightness.}
\end{center}

\end{document}

