\documentclass[11pt]{article}

% --- Page + typography ---
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{amsmath} % <-- added for \tfrac

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% --- Section spacing ---
\titlespacing*{\section}{0pt}{8pt plus 2pt}{6pt}
\titlespacing*{\subsection}{0pt}{6pt}{4pt}

% --- Simple helpers ---
\newcommand{\RecipeTitle}{The Best Chili Recipe}
\newcommand{\RecipeAuthor}{Holly Nilsson}
\newcommand{\RecipeDate}{July 10, 2020}
\newcommand{\RecipeYield}{8 servings}
\newcommand{\PrepTime}{20 minutes}
\newcommand{\CookTime}{45--60 minutes}
\newcommand{\TotalTime}{1 hour 5 minutes}
\newcommand{\Rating}{4.96/5 (2{,}892 votes)}

\setlist[itemize]{itemsep=4pt, topsep=4pt, leftmargin=1.2em}
\setlist[enumerate]{itemsep=6pt, topsep=6pt, leftmargin=1.4em}

\begin{document}
\begin{center}
  {\LARGE \bfseries \RecipeTitle}\\[6pt]
  {\small by \RecipeAuthor\quad|\quad \RecipeDate}
\end{center}

\vspace{0.75em}

\noindent\textbf{At a Glance}\par
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
\textbf{Yield} & \RecipeYield \\
\textbf{Prep Time} & \PrepTime \\
\textbf{Cook Time} & \CookTime \\
\textbf{Total Time} & \TotalTime \\
\textbf{Rating} & \Rating \\
\bottomrule
\end{tabularx}

\section*{Overview}
A hearty, weeknight-friendly pot of chili loaded with ground beef and beans, gently simmered with chili powder and cumin. Serve with your favorite toppings and a side of cornbread, buttered toast, or buttermilk biscuits.

\section*{Ingredients}
\textit{Note: The chili powder is used in two additions (``divided'').}
\begin{itemize}
  \item 2 pounds lean ground beef
  \item 1 onion, diced
  \item 1 jalape\~no, seeded and finely diced
  \item 4 cloves garlic, minced
  \item 2\,$\tfrac{1}{2}$ tablespoons chili powder, \emph{divided}, or to taste
  \item 1 teaspoon ground cumin
  \item 1 green bell pepper, seeded and diced
  \item 14.5 ounces crushed tomatoes (1 can)
  \item 19 ounces canned red kidney beans, drained and rinsed
  \item 14.5 ounces canned diced tomatoes (with juices)
  \item 1\,$\tfrac{1}{2}$ cups beef broth
  \item 1 cup beer (optional; see Variations)
  \item 1 tablespoon tomato paste
  \item 1 tablespoon brown sugar (optional)
  \item Kosher salt and freshly ground black pepper, to taste
\end{itemize}

\section*{Instructions}
\begin{enumerate}
  \item \textbf{Season \& brown.} In a bowl, combine the ground beef with 1\,$\tfrac{1}{2}$ tablespoons of the chili powder. In a large heavy pot over medium--high heat, brown the seasoned beef with the onion, jalape\~no, and garlic until the beef is no longer pink. Drain excess fat.
  \item \textbf{Simmer.} Stir in the remaining chili powder, cumin, bell pepper, crushed tomatoes, kidney beans, diced tomatoes (with juices), beef broth, beer (if using), tomato paste, brown sugar (if using), and a pinch of salt and pepper. Bring to a boil, then reduce heat and simmer \emph{uncovered} for 45--60 minutes, stirring occasionally, until thickened to your liking.
  \item \textbf{Taste \& serve.} Adjust seasoning with additional salt, pepper, or chili powder as desired. Ladle into bowls and finish with your favorite toppings.
\end{enumerate}

\section*{Suggested Toppings \& Sides}
Shredded cheddar, sliced green onions, chopped cilantro, diced onion, sour cream, sliced jalape\~nos, hot sauce, crushed tortilla chips. Serve with cornbread, buttermilk biscuits, or buttered toast.

\section*{Variations \& Tips}
\begin{itemize}
  \item \textbf{Heat level:} For more heat, keep the jalape\~no seeds or add chili flakes/hot sauce. For milder chili, use half a jalape\~no or omit.
  \item \textbf{Protein swaps:} Ground turkey or chicken work well; drain fat if needed.
  \item \textbf{Beans:} Red kidney beans are classic; pinto or black beans are excellent alternatives. If using \emph{chili beans} (seasoned and sauced), you may reduce added salt.
  \item \textbf{Beer:} Adds depth; feel free to substitute with additional beef broth.
  \item \textbf{Spice profile:} For a Tex--Mex vibe, add a packet of taco seasoning or extra cumin, paprika, and oregano.
\end{itemize}

\section*{How to Thicken}
Simmering uncovered naturally reduces and thickens the chili (preferred). If short on time, add a small sprinkle of cornmeal or a quick slurry (1 tablespoon cornstarch or flour whisked with cold water) and simmer a few minutes more.

\section*{Freezing \& Reheating}
Chili freezes and reheats beautifully. Cool completely, portion into airtight containers or freezer bags, and freeze up to 3 months. Thaw overnight in the refrigerator and reheat gently on the stovetop or in the microwave, adding a splash of broth or water if needed.

\section*{Ingredient Scaling (Quick Reference)}
\noindent\textit{Multiply amounts as needed. ``To taste'' items remain as desired.}
\begin{tabularx}{\linewidth}{@{}lccc@{}}
\toprule
\textbf{Ingredient} & \textbf{1\,$\times$} & \textbf{2\,$\times$} & \textbf{3\,$\times$}\\
\midrule
Lean ground beef & 2 lb & 4 lb & 6 lb \\
Onion (diced) & 1 & 2 & 3 \\
Jalape\~no (finely diced) & 1 & 2 & 3 \\
Garlic (cloves, minced) & 4 & 8 & 12 \\
Chili powder (divided) & 2.5 tbsp & 5 tbsp & 7.5 tbsp \\
Ground cumin & 1 tsp & 2 tsp & 3 tsp \\
Green bell pepper (diced) & 1 & 2 & 3 \\
Crushed tomatoes & 14.5 oz & 29 oz & 43.5 oz \\
Kidney beans (rinsed) & 19 oz & 38 oz & 57 oz \\
Diced tomatoes (with juices) & 14.5 oz & 29 oz & 43.5 oz \\
Beef broth & 1.5 cups & 3 cups & 4.5 cups \\
Beer (optional) & 1 cup & 2 cups & 3 cups \\
Tomato paste & 1 tbsp & 2 tbsp & 3 tbsp \\
Brown sugar (optional) & 1 tbsp & 2 tbsp & 3 tbsp \\
Salt \& black pepper & to taste & to taste & to taste \\
\bottomrule
\end{tabularx}

\vfill
\noindent{\small\textbf{Disclosure:} This recipe text is adapted from the user-provided source. Mentions of affiliate links in the original have been retained here as a general disclosure note.}

\end{document}
