\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[table]{xcolor}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage{ragged2e}
\usepackage{titlesec}
\usepackage{parskip}
\usepackage{graphicx}

\definecolor{SBABlue}{HTML}{163A63}
\definecolor{SBALight}{HTML}{EFF4F9}
\definecolor{HeaderFill}{HTML}{D9E6F2}

\pagestyle{fancy}
\fancyhf{}
\lhead{SBA Support Intake Summary}
\rhead{ZIP 30168}
\cfoot{\thepage}
\setlength{\headheight}{14pt}

\titleformat{\section}{\large\bfseries\color{SBABlue}}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries\color{SBABlue}}{}{0em}{}

\renewcommand{\arraystretch}{1.22}
\newcolumntype{Y}{>{\RaggedRight\arraybackslash}X}
\newcolumntype{L}[1]{>{\RaggedRight\arraybackslash}p{#1}}

\begin{document}

\begin{center}
{\LARGE\bfseries SBA Local Assistance Intake Summary}\\[3pt]
{\large ZIP Code 30168}\\[3pt]
\small Prepared from the live SBA local assistance finder and linked official office pages.
\end{center}

\rowcolors{2}{white}{SBALight}
\begin{tabularx}{\textwidth}{L{0.33\textwidth}Y}
\rowcolor{HeaderFill}
\textbf{Field} & \textbf{Summary} \\
Best Support Option & The University of Georgia SBDC at Kennesaw State University \\
Support Type & Small Business Development Center (SBDC) \\
Organization & University of Georgia Small Business Development Center at Kennesaw State University \\
Address & 3333 Busbee Drive, Suite 500, Kennesaw, GA 30144 \\
Phone & (470) 578-6450 \\
Website & \href{https://georgiasbdc.org/locations/kennesaw-office/}{Georgia SBDC Kennesaw office page} \\
Primary Contact Person (if listed; otherwise say ``Not listed'') & Drew Tonsmeire, Area Director. For first outreach, the practical intake path is the office line or \href{mailto:kennesaw@georgiasbdc.org}{kennesaw@georgiasbdc.org}. \\
Best Time to Contact & Recommendation: weekdays from 9:00 AM to 11:30 AM ET or 1:30 PM to 4:00 PM ET. The office page lists phone and email but does not publish office hours on the page used here, so this is a practical contact window. \\
Why This Is My Best First Option & The live SBA finder for ZIP 30168 shows only two local SBA Resource Partners in Cobb County, both in Kennesaw. For a general startup founder, the SBDC is the strongest first stop because SBA describes SBDCs as providers of counseling and training for business start-up and expansion, and the Kennesaw office has direct intake details plus a full consulting team. \\
First Question to Ask & ``I am an early-stage founder in Cobb County. Based on my business model and current stage, what are the first three actions I should take in the next 30 days, and which SBDC service, advisor, or workshop should I start with right now?'' \\
\end{tabularx}

\section*{Top 3 Nearby Options}

{\scriptsize
\setlength{\tabcolsep}{4pt}
\resizebox{\textwidth}{!}{%
\begin{tabular}{L{0.06\textwidth}L{0.20\textwidth}L{0.18\textwidth}L{0.18\textwidth}L{0.22\textwidth}L{0.20\textwidth}}
\rowcolor{HeaderFill}
\textbf{Rank} & \textbf{Organization} & \textbf{Support Type} & \textbf{Distance/Location Clue} & \textbf{Contact Info} & \textbf{Why It Matters} \\
1 & UGA SBDC at Kennesaw State University & Small Business Development Center & Kennesaw, GA 30144; one of only two local partner results shown for Cobb County & Phone: (470) 578-6450\\ Email: \href{mailto:kennesaw@georgiasbdc.org}{kennesaw@georgiasbdc.org}\\ Site: \href{https://georgiasbdc.org/locations/kennesaw-office/}{Kennesaw office} & Best all-purpose first stop for startup guidance, planning, and growth support. \\
2 & Greater Wealth Works Women's Business Center & Women's Business Center & Kennesaw, GA 30144; the other local partner result shown for Cobb County & Phone: (770) 694-6593\\ Email: \href{mailto:wbc@greaterwealthworks.org}{wbc@greaterwealthworks.org}\\ Site: \href{https://greaterwealthworks.org/wbc/}{WBC page} & Strong option for one-on-one consultations, training, and Tuesday evening availability. \\
3 & SBA Georgia District Office - Atlanta & SBA District Office & Atlanta, GA 30303; district office shown on the same finder page & Phone: (404) 331-0100\\ Site: \href{https://www.sba.gov/district/georgia}{Georgia District page} & Best escalation path if you need routing, broader SBA program guidance, or help choosing the right partner. \\
\end{tabular}%
}
}

\section*{Recommended First Outreach Script}

\subsection*{Short phone script}
Hello, my name is Jordan Suber. I am an early-stage founder in Cobb County and I found your office through the SBA local assistance finder for ZIP 30168. I would like help figuring out the right first steps for my startup and which SBA-backed counseling, training, or advisor support I should begin with. Could I schedule an initial conversation, and could you tell me the first three actions you recommend I take in the next 30 days?

\subsection*{Short email draft}
\textbf{Subject:} Request for startup guidance and first-step appointment

Hello,

I am an early-stage founder in Cobb County and found your office through the SBA local assistance finder for ZIP 30168. I am looking for practical startup guidance on the right first steps to take, the best counseling or training resources to start with, and whether there is an advisor or workshop you recommend for someone at my stage.

My main question is: what are the first three actions I should take in the next 30 days based on my business model and current stage?

Please let me know the best way to schedule an initial conversation. Thank you.

Jordan Suber

\section*{Notes on ranking}
The live SBA local assistance finder page for ZIP 30168 displays two SBA Resource Partners in Cobb County: Greater Wealth Works Women's Business Center and the University of Georgia SBDC at Kennesaw State University. The same page also displays the Atlanta district office. The result cards do not publish numeric mileage, so the ranking above uses the visible local location clues together with startup-founder relevance.

\section*{Sources used}
\begin{itemize}[leftmargin=1.25em]
\item \href{https://www.sba.gov/local-assistance/find?zip=30168&type=All}{SBA Local Assistance Finder for ZIP 30168}
\item \href{https://georgiasbdc.org/locations/kennesaw-office/}{Georgia SBDC Kennesaw office page}
\item \href{https://greaterwealthworks.org/wbc/}{Greater Wealth Works Women's Business Center page}
\item \href{https://www.sba.gov/district/georgia}{SBA Georgia District page}
\end{itemize}

\end{document}
