% -------------------------------------------------------------------------
% File: src/security/standards/cis/AutomatingCISBenchmarks.tex
% Style: latex-docs-style v1.0 (see tooling/latex/latex-docs-style.sty)
% Build: pdflatex via latexmk; minted-aware with listings fallback.
% -------------------------------------------------------------------------
\documentclass[12pt]{article}
\usepackage{ltechnical-design-spec}
\geometry{a4paper, margin=1in}
\usepackage{listings}

% Define colors for code examples.
\definecolor{codebg}{RGB}{245,245,245}
\definecolor{commentcolor}{rgb}{0,0.5,0}
\definecolor{keywordcolor}{rgb}{0,0,0.8}

\lstdefinestyle{mystyle}{
  backgroundcolor=\color{codebg},
  basicstyle=\ttfamily\small,
  commentstyle=\color{commentcolor},
  keywordstyle=\color{keywordcolor},
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  numbersep=10pt,
  breaklines=true,
  breakatwhitespace=true,
  showstringspaces=false,
  frame=single,
}

\lstdefinelanguage{Rego}{
  morekeywords={package, deny, some, in, input},
  sensitive=true,
  morecomment=[l]{\#},
  morestring=[b]"}

\lstdefinelanguage{yaml}{
  keywords={true,false,null,y,n},
  keywordstyle=\color{blue},
  basicstyle=\small\ttfamily,
  sensitive=false,
  comment=[l]{\#},
  morestring=[b]',
  morestring=[b]"
}

\lstdefinelanguage{HCL}{
  keywords={resource,provider,variable,output,module,locals,data},
  keywordstyle=\color{blue},
  basicstyle=\small\ttfamily,
  sensitive=true,
  comment=[l]{\#},
  morestring=[b]"
}

\lstdefinelanguage{Docker}{
  keywords={FROM, RUN, CMD, LABEL, EXPOSE, ENV, ADD, COPY, ENTRYPOINT, VOLUME, USER, WORKDIR, ARG, ONBUILD, STOPSIGNAL, HEALTHCHECK, SHELL},
  sensitive=true,
  morecomment=[l]{\#},
  morestring=[b]",
  morestring=[b]'
}

\lstset{style=mystyle}

\title{Organizing and Automating CIS Benchmarks: A Strategy for Automated Security Hardening}
\author{Your Name}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This document outlines a comprehensive strategy to organize CIS benchmarks and automate security hardening for system configurations and network infrastructure. The proposed solution leverages open source tools such as Open Policy Agent (OPA), Cloud Custodian, Terraform, and Docker, and integrates these solutions into a CI/CD pipeline. This strategy includes categorization of benchmarks, translation of recommendations into code, version-controlled repository structure, and continuous compliance monitoring.
\end{abstract}

\section{Introduction}
Organizations are increasingly moving towards a "compliance as code" approach to security. With the availability of numerous CIS benchmarks for diverse technologies, a well-organized repository and an automated pipeline are vital to enforce and maintain secure configurations. This document describes how to:
\begin{itemize}
    \item Categorize the CIS benchmarks by cloud provider, operating system, and technology.
    \item Translate CIS recommendations into machine-readable policies and infrastructure as code.
    \item Set up a structured repository and integrate automation into CI/CD pipelines.
    \item Establish governance with traceability and version control.
\end{itemize}

\section{Organizing the CIS Benchmarks}
\subsection{Categorize by Technology and Domain}
Organize the benchmarks into folders based on cloud providers and platforms. For example:
\begin{itemize}
    \item \textbf{Cloud Providers:}
    \begin{itemize}
        \item AWS (e.g., Compute Services, Database Services, Foundations)
        \item Microsoft Azure (e.g., Compute, Storage, Database, Foundations)
    \end{itemize}
    \item \textbf{Operating Systems:} Linux (Debian, Red Hat, etc.)
    \item \textbf{Application Servers and Containers:} Apache, Tomcat, NGINX, Docker, Kubernetes
    \item \textbf{Developer Platforms:} GitHub Benchmark, etc.
    \item \textbf{DNS and Infrastructure:} ISC BIND DNS Server
\end{itemize}

Create a mapping document (e.g., \texttt{benchmarks\_mapping.md}) that details:
\begin{itemize}
    \item Recommendation identifiers/titles.
    \item Brief descriptions.
    \item Corresponding automation code (e.g., OPA policies, Terraform modules, Cloud Custodian rules).
\end{itemize}

\subsection{Repository Structure}
A suggested repository structure is as follows:
\begin{lstlisting}[language={}, caption={Repository Structure}]
/CIS-benchmarks
    /pdfs
        /AWS
            CIS_AWS_Compute_Services_Benchmark_v1.1.0.pdf
            CIS_AWS_Database_Services_Benchmark_v1.0.0.pdf
            CIS_Amazon_Web_Services_Foundations_Benchmark_v4.0.1.pdf
        /Azure
            CIS_Microsoft_Azure_Storage_Services_Benchmark_v1.0.0.pdf
            CIS_Microsoft_Azure_Database_Services_Benchmark_v1.0.0.pdf
            CIS_Microsoft_Azure_Foundations_Benchmark_v4.0.0.pdf
        /Linux
            CIS_Debian_Linux_12_Benchmark_v1.1.0.pdf
            CIS_Debian_Linux_11_Benchmark_v1.0.0_ARCHIVE.pdf
        /Other
            CIS_ISC_BIND_DNS_Server_9.11_Benchmark_v1.0.0_ARCHIVE.pdf
            ...
    /mapping
        benchmarks_mapping.md
    /automation
        /opa
            aws.rego
            azure.rego
            linux.rego
        /terraform
            /modules
                aws_ec2_hardened
                aws_vpc_security
                ...
        /cloud_custodian
            aws_s3_policies.yml
            aws_ec2_policies.yml
        /docker
            Dockerfile
            compliance_scanner.sh
        /ci-cd
            pipeline_config.yml
    README.md
\end{lstlisting}

\section{Translating CIS Recommendations into Code}
Below are examples that demonstrate the conversion of CIS recommendations into automated code.

\subsection{Open Policy Agent (OPA)}
Develop OPA policies in Rego that validate configurations against CIS guidelines. For example, to check that no AWS security groups allow unrestricted access:
\begin{lstlisting}[language=Rego, caption={CIS Security Group Check in Rego}]
package cis.aws

deny[msg] {
  some rule in input.aws_ec2.security_group_rules
  rule == "0.0.0.0/0"
  msg = "Security group rule allowing access from anywhere is not allowed."
}
\end{lstlisting}

\subsection{Cloud Custodian}
Write Cloud Custodian policies to enforce and monitor compliance. For example, to ensure that S3 buckets do not have public access:
\begin{lstlisting}[language=YAML, caption={Cloud Custodian Policy for S3 Compliance}]
policies:
  - name: aws-s3-no-public-access
    resource: aws.s3
    filters:
      - type: value
        key: "PublicAccessBlockConfiguration.BlockPublicAcls"
        value: false
    actions:
      - type: mark-for-op
        op: delete
        days: 7
\end{lstlisting}

\subsection{Terraform Modules}
Create Terraform modules to provision resources with CIS-compliant configurations. For example, a module for securing an EC2 instance:
\begin{lstlisting}[language=HCL, caption={Terraform Module Example for a Hardened EC2 Instance}]
variable "ami_id" {}
variable "instance_type" {}

resource "aws_instance" "secure_ec2" {
  ami           = var.ami_id
  instance_type = var.instance_type

  vpc_security_group_ids = [aws_security_group.sg.id]
  tags = {
    Name = "CIS_Hardened_EC2"
  }

  # Additional hardened configuration according to CIS benchmarks.
}

resource "aws_security_group" "sg" {
  name        = "cis_hardened_sg"
  description = "Security group for CIS hardened EC2 instance"
  vpc_id      = var.vpc_id

  ingress {
    from_port   = 22
    to_port     = 22
    protocol    = "tcp"
    cidr_blocks = ["10.0.0.0/16"]
  }

  egress {
    from_port   = 0
    to_port     = 0
    protocol    = "-1"
    cidr_blocks = ["0.0.0.0/0"]
  }
}
\end{lstlisting}

\subsection{Docker}
Build Docker images that include compliance scanning tools. For example, a Dockerfile to run a compliance script:
\begin{lstlisting}[language=Docker, caption={Dockerfile for Compliance Scanner}]
FROM alpine:latest

RUN apk add --no-cache bash ca-certificates

# Copy the compliance scanner script into the container
COPY compliance_scanner.sh /usr/local/bin/
RUN chmod +x /usr/local/bin/compliance_scanner.sh

CMD ["/usr/local/bin/compliance_scanner.sh"]
\end{lstlisting}

\subsection{CI/CD Pipeline Integration}
Integrate the above code into a CI/CD pipeline using open source tools. For example, a GitHub Actions YAML file:
\begin{lstlisting}[language=YAML, caption={GitHub Actions Pipeline for CIS Compliance}]
name: CI/CD Pipeline

on:
  push:
    branches: [ main ]
  pull_request:
    branches: [ main ]

jobs:
  terraform:
    runs-on: ubuntu-latest
    steps:
      - uses: actions/checkout@v2
      - name: Setup Terraform
        uses: hashicorp/setup-terraform@v1
      - name: Terraform Init
        run: terraform init
      - name: Terraform Plan
        run: terraform plan

  opa_test:
    runs-on: ubuntu-latest
    steps:
      - uses: actions/checkout@v2
      - name: Run OPA tests
        run: opa test ./automation/opa

  cloud_custodian:
    runs-on: ubuntu-latest
    steps:
      - uses: actions/checkout@v2
      - name: Run Cloud Custodian policies
        run: custodian run -s . automation/cloud_custodian/aws_s3_policies.yml
\end{lstlisting}

\section{Governance and Traceability}
\subsection{Mapping and Traceability Document}
Develop a mapping document (e.g., \texttt{benchmarks\_mapping.md}) that:
\begin{itemize}
    \item Lists each CIS recommendation with an identifier.
    \item Provides a summary of the required configuration.
    \item Maps the recommendation to the corresponding automation tool/code snippet.
\end{itemize}

\subsection{Versioning and Audit}
Use Git for version control to track changes to automation code and policies. Additionally, use periodic automated audits (via tools such as CIS-CAT Pro Assessor or your open source alternatives) to validate that the deployed system is in compliance with the CIS benchmarks.

\section{Conclusion}
Organizing your CIS benchmarks into a structured repository and converting them into machine-readable policies and infrastructure code allows you to automate the hardening process. By using Open Policy Agent, Cloud Custodian, Terraform, Docker, and integrating these into a CI/CD pipeline, you create a continuous compliance system that is both efficient and auditable. This strategy ensures your systems meet rigorous security guidelines and that any changes are automatically validated and enforced.

\bigskip

This approach is supported by the detailed guidance provided in benchmarks such as the CIS AWS Compute Services Benchmark~\cite{turn2file0} and the associated AWS Foundations Benchmark~\cite{turn2file2}.

\bibliographystyle{plain}
\bibliography{references}

\end{document}



