\documentclass[11pt]{article}

% ---------- Stable packages ----------
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{hyperref}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue
}

% ---------- Helpers ----------
\newcommand{\Section}[1]{\vspace{0.75em}\noindent{\Large\bfseries #1}\par\vspace{0.3em}}
\newcommand{\Subsection}[1]{\vspace{0.4em}\noindent{\large\bfseries #1}\par\vspace{0.15em}}
\setlist[itemize]{itemsep=3pt, topsep=4pt}
\setlist[enumerate]{itemsep=5pt, topsep=6pt}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}

\begin{document}

\begin{center}
  {\LARGE \textbf{Grilled Shrimp Kabobs}}\\[2pt]
  \small By Imma Adamu \quad\textbullet\quad Published: April 14, 2025 \quad\textbullet\quad Updated: May 17, 2025
\end{center}

\Section{Overview}
Juicy shrimp marinated with herbs, spices, olive oil, and lemon, then skewered with colorful veggies and grilled to smoky perfection. Quick to prep, fast to cook, and easy to customize—perfect for busy weeknights or a backyard cookout. Grill outdoors or use a grill pan indoors.

\Section{At a Glance}
\begin{tabularx}{\textwidth}{@{} l L @{}}
\toprule
\textbf{Yield} & Serves 6 (about 6 kabobs)\\
\textbf{Prep Time} & 20 minutes \\
\textbf{Cook Time} & 6 minutes \\
\textbf{Total Time} & 26 minutes \\
\textbf{Method} & Grill (outdoor or grill pan) \\
\textbf{Cuisine} & American \\
\textbf{Equipment} & Skewers (metal or soaked wooden), grill or grill pan, mixing bowls \\
\textbf{Nutrition (approx.)} & \(\sim\)270 kcal per kabob (see Nutrition) \\
\bottomrule
\end{tabularx}

\Section{Ingredients}

\Subsection{Marinated Shrimp}
\begin{itemize}
  \item 2 lb (950 g) large shrimp, peeled and deveined
  \item 2 tsp fresh thyme (leaves)
  \item 1--2 tsp fresh oregano (or \(1/2\) tsp dried)
  \item 1 tsp freshly ground black pepper
  \item 1 tsp ground cumin
  \item 1 Tbsp (15 g) minced garlic (about 3 cloves)
  \item 1 tsp red pepper flakes
  \item \(1/2\) tsp Italian seasoning
  \item 3--4 Tbsp (45--60 ml) olive oil
  \item 3--4 Tbsp (45--60 ml) lemon juice
  \item Salt, to taste
\end{itemize}

\Subsection{Veggie Options (mix \& match)}
\begin{itemize}
  \item 2 bell peppers, cut into large chunks
  \item 1 red onion, quartered lengthwise, then halved crosswise; separate layers
  \item 1 zucchini, cut into thick slices
  \item 1 yellow squash, peeled and cut into thick slices
  \item (Optional) Grape tomatoes, button mushrooms, small corn rounds, pineapple chunks
\end{itemize}

\Section{Method}

\Subsection{Prep Skewers \& Grill}
\begin{enumerate}
  \item If using wooden skewers, soak in water 20--30 minutes to prevent scorching. (Skip if using metal skewers.)
  \item Preheat an outdoor grill (gas or charcoal) to medium–high. For indoor cooking, preheat a lightly oiled grill pan over medium–high.
\end{enumerate}

\Subsection{Make Marinade}
\begin{enumerate}
  \item In a medium bowl, whisk together thyme, oregano, black pepper, cumin, garlic, red pepper flakes, Italian seasoning, olive oil, and lemon juice.
\end{enumerate}

\Subsection{Season Shrimp \& Veggies}
\begin{enumerate}
  \item Place shrimp in a large bowl, season with salt, and pour most of the marinade over them; toss to coat. Reserve a few tablespoons of marinade for the vegetables.
  \item Toss vegetables with the reserved marinade; salt to taste.
\end{enumerate}

\Subsection{Skewer \& Grill}
\begin{enumerate}
  \item Thread shrimp and vegetables onto skewers, alternating and covering most of each skewer.
  \item Grill 2--3 minutes per side until shrimp are opaque and just cooked through and vegetables are tender with light char.
  \item Serve hot with salad, rice, or creamy orzo.
\end{enumerate}

\Section{Notes \& Tips}
\begin{itemize}
  \item \textbf{Marinade timing:} 10--30 minutes is ideal for shrimp; extended marinating (hours/overnight) can make shrimp mealy.
  \item \textbf{Variations:} Swap in teriyaki, jerk, or honey–hot marinades; add lemon zest; splash in your favorite hot sauce for extra heat.
  \item \textbf{Vegan option:} Replace shrimp with meaty mushrooms (portobello, oyster, shiitake) and proceed as directed.
  \item \textbf{Hawaiian twist:} Add pineapple chunks to skewers; grill until caramelized.
  \item \textbf{Frozen shrimp:} Works fine—ideally thawed; if grilling from frozen, cook gently and check doneness frequently. Ensure deveined.
  \item \textbf{Beer soak (optional):} A 10-minute beer soak imparts malty notes; do not exceed or shrimp can toughen.
\end{itemize}

\Section{Leftovers \& Storage}
\begin{itemize}
  \item \textbf{Refrigerate:} Store grilled shrimp and veggies airtight up to 3--4 days.
  \item \textbf{Freeze:} Up to 2--3 months; thaw overnight in the refrigerator.
  \item \textbf{Leftover makeovers:}
    \begin{itemize}
      \item Chop and fold into day-old rice for quick fried rice.
      \item Top a green salad with chilled grilled shrimp.
      \item Warm in chicken broth for a light soup.
      \item Quesadillas or tacos: melt Monterey Jack in a tortilla, add chopped shrimp/veggies, cook until the cheese is gooey.
    \end{itemize}
\end{itemize}

\Section{Serving Ideas}
Cilantro-lime rice, simple garden salad, grilled corn, orzo with lemon and herbs, or crusty bread. A squeeze of fresh lemon over the finished kabobs brightens everything.

\Section{Nutrition (Approximate)}
Per serving (about 1 kabob): \textbf{270} kcal; \textbf{Carbohydrates} 25 g; \textbf{Protein} 24 g; \textbf{Fat} 10 g; \textbf{Saturated Fat} 1 g; \textbf{Cholesterol} 191 mg; \textbf{Sodium} 1269 mg; \textbf{Fiber} 5 g; \textbf{Sugar} 6 g; \textbf{Vitamin A} 15,225 IU; \textbf{Vitamin C} 91 mg; \textbf{Calcium} 199 mg; \textbf{Iron} 3 mg.\\
\emph{Estimates vary with exact ingredients, marinade amounts, and skewer size.}

\vfill
\begin{center}
  \small \emph{For indoor cooking, maintain good ventilation; avoid overcooking to keep shrimp juicy.}
\end{center}

\end{document}

