\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{xurl}
\usepackage[hidelinks]{hyperref}
\usepackage{titlesec}
\usepackage{array}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.75em}
\titleformat{\section}{\large\bfseries}{}{0pt}{}
\definecolor{accent}{HTML}{1F4E79}
\newcommand{\handouttitle}[1]{%
  {\Huge\bfseries\color{accent} #1}\par
  \vspace{0.5em}
  {\large Professional Reference Handout}\par
  \vspace{0.75em}
  \hrule
  \vspace{1em}
}

\begin{document}
\sloppy

\handouttitle{Key Uses of Letterhead}

\textbf{Definition.} Letterhead is specialized stationery that presents an organization's logo, name, address, and contact details in a consistent, branded format. It is used to strengthen professionalism, credibility, and brand recognition in official correspondence.

\section*{Why Letterhead Matters}
Letterhead helps an organization present formal communication in a clear and recognizable way. It signals legitimacy, supports brand consistency, and gives routine documents a more official and trustworthy appearance.

\section*{Key Uses of Letterhead}
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Professional correspondence:} Adds a formal, branded appearance to letters sent to clients, partners, vendors, and other stakeholders.
    \item \textbf{Branding and identification:} Makes the sender immediately recognizable through the use of the company name, logo, and visual identity.
    \item \textbf{Legal and official documents:} Supports a more official presentation for contracts, proposals, quotations, and invoices.
    \item \textbf{Internal communication:} Gives additional importance to internal memos, notices, and formal announcements.
    \item \textbf{Credibility and trust:} Reinforces legitimacy and professionalism, which can improve how recipients perceive the organization.
    \item \textbf{Marketing value:} Functions as a recurring brand touchpoint each time the organization sends written communication.
\end{itemize}

\section*{Common Elements on Letterhead}
A standard letterhead commonly includes the following elements:
\begin{itemize}[leftmargin=1.5em]
    \item Company logo
    \item Company name
    \item Physical or mailing address
    \item Contact information, such as phone number, email address, and website
\end{itemize}

\section*{Typical Formats}
Letterhead can be used in both physical and digital formats:
\begin{itemize}[leftmargin=1.5em]
    \item \textbf{Printed letterhead:} Used for hard-copy letters and signed documents.
    \item \textbf{Digital letterhead template:} Used in word processors, PDF templates, and electronic correspondence.
\end{itemize}

\section*{Practical Guidance}
Use letterhead for communication that represents the organization in an official capacity. It is most appropriate when the document needs to look formal, credible, and clearly tied to the business or institution issuing it.

\section*{Source List}
\begin{enumerate}[leftmargin=1.75em]
    \item \url{https://www.theroyalstore.com/blog/what-is-letterhead/}
    \item \url{https://www.template.net/graphic-design/letterheads-designs/}
    \item \url{https://www.jezweb.com.au/importance-of-letterheads-to-your-business/}
    \item \url{https://www.quora.com/What-is-a-letterhead-How-does-it-differ-from-other-types-of-stationery}
    \item \url{https://westpress.com/creating-professional-letterhead/}
    \item \url{https://www.ibcscorp.com/blog/letterheads-unfolded/}
    \item \url{https://www.printsafari.com/pages/how-to/design-perfect-letterheads/}
    \item \url{https://kearneycopycat.com/2021/01/the-importance-of-letterhead/}
    \item \url{https://boldcolours.co.uk/blog/essentials-things-to-include-on-your-letterhead}
    \item \url{https://www.printsafari.com/pages/printwiki/importance-of-letterheads/}
    \item \url{https://help.printulu.co.za/en/articles/6557334-what-is-a-letterhead}
\end{enumerate}

\vfill
{\footnotesize Note: This handout is a cleaned and formatted restatement of the source material provided by the user.}

\end{document}



