\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{titlesec}
\usepackage{parskip}
\pagestyle{empty}

\definecolor{Accent}{HTML}{1F4E79}
\definecolor{LightGray}{HTML}{F3F5F7}
\definecolor{RuleGray}{HTML}{D9DEE3}

\titleformat{\section}{\large\bfseries\color{Accent}}{}{0pt}{}
\titleformat{\subsection}{\normalsize\bfseries\color{Accent}}{}{0pt}{}
\setlist[itemize]{leftmargin=1.4em,itemsep=0.35em,topsep=0.35em}
\renewcommand{\arraystretch}{1.25}

\begin{document}

{\LARGE\bfseries Professional Letterhead Checklist}\par
\vspace{0.4em}
{\large Review and approval checklist for business correspondence assets}\par
\vspace{0.8em}
\hrule height 1pt
\vspace{1em}

\section*{Purpose}
Use this checklist to verify that a company letterhead is complete, legible, brand-aligned, and ready for professional use in print or digital correspondence.

\section*{Required identity elements}
\begin{tabularx}{\textwidth}{>{\bfseries}p{0.33\textwidth} X >{\centering\arraybackslash}p{0.12\textwidth}}
\toprule
Item & Verification standard & Status \\
\midrule
Company logo & High-resolution mark or vector asset is present and clearly rendered. & \fbox{\rule{0pt}{1.2ex}\hspace{1.2ex}} \\
Business name & Full official business or organization name is shown exactly as approved. & \fbox{\rule{0pt}{1.2ex}\hspace{1.2ex}} \\
Address & Mailing or physical business address is present and current. & \fbox{\rule{0pt}{1.2ex}\hspace{1.2ex}} \\
Phone number & Primary business phone number is included and formatted consistently. & \fbox{\rule{0pt}{1.2ex}\hspace{1.2ex}} \\
Email address & Professional contact email is present and uses the company domain when applicable. & \fbox{\rule{0pt}{1.2ex}\hspace{1.2ex}} \\
Website URL & Official website is included and spelled correctly. & \fbox{\rule{0pt}{1.2ex}\hspace{1.2ex}} \\
\bottomrule
\end{tabularx}

\section*{Branding and design review}
\begin{itemize}
  \item Brand colors match the approved company palette.
  \item Fonts are professional, readable, and consistent with the organization's visual identity.
  \item Logo size is prominent enough for recognition without overpowering the page.
  \item Branding elements appear primarily in the header, footer, or both.
  \item Visual design is clean and does not crowd the letter body area.
  \item Alignment, spacing, and margins are consistent across the page.
  \item Printed and on-screen versions remain clear and easy to read.
\end{itemize}

\section*{Content quality checks}
\begin{itemize}
  \item All contact information is current and verified.
  \item Spelling, punctuation, and capitalization are correct.
  \item Telephone, email, and website formatting follow a consistent style.
  \item Optional tagline or slogan, if used, is short and brand-appropriate.
  \item Optional legal or registration details are present if required by jurisdiction or business type.
\end{itemize}

\section*{Layout best practices}
\begin{itemize}
  \item Sufficient white space remains for the body of the letter.
  \item Decorative elements do not interfere with readability.
  \item The design works on both first page and continuation pages, if a second-sheet format is required.
  \item Footer content, if used, is subtle and does not compete with the main message.
\end{itemize}

\section*{Optional enhancements}
\begin{itemize}
  \item Social media handles or icons are included only when relevant to business communication.
  \item A second-page variation has been prepared for longer correspondence.
  \item Separate print and digital export versions have been tested if needed.
\end{itemize}

\section*{Approval record}
\begin{tabularx}{\textwidth}{|p{0.31\textwidth}|X|}
\hline
Prepared by & \\
\hline
Reviewed by & \\
\hline
Approved by & \\
\hline
Version / date & \\
\hline
Notes & \\
\hline
\end{tabularx}

\end{document}
