\documentclass[10pt]{article}
\usepackage[margin=0.55in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{parskip}
\pagestyle{empty}

\definecolor{Accent}{HTML}{1F4E79}
\definecolor{SoftFill}{HTML}{F4F7FA}
\definecolor{SoftRule}{HTML}{D7DEE5}

\titleformat{\section}{\small\bfseries\color{Accent}}{}{0pt}{}
\titlespacing*{\section}{0pt}{0.45em}{0.2em}
\setlist[itemize]{leftmargin=1.1em,itemsep=0.08em,topsep=0.08em,parsep=0pt,partopsep=0pt}
\renewcommand{\arraystretch}{1.06}
\setlength{\parskip}{0.25em}

\begin{document}
\small

{\Large\bfseries Professional Letterhead One-Page Style Guide}\par
{\normalsize Visual and editorial standard for branded correspondence}\par
\vspace{0.35em}
\hrule height 1pt
\vspace{0.45em}

\begin{tabularx}{\textwidth}{>{\bfseries\raggedright\arraybackslash}p{0.21\textwidth} X}
Brand objective & Present a credible, recognizable, and uncluttered business identity while preserving space for the body of the letter. \\
Primary placement & Use the letterhead in the header, footer, or a restrained combination of both. Keep branding out of the main writing area. \\
Visual principle & Prioritize clarity, restraint, alignment, and consistency over decoration. \\
\end{tabularx}

\section*{Required content}
\begin{tabularx}{\textwidth}{>{\bfseries\raggedright\arraybackslash}p{0.23\textwidth} X}
\toprule
Element & Standard \\
\midrule
Logo & Approved high-resolution or vector logo, placed prominently. \\
Business name & Full official business name, shown exactly as approved. \\
Address & Current mailing or physical address. \\
Phone & Main business phone number in a consistent format. \\
Email & Professional business email address. \\
Website & Official website URL. \\
\bottomrule
\end{tabularx}

\section*{Optional content}
\begin{tabularx}{\textwidth}{>{\bfseries\raggedright\arraybackslash}p{0.23\textwidth} X}
\toprule
Element & Use rule \\
\midrule
Tagline & Include only if brief, approved, and visually restrained. \\
Social media & Include only when relevant to professional communication. \\
Legal details & Add when required by law, policy, or industry practice. \\
Second sheet & Use a simplified continuation-page treatment for long letters. \\
\bottomrule
\end{tabularx}

\section*{Typography and color}
\begin{itemize}
  \item Use official brand fonts or approved substitutes.
  \item Favor highly legible, print-friendly type.
  \item Apply brand colors sparingly and maintain strong contrast.
  \item Avoid decorative type, heavy backgrounds, and low-contrast combinations.
\end{itemize}

\section*{Layout rules}
\begin{itemize}
  \item Preserve generous white space for the body.
  \item Align logo, business name, and contact block to a clear grid.
  \item Keep the letterhead clear of dates, addresses, subject lines, and signatures.
  \item Keep footer content subtle and easy to scan.
\end{itemize}

\section*{Production standard}
\begin{itemize}
  \item Test in both digital PDF output and print.
  \item Verify all contact details before release.
  \item Use consistent assets across templates and export formats.
\end{itemize}

\vspace{0.25em}
\begin{center}
\fcolorbox{SoftRule}{SoftFill}{\parbox{0.95\textwidth}{\textbf{Working rule:} brand clearly, then get out of the way. The letterhead should identify the sender immediately without competing with the letter itself.}}
\end{center}

\end{document}
