\documentclass[10pt]{article}
\usepackage[letterpaper,margin=0.45in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{array}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{longtable}
\usepackage{ragged2e}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{tabularx}
\usepackage{colortbl}
\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{4pt}
\titleformat{\section}{\large\bfseries}{}{0pt}{}
\titlespacing*{\section}{0pt}{8pt}{4pt}
\setlist[itemize]{leftmargin=1.1em,itemsep=1pt,topsep=2pt,parsep=0pt,partopsep=0pt}
\definecolor{HeaderBlue}{HTML}{1F4E79}
\definecolor{RowGray}{HTML}{F6F8FA}
\definecolor{StatusGray}{HTML}{EAECEF}
\newcolumntype{L}[1]{>{\RaggedRight\arraybackslash}p{#1}}
\newcommand{\statusbox}[1]{\textbf{#1}}
\newcommand{\tightlist}[1]{\begin{itemize}#1\end{itemize}}
\begin{document}
\small

{\LARGE\bfseries FAST Animation Channels Startup Roadmap}\\[-1pt]
{\normalsize SBA-aligned founder handout for turning a six-channel FAST animation strategy into an operating company}

\vspace{6pt}
\hrule
\vspace{6pt}

This roadmap mirrors the current SBA ten-step startup sequence and tailors each step to FAST Animation Channels as an early-stage FAST animation venture. Step framing follows the live SBA page; business-specific interpretation and status calls are based on the FAST working file and prior chat context, with conservative treatment where legal, tax, licensing, and banking evidence is not yet established.

\setlength{\LTpre}{6pt}
\setlength{\LTpost}{6pt}
\renewcommand{\arraystretch}{1.14}
\setlength{\tabcolsep}{3pt}
\footnotesize

\begin{longtable}{@{}L{0.05\textwidth}L{0.16\textwidth}L{0.28\textwidth}L{0.10\textwidth}L{0.25\textwidth}@{}}
\rowcolor{HeaderBlue}
\color{white}\textbf{Step} & \color{white}\textbf{SBA Step Name} & \color{white}\textbf{FAST Animation Channels Interpretation} & \color{white}\textbf{Status} & \color{white}\textbf{Next Actions} \\
\endfirsthead
\rowcolor{HeaderBlue}
\color{white}\textbf{Step} & \color{white}\textbf{SBA Step Name} & \color{white}\textbf{FAST Animation Channels Interpretation} & \color{white}\textbf{Status} & \color{white}\textbf{Next Actions} \\
\endhead
\rowcolor{RowGray}
1 & Conduct market research & Validate demand with FAST distributors, aggregators, advertisers, and audience segments for the six-channel bouquet; benchmark competing animation and kids FAST offerings. & \statusbox{In Progress} & \tightlist{
\item Complete buyer and competitor map.
\item Pressure-test audience promise for each channel.
\item Select the strongest first one to two launch channels.
} \\
2 & Write your business plan & Convert the working file and programming bible into a company-level plan covering channel launch order, rights model, revenue assumptions, metadata operations, and pilot rollout. & \statusbox{In Progress} & \tightlist{
\item Publish a concise operating business plan.
\item Set 12-month launch economics and KPI targets.
\item Define the minimum viable operating model for pilot launch.
} \\
\rowcolor{RowGray}
3 & Fund your business & Decide how the startup will finance licensing, branding, metadata, QC, legal work, and distributor launch costs before ad revenue becomes dependable. & \statusbox{Not Started} & \tightlist{
\item Build a startup cost model for pilot launch.
\item Set minimum capital required for the first channels.
\item Choose founder-funded, partner-funded, or outside-capital path.
} \\
4 & Pick your business location & Choose the legal and operating base for entity formation, contracts, banking, and tax administration, even if channel operations remain digitally distributed. & \statusbox{Not Started} & \tightlist{
\item Select state and jurisdiction for formation.
\item Decide home-office versus registered-agent setup.
\item Check state-level tax and filing implications.
} \\
\rowcolor{RowGray}
5 & Choose a business structure & Select the entity form that best supports founder liability protection, media contracts, tax treatment, and possible future investor participation. & \statusbox{Not Started} & \tightlist{
\item Compare LLC versus corporation with professional advice.
\item Decide founder ownership and governance basics.
\item Prepare formation-ready entity information.
} \\
6 & Choose your business name & The umbrella name FAST Animation Channels is established in working materials, but parent-brand clearance and the six-channel naming architecture still need to be locked down. & \statusbox{In Progress} & \tightlist{
\item Finalize parent brand and channel naming system.
\item Run domain, state, and trademark screening.
\item Freeze naming rules for decks, metadata, and distributor materials.
} \\
\rowcolor{RowGray}
7 & Register your business & Form the legal entity once the structure and naming decisions are finalized so the company can sign agreements and operate formally. & \statusbox{Not Started} & \tightlist{
\item File entity formation documents.
\item Register any required DBA names.
\item Create core company records and formation files.
} \\
8 & Get federal and state tax IDs & Obtain the identifiers needed for banking, taxes, vendor onboarding, and routine business administration. & \statusbox{Not Started} & \tightlist{
\item Apply for an EIN.
\item Confirm whether state tax registration is required.
\item Set a tax and bookkeeping calendar.
} \\
\rowcolor{RowGray}
9 & Apply for licenses and permits & Beyond basic business licensing, the operating model must be rights-aware and compliance-ready for content sourcing, ad operations, and distributor delivery requirements. & \statusbox{In Progress} & \tightlist{
\item Confirm required local or state business licenses.
\item Build a proof-of-rights and license-tracking repository.
\item Document platform delivery, compliance, and ad-operations requirements.
} \\
10 & Open a business bank account & Separate company funds from founder activity and create basic financial controls for licensing spend, operations, and incoming platform revenue. & \statusbox{Not Started} & \tightlist{
\item Choose a bank and account structure.
\item Open checking after formation and EIN setup.
\item Define approval controls for contracts and spend.
} \\
\end{longtable}

\normalsize
\section*{Immediate Founder Focus}
\begin{itemize}
    \item Finalize launch order, audience promise, and naming architecture for the first one to two channels.
    \item Turn the strategy into a legal operating shell by choosing structure, registering the entity, obtaining tax IDs, and opening banking.
    \item Build a rights-cleared pilot package with starter libraries, metadata standards, schedules, and distributor-ready materials.
\end{itemize}

\section*{Status Key}
\begin{itemize}
    \item \textbf{Not Started} - no clear evidence yet in the available file or chat context.
    \item \textbf{In Progress} - materially outlined or partially underway, but not yet operationalized.
    \item \textbf{Done} - clearly established by the available evidence.
\end{itemize}

\vfill
{\footnotesize SBA sequence aligned to the live SBA ``10 steps to start your business'' page scanned on May 6, 2026.}

\end{document}
