
\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=0.75in]{geometry}
\usepackage{array,booktabs,longtable,tabularx,multicol}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{pdflscape}
\usepackage{tcolorbox}
\tcbuselibrary{breakable,skins}
\hypersetup{hidelinks}
\definecolor{Navy}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{EAF3F8}
\definecolor{LightGray}{HTML}{F5F7FA}
\definecolor{RuleGray}{HTML}{D9E2EC}
\definecolor{GoodGreen}{HTML}{2E7D32}
\definecolor{RiskAmber}{HTML}{B26A00}
\definecolor{RiskRed}{HTML}{B00020}
\pagestyle{fancy}
\fancyhf{}
\lhead{GLPI + iTop Integration}
\rhead{\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\headheight}{14pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlist[itemize]{leftmargin=*,topsep=2pt,itemsep=2pt}
\setlist[enumerate]{leftmargin=*,topsep=2pt,itemsep=2pt}
\titleformat{\section}{\Large\bfseries\color{Navy}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries\color{Navy}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{Navy}}{}{0em}{}
\newcommand{\docversion}{Version 1.0}
\newcommand{\checkbox}{\fbox{\rule{0pt}{1.1ex}\rule{1.1ex}{0pt}}}
\newcommand{\statusgood}{\textcolor{GoodGreen}{Recommended}}
\newcommand{\statuswarn}{\textcolor{RiskAmber}{Conditional}}
\newcommand{\statusrisk}{\textcolor{RiskRed}{Risk}}
\newtcolorbox{purposebox}{enhanced,breakable,colback=LightBlue,colframe=Navy,boxrule=0.6pt,arc=2mm,left=6pt,right=6pt,top=6pt,bottom=6pt}
\newtcolorbox{calloutbox}{enhanced,breakable,colback=LightGray,colframe=RuleGray,boxrule=0.5pt,arc=2mm,left=6pt,right=6pt,top=6pt,bottom=6pt}
\newcommand{\sourcesection}{%
\section*{Reference Sources}
\small
\begin{itemize}
  \item GLPI Help Center, user documentation: \url{https://help.glpi-project.org/documentation/readme-1}
  \item GLPI Help Center, RESTful API V2: \url{https://help.glpi-project.org/documentation/modules/configuration/general/api/restful-api-v2}
  \item GLPI Project, native inventory announcement and inventory endpoint context: \url{https://www.glpi-project.org/en/discover-native-glpi-inventory/}
  \item GLPI Project, GLPI Inventory integration overview: \url{https://www.glpi-project.org/en/integration/glpi-inventory/}
  \item GLPI Agent documentation: \url{https://glpi-agent.readthedocs.io/}
  \item iTop Documentation, REST/JSON services: \url{https://www.itophub.io/wiki/page?id=latest\%3Aadvancedtopics\%3Arest_json}
  \item iTop Documentation, CMDB data model: \url{https://www.itophub.io/wiki/page?id=latest\%3Adatamodel\%3Astart}
  \item iTop Documentation, change management module: \url{https://www.itophub.io/wiki/page?id=latest\%3Adatamodel\%3Aitop-change-mgmt}
  \item iTop Documentation, integrating iTop: \url{https://www.itophub.io/wiki/page?id=latest\%3Aadvancedtopics\%3Astart}
\end{itemize}
\normalsize
}


\title{\bfseries\color{Navy} GLPI + iTop ITAM / CMDB / ITSM Comparison Table\\\large Tool Responsibility and Integration Fit}
\author{Prepared for Architecture and Operations Review}
\date{\docversion}
\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{purposebox}
\textbf{Purpose.} This comparison table separates GLPI and iTop responsibilities across ITAM, CMDB, and ITSM concerns so integration decisions can be made without assigning overlapping ownership to both tools.
\end{purposebox}

\begin{landscape}
\section{Capability Comparison}
\small
\begin{longtable}{p{0.16\linewidth}p{0.25\linewidth}p{0.25\linewidth}p{0.23\linewidth}}
\toprule
\rowcolor{LightBlue}\textbf{Capability Area} & \textbf{GLPI Fit} & \textbf{iTop Fit} & \textbf{Recommended Integration Decision} \\
\midrule
IT asset inventory & Strong fit for inventory, hardware, software, licenses, and asset lifecycle tracking. & Useful only when asset records become governed CIs needed by ITSM processes. & Make GLPI the ITAM/inventory source. Send only approved, actionable assets to iTop. \\
\addlinespace
Automated inventory collection & Strong fit through GLPI Inventory, GLPI Agent, and compatible inventory payloads. & Not the preferred discovery collection layer. & Collect in GLPI first; normalize and validate before CMDB synchronization. \\
\addlinespace
Asset lifecycle management & Strong fit for stock, assignment, acquisition, retirement, contracts, licenses, and operational asset tracking. & Fit for CI lifecycle state when the item participates in service operations. & GLPI owns asset lifecycle facts; iTop owns CI lifecycle state used by incident/change workflows. \\
\addlinespace
CMDB classes and relationships & Can represent assets and related data, but should not become the primary governed CMDB if iTop is selected for ITSM/CMDB. & Strong fit for physical devices, logical elements, application solutions, business processes, organizations, locations, persons, and teams. & Use iTop as the governed CMDB. Build service relationships in iTop after CI identity is stable. \\
\addlinespace
Service dependency modeling & Limited fit unless customized for service relationship modeling. & Strong fit for relationship-driven impact analysis and service context. & Model business services, application solutions, and dependencies in iTop. \\
\addlinespace
Incident management & GLPI can provide helpdesk/ticketing capability and may be useful for local support workflows. & Strong fit when incident workflows need governed CI relationships and ITSM process context. & Choose one operational incident front door per support domain. Avoid unmanaged duplicate ticketing. \\
\addlinespace
Change management & Not the preferred system for relationship-driven change impact if iTop is selected. & Strong fit for change tickets, change roles, and automatic impact analysis of infrastructure and application solutions. & Use iTop for change management when CMDB relationships are central to approval and risk analysis. \\
\addlinespace
Impact analysis & Limited compared with relationship-rich CMDB impact modeling. & Strong fit for analyzing change or incident impact across CI dependencies. & Treat impact analysis as an iTop outcome. Synchronize only CIs that improve this outcome. \\
\addlinespace
Software and license management & Strong fit for software inventory and license tracking. & Can represent software instances and licenses where they are part of the CMDB model. & Keep detailed license operations in GLPI; synchronize license-related CIs only when needed for service, compliance, or risk. \\
\addlinespace
Contracts and warranties & Strong fit for asset contracts, warranty context, and procurement-oriented information. & Useful where contract data is required for service/provider relationships. & GLPI owns detailed contract/warranty data; iTop receives only fields needed for ITSM decisions. \\
\addlinespace
Data governance & Requires clear rules to distinguish inventory facts from CMDB facts. & Stronger fit for governed CMDB relationships, audit checks, and ITSM process data. & Use an explicit field ownership matrix. \\
\addlinespace
Integration interface & REST API, inventory endpoints, exports, and plugins can support extraction. & REST/JSON interface and import mechanisms can support create/update of objects. & Prefer API-based scheduled sync for repeatability; use CSV imports for initial baseline or controlled manual phase. \\
\addlinespace
Operational reporting & Strong fit for inventory, stock, licenses, and asset status reporting. & Strong fit for CI/service/process reporting, impact, and ITSM process health. & Report asset completeness from GLPI and service/CI process readiness from iTop. \\
\addlinespace
Best use in combined architecture & Operational ITAM and inventory authority. & Governed CMDB and ITSM process authority. & Integrate GLPI to iTop through controlled, scoped, validated synchronization. \\
\bottomrule
\end{longtable}
\normalsize
\end{landscape}

\section{System of Record Decision Matrix}

\begin{longtable}{p{0.31\linewidth}p{0.27\linewidth}p{0.32\linewidth}}
\toprule
\rowcolor{LightBlue}\textbf{Data Element} & \textbf{Recommended Owner} & \textbf{Rationale} \\
\midrule
Asset tag, serial number, discovered hardware facts & GLPI & These are inventory and asset-management facts best maintained at the ITAM layer. \\
Installed software inventory & GLPI & GLPI inventory and license workflows are the natural source for detailed software facts. \\
Contract and warranty detail & GLPI & Procurement and asset lifecycle teams usually maintain this information with assets. \\
CI identity and class & iTop, derived from GLPI input & iTop must control how operational objects are represented in the CMDB. \\
CI lifecycle state & Shared with explicit mapping & GLPI provides asset lifecycle; iTop uses a translated state for service processes. \\
CI-to-service relationship & iTop & Service dependency and impact analysis should be maintained in the CMDB/process system. \\
Support team and service owner & iTop, with directory or GLPI input as needed & ITSM routing and change approval depend on process-accountable owners. \\
Incident and change ticket state & iTop when iTop is selected for ITSM & Ticket workflow state should not be synchronized from inventory tooling. \\
Synchronization status and exceptions & Integration layer & The connector should own run state, failures, and exception evidence. \\
\bottomrule
\end{longtable}

\section{Recommended Architecture Pattern}

\begin{calloutbox}
\textbf{Best-fit pattern.} GLPI should publish normalized, approved asset records. The integration layer should enforce mapping, validation, reconciliation, and audit logging. iTop should receive governed CIs and maintain service relationships used by incident, change, and impact-analysis workflows.
\end{calloutbox}

\sourcesection
\end{document}
