
\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=0.75in]{geometry}
\usepackage{array,booktabs,longtable,tabularx,multicol}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{pdflscape}
\usepackage{tcolorbox}
\tcbuselibrary{breakable,skins}
\hypersetup{hidelinks}
\definecolor{Navy}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{EAF3F8}
\definecolor{LightGray}{HTML}{F5F7FA}
\definecolor{RuleGray}{HTML}{D9E2EC}
\definecolor{GoodGreen}{HTML}{2E7D32}
\definecolor{RiskAmber}{HTML}{B26A00}
\definecolor{RiskRed}{HTML}{B00020}
\pagestyle{fancy}
\fancyhf{}
\lhead{GLPI + iTop Integration}
\rhead{\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\headheight}{14pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlist[itemize]{leftmargin=*,topsep=2pt,itemsep=2pt}
\setlist[enumerate]{leftmargin=*,topsep=2pt,itemsep=2pt}
\titleformat{\section}{\Large\bfseries\color{Navy}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries\color{Navy}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{Navy}}{}{0em}{}
\newcommand{\docversion}{Version 1.0}
\newcommand{\checkbox}{\fbox{\rule{0pt}{1.1ex}\rule{1.1ex}{0pt}}}
\newcommand{\statusgood}{\textcolor{GoodGreen}{Recommended}}
\newcommand{\statuswarn}{\textcolor{RiskAmber}{Conditional}}
\newcommand{\statusrisk}{\textcolor{RiskRed}{Risk}}
\newtcolorbox{purposebox}{enhanced,breakable,colback=LightBlue,colframe=Navy,boxrule=0.6pt,arc=2mm,left=6pt,right=6pt,top=6pt,bottom=6pt}
\newtcolorbox{calloutbox}{enhanced,breakable,colback=LightGray,colframe=RuleGray,boxrule=0.5pt,arc=2mm,left=6pt,right=6pt,top=6pt,bottom=6pt}
\newcommand{\sourcesection}{%
\section*{Reference Sources}
\small
\begin{itemize}
  \item GLPI Help Center, user documentation: \url{https://help.glpi-project.org/documentation/readme-1}
  \item GLPI Help Center, RESTful API V2: \url{https://help.glpi-project.org/documentation/modules/configuration/general/api/restful-api-v2}
  \item GLPI Project, native inventory announcement and inventory endpoint context: \url{https://www.glpi-project.org/en/discover-native-glpi-inventory/}
  \item GLPI Project, GLPI Inventory integration overview: \url{https://www.glpi-project.org/en/integration/glpi-inventory/}
  \item GLPI Agent documentation: \url{https://glpi-agent.readthedocs.io/}
  \item iTop Documentation, REST/JSON services: \url{https://www.itophub.io/wiki/page?id=latest\%3Aadvancedtopics\%3Arest_json}
  \item iTop Documentation, CMDB data model: \url{https://www.itophub.io/wiki/page?id=latest\%3Adatamodel\%3Astart}
  \item iTop Documentation, change management module: \url{https://www.itophub.io/wiki/page?id=latest\%3Adatamodel\%3Aitop-change-mgmt}
  \item iTop Documentation, integrating iTop: \url{https://www.itophub.io/wiki/page?id=latest\%3Aadvancedtopics\%3Astart}
\end{itemize}
\normalsize
}


\title{\bfseries\color{Navy} GLPI + iTop Integration Implementation Checklist\\\large Minimum Readiness and Deployment Controls}
\author{Prepared for Implementation Teams}
\date{\docversion}
\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{purposebox}
\textbf{Purpose.} This checklist provides a minimum implementation standard for connecting GLPI and iTop. Use it before a pilot, before production deployment, and during periodic operational reviews.
\end{purposebox}

\section{Readiness Checklist}

\subsection{1. Scope and Operating Model}
\begin{itemize}
  \item[\checkbox] Define the primary business reason for integration: incident routing, change impact analysis, service mapping, inventory visibility, compliance, or reporting.
  \item[\checkbox] Confirm GLPI is the ITAM/inventory source for the selected asset classes.
  \item[\checkbox] Confirm iTop is the CMDB/ITSM source for governed CIs, relationships, and service process workflows.
  \item[\checkbox] Identify in-scope asset classes for the first release.
  \item[\checkbox] Identify out-of-scope asset classes and document why they are excluded.
  \item[\checkbox] Define the pilot population by business service, location, entity, team, or asset type.
  \item[\checkbox] Assign a GLPI asset owner, iTop CMDB owner, integration owner, security owner, and process owner.
\end{itemize}

\subsection{2. Data Model and Mapping}
\begin{itemize}
  \item[\checkbox] Create a GLPI-to-iTop field mapping matrix.
  \item[\checkbox] Identify mandatory fields for each iTop CI class.
  \item[\checkbox] Define lifecycle state translations, including active, stock, retired, disposed, test, and unknown states.
  \item[\checkbox] Define organization, location, team, and person mapping rules.
  \item[\checkbox] Define how software inventory and licenses will be synchronized or intentionally excluded.
  \item[\checkbox] Define how GLPI asset tags, serial numbers, hostnames, UUIDs, and MAC addresses will be used for identity matching.
  \item[\checkbox] Define which fields are owned by GLPI and which are owned by iTop.
  \item[\checkbox] Define how service relationships, application solutions, and business process links will be created in iTop.
\end{itemize}

\subsection{3. Source Data Quality}
\begin{itemize}
  \item[\checkbox] Confirm GLPI inventory collection is functioning for the target asset population.
  \item[\checkbox] Confirm the target assets have stable identifiers.
  \item[\checkbox] Confirm target assets have owners or accountable support teams.
  \item[\checkbox] Confirm target assets have location and organization data.
  \item[\checkbox] Confirm lifecycle states are accurate enough for CMDB synchronization.
  \item[\checkbox] Identify duplicate records in GLPI before the first synchronization.
  \item[\checkbox] Identify assets with missing mandatory iTop attributes.
  \item[\checkbox] Create a remediation queue for incomplete or conflicting records.
\end{itemize}

\subsection{4. Integration Method}
\begin{itemize}
  \item[\checkbox] Select the integration pattern: scheduled API synchronization, CSV import, or supported connector.
  \item[\checkbox] Document why the selected pattern is appropriate for the first release.
  \item[\checkbox] Define synchronization frequency.
  \item[\checkbox] Define whether the process is full-refresh, incremental, or event-driven.
  \item[\checkbox] Define retry behavior and failure handling.
  \item[\checkbox] Define dry-run behavior before production updates.
  \item[\checkbox] Define rollback procedures for incorrect bulk updates.
  \item[\checkbox] Define logging requirements for each synchronization run.
\end{itemize}

\subsection{5. Security and Access Control}
\begin{itemize}
  \item[\checkbox] Create dedicated service accounts for GLPI and iTop integration access.
  \item[\checkbox] Remove dependency on personal administrator accounts.
  \item[\checkbox] Apply least privilege to source reads and target writes.
  \item[\checkbox] Store credentials in an approved secret-management location.
  \item[\checkbox] Require TLS for API traffic.
  \item[\checkbox] Restrict integration network paths where feasible.
  \item[\checkbox] Redact secrets from logs.
  \item[\checkbox] Review whether personal data is necessary before synchronizing it.
\end{itemize}

\subsection{6. Validation and Reconciliation}
\begin{itemize}
  \item[\checkbox] Implement mandatory field validation before iTop updates.
  \item[\checkbox] Implement duplicate detection before CI creation.
  \item[\checkbox] Implement class-specific validation rules.
  \item[\checkbox] Implement exception routing for rejected records.
  \item[\checkbox] Confirm the integration does not overwrite iTop-owned fields.
  \item[\checkbox] Confirm retired GLPI assets are retired or marked inactive in iTop according to policy.
  \item[\checkbox] Confirm deleted assets are not automatically deleted from iTop without governance approval.
  \item[\checkbox] Confirm each integration run produces a reconciliation report.
\end{itemize}

\subsection{7. Testing}
\begin{itemize}
  \item[\checkbox] Test using non-production GLPI and iTop environments where available.
  \item[\checkbox] Test create, update, no-change, reject, duplicate, retirement, and error scenarios.
  \item[\checkbox] Test malformed data and missing mandatory attributes.
  \item[\checkbox] Test authentication failure, API timeout, and rate/throughput limits.
  \item[\checkbox] Test bulk-load behavior with realistic data volume.
  \item[\checkbox] Test whether synchronized CIs support incident and change workflows in iTop.
  \item[\checkbox] Test reporting and audit-log review.
  \item[\checkbox] Obtain sign-off from GLPI owner, CMDB owner, integration owner, security owner, and ITSM process owner.
\end{itemize}

\subsection{8. Production Deployment}
\begin{itemize}
  \item[\checkbox] Schedule the initial load during an approved maintenance or change window if required.
  \item[\checkbox] Back up or export relevant iTop data before the first bulk update.
  \item[\checkbox] Run the integration in dry-run mode and review expected changes.
  \item[\checkbox] Execute the first production synchronization with monitoring enabled.
  \item[\checkbox] Review created, updated, skipped, failed, and rejected counts.
  \item[\checkbox] Validate a sample of CIs in iTop after synchronization.
  \item[\checkbox] Confirm incident and change workflows can reference synchronized CIs.
  \item[\checkbox] Open remediation tickets for unresolved exceptions.
\end{itemize}

\subsection{9. Operations and Continuous Improvement}
\begin{itemize}
  \item[\checkbox] Review synchronization logs on a defined cadence.
  \item[\checkbox] Track rejected-record rate, duplicate rate, stale-CI rate, and missing-owner rate.
  \item[\checkbox] Review field ownership after tool upgrades or process changes.
  \item[\checkbox] Re-test the connector after GLPI, iTop, plugin, or API upgrades.
  \item[\checkbox] Maintain a runbook for common failures and remediation steps.
  \item[\checkbox] Maintain an approved mapping document under version control.
  \item[\checkbox] Review access permissions periodically.
  \item[\checkbox] Expand scope only after data quality and operational value are proven.
\end{itemize}

\section{Minimum Acceptance Criteria}

\begin{longtable}{p{0.28\linewidth}p{0.62\linewidth}}
\toprule
\rowcolor{LightBlue}\textbf{Area} & \textbf{Minimum Acceptance Criterion} \\
\midrule
Scope & In-scope and out-of-scope asset classes are documented and approved. \\
Data mapping & Every synchronized field has source, target, transformation rule, and owner. \\
Identity matching & The connector can update existing CIs without creating duplicates. \\
Data quality & Records with missing mandatory fields are rejected before iTop update. \\
Security & Integration accounts are least-privilege and secrets are protected. \\
Auditability & Every run produces counts, errors, rejected records, and timestamps. \\
Operational value & Synchronized CIs are usable in at least one incident, change, service, or impact-analysis workflow. \\
Rollback & Incorrect bulk changes can be stopped, reverted, or corrected through a documented process. \\
\bottomrule
\end{longtable}

\section{Go/No-Go Review}

\begin{longtable}{p{0.36\linewidth}p{0.18\linewidth}p{0.36\linewidth}}
\toprule
\rowcolor{LightBlue}\textbf{Question} & \textbf{Decision} & \textbf{Evidence / Notes} \\
\midrule
Are the integration scope and field owners approved? & Go / No-Go & \\
Are source data quality defects acceptable for the pilot? & Go / No-Go & \\
Has duplicate prevention been tested? & Go / No-Go & \\
Has security reviewed integration access? & Go / No-Go & \\
Has dry-run output been reviewed by data owners? & Go / No-Go & \\
Is rollback or correction documented? & Go / No-Go & \\
Are operational owners ready to handle exceptions? & Go / No-Go & \\
\bottomrule
\end{longtable}

\sourcesection
\end{document}



