
\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=0.75in]{geometry}
\usepackage{array,booktabs,longtable,tabularx,multicol}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{pdflscape}
\usepackage{tcolorbox}
\tcbuselibrary{breakable,skins}
\hypersetup{hidelinks}
\definecolor{Navy}{HTML}{1F4E79}
\definecolor{LightBlue}{HTML}{EAF3F8}
\definecolor{LightGray}{HTML}{F5F7FA}
\definecolor{RuleGray}{HTML}{D9E2EC}
\definecolor{GoodGreen}{HTML}{2E7D32}
\definecolor{RiskAmber}{HTML}{B26A00}
\definecolor{RiskRed}{HTML}{B00020}
\pagestyle{fancy}
\fancyhf{}
\lhead{GLPI + iTop Integration}
\rhead{\thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\headheight}{14pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlist[itemize]{leftmargin=*,topsep=2pt,itemsep=2pt}
\setlist[enumerate]{leftmargin=*,topsep=2pt,itemsep=2pt}
\titleformat{\section}{\Large\bfseries\color{Navy}}{}{0em}{}[\titlerule]
\titleformat{\subsection}{\large\bfseries\color{Navy}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{Navy}}{}{0em}{}
\newcommand{\docversion}{Version 1.0}
\newcommand{\checkbox}{\fbox{\rule{0pt}{1.1ex}\rule{1.1ex}{0pt}}}
\newcommand{\statusgood}{\textcolor{GoodGreen}{Recommended}}
\newcommand{\statuswarn}{\textcolor{RiskAmber}{Conditional}}
\newcommand{\statusrisk}{\textcolor{RiskRed}{Risk}}
\newtcolorbox{purposebox}{enhanced,breakable,colback=LightBlue,colframe=Navy,boxrule=0.6pt,arc=2mm,left=6pt,right=6pt,top=6pt,bottom=6pt}
\newtcolorbox{calloutbox}{enhanced,breakable,colback=LightGray,colframe=RuleGray,boxrule=0.5pt,arc=2mm,left=6pt,right=6pt,top=6pt,bottom=6pt}
\newcommand{\sourcesection}{%
\section*{Reference Sources}
\small
\begin{itemize}
  \item GLPI Help Center, user documentation: \url{https://help.glpi-project.org/documentation/readme-1}
  \item GLPI Help Center, RESTful API V2: \url{https://help.glpi-project.org/documentation/modules/configuration/general/api/restful-api-v2}
  \item GLPI Project, native inventory announcement and inventory endpoint context: \url{https://www.glpi-project.org/en/discover-native-glpi-inventory/}
  \item GLPI Project, GLPI Inventory integration overview: \url{https://www.glpi-project.org/en/integration/glpi-inventory/}
  \item GLPI Agent documentation: \url{https://glpi-agent.readthedocs.io/}
  \item iTop Documentation, REST/JSON services: \url{https://www.itophub.io/wiki/page?id=latest\%3Aadvancedtopics\%3Arest_json}
  \item iTop Documentation, CMDB data model: \url{https://www.itophub.io/wiki/page?id=latest\%3Adatamodel\%3Astart}
  \item iTop Documentation, change management module: \url{https://www.itophub.io/wiki/page?id=latest\%3Adatamodel\%3Aitop-change-mgmt}
  \item iTop Documentation, integrating iTop: \url{https://www.itophub.io/wiki/page?id=latest\%3Aadvancedtopics\%3Astart}
\end{itemize}
\normalsize
}


\title{\bfseries\color{Navy} GLPI + iTop Integration Handout\\\large Practical ITAM-to-CMDB Operating Pattern}
\author{Prepared for IT Operations, CMDB, and ITSM Teams}
\date{\docversion}
\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{purposebox}
\textbf{Purpose.} Use this handout as a concise reference for combining GLPI and iTop without creating duplicate systems of record. The recommended model is GLPI for ITAM and inventory, iTop for governed CMDB and ITSM workflows, and a controlled integration layer between them.
\end{purposebox}

\section{Recommended Positioning}

\begin{longtable}{p{0.23\linewidth}p{0.33\linewidth}p{0.34\linewidth}}
\toprule
\rowcolor{LightBlue}\textbf{Tool} & \textbf{Best Role} & \textbf{Primary Value} \\
\midrule
GLPI & ITAM and inventory source & Tracks hardware, software, licenses, contracts, asset lifecycle, inventory collection, and operational support details. \\
iTop & CMDB and ITSM process layer & Models CIs, relationships, organizations, teams, services, incidents, changes, and service impact analysis. \\
Integration layer & Policy enforcement and synchronization control & Maps, validates, reconciles, logs, and synchronizes only approved records from GLPI to iTop. \\
\bottomrule
\end{longtable}

\section{Target Architecture}

\begin{enumerate}
  \item \textbf{Discover and manage assets in GLPI.} GLPI Agent, inventory plugins, OCS-style inputs, manual updates, or APIs populate the ITAM repository.
  \item \textbf{Approve assets for CMDB use.} Asset owners confirm identity, lifecycle status, location, organization, and ownership.
  \item \textbf{Transform assets into CIs.} The integration layer maps GLPI records to iTop classes and attributes.
  \item \textbf{Validate and reconcile.} The connector rejects incomplete records, prevents duplicates, and preserves field ownership.
  \item \textbf{Create or update iTop CIs.} iTop receives only actionable CIs needed for service management, impact analysis, or change/incident workflows.
  \item \textbf{Use iTop for process execution.} Incidents, changes, service catalog decisions, and impact analysis use the governed CI model.
\end{enumerate}

\section{What Should Synchronize?}

\subsection{Synchronize}
\begin{itemize}
  \item Servers, endpoints, and network devices that support business services.
  \item Assets with stable identifiers, owners, locations, and lifecycle state.
  \item Software or license information needed for compliance, service support, or risk decisions.
  \item CIs required for incident routing, change impact analysis, or service dependency mapping.
\end{itemize}

\subsection{Do Not Synchronize by Default}
\begin{itemize}
  \item Temporary, test-only, unmanaged, or low-value assets.
  \item Records without reliable identifiers or accountable ownership.
  \item Raw inventory noise that does not improve ITSM workflows.
  \item Sensitive personal data unless it is required for an approved operational purpose.
\end{itemize}

\section{Minimum Implementation Checklist}

\begin{multicols}{2}
\begin{itemize}
  \item[\checkbox] Define GLPI and iTop system-of-record boundaries.
  \item[\checkbox] Identify in-scope asset classes.
  \item[\checkbox] Define GLPI-to-iTop field mappings.
  \item[\checkbox] Define CI identity and duplicate-prevention rules.
  \item[\checkbox] Define lifecycle state translations.
  \item[\checkbox] Define mandatory iTop CI attributes.
  \item[\checkbox] Create dedicated integration accounts.
  \item[\checkbox] Store API secrets securely.
  \item[\checkbox] Run a dry-run synchronization.
  \item[\checkbox] Review rejected records and duplicate candidates.
  \item[\checkbox] Pilot with a limited CI class or service.
  \item[\checkbox] Track synchronization and data-quality KPIs.
\end{itemize}
\end{multicols}

\section{Recommended Controls}

\begin{longtable}{p{0.27\linewidth}p{0.63\linewidth}}
\toprule
\rowcolor{LightBlue}\textbf{Control} & \textbf{Requirement} \\
\midrule
Field ownership & GLPI owns inventory facts; iTop owns CI relationships and ITSM process context. \\
Reconciliation & Use serial number, asset tag, hostname, UUID, MAC address, or composite keys before creating a CI. \\
Validation & Reject missing mandatory attributes before updating iTop. \\
Exception handling & Route rejected records to accountable asset or CMDB owners. \\
Auditability & Log every run: timestamp, created, updated, skipped, rejected, failed, and error reasons. \\
Security & Use least-privilege service accounts, TLS, protected secrets, and no personal admin credentials. \\
Change control & Treat mapping, connector, and bulk update changes as controlled releases. \\
\bottomrule
\end{longtable}

\section{Recommended Decision}

\begin{calloutbox}
Adopt a controlled one-way synchronization model from GLPI to iTop for approved CI-relevant asset data. Start with a small pilot, measure data quality, validate service-management value, then expand by asset class and business service.
\end{calloutbox}

\sourcesection
\end{document}
