\documentclass[11pt]{article}

\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{xurl}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{longtable}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{lastpage}

\definecolor{Navy}{HTML}{1F4E79}
\definecolor{DarkGray}{HTML}{333333}
\definecolor{LightGray}{HTML}{F5F7FA}
\definecolor{RuleGray}{HTML}{D9DEE7}
\definecolor{CodeBg}{HTML}{F7F7F7}
\definecolor{CodeFrame}{HTML}{DDDDDD}

\hypersetup{
  colorlinks=true,
  linkcolor=Navy,
  citecolor=Navy,
  urlcolor=Navy,
  pdftitle={Getting Started with CWEB Literate Programming for C, C++, and Java Projects},
  pdfauthor={Prepared for Jordan Suber},
  pdfsubject={CWEB literate programming onboarding handout},
  pdfkeywords={CWEB, ctangle, cweave, literate programming, C, C++, Java, noweb, Org Babel}
}

\setlength{\headheight}{14pt}
\pagestyle{fancy}
\fancyhf{}
\lhead{CWEB Literate Programming}
\rhead{Getting Started}
\cfoot{Page \thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

\titleformat{\section}{\large\bfseries\color{Navy}}{\thesection}{0.7em}{}
\titleformat{\subsection}{\normalsize\bfseries\color{DarkGray}}{\thesubsection}{0.7em}{}
\titlespacing*{\section}{0pt}{1.1em}{0.45em}
\titlespacing*{\subsection}{0pt}{0.85em}{0.35em}

\setlist[itemize]{leftmargin=1.35em,itemsep=0.2em,topsep=0.25em}
\setlist[enumerate]{leftmargin=1.55em,itemsep=0.25em,topsep=0.25em}
\renewcommand{\arraystretch}{1.18}

\lstdefinestyle{handout}{
  backgroundcolor=\color{CodeBg},
  frame=single,
  rulecolor=\color{CodeFrame},
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries\color{Navy},
  commentstyle=\itshape\color{DarkGray},
  stringstyle=\color{DarkGray},
  showstringspaces=false,
  breaklines=true,
  columns=fullflexible,
  keepspaces=true,
  tabsize=2,
  numbers=none,
  xleftmargin=0.5em,
  xrightmargin=0.5em,
  aboveskip=0.55em,
  belowskip=0.65em
}
\lstset{style=handout}

\newcommand{\callout}[2]{%
  \vspace{0.45em}
  \noindent\fcolorbox{RuleGray}{LightGray}{%
    \begin{minipage}{0.96\linewidth}
    \textbf{#1.} #2
    \end{minipage}}
  \vspace{0.45em}
}

\begin{document}

\begin{center}
  {\LARGE\bfseries\color{Navy} Getting Started with CWEB Literate Programming}\\[0.25em]
  {\large For C, C++, and Java Projects}\\[0.4em]
  {\normalsize Repository-grounded onboarding handout}\par
\end{center}

\vspace{0.5em}

\noindent\textbf{Audience.} Software engineers who want to adopt literate programming without losing normal compiler, build, and CI/CD workflows.\\
\textbf{Primary repository.} The user-provided \texttt{ascherer/cweb} repository snapshot, containing CWEB 4.12.2-era source files, examples, change files, macro files, and manual material.\\
\textbf{Core recommendation.} Use CWEB directly for C and C++; use a language-neutral literate workflow for Java unless you are deliberately experimenting.

\callout{Bottom line}{CWEB is the right starting point for literate C and C++ projects. It is not the right default for Java. For Java, keep the literate-programming discipline, but use tools such as noweb or Org Babel, and keep Javadoc for API reference output.}

\section{What CWEB Gives You}

CWEB is a literate-programming system derived from Knuth's WEB idea and adapted by Donald E. Knuth and Silvio Levy for C-family programming. CTAN describes CWEB as a system for literate programming, also called structured software documentation, with code written in C and C++ \cite{ctan-cweb}. Knuth's CWEB page identifies \emph{The CWEB System of Structured Documentation} as the canonical CWEB work by Knuth and Levy \cite{knuth-cweb}. Knuth's earlier literate-programming paper introduced the central idea: write programs as explanations for humans, while still generating source that machines can compile \cite{knuth-lp-paper,knuth-lp-page}.

In practical terms, one \texttt{.w} source file produces two outputs:

\begin{center}
\begin{tabularx}{0.96\linewidth}{>{\bfseries}p{0.17\linewidth} X X}
\toprule
Tool & Input & Output \\
\midrule
\texttt{ctangle} & CWEB \texttt{.w} file & C/C++ source file for the compiler \\
\texttt{cweave} & Same CWEB \texttt{.w} file & TeX documentation source for typesetting \\
TeX/PDFTeX & Generated \texttt{.tex} file & Human-readable documentation, usually PDF \\
\bottomrule
\end{tabularx}
\end{center}

\section{How to Read This Repository}

Treat the repository as a \textbf{toolchain source repository}, not as an application framework. Your first objective is to build or obtain \texttt{ctangle} and \texttt{cweave}; your second objective is to practice on a small \texttt{.w} file.

The repository README identifies these important files and directories:

\begin{longtable}{>{\ttfamily}p{0.22\linewidth} p{0.69\linewidth}}
\toprule
\normalfont\textbf{Path} & \textbf{Why it matters} \\
\midrule
README & Build and installation guidance for the repository snapshot. It states that \texttt{cwebman.tex} is the user manual, \texttt{examples/} contains examples, \texttt{common.c} and \texttt{ctangle.c} are used for bootstrapping, and \texttt{c++lib.w}/\texttt{iso\_types.w} are C++ helper include files. \cite{repo-readme} \\
README.md & Project-level note for CWEB 4.12.2 and contemporary development status in the GitHub project. \cite{repo-readme-md} \\
Makefile & Build rules for \texttt{ctangle}, \texttt{cweave}, generated \texttt{.c}, generated \texttt{.tex}, and PDF output. \cite{repo-makefile} \\
cweb.1 & Manual-page style usage reference for command-line behavior, options, change files, and output-file override behavior. \cite{repo-manpage} \\
cwebman.tex & User manual source. Start here after the first working example, not before. \cite{repo-readme} \\
examples/ & Small examples. Use this before attempting to understand \texttt{common.w}, \texttt{ctangle.w}, or \texttt{cweave.w}. \cite{repo-readme} \\
c++lib.w & C++ identifier support definitions for common standard-library names. \cite{repo-readme,repo-cpplib} \\
iso\_types.w & Additional type support intended for C/C++ users in this repository snapshot. \cite{repo-readme} \\
*.ch & Change files. Use them later for platform customization or variant builds, not for your first project. \cite{repo-readme,repo-manpage} \\
\bottomrule
\end{longtable}

\section{First Build: Get \texttt{ctangle} and \texttt{cweave}}

From the repository root, run:

\begin{lstlisting}[language=bash]
touch *.c
make all
make cautiously
\end{lstlisting}

The README specifically instructs users to touch the generated \texttt{.c} files before proceeding, edit the opening lines of the Makefile for local directory information, run \texttt{make all}, and use \texttt{make cautiously} for reassurance that \texttt{CTANGLE} reproduces itself \cite{repo-readme}. The Makefile also exposes direct suffix rules for converting \texttt{.w} files to \texttt{.c}, \texttt{.tex}, \texttt{.dvi}, and \texttt{.pdf} outputs \cite{repo-makefile}.

A successful build gives you at least:

\begin{lstlisting}
./ctangle
./cweave
\end{lstlisting}

Install globally only after the local build works:

\begin{lstlisting}[language=bash]
sudo make install
\end{lstlisting}

\section{First Experiment: Process an Example}

Start with the example directory, because the repository itself says that \texttt{examples/} contains additional examples of CWEB usage \cite{repo-readme}. A typical first experiment is:

\begin{lstlisting}[language=bash]
cd examples
../ctangle wc.w
../cweave wc.w
pdftex wc.tex
cc -Wall -Wextra -O2 -o wc wc.c
./wc < README
\end{lstlisting}

The flow is:

\begin{lstlisting}
wc.w
  |-- ctangle --> wc.c   --> cc/gcc/clang --> executable
  |-- cweave  --> wc.tex --> tex/pdftex   --> readable documentation
\end{lstlisting}

\callout{Working rule}{Edit the \texttt{.w} file. Treat tangled \texttt{.c} or \texttt{.cpp} files as generated artifacts. Do not hand-edit generated output unless you are debugging the toolchain itself.}

\section{Minimum CWEB Syntax You Need First}

A CWEB file interleaves TeX prose and C/C++ code. For a first project, learn only five concepts:

\begin{center}
\begin{tabularx}{0.96\linewidth}{>{\ttfamily}p{0.18\linewidth} X}
\toprule
\normalfont\textbf{Syntax} & \textbf{Meaning} \\
\midrule
@* & Begin a major documented section. \\
@ & Begin an ordinary section. \\
@c & Begin code that is emitted into the generated C/C++ file. \\
@<name@>= & Define a named code section. \\
@<name@> & Use a named code section at this point. \\
@i file.w & Include another CWEB file. Useful for \texttt{c++lib.w} and \texttt{iso\_types.w}. \\
\bottomrule
\end{tabularx}
\end{center}

A minimal C example:

\begin{lstlisting}
\datethis

@* Hello program.
This small program demonstrates the CWEB split between exposition and code.

@c
#include <stdio.h>

int main(void)
{
  @<Print the greeting@>;
  return 0;
}

@ The greeting is separated from the control flow so the program reads as a
short explanation instead of a compiler-ordered listing.

@<Print the greeting@>=
printf("Hello from CWEB.\n");
\end{lstlisting}

Build it:

\begin{lstlisting}[language=bash]
ctangle hello.w
cweave hello.w
pdftex hello.tex
cc -Wall -Wextra -O2 -o hello hello.c
./hello
\end{lstlisting}

\section{Recommended Project Layout for C}

Use a layout that keeps literate sources, generated sources, documentation, and build products distinct:

\begin{lstlisting}
my-c-project/
  Makefile
  src-lit/
    app.w
    parser.w
    graph.w
  generated/
    app.c
    parser.c
    graph.c
  include/
    project.h
  docs/
    app.pdf
    parser.pdf
    graph.pdf
  build/
\end{lstlisting}

A minimal Makefile pattern:

\begin{lstlisting}[language=make]
CTANGLE := ./tools/cweb/ctangle
CWEAVE  := ./tools/cweb/cweave
PDFTEX  := pdftex
CC      := cc
CFLAGS  := -Wall -Wextra -O2

all: build/app docs/app.pdf

generated/app.c: src-lit/app.w
	mkdir -p generated
	$(CTANGLE) src-lit/app.w - generated/app.c

build/app: generated/app.c
	mkdir -p build
	$(CC) $(CFLAGS) -o build/app generated/app.c

docs/app.tex: src-lit/app.w
	mkdir -p docs
	$(CWEAVE) src-lit/app.w - docs/app.tex

docs/app.pdf: docs/app.tex
	cd docs && $(PDFTEX) app.tex

clean:
	rm -rf build generated docs/*.tex docs/*.pdf docs/*.idx docs/*.scn docs/*.log
\end{lstlisting}

\section{Recommended Project Layout for C++}

For C++, begin each major \texttt{.w} source with the repository-provided C++ support includes:

\begin{lstlisting}
@i c++lib.w
@i iso_types.w
\end{lstlisting}

The repository README explicitly says these files are for C++ users and should be included at the beginning of the program when needed \cite{repo-readme}. The \texttt{c++lib.w} file itself defines standard C++ namespace and class names for CWEB formatting support \cite{repo-cpplib}.

Minimal C++ example:

\begin{lstlisting}
@i c++lib.w
@i iso_types.w

@* A small C++ program.
This example demonstrates CWEB with ordinary C++ library types.

@c
#include <iostream>
#include <string>
#include <vector>

int main()
{
  @<Create names@>;
  @<Print names@>;
  return 0;
}

@<Create names@>=
std::vector<std::string> names = {"Ada", "Grace", "Barbara"};

@<Print names@>=
for (const auto& name : names) {
  std::cout << name << '\n';
}
\end{lstlisting}

Use the optional third command-line argument to name the generated file \texttt{.cpp}. The CWEB manual-page material says that if a third file name is present, it overrides the default output file name \cite{repo-manpage}.

\begin{lstlisting}[language=bash]
ctangle names.w - generated/names.cpp
cweave names.w - docs/names.tex
cd docs && pdftex names.tex
c++ -std=c++20 -Wall -Wextra -O2 -o build/names generated/names.cpp
\end{lstlisting}

\section{Java: Do Not Force Java into CWEB First}

CWEB is a C/C++ literate-programming system. CTAN describes its target code as C and C++ \cite{ctan-cweb}. That makes CWEB a poor default for Java because \texttt{cweave} expects C-like lexical and formatting conventions. You can still adopt the literate-programming discipline for Java, but use a language-neutral tangler.

Recommended Java path:

\begin{center}
\begin{tabularx}{0.96\linewidth}{>{\bfseries}p{0.24\linewidth} X}
\toprule
Need & Recommended approach \\
\midrule
Literate Java source & Use noweb or Org Babel source blocks. Noweb emphasizes simplicity, extensibility, and language independence \cite{noweb}. Org tangling expands, merges, and writes source blocks into one or more files \cite{org-tangle}. \\
Java API documentation & Use Javadoc. Oracle's documentation states that the \texttt{javadoc} command parses declarations and documentation comments and produces corresponding HTML pages \cite{oracle-javadoc-tool}. \\
Unified engineering standard & Use CWEB for C/C++, noweb or Org Babel for Java, and enforce the same repository conventions, review gates, and CI checks across all languages. \\
\bottomrule
\end{tabularx}
\end{center}

A simple Java literate file using Markdown-style fenced blocks might look like this:

\begin{lstlisting}
# Main program

This class demonstrates the application entry point.

```java file=Main.java
public class Main {
    public static void main(String[] args) {
        System.out.println("Hello from literate Java.");
    }
}
```
\end{lstlisting}

Then a small tangling script extracts fenced Java blocks into \texttt{generated/Main.java}, after which Java builds normally:

\begin{lstlisting}[language=bash]
javac generated/Main.java
java -cp generated Main
javadoc -d docs/api generated/Main.java
\end{lstlisting}

\section{Adoption Roadmap}

\begin{longtable}{>{\bfseries}p{0.13\linewidth} p{0.37\linewidth} p{0.38\linewidth}}
\toprule
Phase & Goal & Exit criteria \\
\midrule
1 & Build CWEB locally. & \texttt{ctangle}, \texttt{cweave}, and \texttt{make cautiously} work. \\
2 & Process one repository example. & A sample \texttt{.w} file produces source, TeX, PDF, and an executable. \\
3 & Write one small C utility as \texttt{.w}. & The \texttt{.w} file is the only edited source; generated C is reproducible. \\
4 & Convert one C++ utility. & \texttt{c++lib.w} and \texttt{iso\_types.w} are included when useful; generated output is named \texttt{.cpp}. \\
5 & Add CI validation. & CI runs tangle, weave, compile, unit tests, and documentation build. \\
6 & Add Java literate flow separately. & Java source is tangled by a language-neutral tool and documented by Javadoc. \\
\bottomrule
\end{longtable}

\section{CI/CD Quality Gates}

A literate-programming repository should fail the build if generated artifacts are stale. Suggested CI stages:

\begin{enumerate}
  \item \textbf{Tangle:} regenerate \texttt{.c}, \texttt{.cpp}, or Java source from literate inputs.
  \item \textbf{Diff check:} verify generated files are either ignored or exactly reproducible.
  \item \textbf{Compile:} build with normal compiler flags.
  \item \textbf{Test:} run unit and integration tests.
  \item \textbf{Weave:} produce TeX documentation.
  \item \textbf{Typeset:} generate PDF or HTML documentation.
  \item \textbf{Publish:} publish the generated documentation as a build artifact.
\end{enumerate}

\section{Practical Rules for Writing Literate Code}

\begin{enumerate}
  \item \textbf{Lead with intent.} Begin each major section by explaining the problem, not the syntax.
  \item \textbf{Name sections by behavior.} Prefer \texttt{@<Validate input record@>} over \texttt{@<Loop body@>}.
  \item \textbf{Keep chunks small.} A code section should usually explain one decision or operation.
  \item \textbf{Keep generated files disposable.} The authoritative source is the literate file.
  \item \textbf{Use normal compilers and analyzers.} CWEB does not replace \texttt{cc}, \texttt{clang-tidy}, CodeQL, tests, or CI.
  \item \textbf{Separate API reference from design narrative.} CWEB/noweb/Org explain design and implementation; Javadoc explains Java APIs.
\end{enumerate}

\section{Troubleshooting Checklist}

\begin{center}
\begin{tabularx}{0.96\linewidth}{>{\bfseries}p{0.28\linewidth} X}
\toprule
Problem & Likely fix \\
\midrule
\texttt{ctangle} cannot open an include file & Check the current working directory and \texttt{CWEBINPUTS}. The repository build examples compile \texttt{common.c} with \texttt{-DCWEBINPUTS=...} to define the include path \cite{repo-readme}. \\
Generated C++ file has \texttt{.c} extension & Pass an explicit output file: \texttt{ctangle source.w - generated/source.cpp}. \\
CWEAVE output looks odd for C++ identifiers & Include \texttt{c++lib.w} and \texttt{iso\_types.w} where appropriate. \\
Java formatting is poor under CWEB & Do not use CWEB as the Java default; use noweb or Org Babel instead. \\
Generated files drift from source & Add a CI diff check after tangling. \\
Large codebase feels hard to read & Split literate files by bounded modules, not by arbitrary compiler order. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Starting Recommendation}

Begin with one small C program and one small C++ program. Do not begin by converting an entire repository. The first milestone is simple and concrete:

\begin{quote}
One literate source file should produce one compilable source file, one executable, and one readable PDF document from a reproducible build command.
\end{quote}

Once that works, expand into module-level files, Makefile automation, CI checks, and documentation publishing.

\begin{thebibliography}{99}

\bibitem{repo-readme}
User-provided \texttt{ascherer/cweb} repository snapshot, \texttt{README}. Notes on repository contents, \texttt{cwebman.tex}, examples, bootstrapping files, C++ helper includes, \texttt{make all}, \texttt{make cautiously}, and installation guidance.

\bibitem{repo-readme-md}
User-provided \texttt{ascherer/cweb} repository snapshot, \texttt{README.md}. Project summary for CWEB 4.12.2 and contemporary development notes.

\bibitem{repo-makefile}
User-provided \texttt{ascherer/cweb} repository snapshot, \texttt{Makefile}. Build rules for \texttt{ctangle}, \texttt{cweave}, and suffix rules for \texttt{.w} to \texttt{.c}, \texttt{.tex}, \texttt{.dvi}, and \texttt{.pdf}.

\bibitem{repo-manpage}
User-provided \texttt{ascherer/cweb} repository snapshot, \texttt{cweb.1}. Command-line usage, options, output-file override behavior, TeX macro files, and related references.

\bibitem{repo-cpplib}
User-provided \texttt{ascherer/cweb} repository snapshot, \texttt{c++lib.w}. Standard C++ namespace and class-name support definitions for CWEB formatting.

\bibitem{knuth-cweb}
Donald E. Knuth, \emph{The CWEB System of Structured Documentation}. Stanford page for Knuth and Levy's CWEB book. Available at: \url{https://www-cs-faculty.stanford.edu/~knuth/cweb.html}.

\bibitem{ctan-cweb}
CTAN, \emph{cweb - A Web system in C}. Available at: \url{https://ctan.org/pkg/cweb}.

\bibitem{knuth-lp-paper}
Donald E. Knuth, ``Literate Programming,'' \emph{The Computer Journal}, 27(2), 97--111, 1984. Publisher page: \url{https://academic.oup.com/comjnl/article-pdf/27/2/97/981657/270097.pdf}.

\bibitem{knuth-lp-page}
Donald E. Knuth, \emph{Literate Programming}. Stanford page for the collected essays. Available at: \url{https://www-cs-faculty.stanford.edu/~knuth/lp.html}.

\bibitem{noweb}
Norman Ramsey, \emph{noweb - A Simple, Extensible Tool for Literate Programming}. Tufts University. Available at: \url{https://www.cs.tufts.edu/~nr/noweb/}.

\bibitem{org-tangle}
GNU Emacs Org Manual, \emph{Extracting Source Code}. Available at: \url{https://orgmode.org/manual/Extracting-Source-Code.html}.

\bibitem{oracle-javadoc-tool}
Oracle, \emph{javadoc - The Java API Documentation Generator}. Available at: \url{https://docs.oracle.com/javase/8/docs/technotes/tools/windows/javadoc.html}.

\end{thebibliography}

\end{document}
