\documentclass[11pt]{article}

\usepackage[letterpaper,margin=0.75in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{longtable}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{lastpage}

\definecolor{TitleBlue}{HTML}{1F4E79}
\definecolor{RuleBlue}{HTML}{2F75B5}
\definecolor{LightBlue}{HTML}{EAF3F8}
\definecolor{DarkGray}{HTML}{333333}
\definecolor{CodeBg}{HTML}{F6F8FA}
\definecolor{CodeRule}{HTML}{D0D7DE}

\hypersetup{
  colorlinks=true,
  linkcolor=TitleBlue,
  urlcolor=TitleBlue,
  citecolor=TitleBlue,
  pdftitle={WEB and Literate Programming: A Technical Handout},
  pdfauthor={Jordan Suber / ChatGPT},
  pdfsubject={Donald Knuth's WEB, CWEB, TANGLE, WEAVE, and literate programming},
  pdfkeywords={WEB, CWEB, literate programming, Donald Knuth, TANGLE, WEAVE, CTANGLE, CWEAVE}
}

\pagestyle{fancy}
\fancyhf{}
\lhead{WEB and Literate Programming}
\rhead{Technical Handout}
\cfoot{Page \thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\setlength{\headheight}{14pt}

\titleformat{\section}{\Large\bfseries\color{TitleBlue}}{\thesection}{0.6em}{}
\titleformat{\subsection}{\large\bfseries\color{DarkGray}}{\thesubsection}{0.6em}{}
\titlespacing*{\section}{0pt}{1.2em}{0.55em}
\titlespacing*{\subsection}{0pt}{0.85em}{0.35em}

\setlist[itemize]{leftmargin=1.4em, itemsep=0.25em, topsep=0.25em}
\setlist[enumerate]{leftmargin=1.6em, itemsep=0.25em, topsep=0.25em}
\renewcommand{\arraystretch}{1.25}

\lstdefinestyle{handoutcode}{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{CodeBg},
  frame=single,
  rulecolor=\color{CodeRule},
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  breaklines=true,
  tabsize=2,
  aboveskip=0.75em,
  belowskip=0.75em
}
\lstset{style=handoutcode}

\newcommand{\callout}[2]{%
  \vspace{0.4em}
  \noindent\fcolorbox{RuleBlue}{LightBlue}{%
    \begin{minipage}{0.965\linewidth}
      \textbf{#1}\par\vspace{0.2em}#2
    \end{minipage}}
  \vspace{0.6em}
}

\newcommand{\source}[1]{\textsuperscript{[#1]}}

\begin{document}

\begin{center}
  {\LARGE\bfseries\color{TitleBlue} WEB and Literate Programming}\par
  \vspace{0.25em}
  {\large A polished technical handout using primary and authoritative sources}\par
  \vspace{0.5em}
  {\small Focus: Donald Knuth's WEB, CWEB, TANGLE, WEAVE, and the engineering rationale for literate programming}
\end{center}

\vspace{0.8em}
\hrule height 0.8pt
\vspace{1em}

\callout{Executive summary}{%
\texttt{WEB} is Donald E. Knuth's original system for literate programming. It keeps explanatory documentation and program fragments in one structured source file, then processes that source in two directions: \texttt{TANGLE} extracts compiler-ready program text, while \texttt{WEAVE} produces typeset documentation. The later \texttt{CWEB} system, developed by Knuth and Silvio Levy, adapts WEB for C, C++, and related languages.\source{1}\source{2}\source{3}}

\section{What WEB Is}

\texttt{WEB} is both a software system and a documentation discipline. Knuth's central premise is that a program should be written as a communication to human beings rather than merely as an instruction stream for a computer. His Stanford description of literate programming says that the method combines a programming language with a documentation language and treats the program as a piece of literature addressed to human readers.\source{1}

In practical terms, a WEB file is the authoritative source for both:

\begin{itemize}
  \item a machine-oriented program file suitable for compilation or downstream translation; and
  \item a reader-oriented document containing prose explanation, program structure, cross-references, and formatted code.
\end{itemize}

This is the reverse of ordinary documentation extraction. In conventional documentation generation, code is primary and prose is usually embedded as comments. In literate programming, the explanation is primary, and code fragments are embedded where they best support the narrative.

\section{The Two-Processor Model}

WEB operationalizes literate programming through two complementary processors.

\begin{center}
\begin{tabularx}{0.95\linewidth}{>{\bfseries}p{0.18\linewidth} X X}
\toprule
Processor & Main output & Function \\
\midrule
\texttt{TANGLE} & Pascal source in original WEB & Reorders and expands the WEB program fragments into the sequence required by the compiler or downstream build chain.\source{4} \\
\texttt{WEAVE} & TeX documentation & Converts the same WEB source into a typeset document with readable program structure and supporting explanation.\source{4} \\
\texttt{CTANGLE} & C or C++ source in CWEB & Converts a CWEB source document into a compilable C/C++ program.\source{3}\source{5} \\
\texttt{CWEAVE} & TeX documentation in CWEB & Converts the same CWEB file into TeX, adding typographic formatting and cross-index information.\source{3}\source{5} \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Processing flow}

\begin{center}
\fcolorbox{RuleBlue}{LightBlue}{\begin{minipage}{0.88\linewidth}
\centering
\texttt{WEB / CWEB source (.web or .w)}\\[0.6em]
\begin{tabular}{ccc}
& \texttt{TANGLE / CTANGLE} & $\longrightarrow$ compiler-ready source \\
& \texttt{WEAVE / CWEAVE} & $\longrightarrow$ TeX documentation \\
\end{tabular}
\end{minipage}}
\end{center}

The key architectural point is that the same source file is processed into two products. A single literate source therefore becomes the source of truth for both implementation and explanation.

\section{Modular Structure}

A WEB or CWEB program is divided into numbered sections or modules. Each module can contain prose explanation, definitions, and code fragments. Modules may be presented in the order that best teaches or explains the system, even if that order differs from the order required by the compiler.

\begin{center}
\begin{tabularx}{0.95\linewidth}{>{\bfseries}p{0.22\linewidth} X}
\toprule
Element & Purpose \\
\midrule
Documentation text & Explains the algorithm, design constraints, invariants, data structures, and reasoning behind the code. \\
Definitions & Introduces constants, macros, types, and local vocabulary used by the program fragments. \\
Code fragments & Supplies the executable statements or declarations that are later assembled by tangling. \\
Cross-references & Help readers navigate where names and modules are defined and used in the woven output. \\
\bottomrule
\end{tabularx}
\end{center}

The CWEB documentation describes the programmer's source file as a plain text file, conventionally with a \texttt{.w} extension, that contains both program fragments and commentary. CWEB's \texttt{ctangle} and \texttt{cweave} utilities are then applied to the same source file to create executable program text and documentation respectively.\source{5}

\section{WEB vs. CWEB}

The original WEB system targeted Pascal and TeX. CWEB is the better-known modern descendant for C-family projects. Knuth and Levy's official CWEB page describes CWEB as a version of WEB for documenting C, C++, and Java programs, noting that Levy adapted WEB to C in 1987 and that later revisions added support for C++ and ANSI C.\source{2}

\begin{center}
\begin{tabularx}{0.95\linewidth}{>{\bfseries}p{0.18\linewidth} X X}
\toprule
System & Language orientation & Processors \\
\midrule
WEB & Pascal plus TeX documentation & \texttt{TANGLE} and \texttt{WEAVE} \\
CWEB & C, C++, and related C-family documentation workflows & \texttt{CTANGLE} and \texttt{CWEAVE} \\
\bottomrule
\end{tabularx}
\end{center}

For modern C and C++ projects, the usual entry point is CWEB rather than original Pascal-oriented WEB.

\section{Minimal CWEB-Style Example}

The following illustrative fragment is not intended to be a full CWEB tutorial, but it shows the basic style: prose introduces the idea first, then the code fragment is named and connected to other fragments.

\begin{lstlisting}[language=C,caption={Illustrative CWEB-style structure}]
@* A tiny counting program.
This section introduces the top-level structure before showing low-level details.
The program opens an input stream, scans characters, and counts lines.

@c
#include <stdio.h>

int main(void) {
  int c;
  long lines = 0;
  @<Scan input and count newlines@>;
  printf("%ld\n", lines);
  return 0;
}

@ The scan is isolated because it is easier to discuss separately.
@<Scan input and count newlines@>=
while ((c = getchar()) != EOF) {
  if (c == '\n') {
    ++lines;
  }
}
\end{lstlisting}

The human-facing document can explain the scan before or after the top-level outline. The tangled program, however, is assembled into the order expected by the compiler.

\section{Engineering Value}

Literate programming is valuable when a program's correctness depends on understanding nontrivial design choices, algorithms, invariants, or domain rules. It is especially useful for:

\begin{itemize}
  \item algorithm-heavy systems where the proof or reasoning is as important as the implementation;
  \item security-sensitive or safety-sensitive code where assumptions and invariants must be explicit;
  \item research software where reproducibility depends on both explanation and executable artifacts;
  \item libraries whose internal architecture must be understood by future maintainers; and
  \item legacy systems where documentation drift is a major risk.
\end{itemize}

\callout{Practical interpretation}{%
WEB is not merely a pretty-printer or comment extractor. Its central discipline is narrative control: the author chooses the explanatory order first, then lets the tool recover the compiler order.}

\section{Where WEB Fits Relative to Modern Tools}

Modern systems such as Markdown notebooks, executable notebooks, documentation generators, and static-site documentation pipelines borrow some literate-programming ideas, but they are not always equivalent to WEB.

\begin{center}
\begin{tabularx}{0.95\linewidth}{>{\bfseries}p{0.24\linewidth} X X}
\toprule
Approach & Similarity to WEB & Important difference \\
\midrule
Documentation generators & Produce formatted documentation from source-related material. & Usually follow the code's structure; WEB lets the narrative structure control presentation. \\
Executable notebooks & Combine prose, code, and output in one document. & Usually emphasize interactive execution and analysis, not compiler-order reconstruction through named modules. \\
Markdown plus code blocks & Easy to read and write, often used in docs-as-code pipelines. & Code blocks are often illustrative unless extra tooling extracts and validates them. \\
CWEB / WEB & One source produces both implementation and documentation. & Requires a stricter authoring model and specialized processing tools. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Recommended Modern Workflow for C/C++}

For a small C or C++ component, a disciplined CWEB workflow can look like this:

\begin{enumerate}
  \item Write a \texttt{component.w} source file in explanatory order.
  \item Use named modules for major algorithmic steps, state transitions, invariants, and boundary cases.
  \item Run \texttt{ctangle component.w} to generate compiler-ready source.
  \item Compile the generated source with the normal C/C++ toolchain.
  \item Run \texttt{cweave component.w} to produce TeX documentation.
  \item Format the generated TeX output and publish it alongside the build artifacts.
  \item Treat both generated source and generated documentation as build outputs; treat the \texttt{.w} file as the authoritative source.
\end{enumerate}

\section{Source Quality Guidance}

For citation-clean technical writing about WEB and literate programming, prefer primary or authoritative sources.

\begin{center}
\begin{tabularx}{0.95\linewidth}{>{\bfseries}p{0.28\linewidth} X}
\toprule
Preferred sources & Why \\
\midrule
Knuth's Stanford literate-programming page & Direct statement from Knuth on the methodology and its purpose. \\
Knuth and Levy's Stanford CWEB page & Direct source for what CWEB is and which languages it targets. \\
The 1984 \emph{Computer Journal} article & Canonical peer-reviewed publication that introduced and explained literate programming. \\
CTAN/TUG WEB and CWEB materials & Authoritative distribution and manual sources for WEB/CWEB tooling. \\
\bottomrule
\end{tabularx}
\end{center}

Avoid relying on Wikipedia, YouTube videos, or generic blog posts for core claims. Those sources can be useful for orientation, but they should not be the evidentiary base for a technical handout.

\section{Corrected One-Paragraph Definition}

\callout{Clean definition}{%
\texttt{WEB} is Donald E. Knuth's original literate-programming system. It combines explanatory prose, definitions, and program fragments in a single structured source file. \texttt{TANGLE} transforms that source into compiler-ready program text, while \texttt{WEAVE} transforms it into typeset TeX documentation. The later \texttt{CWEB} system, developed by Knuth and Silvio Levy, adapts the same idea to C-family programming. WEB's core principle is that software should be organized for human understanding first, while still generating exact machine-oriented source code.}

\section*{References}

\begingroup
\small
\setlength{\emergencystretch}{3em}
\begin{enumerate}[label={[{\arabic*}]},leftmargin=2.2em]
  \item Donald E. Knuth. ``Literate Programming.'' Stanford University. \url{https://www-cs-faculty.stanford.edu/~knuth/lp.html}
  \item Donald E. Knuth and Silvio Levy. ``CWEB.'' Stanford University. \url{https://www-cs-faculty.stanford.edu/~knuth/cweb.html}
  \item Donald E. Knuth. ``Literate Programming.'' \emph{The Computer Journal}, Volume 27, Issue 2, 1984, pages 97--111. DOI: \url{https://doi.org/10.1093/comjnl/27.2.97}
  \item Donald E. Knuth. \emph{The WEB System of Structured Documentation}. CTAN/TUG PDF archive. \url{https://tug.ctan.org/info/knuth-pdf/web/webman.pdf}
  \item CTAN. ``Directory \texttt{web/c\_cpp/cweb}.'' Includes CWEB user manual and tool materials. \url{https://ctan.org/tex-archive/web/c_cpp/cweb}
\end{enumerate}
\endgroup

\end{document}
