
\documentclass[11pt,a4paper]{article}

% --- Page + typography ---
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{parskip}
\usepackage{setspace}
\setstretch{1.05}

% --- Colors + links ---
\usepackage[dvipsnames]{xcolor}
\definecolor{cardframe}{RGB}{33,37,41}
\definecolor{cardback}{RGB}{248,249,250}
\definecolor{accent}{named}{MidnightBlue}

\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=accent,urlcolor=accent,citecolor=accent}

% --- Structure + lists ---
\usepackage{enumitem}
\setlist{itemsep=3pt, topsep=4pt, leftmargin=1.3em,labelsep=0.6em}

% --- Boxes + tables ---
\usepackage[most]{tcolorbox}
\tcbset{enhanced, breakable}
\usepackage{tabularx}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{amsmath,amssymb}

% --- Column helpers ---
\newcolumntype{L}[1]{>{\RaggedLeft\arraybackslash\bfseries}p{#1}}
\newcolumntype{Y}{>{\RaggedRight\arraybackslash}X}

% --- Pills / badges ---
\newtcbox{\pill}{on line, arc=3pt, boxsep=0.7pt, left=4pt,right=4pt,top=1pt,bottom=1pt,
  colframe=gray!60, colback=gray!15, boxrule=0.3pt}
\newcommand{\badge}[1]{\pill{\footnotesize #1}}

% --- Checkbox ---
\newcommand{\cb}{\(\square\)}

% --- Reusable captions ---
\newcommand{\DoR}{\textbf{Definition of Ready:} Persona clear; AC drafted; Dependencies known; Estimate set.}
\newcommand{\DoD}{\textbf{Definition of Done:} All ACs pass; Tests green; Security/a11y checks; Docs updated; Deployed/flagged.}

% --- StoryCard environment ---
\newenvironment{StoryCard}[9]{%
  \begin{tcolorbox}[
    colback=white,
    colframe=gray!60,
    boxrule=0.6pt,
    arc=2pt,
    title={\large \textbf{#1} --- \textbf{#2}},
    fonttitle=\sffamily,
    attach boxed title to top left={xshift=6pt,yshift*=-2pt},
    boxed title style={colback=white, colframe=white, boxrule=0pt, left=2pt, right=2pt},
    borderline west={2pt}{0pt}{accent},
    before skip=6pt, after skip=10pt
  ]
  \small
  \begin{tabularx}{\textwidth}{@{}L{3.2cm}Y@{}}
    Epic / Feature          & #3 \\
    Business Value          & #4 \\
    Priority / Estimate     & \badge{Priority: #5}\ \badge{SP: #6} \\
    Persona                 & #7 \\
    Dependencies            & #8 \\
    Assumptions / Risks     & #9 \\
  \end{tabularx}

  \medskip
  \textbf{Story}\quad
  \emph{As a #7, I want to #2 so that #4.}

  \medskip
}{%
  \end{tcolorbox}
}

% --- Non-functional tag line ---
\newcommand{\NonFunctional}[1]{%
  \textbf{Non-Functional}\quad #1
}

% --- Acceptance block ---
\newenvironment{Acceptance}{%
  \medskip
  \textbf{Acceptance Criteria (BDD)}\\
}{}

% --- Scenario helper ---
\newcommand{\Scenario}[4]{%
  \textbf{Scenario}\quad #1\\
  \textbf{Given}\quad #2\\
  \textbf{When}\quad #3\\
  \textbf{Then}\quad #4\\
}

% --- Tasks box ---
\newenvironment{TasksBox}{%
  \begin{tcolorbox}[
    title={\normalsize \textbf{Tasks}},
    colback=white,
    colframe=gray!60,
    borderline west={1.8pt}{0pt}{accent},
    arc=2pt, boxrule=0.4pt,
    left=10pt,right=10pt,top=6pt,bottom=6pt,
    before skip=6pt, after skip=10pt
  ]
  \small
  \begin{itemize}[label=\cb, leftmargin=*, itemsep=4pt, topsep=2pt, parsep=0pt]
}{%
  \end{itemize}
  \end{tcolorbox}
}

\title{\textbf{Dynamic Project Management Board — Micro Frontends}\\Tailored User Stories (Card Style)}
\author{}
\date{\today}

\begin{document}
\maketitle
\tableofcontents
\newpage
\section*{A. Board View Micro Frontend}\addcontentsline{toc}{section}{A. Board View Micro Frontend}

% --- MFE-BV-001 ---
\begin{StoryCard}
{MFE-BV-001}{Create Columns \& Define WIP Limits}
{Board Configuration}
{Establish workflow visualization with enforced WIP to reduce overload and cycle time}
{Must}{3}
{Board Admin}
{Build tooling, column schema, policy checks}
{Duplicate names avoided; risk of initial misconfiguration mitigated with rename/archive}
  \NonFunctional{\badge{Performance} \badge{Security} \badge{Reliability} \badge{Accessibility}}

  \begin{Acceptance}
  \Scenario{Happy path}
    {an empty board is available}
    {I add a column named \emph{In Progress} and set WIP=3}
    {the column appears; WIP badge displays; reorder persists after refresh}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Schema: column model (id, name, order, wipLimit, archivedAt).
  \item UI: header actions (Add, Rename, Reorder via drag, Archive).
  \item Persist: POST/PUT/PATCH endpoints or federated API contract.
  \item Validation: prevent duplicates within a board; soft delete columns.
  \item WIP indicator: visual badge and ARIA announcement on exceed.
\end{TasksBox}

% --- MFE-BV-002 ---
\begin{StoryCard}
{MFE-BV-002}{Drag-and-Drop Cards Across Columns}
{Card Movement}
{Fast status changes with optimistic updates to improve collaboration}
{Must}{5}
{Team Member}
{Server PATCH for status; event bus for broadcasts}
{WIP enforcement is advisory; rollback on failure to avoid inconsistency}
  \NonFunctional{\badge{Performance} \badge{Reliability} \badge{Accessibility}}

  \begin{Acceptance}
  \Scenario{Optimistic update}
    {a card is in \emph{Todo}}
    {I drop it in \emph{In Progress}}
    {status updates immediately; on failure the UI rolls back with an error}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Integrate drag library with keyboard support (Space/Enter to pick/place).
  \item Optimistic mutation and rollback logic.
  \item PATCH \texttt{/cards/\{id\}} status change.
  \item Real-time broadcast to other clients.
\end{TasksBox}

% --- MFE-BV-003 ---
\begin{StoryCard}
{MFE-BV-003}{Board Filters \& Saved Views}
{Discoverability}
{Quickly surface relevant work for stakeholders; sharable default views}
{Should}{5}
{Product Manager}
{Saved views API; permissions}
{Links need stability across releases; avoid leaking private filters}
  \NonFunctional{\badge{Performance} \badge{Security} \badge{Reliability} \badge{Accessibility}}

  \begin{Acceptance}
  \Scenario{Saved view}
    {filters are applied}
    {I click ``Save View'' and share the link}
    {others opening the link see the same filters if they have access}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Filter builder component; URL-encoded query state.
  \item Saved views API (create/update/delete, permissions).
  \item Share links with short IDs and access checks.
\end{TasksBox}

% --- MFE-BV-004 ---
\begin{StoryCard}
{MFE-BV-004}{High-Density Virtualized Card List}
{Performance}
{Maintain smooth scrolling on columns with hundreds of cards}
{Must}{8}
{Team Member}
{Virtualization library; measurement utilities}
{Risk of jank on resize; recycle DOM nodes}
  \NonFunctional{\badge{Performance} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Virtualized scroll}
    {1{,}500 cards exist in a column}
    {I scroll continuously}
    {frames remain $\ge 50$ FPS; virtualization recalculates on resize}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Windowed list + item measurement; shimmer placeholders for offscreen items.
  \item Prefetch next viewport page; recycle DOM nodes.
  \item Performance budget doc and profiler checkpoints.
\end{TasksBox}

% --- MFE-BV-005 ---
\begin{StoryCard}
{MFE-BV-005}{Keyboard Navigation \& A11y}
{Accessibility}
{Enable efficient mouseless use and screen reader support}
{Must}{3}
{Keyboard/Screen Reader User}
{Shortcut map; ARIA roles; focus management}
{Accessibility regressions prevented via automated checks}
  \NonFunctional{\badge{Accessibility} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Keyboard flow}
    {focus is on a card}
    {I use Arrow keys and shortcuts}
    {focus moves appropriately; quick edit opens with ``E''; actions are announced}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Focus ring management; roving \texttt{tabindex}.
  \item ARIA roles for lists, items, and drag regions.
  \item Keyboard help overlay; automated a11y tests.
\end{TasksBox}

% --- MFE-BV-006 ---
\begin{StoryCard}
{MFE-BV-006}{Swimlanes (Group by Assignee/Priority)}
{Visualization}
{Scan workload distribution via lane grouping}
{Could}{3}
{Scrum Master}
{Group-by transformer; totals per lane}
{Consistency maintained under concurrent updates}
  \NonFunctional{\badge{Performance} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Lane toggle}
    {a board is open}
    {I toggle swimlanes by Assignee}
    {columns render rows per assignee; filters remain respected}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Group-by transformer with virtualization across lanes.
  \item Lane headers with totals and collapse/expand.
  \item Update counts on events.
\end{TasksBox}

\section*{B. Task Details Micro Frontend}\addcontentsline{toc}{section}{B. Task Details Micro Frontend}

% --- MFE-TD-001 ---
\begin{StoryCard}
{MFE-TD-001}{Open Task in Modal/Side Panel}
{Task Viewer}
{Quick context without leaving the board via modal/side panel}
{Must}{3}
{Team Member}
{Router integration; deep link patterns}
{Return focus to invoking card; ensure responsive layout}
  \NonFunctional{\badge{Performance} \badge{Accessibility} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Open panel}
    {a card exists}
    {I click the card or navigate to /task/\{id\}}
    {panel opens with title, status, assignee, tags, due date}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Router state + deep linking.
  \item Fetch task + relations; skeleton loading.
  \item Close/escape handling and focus management.
\end{TasksBox}

% --- MFE-TD-002 ---
\begin{StoryCard}
{MFE-TD-002}{Edit Core Fields \& Markdown Description}
{Task Editing}
{Keep task details accurate with markdown preview}
{Must}{5}
{Team Member}
{PATCH endpoints; sanitizer}
{Prevent XSS via markdown; enforce CSP}
  \NonFunctional{\badge{Security} \badge{Reliability} \badge{Performance}}

  \begin{Acceptance}
  \Scenario{Inline edit}
    {edit mode is enabled}
    {I update title/status and toggle preview}
    {changes save; markdown renders (headings, code, links)}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Inline editable fields with validation.
  \item Markdown editor with preview; sanitize HTML.
  \item Optimistic update and error handling.
\end{TasksBox}

% --- MFE-TD-003 ---
\begin{StoryCard}
{MFE-TD-003}{Checklists \& Subtasks}
{Task Structure}
{Track granular work through subtasks and checklists}
{Should}{5}
{QA Engineer}
{Subtask model; progress calc}
{Concurrent edits must resolve deterministically}
  \NonFunctional{\badge{Reliability} \badge{Performance}}

  \begin{Acceptance}
  \Scenario{Progress}
    {a task exists}
    {I add subtasks and mark items complete}
    {parent shows progress \% on card and panel}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Subtask CRUD and ordering.
  \item Checklist component with keyboard support.
  \item Progress calculation and display.
\end{TasksBox}

% --- MFE-TD-004 ---
\begin{StoryCard}
{MFE-TD-004}{Comments with @Mentions \& Attachments}
{Collaboration}
{Enable in-context collaboration via mentions and uploads}
{Should}{3}
{Developer}
{Mention typeahead; upload pipeline}
{GDPR/PII controls for images and text}
  \NonFunctional{\badge{Security} \badge{Reliability} \badge{Performance}}

  \begin{Acceptance}
  \Scenario{Mention \& upload}
    {the comment box is focused}
    {I type @ and select a user; I upload a file}
    {mention resolves and notifies; file appears with preview/download}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Rich comment input with typeahead.
  \item Upload component with progress; resumable on blips.
  \item Webhooks to Notifications for mentions.
\end{TasksBox}

% --- MFE-TD-005 ---
\begin{StoryCard}
{MFE-TD-005}{Relations: Dependencies, Duplicates, Links}
{Task Relationships}
{Coordinate sequencing and avoid rework}
{Could}{3}
{Project Manager}
{Relation types \texttt{blocks/blockedBy/duplicateOf/relatesTo}}
{Integrity checks to prevent circular relations}
  \NonFunctional{\badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Relations}
    {a task is open}
    {I add a ``blocked by'' relation and a duplicate link}
    {badges show with tooltips; both tasks cross-link}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Badges with tooltips and click-through.
  \item Relation CRUD and integrity checks.
  \item Analytics hook to exclude blocked time (later).
\end{TasksBox}

\section*{C. Analytics Dashboard Micro Frontend}\addcontentsline{toc}{section}{C. Analytics Dashboard Micro Frontend}

% --- MFE-AD-001 ---
\begin{StoryCard}
{MFE-AD-001}{Throughput \& Velocity Charts}
{Core Charts}
{Forecast capacity with completed-per-week and sprint velocity}
{Should}{5}
{Engineering Manager}
{Aggregation API; chart components}
{Timezone/locale aware bucketing; export options}
  \NonFunctional{\badge{Performance} \badge{Accessibility} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Throughput over 12 weeks}
    {a date range is selected}
    {I choose last 12 weeks}
    {bar chart renders weekly totals; velocity shows points per sprint when enabled}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Completed tasks by week aggregation.
  \item Chart export (PNG/CSV); data table for screen readers.
  \item Access control by board.
\end{TasksBox}

% --- MFE-AD-002 ---
\begin{StoryCard}
{MFE-AD-002}{Lead \& Cycle Time Analysis}
{Flow Metrics}
{Spot variability and bottlenecks using distributions and percentiles}
{Must}{8}
{Process Lead}
{Event timestamps per status; histogram/scatter charts}
{Outlier toggles should recompute metrics accurately}
  \NonFunctional{\badge{Performance} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Cycle time P50/P85/P95}
    {statuses are selected}
    {I compute cycle time from ``In Progress'' to ``Done''}
    {percentiles display; outlier toggle recomputes}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Cycle time computation service.
  \item Charts for histogram and scatter.
  \item Validation against sample dataset.
\end{TasksBox}

% --- MFE-AD-003 ---
\begin{StoryCard}
{MFE-AD-003}{Cumulative Flow Diagram (CFD)}
{Flow Metrics}
{Monitor WIP stability across statuses over time}
{Should}{5}
{Scrum Master}
{Daily snapshots or backfilled timeseries}
{Colorblind-safe palette and print-friendly mode}
  \NonFunctional{\badge{Performance} \badge{Accessibility}}

  \begin{Acceptance}
  \Scenario{90-day CFD}
    {a board is selected}
    {I view CFD for last 90 days}
    {stacked areas show counts per status; toggling a status updates chart and legend}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Timeseries generation from events.
  \item Stacked area chart with hover and legend.
  \item Export and print styles.
\end{TasksBox}

% --- MFE-AD-004 ---
\begin{StoryCard}
{MFE-AD-004}{Bottleneck Detector \& WIP Alerts}
{Insights}
{Raise automated flags when thresholds are exceeded}
{Could}{3}
{Delivery Lead}
{Rule engine for WIP\% and stagnation hours}
{False positives must be $<5\%$ on benchmark}
  \NonFunctional{\badge{Reliability} \badge{Performance}}

  \begin{Acceptance}
  \Scenario{WIP threshold}
    {WIP limit=3}
    {a column has 5 items for $>24$h}
    {an alert appears and links to offending items}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Rule engine with thresholds and tuning.
  \item Alert list and drill-through.
  \item Push summaries to Notifications.
\end{TasksBox}

% --- MFE-AD-005 ---
\begin{StoryCard}
{MFE-AD-005}{Stakeholder Reports \& Scheduling}
{Reporting}
{Deliver scheduled PDFs/CSVs with key metrics securely}
{Could}{5}
{Executive}
{Report templates; scheduler; secure links}
{Role-restricted; links expire in 7 days}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Weekly delivery}
    {a saved report exists}
    {I schedule weekly delivery}
    {recipients receive the email with attachments; restricted boards are excluded}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Render reports; email via trusted provider.
  \item DMARC/DKIM alignment; unsub links and GDPR compliance.
  \item Job logs and failure notifications.
\end{TasksBox}

\section*{D. User Management Micro Frontend}\addcontentsline{toc}{section}{D. User Management Micro Frontend}

% --- MFE-UM-001 ---
\begin{StoryCard}
{MFE-UM-001}{Invite Members \& Assign Roles}
{Access Control}
{Control who can see and modify boards via role assignment}
{Must}{3}
{Org Admin}
{Invite flow; tokenized links; role matrix}
{Audit all grants/revokes with actor and timestamp}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Invitation}
    {an email address is provided}
    {I send an invite with a role}
    {the user receives a link to join; role takes effect immediately on acceptance}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Invite flow with tokenized links and expiry.
  \item Role definitions and policy checks across MFEs.
  \item Audit log view.
\end{TasksBox}

% --- MFE-UM-002 ---
\begin{StoryCard}
{MFE-UM-002}{Board-Level Permissions}
{Access Control}
{Enforce governance for create/edit/move/delete actions}
{Must}{5}
{Board Admin}
{Policy matrix (role $\times$ action); server enforcement}
{No client-only enforcement; cache policy checks}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Policy gating}
    {a Contributor lacks column-edit rights}
    {they view the board}
    {column actions are disabled/hidden; attempted move is blocked with rationale}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Policy matrix and conditional rendering.
  \item Server-side checks; client flags driven by policy.
  \item E2E tests for blocked actions.
\end{TasksBox}

% --- MFE-UM-003 ---
\begin{StoryCard}
{MFE-UM-003}{SSO Integration (OIDC/SAML)}
{Identity}
{Centralize authentication and provisioning}
{Should}{8}
{IT Admin}
{OIDC/SAML config UI; SCIM jobs}
{Break-glass admin accounts for IdP failures}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{SSO login}
    {OIDC config is enabled}
    {a user signs in}
    {authentication occurs via IdP; SCIM deprovision revokes access within 15 minutes}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Metadata upload and validation.
  \item SCIM sync jobs; deprovision hooks.
  \item Admin alerts on failures.
\end{TasksBox}

% --- MFE-UM-004 ---
\begin{StoryCard}
{MFE-UM-004}{Field-Level Permissions}
{Governance}
{Restrict edits to sensitive fields and record audits}
{Should}{5}
{Compliance Lead}
{Field policy registry; audit trail}
{PII masking where required}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Restricted field}
    {a field is restricted}
    {a user without rights attempts to edit}
    {field renders read-only; change attempts are rejected and logged}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Field policy registry and UI enforcement.
  \item Audit emission on changes.
  \item Summaries available to Analytics.
\end{TasksBox}

% --- MFE-UM-005 ---
\begin{StoryCard}
{MFE-UM-005}{Team \& Board Sharing}
{Collaboration}
{Share boards with teams and manage access at scale}
{Could}{3}
{Team Lead}
{Team directory; share dialog; membership listeners}
{Propagation latency $\le$ 5 minutes across MFEs}
  \NonFunctional{\badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Share with team}
    {a team is selected}
    {I share a board with role=Viewer}
    {all members can view; membership changes update access within 5 minutes}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Share dialog with search and team selection.
  \item Membership listeners to update effective permissions.
  \item Board view permission enforcement.
\end{TasksBox}

\section*{E. Notifications Micro Frontend}\addcontentsline{toc}{section}{E. Notifications Micro Frontend}

% --- MFE-NF-001 ---
\begin{StoryCard}
{MFE-NF-001}{Real-Time Toasts for Key Events}
{Realtime UX}
{Stay aware of changes without losing context}
{Must}{3}
{Any User}
{WebSocket/SSE client; event bus}
{Rate limit to avoid noise; screen reader announcements}
  \NonFunctional{\badge{Performance} \badge{Accessibility} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Toast on card move}
    {another user moves a card while I'm viewing the board}
    {an event is received}
    {a toast appears with a deep link to the card}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Event bus integration.
  \item Toast queue with dedupe and focus management.
  \item Test end-to-end latency $\le$ 2s p95.
\end{TasksBox}

% --- MFE-NF-002 ---
\begin{StoryCard}
{MFE-NF-002}{Activity Feed with Filters}
{Audit \& Review}
{Audit changes and review progress by actor/action/date/board}
{Should}{5}
{Project Manager}
{Feed API with pagination and filters}
{Immutable event storage once written}
  \NonFunctional{\badge{Performance} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Filter feed}
    {filters are set to action=`status-changed'}
    {I scroll the list}
    {only matching entries show; older entries load as I scroll}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Endpoint with pagination and indexes.
  \item UI with saved filters and deep links.
  \item Access-scoped results.
\end{TasksBox}

% --- MFE-NF-003 ---
\begin{StoryCard}
{MFE-NF-003}{Email \& Digest Notifications}
{Channels}
{Immediate emails for mentions and daily digest of changes}
{Could}{3}
{Stakeholder}
{Templates; scheduler; unsubscribe links}
{Comply with GDPR and CAN-SPAM; DMARC aligned}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Digest}
    {preference is set to Daily 9am}
    {the scheduler runs}
    {a summary email is delivered with links to items}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Localized templates (mention, digest).
  \item Scheduler for digests; secure downloads with expiry.
  \item Delivery monitoring and logs.
\end{TasksBox}

% --- MFE-NF-004 ---
\begin{StoryCard}
{MFE-NF-004}{Slack/MS Teams Integrations}
{Channels}
{Post board updates to team chat and provide slash commands}
{Should}{5}
{Team Member}
{OAuth install; outgoing webhooks; slash command endpoint}
{Verify signing secrets; conform to rate limits}
  \NonFunctional{\badge{Security} \badge{Reliability}}

  \begin{Acceptance}
  \Scenario{Assignment message}
    {a channel is connected}
    {a card is assigned to me}
    {a message posts with a deep link; /board search returns results}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item App manifest and credentials.
  \item Webhook handlers and slash commands.
  \item Retry with backoff and signature verification.
\end{TasksBox}

% --- MFE-NF-005 ---
\begin{StoryCard}
{MFE-NF-005}{Notification Preferences \& Rules}
{Personalization}
{Let users configure events and channels to minimize noise}
{Could}{5}
{Any User}
{Preferences storage; rule builder}
{Prevent rule loops; sync across sessions within 5s}
  \NonFunctional{\badge{Reliability} \badge{Performance}}

  \begin{Acceptance}
  \Scenario{Preference toggle}
    {preferences are open}
    {I disable ``card moved'' toasts and add a rule ``due date change on my tasks $\to$ email''}
    {I no longer see those toasts; emails trigger on rule}
  \end{Acceptance}

  \small \DoR\ \ \textbullet\ \ \DoD

\end{StoryCard}

\begin{TasksBox}
  \item Per-channel toggles; cached reads, debounced writes.
  \item Rule builder: conditions (actor, field, board, label) and actions (toast, email, webhook).
  \item Sandbox rule execution to avoid loops.
\end{TasksBox}

\end{document}