\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[margin=0.7in]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{ragged2e}
\usepackage{titlesec}

\hypersetup{hidelinks,pdfauthor={OpenAI},pdftitle={Algorithmic Stock Trading System Startup Roadmap}}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlength{\tabcolsep}{4pt}
\setlength{\LTpre}{0pt}
\setlength{\LTpost}{0pt}
\setlength{\emergencystretch}{3em}
\pagestyle{plain}
\renewcommand{\arraystretch}{1.15}
\titleformat{\section}{\large\bfseries}{}{0pt}{}
\titlespacing*{\section}{0pt}{10pt}{4pt}

\newcolumntype{Y}{>{\RaggedRight\arraybackslash}X}
\newcommand{\statusdone}{Done}
\newcommand{\statusprogress}{In Progress}
\newcommand{\statusnotstarted}{Not Started}

\begin{document}

\begin{center}
    {\LARGE\bfseries Algorithmic Stock Trading System Startup Roadmap}\\[4pt]
    {\large SBA 10-Step Startup Sequence}\\[6pt]
    {\normalsize Founder planning handout for building a disciplined trading-system business}
\end{center}

\vspace{4pt}
\section*{Business Snapshot}
{\small
\textbf{Startup Name:} Algorithmic Stock Trading System\par
\textbf{Working Business Idea:} Build a disciplined algorithmic stock trading platform that unifies market data ingestion, research, backtesting, paper trading, live execution, and risk controls into one repeatable operating system.\par
\textbf{Problem:} Independent systematic traders often rely on disconnected scripts, spreadsheets, broker dashboards, and notebooks, which makes strategy validation, execution control, and risk management inconsistent.\par
\textbf{Target Customer:} Initial target customer is the founder-operator. Later expansion could serve advanced independent traders, small quantitative research teams, and boutique trading operations that need disciplined execution and research tooling.\par
\textbf{Value Proposition:} Provide a single workflow from data to signal to order to review so strategy development is more reproducible, execution is more controlled, and risk management is enforced by system design instead of manual discipline alone.\par
\textbf{Revenue Model:} Near-term value comes from founder-operated trading-system capability. Longer-term options may include subscription software, premium analytics, hosted tooling, or research and execution support services.\par
\textbf{Current Stage:} Validation, MVP definition, and founder-operated build.\par
\textbf{Current Focus:} Define the first strategy wedge, build the research and backtesting foundation, stand up paper trading, and codify hard risk controls before deploying real capital.\par
}

\section*{SBA-Aligned Startup Roadmap}
{\footnotesize
\rowcolors{2}{gray!6}{white}
\begin{longtable}{@{}>{\RaggedRight\arraybackslash}p{0.15\textwidth}>{\RaggedRight\arraybackslash}p{0.10\textwidth}>{\RaggedRight\arraybackslash}p{0.22\textwidth}>{\RaggedRight\arraybackslash}p{0.21\textwidth}>{\RaggedRight\arraybackslash}p{0.24\textwidth}@{}}
\rowcolor{gray!20}
\textbf{SBA Step} & \textbf{Status} & \textbf{Algorithmic Trading System Application} & \textbf{Immediate Next Action} & \textbf{Evidence / Output to Produce} \\
\endfirsthead
\rowcolor{gray!20}
\textbf{SBA Step} & \textbf{Status} & \textbf{Algorithmic Trading System Application} & \textbf{Immediate Next Action} & \textbf{Evidence / Output to Produce} \\
\endhead
Conduct market research & \statusprogress & Define the specific operating pain for systematic traders: fragmented research, weak execution discipline, and inconsistent risk controls. Narrow the first user profile and workflow. & Interview or self-document the founder workflow, compare current tools, and identify the minimum differentiators for the first release. & Problem statement, user profile, workflow map, competitor comparison, and MVP wedge definition. \\
Write your business plan & \statusprogress & Convert the working idea into a concise operating plan covering product scope, customer progression, delivery model, cost assumptions, and go-to-market sequencing. & Draft a lean business plan anchored on one strategy-development workflow and one target operator persona. & Lean business plan, one-page operating model, milestone map, and startup cost assumptions. \\
Fund your business & \statusnotstarted & Determine how the build will be financed without mixing platform planning with trading recommendations. Separate software build costs from any eventual trading capital and compliance costs. & Estimate the founder build budget, operating runway, data and infrastructure costs, and any professional-service costs. & Startup budget, funding plan, expense categories, and runway worksheet. \\
Pick your business location & \statusnotstarted & Decide the operating jurisdiction for the business entity, banking, tax nexus, contractor use, and software hosting footprint. A remote-first software business can still require a deliberate home-state decision. & Select the initial state of formation and operating address strategy, then record why that location fits the business model. & Location decision memo, jurisdiction checklist, and operating-address plan. \\
Choose a business structure & \statusnotstarted & Select the legal form that best fits a founder-operated software business that may later evolve into SaaS, analytics, or execution-support services. & Review sole proprietorship, LLC, and corporation options with counsel and decide the preferred initial structure. & Structure comparison sheet, liability and tax questions list, and formation decision record. \\
Choose your business name & \statusprogress & The project has a working name, but brand and legal clearance still need to be completed before formal launch. The name should support disciplined trading-system tooling rather than imply advisory promises. & Decide whether to keep the current working name or adopt a product brand, then run name, domain, and trademark screening. & Naming decision, short brand rationale, domain shortlist, and clearance checklist. \\
Register your business & \statusnotstarted & Formal registration is required before contracting, banking, and vendor setup under a separate business identity. & File the entity after confirming location, structure, and naming choices. & Formation filing receipt, organizational documents, and registered-business record. \\
Get federal and state tax IDs & \statusnotstarted & Tax IDs are needed for banking, vendor onboarding, payroll or contractor administration, and clean separation between founder and business operations. & Obtain the EIN after registration and determine whether any state tax IDs are required for the chosen jurisdiction. & EIN confirmation, state tax registration notes, and records checklist. \\
Apply for licenses and permits & \statusnotstarted & Determine what licenses, permits, registrations, or professional reviews may apply to the actual business model, especially if the platform later expands beyond founder use into commercial services. & Build a counsel-supported compliance checklist for business operations, data use, and any sector-specific obligations before launch. & License and permit decision log, assumptions register, and compliance review checklist. \\
Open a business bank account & \statusnotstarted & A separate account is needed for bookkeeping discipline, vendor payments, software expenses, and auditability of the business itself. & Open a dedicated business checking account once registration and tax ID steps are complete. & Bank account opened, signer policy, expense categories, and bookkeeping setup checklist. \\
\end{longtable}
}

\section*{Founder Priorities}
\begin{enumerate}[leftmargin=1.5em,itemsep=3pt]
    \item Finish the minimum viable research and backtesting environment with reproducible data, signals, and performance reports.
    \item Define non-negotiable risk rules for exposure, position sizing, drawdown, and kill-switch conditions.
    \item Validate one or two strategies in paper trading with realistic assumptions for slippage, costs, and execution quality.
\end{enumerate}

\section*{Practical First-Week Action Plan}
\begin{enumerate}[leftmargin=1.5em,itemsep=3pt]
    \item Write a one-page problem memo describing the current fragmented founder workflow and the exact pain points the platform must eliminate.
    \item Choose one initial strategy wedge and document its required data inputs, research loop, backtest outputs, and paper-trading acceptance criteria.
    \item Create a reproducible repository structure for data ingestion, strategy research, backtesting, reporting, and environment configuration.
    \item Define hard risk rules for maximum position size, portfolio exposure, daily loss, drawdown threshold, and emergency kill-switch triggers.
    \item Run one baseline paper-trading simulation or dry-run workflow using realistic assumptions for slippage, fees, and execution timing.
    \item Draft the lean business plan and startup cost worksheet so legal, banking, and formation decisions can be sequenced cleanly.
    \item Start a formal naming and entity-setup checklist covering brand clearance, business structure review, registration, EIN, and banking prerequisites.
\end{enumerate}

\section*{Notes and Boundaries}
This handout is for startup planning and operational execution. It is not investment advice, legal advice, tax advice, or securities compliance advice. It does not provide stock picks, trading recommendations, or performance claims. The SBA step order reflected here follows the live SBA ``10 steps to start your business'' sequence as reviewed on May 6, 2026.

\vfill
{\footnotesize\textit{Status interpretation:} ``In Progress'' indicates documented planning or active founder work already underway. ``Not Started'' indicates steps that depend on formal filing, legal setup, tax registration, banking, licensing, or other actions not clearly completed.}

\end{document}
