\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{titlesec}

\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}
\setlength{\headheight}{14pt}
\emergencystretch=3em
\pagestyle{fancy}
\fancyhf{}
\lhead{SuiteCRM Self-Hosting Handout}
\rhead{Internal Operations}
\cfoot{\thepage}

\definecolor{lightgray}{gray}{0.95}
\lstset{basicstyle=\ttfamily\small, backgroundcolor=\color{lightgray}, breaklines=true, frame=single, columns=fullflexible}
\setlist[itemize]{topsep=3pt,itemsep=2pt,parsep=0pt}
\setlist[enumerate]{topsep=3pt,itemsep=2pt,parsep=0pt}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\titleformat{\section}{\Large\bfseries}{\thesection}{0.5em}{}
\titleformat{\subsection}{\large\bfseries}{\thesubsection}{0.5em}{}

\title{SuiteCRM Self-Hosting Operations Handout}
\author{Prepared for deployment planning and operational runbooks}
\date{May 2026}

\begin{document}
\maketitle

\section*{Scope}
This handout summarizes a production-oriented self-hosting model for SuiteCRM 8.x. The default target is SuiteCRM 8.10.x on Linux with Apache 2.4, PHP 8.2--8.4, and MariaDB or MySQL. Nginx/LEMP can be used as an operator pattern only after route, rewrite, API, upload, scheduler, and upgrade validation, because the current SuiteCRM 8.10.x compatibility matrix centers Apache 2.4.

\section{Reference Architecture}
\begin{itemize}
  \item Users access SuiteCRM over HTTPS through DNS and a web server or reverse proxy.
  \item The web server should expose SuiteCRM's \texttt{public} directory, not the full application root.
  \item PHP executes the SuiteCRM application and required extensions.
  \item MariaDB or MySQL stores CRM records, users, metadata, and module data.
  \item The filesystem stores uploads, cache, logs, and customizations.
  \item The scheduler runs recurring background jobs such as workflows, email checks, report generation, and indexing.
  \item Backups must include both the database and application file tree.
\end{itemize}

\section{Version and Runtime Baseline}
\begin{longtable}{L{0.30\linewidth} L{0.60\linewidth}}
\textbf{Area} & \textbf{Default Baseline} \\
\hline
SuiteCRM & SuiteCRM 8.10.x for new SuiteCRM 8 deployments. \\
PHP & PHP 8.2, 8.3, or 8.4 for SuiteCRM 8.10.x. \\
Web server & Apache 2.4 for strict official compatibility alignment. \\
Database & MariaDB 10.6, 10.11, 11.4, 11.8 or MySQL 8.0, 8.4 for SuiteCRM 8.10.x. \\
PHP modules & cli, curl, common, intl, json, gd, mbstring, mysqli, pdo\_mysql, openssl, soap, xml, zip; imap and ldap if needed. \\
Performance & Enable OPcache in production after validation. \\
\end{longtable}

\section{Deployment Procedure}
\subsection{Install runtime packages}
\begin{lstlisting}[language=bash]
sudo apt update
sudo apt install -y apache2 mariadb-server unzip curl ca-certificates
sudo apt install -y php php-cli php-common php-curl php-intl php-gd php-mbstring \
  php-mysql php-soap php-xml php-zip php-opcache
sudo apt install -y php-imap php-ldap   # optional
sudo a2enmod rewrite headers ssl
sudo systemctl restart apache2
\end{lstlisting}

\subsection{Create database}
\begin{lstlisting}[language=SQL]
CREATE DATABASE suitecrm CHARACTER SET utf8mb4 COLLATE utf8mb4_unicode_ci;
CREATE USER 'suitecrm'@'localhost' IDENTIFIED BY 'REPLACE_WITH_STRONG_PASSWORD';
GRANT ALL PRIVILEGES ON suitecrm.* TO 'suitecrm'@'localhost';
FLUSH PRIVILEGES;
\end{lstlisting}

\subsection{Configure Apache vhost}
\begin{lstlisting}
<VirtualHost *:80>
    ServerName crm.example.com
    DocumentRoot /var/www/suitecrm/public

    <Directory /var/www/suitecrm/public>
        AllowOverride All
        Require all granted
    </Directory>

    ErrorLog ${APACHE_LOG_DIR}/suitecrm_error.log
    CustomLog ${APACHE_LOG_DIR}/suitecrm_access.log combined
</VirtualHost>
\end{lstlisting}

\subsection{Run installer}
The UI installer is appropriate for first-time manual setup. The CLI installer is better for repeatable builds. Example CLI pattern:

\begin{lstlisting}[language=bash]
cd /var/www/suitecrm
sudo -u www-data ./bin/console suitecrm:app:install \
  -u "admin" \
  -p "REPLACE_WITH_ADMIN_PASSWORD" \
  -U "suitecrm" \
  -P "REPLACE_WITH_DB_PASSWORD" \
  -H "127.0.0.1" \
  -N "suitecrm" \
  -S "https://crm.example.com" \
  -d "no"
\end{lstlisting}

\section{Security Hardening Gate}
\begin{longtable}{L{0.08\linewidth} L{0.82\linewidth}}
\textbf{Done} & \textbf{Control} \\
\hline
$\square$ & HTTPS is enabled and HTTP redirects to HTTPS. \\
$\square$ & The web server document root points to \texttt{/var/www/suitecrm/public}. \\
$\square$ & Database ports are not exposed publicly. \\
$\square$ & SuiteCRM database user is scoped to the SuiteCRM database. \\
$\square$ & Login throttling remains enabled or is explicitly tuned through approved configuration. \\
$\square$ & Trusted hosts are configured for the production hostname. \\
$\square$ & Administrative users have 2FA enabled. \\
$\square$ & Scheduler runs as an allowed non-root user. \\
$\square$ & Backups are off-host, encrypted, and restore-tested. \\
\end{longtable}

\section{Scheduler Runbook}
SuiteCRM scheduled jobs must be manually enabled. Use cron or a systemd timer, and run as an allowed non-root user. Example cron entry:

\begin{lstlisting}[language=bash]
sudo crontab -u www-data -e
\end{lstlisting}

\begin{lstlisting}
* * * * * cd /var/www/suitecrm && ./bin/console schedulers:run >> /var/log/suitecrm-scheduler.log 2>&1
\end{lstlisting}

Verification:
\begin{lstlisting}[language=bash]
cd /var/www/suitecrm
sudo -u www-data ./bin/console schedulers:run -vvv
sudo tail -n 100 /var/log/suitecrm-scheduler.log
\end{lstlisting}

\section{Backup and Restore Runbook}
\subsection{Backup scope}
\begin{itemize}
  \item Database dump or database snapshot.
  \item Application file tree, including configuration, customizations, uploads, cache, and logs as required.
  \item Web server, PHP, cron, and systemd configuration.
\end{itemize}

\subsection{Database backup example}
\begin{lstlisting}[language=bash]
BACKUP_DIR=/srv/backups/suitecrm
DATE=$(date +%Y%m%d-%H%M%S)
mkdir -p "$BACKUP_DIR"
mysqldump --single-transaction --routines --triggers suitecrm \
  | gzip > "$BACKUP_DIR/suitecrm-db-$DATE.sql.gz"
sha256sum "$BACKUP_DIR/suitecrm-db-$DATE.sql.gz" \
  > "$BACKUP_DIR/suitecrm-db-$DATE.sql.gz.sha256"
\end{lstlisting}

\subsection{File backup example}
\begin{lstlisting}[language=bash]
BACKUP_DIR=/srv/backups/suitecrm
DATE=$(date +%Y%m%d-%H%M%S)
tar --acls --xattrs -czf "$BACKUP_DIR/suitecrm-files-$DATE.tar.gz" \
  -C /var/www suitecrm
sha256sum "$BACKUP_DIR/suitecrm-files-$DATE.tar.gz" \
  > "$BACKUP_DIR/suitecrm-files-$DATE.tar.gz.sha256"
\end{lstlisting}

\section{Operational Acceptance Criteria}
\begin{itemize}
  \item SuiteCRM is reachable at the canonical HTTPS URL.
  \item Installer pre-checks are passed or approved warnings are documented.
  \item A test record can be created and retrieved.
  \item Scheduler last-run timestamps update.
  \item Admin 2FA is enabled.
  \item Database and file backups complete.
  \item Restore has been tested in a non-production environment.
\end{itemize}

\section{References}
\begin{itemize}
  \item SuiteCRM 8.x Compatibility Matrix: \url{https://docs.suitecrm.com/8.x/admin/compatibility-matrix/}
  \item SuiteCRM 8.x Webserver Setup Guide: \url{https://docs.suitecrm.com/8.x/admin/installation-guide/webserver-setup-guide/}
  \item SuiteCRM 8.x Downloading and Installing: \url{https://docs.suitecrm.com/8.x/admin/installation-guide/downloading-installing/}
  \item SuiteCRM 8.x UI Installer: \url{https://docs.suitecrm.com/8.x/admin/installation-guide/running-the-ui-installer/}
  \item SuiteCRM 8.x CLI Installer: \url{https://docs.suitecrm.com/8.x/admin/installation-guide/running-the-cli-installer/}
  \item SuiteCRM 8.x Schedulers: \url{https://docs.suitecrm.com/8.x/admin/administration-panel/schedulers/}
  \item SuiteCRM 8.x Performance: \url{https://docs.suitecrm.com/8.x/admin/installation-guide/performance/}
  \item SuiteCRM 8.x Login Throttling: \url{https://docs.suitecrm.com/8.x/admin/configuration/login-throttling-configuration/}
  \item SuiteCRM 8.x Trusted Hosts: \url{https://docs.suitecrm.com/8.x/admin/configuration/trusted-hosts-configuration/}
  \item SuiteCRM 8.x Two-Factor Authentication: \url{https://docs.suitecrm.com/8.x/features/two-factor/two-factor/}
\end{itemize}

\end{document}
