% =============================================================================
% example-technical-design-spec.tex
% Style: ltechnical-design-spec
% Build: pdflatex via latexmk; minted-aware with listings fallback.
% =============================================================================
\documentclass[11pt]{article}
\usepackage{ltechnical-design-spec}

\setDocTitle{AetherForge Render Pipeline -- Design Specification}
\setDocSubtitle{Forward + tile-binned lighting; native and WebGL2 backends}
\setDocVersion{v0.4}
\setDocStatus{Draft}
\setDocOwner{Engine Architecture}
\setDocAuthor{Jordan Suber}
\setDocDate{May 8, 2026}
\setDocSummary{%
  Reference design specification demonstrating the design-spec subtype:
  numbered design decisions, two architecture decision records with
  status pills, and a requirements-to-components traceability matrix.%
}

\begin{document}
\makedoctocpage

\section{Scope}
This specification covers the runtime render pipeline of the AetherForge
Engine: the order of stages, the data each stage consumes and produces,
and the constraints under which the design must operate on both the
native (OpenGL ES 3.0) and web (WebGL 2.0) backends.

\section{Requirements}
\Req{REQ-RP-001}{The pipeline shall produce identical rendered output
                 (within tolerance) on the native and WebGL backends for
                 the reference scene set.}
\Req{REQ-RP-002}{The pipeline shall scale from a single directional
                 light to at least 256 dynamic point lights without
                 exceeding 16~ms of GPU time on the reference hardware.}
\Req{REQ-RP-003}{The pipeline shall produce a deterministic frame for a
                 given input state, suitable for visual regression tests.}
\Req{REQ-RP-004}{The pipeline shall be implementable on the WebGL 2.0
                 feature set without compute shaders.}

\section{Architectural decision records}

\begin{adr}{Forward + tile-binned lighting over deferred shading}{Accepted}
  \ADRContext{%
    Two viable approaches were considered for many-light support:
    deferred shading and forward + tile-binned lighting. WebGL 2.0
    lacks compute shaders, which would limit a deferred design to
    fragment-shader light culling. The render-target memory cost of a
    deferred G-buffer is also significant on integrated GPUs targeted
    by the WebGL build.%
  }
  \ADRDecision{%
    Adopt forward + tile-binned lighting. CPU-side tile binning produces
    a per-tile light list, uploaded once per frame. The fragment shader
    iterates only the lights touching its tile.%
  }
  \ADRConsequences{%
    Native and WebGL backends share a single shading model. Memory
    footprint stays modest. The CPU bin step adds ~0.4~ms on the
    reference scene, well within budget.%
  }
\end{adr}

\begin{adr}{Use linear-space lighting end-to-end}{Accepted}
  \ADRContext{%
    Mixed-space lighting (sRGB inputs, linear shading, sRGB output) is a
    common source of subtle rendering bugs. WebGL 2.0 supports
    \texttt{EXT\_sRGB} on textures and framebuffers; the native backend
    has had this support since OpenGL ES 3.0.%
  }
  \ADRDecision{%
    Author all colour textures in sRGB space; the GPU converts to linear
    on sample. Shading happens in linear space; final output is written
    to an sRGB framebuffer.%
  }
  \ADRConsequences{%
    Output matches across backends. \Component{ContentLoader} must mark
    colour textures with the sRGB internal format; this is enforced by a
    validation pass at content-build time.%
  }
\end{adr}

\section{Numbered design decisions}

\DesignDecision{Tile size}{%
  Tiles are 16x16 pixels. Smaller tiles reduce wasted shading from
  light-count overestimation; 16x16 is the empirical sweet spot on the
  reference hardware and aligns with WebGL2 fragment-derivative
  granularity.}

\DesignDecision{Light list upload}{%
  Per-tile light lists are uploaded as a single buffer texture each
  frame. Buffer textures avoid the per-tile uniform setup that would
  otherwise cost a CPU-side draw-call dispatch.}

\DesignDecision{Shadow map cascading}{%
  Directional shadows use four cascades; the cascade split distances
  are computed once per frame on the CPU and uploaded as a uniform.
  Per-pixel cascade selection happens in the fragment shader.}

\section{Components}

\Component{Frame Graph} -- the orchestration layer that sequences passes
and tracks their resource dependencies.\\
\Component{Tile Binner} -- CPU-side per-frame stage that produces the
per-tile light list.\\
\Component{Shadow Pipeline} -- the cascaded directional and point-light
shadow maps producer.\\
\Component{Forward Shader} -- the per-fragment shader that consumes the
tile light list and shadow maps to produce the final colour.

\section{Traceability}

\begin{traceability}
  \Trace{REQ-RP-001}{Backend output parity}%
        {Frame Graph, Forward Shader}{TC-PARITY-{01..06}}
  \Trace{REQ-RP-002}{Many-light scalability}%
        {Tile Binner, Forward Shader}{TC-LIGHTS-{01,02,03}}
  \Trace{REQ-RP-003}{Frame determinism}%
        {Frame Graph}{TC-DET-{01,02}}
  \Trace{REQ-RP-004}{WebGL 2.0 implementability}%
        {Tile Binner, Forward Shader}{TC-WEB-{01,02}}
\end{traceability}

\end{document}



