\documentclass[11pt]{article}

% ---------- Page and typography ----------
\usepackage[letterpaper,margin=0.78in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{listings}
\usepackage[most]{tcolorbox}
\usepackage{hyperref}

% ---------- Colors ----------
\definecolor{HouseNavy}{HTML}{17365D}
\definecolor{HouseBlue}{HTML}{1F4E79}
\definecolor{HouseLightBlue}{HTML}{EAF2F8}
\definecolor{HouseGray}{HTML}{F3F5F7}
\definecolor{HouseDarkGray}{HTML}{333333}
\definecolor{HouseGreen}{HTML}{2E7D32}
\definecolor{HouseOrange}{HTML}{B45F06}
\definecolor{CodeBack}{HTML}{F7F7F7}

% ---------- Hyperlinks ----------
\hypersetup{
  colorlinks=true,
  linkcolor=HouseBlue,
  urlcolor=HouseBlue,
  citecolor=HouseBlue,
  pdftitle={Algorithm Analysis Patterns and Mathematics},
  pdfauthor={Jordan Suber},
  pdfsubject={Systematic handout on analysis of algorithms},
  pdfkeywords={algorithm analysis, recurrences, generating functions, asymptotics, analytic combinatorics}
}

% ---------- Header/footer ----------
\pagestyle{fancy}
\fancyhf{}
\lhead{Algorithm Analysis Patterns and Mathematics}
\rhead{Sedgewick--Flajolet Study Handout}
\cfoot{\thepage}
\setlength{\headheight}{14pt}

% ---------- Section formatting ----------
\titleformat{\section}{\Large\bfseries\color{HouseNavy}}{\thesection}{0.7em}{}
\titleformat{\subsection}{\large\bfseries\color{HouseBlue}}{\thesubsection}{0.7em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{HouseDarkGray}}{\thesubsubsection}{0.7em}{}
\titlespacing*{\section}{0pt}{1.2em}{0.4em}
\titlespacing*{\subsection}{0pt}{0.9em}{0.25em}
\titlespacing*{\subsubsection}{0pt}{0.65em}{0.2em}

% ---------- Tables ----------
\renewcommand{\arraystretch}{1.18}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

% ---------- Lists ----------
\setlist[itemize]{leftmargin=1.25em,itemsep=0.2em,topsep=0.25em}
\setlist[enumerate]{leftmargin=1.45em,itemsep=0.2em,topsep=0.25em}

% ---------- Boxes ----------
\tcbset{
  enhanced,
  boxrule=0.6pt,
  arc=2mm,
  left=1.1mm,
  right=1.1mm,
  top=0.9mm,
  bottom=0.9mm
}
\newtcolorbox{keybox}[1]{
  colback=HouseLightBlue,
  colframe=HouseBlue,
  title=\textbf{#1},
  fonttitle=\bfseries
}
\newtcolorbox{patternbox}[1]{
  colback=HouseGray,
  colframe=HouseDarkGray,
  title=\textbf{#1},
  fonttitle=\bfseries
}
\newtcolorbox{mathbox}[1]{
  colback=white,
  colframe=HouseGreen,
  title=\textbf{#1},
  fonttitle=\bfseries
}
\newtcolorbox{warningbox}[1]{
  colback=orange!6,
  colframe=HouseOrange,
  title=\textbf{#1},
  fonttitle=\bfseries
}

% ---------- Code listings ----------
\lstdefinestyle{handout}{
  backgroundcolor=\color{CodeBack},
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  rulecolor=\color{HouseDarkGray},
  columns=fullflexible,
  showstringspaces=false,
  keepspaces=true
}
\lstset{style=handout}

% ---------- Math shorthand ----------
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\GF}{\mathrm{GF}}
\newcommand{\OGF}{\mathrm{OGF}}
\newcommand{\EGF}{\mathrm{EGF}}
\newcommand{\PGF}{\mathrm{PGF}}
\newcommand{\BGF}{\mathrm{BGF}}
\newcommand{\Coeff}[2]{[#1^{#2}]}
\newcommand{\BigO}{\mathcal{O}}

\begin{document}
\pagenumbering{gobble}

% ---------- Title page ----------
\begin{titlepage}
  \centering
  \vspace*{0.7in}
  {\Huge\bfseries\color{HouseNavy} Algorithm Analysis Patterns\\[0.15em] and Mathematics\par}
  \vspace{0.3in}
  {\Large A systematic study handout based on Sedgewick and Flajolet's\\ \emph{An Introduction to the Analysis of Algorithms}, Second Edition\par}
  \vspace{0.45in}
  \begin{tcolorbox}[colback=HouseLightBlue,colframe=HouseBlue,width=0.86\textwidth]
  \textbf{Purpose.} This handout extracts the reusable analysis patterns, cost models, input models, recurrences, generating-function methods, asymptotic tools, and combinatorial structures that recur throughout the book. It is designed as a practical reference for analyzing concrete algorithms rather than as a chapter-by-chapter summary.
  \end{tcolorbox}
  \vfill
  {\large Prepared for Jordan Suber\par}
  {\large \today\par}
  \vspace{0.3in}
  {\small Source base: Robert Sedgewick and Philippe Flajolet, \emph{An Introduction to the Analysis of Algorithms}, Second Edition, Pearson, 2013.}
\end{titlepage}
\clearpage
\pagenumbering{roman}

\tableofcontents
\newpage
\pagenumbering{arabic}

\section{Executive Summary}

Sedgewick and Flajolet present algorithm analysis as a scientific modeling discipline. The dominant theme is not merely classification by asymptotic order, but the construction of precise mathematical models that predict resource usage for actual algorithm implementations. The book emphasizes average-case and probabilistic analysis while also covering worst-case and complexity-oriented tools. Its mathematical toolkit includes recurrences, generating functions, asymptotic approximations, and analytic combinatorics, then applies those tools to trees, permutations, strings, tries, words, hashing, and random mappings.\cite{SF2013}

\begin{keybox}{Central Thesis}
Algorithm analysis converts program structure plus an input model into a mathematical object. Solving that object yields a performance prediction: exact, asymptotic, distributional, or empirical-model based.
\end{keybox}

\subsection{The Master Pattern}

\begin{center}
\begin{tabularx}{\textwidth}{>{\bfseries\color{HouseBlue}}L{0.20\textwidth} Y}
\toprule
Stage & Meaning \\
\midrule
Program & Concrete implementation whose operations can be counted. \\
Cost model & Selected primitive operations: comparisons, exchanges, probes, bit inspections, recursive calls, pointer traversals, or memory moves. \\
Input model & Worst-case input, random permutation, random bitstring, balls-and-urns model, random mapping, or another distribution. \\
Mathematical model & Recurrence, sum, probability distribution, generating function, symbolic combinatorial specification, or asymptotic expansion. \\
Solution & Exact formula, leading term, lower-order correction, variance, distribution, or bound. \\
Validation & Empirical comparison against implementation behavior. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{How to Use This Handout}

Use this document as a reusable analysis checklist. For any algorithm, identify the cost model, choose the input model, derive the mathematical object, solve or approximate it, and then interpret the result in engineering terms.

\begin{lstlisting}[caption={Reusable analysis workflow}]
1. Define input size N.
2. Select the cost variable C_N: comparisons, exchanges, probes, path length, etc.
3. Choose the input model: worst case, random permutation, random bitstring,
   random keys, random mapping, balls-and-urns, etc.
4. Derive a recurrence, sum, distribution, or generating function.
5. Solve exactly when possible.
6. Extract asymptotics when exact expressions are unwieldy.
7. Validate against empirical measurements.
8. Use the result to compare algorithms, tune parameters, or predict capacity.
\end{lstlisting}

\section{Source Map of the Book}

The book is organized around a mathematical progression: first, the motivation and methodology of algorithm analysis; second, the discrete mathematical machinery; third, applications to combinatorial structures central to algorithms.

\begin{longtable}{L{0.20\textwidth} L{0.26\textwidth} L{0.44\textwidth}}
\toprule
\textbf{Book Area} & \textbf{Primary Topics} & \textbf{Analysis Contribution} \\
\midrule
\endhead
Analysis of algorithms & Motivation, theory, quicksort, average case, randomized algorithms & Establishes the scientific workflow: implement, model costs, model inputs, solve mathematically, validate empirically. \\
Recurrence relations & First-order, nonlinear, higher-order, divide-and-conquer recurrences & Converts recursive program structure into solvable equations. \\
Generating functions & OGF, EGF, PGF, BGF, functional equations, special functions & Converts sequences, counts, recurrences, and distributions into algebraic or analytic objects. \\
Asymptotic approximations & Asymptotic notation, expansions, finite sums, Euler-Maclaurin, Laplace method, normal and Poisson examples & Converts exact or implicit answers into interpretable leading terms and correction terms. \\
Analytic combinatorics & Symbolic method, labelled and unlabelled classes, parameters, coefficient asymptotics & Provides a systematic grammar-to-generating-function method for recursive discrete structures. \\
Trees & Binary trees, Catalan trees, BSTs, path length, height, Lagrange inversion, labelled trees & Connects recursive algorithms to implicit tree structures and path-length parameters. \\
Permutations & Inversions, cycles, records, sorting algorithms, insertion sort, selection sort, in-situ permutation & Shows how sorting costs map to permutation statistics. \\
Strings and tries & String searching, bitstrings, regular expressions, automata, KMP, tries & Links algorithm costs to prefix structure and language models. \\
Words and mappings & Hashing, balls-and-urns, birthday paradox, coupon collector, occupancy, open addressing, random mappings & Provides probabilistic models for hashing, collision behavior, cycle detection, and rho phenomena. \\
\bottomrule
\end{longtable}

\section{Algorithm Analysis Taxonomy}

\subsection{Theory of Algorithms vs. Analysis of Algorithms}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.22\textwidth} Y Y}
\toprule
\textbf{Dimension} & \textbf{Theory of Algorithms} & \textbf{Analysis of Algorithms} \\
\midrule
Primary question & What is the asymptotic difficulty of the problem or algorithm class? & What does this concrete algorithm cost under a realistic model? \\
Typical result & Upper bound, lower bound, tight complexity class. & Exact formula, average cost, variance, distribution, or refined asymptotic estimate. \\
Input emphasis & Often worst case. & Often average case or probabilistic input models. \\
Precision & Constant factors often suppressed. & Constant factors and lower-order terms may matter. \\
Use case & Algorithm design and impossibility guidance. & Prediction, comparison, tuning, and empirical validation. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{mathbox}{Asymptotic Classification}
For functions \(f(N)\) and \(g(N)\):
\[
 g(N) \in O(f(N)) \quad \text{means an eventual upper bound,}
\]
\[
 g(N) \in \Omega(f(N)) \quad \text{means an eventual lower bound,}
\]
\[
 g(N) \in \Theta(f(N)) \quad \text{means matching upper and lower bounds.}
\]
\end{mathbox}

\subsection{Exact, Average-Case, and Distributional Goals}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.24\textwidth} L{0.27\textwidth} Y}
\toprule
\textbf{Goal} & \textbf{Mathematical Object} & \textbf{Engineering Meaning} \\
\midrule
Worst-case bound & \(\max_x C(x)\) & Guarantee for all inputs. \\
Best-case bound & \(\min_x C(x)\) & Lower envelope of behavior. \\
Average-case estimate & \(\E[C_N]\) & Expected performance under a chosen input distribution. \\
Variance / deviation & \(\Var(C_N)\), standard deviation & Stability of the average-case prediction. \\
Distributional result & \(\Prob(C_N=k)\) or \(P_N(u)\) & Complete probabilistic description of cost. \\
Cumulative cost & \(T_N = \sum_x C(x)\) & Total cost over all inputs, divided by the number of inputs. \\
Parameter optimization & \(C_N(M)\), \(C_N(s)\), etc. & Select threshold, sample size, table size, load factor, or pivot strategy. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Pattern Catalog}

\begin{longtable}{L{0.20\textwidth} L{0.23\textwidth} L{0.23\textwidth} L{0.20\textwidth}}
\toprule
\textbf{Pattern} & \textbf{Recognition Signal} & \textbf{Mathematical Form} & \textbf{Typical Tools} \\
\midrule
\endhead
Worst-case classification & Need guarantees for all inputs & Upper, lower, and tight asymptotic notation & Bounds, adversarial inputs, decision trees \\
Exact cost counting & Dominant primitive operation is countable & \(C_N\), \(B_N\), \(A_N\) & Summation, recurrences, induction \\
Average-case analysis & Inputs modeled probabilistically & \(\E[C_N]\) & Random variables, recurrences, counting \\
Distributional analysis & Need full cost distribution & \(\Prob(C_N=k)\), \(P_N(u)\) & PGFs, bivariate GFs, moments \\
Cumulative analysis & Easier to sum total cost over all inputs & \(\bar C_N=T_N/|\Omega_N|\) & Total-cost recurrences, generating functions \\
Recursive decomposition & Algorithm solves subproblems recursively & \(C_N=C_i+C_{N-i-1}+t_N\) & Divide-and-conquer recurrences \\
Sampling-based decomposition & Sample changes subproblem-size distribution & Weighted recurrence over split sizes & Probability distributions, OGFs \\
Symbolic combinatorial modeling & Objects have recursive grammar & \(\mathcal A=\mathcal B+\mathcal C\), \(T=z\Phi(T)\) & Symbolic method, Lagrange inversion \\
Coefficient extraction & Need count or cost at size \(N\) & \([z^N]F(z)\) or \([z^N]F(z)/n!\) & OGF, EGF, transfer, singularity analysis \\
Occupancy analysis & Items placed into buckets or slots & Balls-and-urns probabilities & Binomial, Poisson, extreme values \\
Tree path analysis & Search/insertion creates a tree & Internal/external path length & Catalan numbers, BST recurrence, additive parameters \\
Permutation-statistic analysis & Cost depends on order statistics & Inversions, cycles, records & Stirling numbers, Eulerian numbers, CGFs \\
String/trie analysis & Cost depends on prefixes or bit positions & Prefix trees, automata, languages & Tries, regular expressions, automata, GFs \\
Random mapping analysis & Repeated function iteration creates cycle and tail & Tail length, cycle length, rho length & Mapping decompositions, EGFs, asymptotics \\
\bottomrule
\end{longtable}

\section{The Mathematical Toolkit}

\subsection{Notation and Constants}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.20\textwidth} Y}
\toprule
\textbf{Notation} & \textbf{Meaning and Use} \\
\midrule
\(\lfloor x\rfloor\), \(\lceil x\rceil\) & Floor and ceiling; common in divide-and-conquer splits. \\
\(\lg N\) & Binary logarithm, \(\log_2 N\); natural in comparison trees, bits, and binary recurrences. \\
\(\ln N\) & Natural logarithm; common in harmonic approximations and analytic asymptotics. \\
\(\binom{n}{k}\) & Number of ways to choose \(k\) objects from \(n\); central to sampling and occupancy. \\
\(\gamma\) & Euler's constant; appears in harmonic-number approximations. \\
\(\sqrt{2\pi}\) & Stirling constant factor; appears in factorial approximations. \\
\(\left[{n\atop k}\right]\) & Stirling number of the first kind; counts permutations of \(n\) elements with \(k\) cycles. \\
\(\left\{{n\atop k}\right\}\) & Stirling number of the second kind; counts partitions into \(k\) nonempty subsets. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Core Sequences and Approximations}

\begin{mathbox}{Harmonic Numbers}
\[
H_N=\sum_{k=1}^{N}\frac{1}{k}
\]
A standard approximation used throughout average-case analysis is
\[
H_N = \ln N + \gamma + \frac{1}{2N}+O(N^{-2}).
\]
\end{mathbox}

\begin{mathbox}{Stirling Approximation}
\[
N! \sim \sqrt{2\pi N}\left(\frac{N}{e}\right)^N.
\]
It is used for sorting lower bounds, permutation counts, and labelled combinatorial structures.
\end{mathbox}

\begin{mathbox}{Catalan Numbers}
\[
C_N = \frac{1}{N+1}\binom{2N}{N},
\qquad
C_N \sim \frac{4^N}{\sqrt{\pi}N^{3/2}}.
\]
Catalan numbers enumerate many binary-tree and parenthesization structures.
\end{mathbox}

\section{Cost Models and Input Models}

\subsection{Cost Model Patterns}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.22\textwidth} L{0.28\textwidth} Y}
\toprule
\textbf{Cost Variable} & \textbf{Typical Algorithms} & \textbf{Reason It Matters} \\
\midrule
Comparisons & Sorting, searching, BST insertion & Dominant decision cost in comparison-based algorithms. \\
Exchanges / moves & Sorting, heap algorithms, in-place permutation & Captures data movement cost, especially for large records. \\
Probes & Hashing and symbol tables & Measures table access and collision behavior. \\
Bit inspections & Radix sorting, tries, digital search & Natural for string and bitstring input models. \\
Path length & Trees, tries, search structures & Translates tree shape into search or construction cost. \\
Recursive calls & Divide-and-conquer algorithms & Captures recursion tree size and overhead. \\
Memory allocation & Tree/trie construction, dynamic structures & Critical when allocation dominates CPU cost. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Input Model Patterns}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.24\textwidth} L{0.30\textwidth} Y}
\toprule
\textbf{Input Model} & \textbf{Used For} & \textbf{Typical Mathematical Consequence} \\
\midrule
Worst case & Complexity guarantees & Bounds or adversarial construction. \\
Random permutation & Quicksort, BSTs, insertion sort, permutation algorithms & Uniform distribution over \(N!\) orderings. \\
Random distinct keys & Sorting and search & Often equivalent to random permutation for comparison structure. \\
Random bitstrings & Radix sort, tries, digital search & Binomial splitting and prefix distributions. \\
Balls-and-urns & Hashing, occupancy, collision analysis & Bucket loads, empty cells, maximum load, coupon collector behavior. \\
Random mapping & Cycle detection, Pollard-rho-style behavior, functional graphs & Decomposition into components with cycles and trees. \\
Random labelled structures & Labelled trees, mappings, permutations & EGF and symbolic-method analysis. \\
\bottomrule
\end{tabularx}
\end{center}

\begin{warningbox}{Model Validity}
A mathematically elegant result is only as useful as the cost and input models behind it. Treat each model as a hypothesis: validate it empirically and refine it when the data disagree.
\end{warningbox}

\section{Recurrence Relation Patterns}

Recurrences are the most direct way to encode recursive algorithm structure. They typically arise by conditioning on the first split, root, pivot, or recursive subproblem size.

\subsection{Mergesort Compare Recurrence}

Mergesort divides the input into two nearly equal halves, recursively sorts each half, and merges the results.

\begin{mathbox}{Mergesort}
\[
C_N = C_{\lfloor N/2\rfloor}+C_{\lceil N/2\rceil}+N,
\qquad C_1=0.
\]
For powers of two, this gives
\[
C_N=N\lg N.
\]
A refined exact expression is
\[
C_N = N\lceil \lg N\rceil + N - 2^{\lceil \lg N\rceil}.
\]
\end{mathbox}

\subsection{Quicksort Average Compare Recurrence}

Under a random permutation model, the pivot rank is equally likely to be any of the \(N\) positions. Conditioning on the pivot rank yields:

\begin{mathbox}{Quicksort}
\[
C_N = N+1 + \frac{1}{N}\sum_{j=1}^{N}\left(C_{j-1}+C_{N-j}\right),
\qquad C_0=C_1=0.
\]
The average number of comparisons is
\[
C_N = 2(N+1)H_{N+1}-3N-2.
\]
Using \(H_N\sim \ln N+\gamma\), the leading behavior is
\[
C_N = 2N\ln N + O(N).
\]
\end{mathbox}

\subsection{Median-of-Three Quicksort Recurrence}

Sampling changes the pivot-rank distribution. A median-of-three pivot favors balanced partitions, so the recurrence has weighted split probabilities rather than a uniform split.

\begin{mathbox}{Sampling Alters Split Distribution}
\[
C_N = N+1 + \sum_{k=1}^{N}
\frac{(N-k)(k-1)}{\binom{N}{3}}
\left(C_{k-1}+C_{N-k}\right).
\]
\end{mathbox}

\subsection{Binary Search Tree Construction}

Binary search tree insertion from a random permutation has a recurrence structurally related to quicksort. The root rank plays the same role as the pivot rank.

\begin{mathbox}{BST Internal Path Length}
\[
C_N = N-1 + \frac{1}{N}\sum_{k=1}^{N}\left(C_{k-1}+C_{N-k}\right),
\qquad C_0=0.
\]
The average number of comparisons for constructing a random BST is
\[
2(N+1)(H_{N+1}-1)-2N
\approx 1.386N\lg N - 2.846N.
\]
\end{mathbox}

\subsection{Binary Divide-and-Conquer Recurrence}

Many binary divide-and-conquer algorithms have the pattern
\[
C_N = aC_{N/b}+f(N),
\]
where \(a\) is the number of subproblems, \(N/b\) is the subproblem size, and \(f(N)\) is the toll function. The toll is the nonrecursive work performed at size \(N\). The solution is often interpreted through the recursion tree.

\section{Generating Function Patterns}

Generating functions transform sequences into algebraic or analytic objects. They are useful both for solving recurrences and for counting combinatorial structures directly.

\subsection{Ordinary Generating Functions}

\begin{mathbox}{OGF}
\[
A(z)=\sum_{n\ge0}a_nz^n,
\qquad
 a_n=[z^n]A(z).
\]
Use OGFs for unlabelled structures, ordinary sequences, and many recurrence-derived quantities.
\end{mathbox}

\subsection{Exponential Generating Functions}

\begin{mathbox}{EGF}
\[
A(z)=\sum_{n\ge0}a_n\frac{z^n}{n!}.
\]
Use EGFs for labelled objects such as labelled trees, permutations, and mappings.
\end{mathbox}

\subsection{Probability Generating Functions}

\begin{mathbox}{PGF}
\[
P_N(u)=\sum_k \Prob\{X_N=k\}u^k.
\]
Moments follow from derivatives:
\[
\E[X_N]=P_N'(1),
\qquad
\E[X_N(X_N-1)]=P_N''(1).
\]
\end{mathbox}

\subsection{Bivariate Generating Functions}

\begin{mathbox}{BGF for Parameters}
\[
A(z,u)=\sum_{n,k}a_{n,k}z^nu^k.
\]
When \(u\) marks a cost or parameter,
\[
\E[K_N]
=
\frac{[z^N]\left.\frac{\partial}{\partial u}A(z,u)\right|_{u=1}}
{[z^N]A(z,1)}.
\]
\end{mathbox}

\subsection{Generating-Function Workflow}

\begin{lstlisting}[caption={Generating-function analysis workflow}]
1. Define the target sequence a_n or parameter a_{n,k}.
2. Select OGF, EGF, PGF, or BGF based on the object type.
3. Translate recurrence or combinatorial specification into a functional equation.
4. Solve or transform the functional equation.
5. Extract coefficients.
6. Apply asymptotic tools if the exact coefficient is complex.
\end{lstlisting}

\section{Analytic Combinatorics Patterns}

Analytic combinatorics converts combinatorial specifications into generating functions and then converts generating functions into asymptotic estimates. It is the book's unifying calculus for recursive data structures.

\subsection{Symbolic Method Translation Table}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.24\textwidth} L{0.25\textwidth} Y}
\toprule
\textbf{Construction} & \textbf{Combinatorial Meaning} & \textbf{Generating-Function Effect} \\
\midrule
Disjoint union & Choose one alternative class & Addition \\
Cartesian product & Combine independent components & Multiplication \\
Sequence & Ordered list of components & Geometric-type form \\
Set & Unordered collection & Exponential-type form for labelled classes \\
Cycle & Cyclic arrangement & Logarithmic-type form in labelled settings \\
Marking & Track a statistic or parameter & Introduce a variable \(u\) \\
Recursive class & Object contains objects of same class & Functional equation \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Typical Specifications}

\begin{mathbox}{Binary Tree Example}
A common binary-tree specification is
\[
\mathcal{T}=\mathcal{E}+\mathcal{Z}\times\mathcal{T}\times\mathcal{T},
\]
which translates into
\[
T(z)=1+zT(z)^2.
\]
\end{mathbox}

\begin{mathbox}{Cayley Tree Example}
Rooted labelled trees satisfy an implicit equation of the form
\[
C(z)=ze^{C(z)}.
\]
Such implicit equations are natural candidates for Lagrange inversion.
\end{mathbox}

\begin{mathbox}{Mapping Example}
Random mappings decompose into sets of cycles of rooted trees, leading to a specification of the form
\[
\mathrm{Mappings}=\mathrm{SET}(\mathrm{CYC}(\mathrm{Trees})).
\]
This type of decomposition explains why mappings connect tree enumeration, permutation cycles, and functional-graph behavior.
\end{mathbox}

\section{Asymptotic Approximation Patterns}

Exact answers are often too complex to interpret directly. Asymptotic approximations expose the leading driver of cost and retain correction terms when those constants matter.

\subsection{Finite-Sum Approximation}

\begin{mathbox}{Sum-to-Integral Heuristic}
\[
\sum_{k=1}^{N} f(k) \approx \int_1^N f(x)\,dx.
\]
This is the first approximation step for many cost sums.
\end{mathbox}

\subsection{Euler-Maclaurin Summation}

Euler-Maclaurin summation refines integral approximations by adding endpoint and derivative corrections. It is especially useful for harmonic sums and finite sums arising from expected costs.

\begin{mathbox}{Refinement Pattern}
\[
\sum_{k=a}^{b} f(k)
=\int_a^b f(x)\,dx + \frac{f(a)+f(b)}{2}+\text{correction terms}.
\]
\end{mathbox}

\subsection{Coefficient Asymptotics}

\begin{mathbox}{Transfer-Style Pattern}
If a generating function behaves near its dominant singularity like
\[
F(z)\sim \frac{c}{(1-z)^\alpha},
\]
then the coefficients often satisfy
\[
[z^n]F(z)\sim \frac{c}{\Gamma(\alpha)}n^{\alpha-1}.
\]
\end{mathbox}

\subsection{Laplace Method}

\begin{mathbox}{Dominant Contribution Pattern}
For sums or integrals shaped like
\[
\sum_k e^{N\phi(k/N)},
\]
large \(N\) behavior is dominated by the maximum of \(\phi\). The analysis proceeds by locating the peak and approximating the local shape around it.
\end{mathbox}

\subsection{Normal and Poisson Approximations}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.26\textwidth} Y Y}
\toprule
\textbf{Approximation} & \textbf{When It Appears} & \textbf{Algorithmic Use} \\
\midrule
Normal & Many weakly dependent contributions or central binomial regimes & Distributional approximations, finite sums, occupancy behavior. \\
Poisson & Rare events across many trials & Empty buckets, collisions, pattern counts, sparse occupancy. \\
Ramanujan-style functions & Delicate occupancy and extreme-value approximations & Hashing and related asymptotic estimates. \\
\bottomrule
\end{tabularx}
\end{center}

\section{Applications by Combinatorial Structure}

\subsection{Sorting Algorithms}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.22\textwidth} L{0.27\textwidth} Y}
\toprule
\textbf{Algorithm / Problem} & \textbf{Cost Model} & \textbf{Analysis Pattern} \\
\midrule
Mergesort & Comparisons & Balanced divide-and-conquer recurrence; exact and asymptotic solution. \\
Quicksort & Comparisons, exchanges, partitioning cost & Random pivot-rank recurrence; harmonic-number solution; variance and distribution possible. \\
Median-of-three quicksort & Comparisons and sampling cost & Weighted split distribution from sampling; recurrence solved with generating functions. \\
Insertion sort & Exchanges or comparisons & Inversions in random permutations. \\
Selection sort & Comparisons and exchanges & Left-to-right minima and selection statistics. \\
Non-comparison sorting & Key inspection or digit passes & Escapes comparison lower bound by changing the computational model. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Trees}

\begin{patternbox}{Tree Analysis Pattern}
Recursive programs often induce implicit trees. Search cost, construction cost, and recursive overhead can frequently be interpreted as path length or height in a corresponding tree model.
\end{patternbox}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.22\textwidth} L{0.32\textwidth} Y}
\toprule
\textbf{Tree Class} & \textbf{Mathematical Object} & \textbf{Typical Analysis Target} \\
\midrule
Binary trees & Catalan numbers, \(T=1+zT^2\) & Enumeration, path length, height. \\
Binary search trees & Random permutations and recursive rank splits & Construction cost, internal path length, search cost. \\
Rooted unordered trees & Recursive specifications and generating functions & Enumeration and structural parameters. \\
Labelled trees & EGFs, Cayley-style equations & Labelled enumeration and mapping connections. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Permutations}

\begin{patternbox}{Permutation Statistic Pattern}
Many sorting algorithms analyze more cleanly when their cost is expressed as a statistic on permutations.
\end{patternbox}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.24\textwidth} L{0.30\textwidth} Y}
\toprule
\textbf{Statistic} & \textbf{Meaning} & \textbf{Algorithmic Connection} \\
\midrule
Inversions & Pairs out of order & Insertion-sort movement and exchange count. \\
Cycles & Decomposition of a permutation into cyclic components & In-situ permutation and rearrangement algorithms. \\
Left-to-right minima & New record lows while scanning left to right & Selection-sort behavior and records. \\
Runs & Monotone segments & Sorting adaptivity and permutation structure. \\
Extremal parameters & Longest cycle, shortest cycle, maximum run, etc. & Often requires advanced asymptotic or probabilistic methods. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Strings and Tries}

\begin{patternbox}{Prefix Structure Pattern}
String algorithms often depend less on the number of strings than on the prefix structure of the input. Tries make prefix sharing explicit, so path length and node counts become cost variables.
\end{patternbox}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.25\textwidth} L{0.31\textwidth} Y}
\toprule
\textbf{Area} & \textbf{Mathematical Model} & \textbf{Typical Cost Target} \\
\midrule
String searching & Pattern occurrences, automata, regular languages & Character inspections, automaton transitions. \\
KMP-style algorithms & Finite-state automata and failure functions & Linear scanning cost and preprocessing cost. \\
Bitstrings & Random binary sequences & Bit inspections, prefix collisions. \\
Tries & Digital trees over alphabets & Node counts, path length, successful/unsuccessful search cost. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Hashing, Words, and Occupancy}

\begin{patternbox}{Occupancy Pattern}
Hashing analysis often reduces to placing \(N\) balls into \(M\) urns. Bucket loads, empty slots, collisions, clustering, and maximum load become occupancy statistics.
\end{patternbox}

\begin{mathbox}{Load Factor}
\[
\alpha=\frac{N}{M}.
\]
The load factor is the first-order control parameter for many hashing costs.
\end{mathbox}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.23\textwidth} L{0.30\textwidth} Y}
\toprule
\textbf{Problem} & \textbf{Model} & \textbf{Analysis Target} \\
\midrule
Separate chaining & Balls into buckets & Chain lengths, successful/unsuccessful search cost. \\
Open addressing & Probing in a finite table & Probe count, clustering, occupancy restrictions. \\
Birthday paradox & Collision probability & First collision threshold. \\
Coupon collector & Time to see every bucket/type & Expected coverage time. \\
Occupancy extremes & Maximum or minimum bucket load & Tail estimates and extreme-value behavior. \\
\bottomrule
\end{tabularx}
\end{center}

\subsection{Random Mappings}

A mapping from an \(N\)-element set to itself creates a functional digraph: each component contains a directed cycle with rooted trees feeding into cycle nodes. This is the structural basis for analyzing cycle detection and rho-style behavior.

\begin{mathbox}{Rho-Length Scale}
For a random point in a random mapping, the expected rho length has the square-root scale
\[
\E[\rho] \sim \sqrt{\frac{\pi N}{2}}.
\]
\end{mathbox}

\section{Reusable Analysis Worksheet}

\begin{longtable}{L{0.24\textwidth} L{0.34\textwidth} L{0.32\textwidth}}
\toprule
\textbf{Question} & \textbf{What to Fill In} & \textbf{Example} \\
\midrule
\endhead
What is the input size? & Define \(N\), plus secondary parameters such as \(M\), alphabet size, or key length. & \(N\) keys; \(M\) hash buckets. \\
What operation dominates? & Select cost variable. & Comparisons, probes, bit inspections. \\
What is the input model? & Worst case, random permutation, random keys, random bitstrings, balls-and-urns, random mapping. & Quicksort on random permutation. \\
Can the algorithm split recursively? & Identify subproblem sizes and toll cost. & \(C_N=C_i+C_{N-i-1}+N+1\). \\
Is the split deterministic or random? & Fixed halves, pivot-rank distribution, bucket distribution, prefix split. & Mergesort fixed; quicksort random. \\
Is a generating function natural? & Pick OGF, EGF, PGF, or BGF. & BGF when size and cost both matter. \\
Is exact solution useful? & Solve exactly if simple enough. & \(C_N=N\lg N\) for mergesort powers of two. \\
What asymptotic form matters? & Keep leading term and significant correction terms. & \(2N\ln N + O(N)\). \\
Does variance matter? & Compute or estimate deviation. & Important when average prediction may be unstable. \\
How will you validate? & Compare formula against measured runtime or operation counts. & Count compares and plot \(C_N/(N\log N)\). \\
\bottomrule
\end{longtable}

\section{Formula Quick Reference}

\begin{center}
\begin{tabularx}{\textwidth}{L{0.28\textwidth} Y}
\toprule
\textbf{Topic} & \textbf{Formula / Pattern} \\
\midrule
Harmonic numbers & \(H_N=\sum_{k=1}^{N}1/k\), \quad \(H_N=\ln N+\gamma+1/(2N)+O(N^{-2})\). \\
Stirling approximation & \(N!\sim\sqrt{2\pi N}(N/e)^N\). \\
Mergesort compares & \(C_N=C_{\lfloor N/2\rfloor}+C_{\lceil N/2\rceil}+N\); for \(N=2^n\), \(C_N=N\lg N\). \\
Sorting lower bound & Comparison sorting needs at least \(\lceil \lg N!\rceil\) comparisons for some input. \\
Quicksort average compares & \(C_N=2(N+1)H_{N+1}-3N-2\). \\
BST construction average & \(2(N+1)(H_{N+1}-1)-2N\). \\
OGF & \(A(z)=\sum_{n\ge0}a_nz^n\). \\
EGF & \(A(z)=\sum_{n\ge0}a_nz^n/n!\). \\
PGF mean & \(\E[X]=P'(1)\). \\
BGF parameter mean & \(\E[K_N]=[z^N]A_u(z,1)/[z^N]A(z,1)\). \\
Catalan numbers & \(C_N=\frac{1}{N+1}\binom{2N}{N}\sim 4^N/(\sqrt{\pi}N^{3/2})\). \\
Hash load factor & \(\alpha=N/M\). \\
Random mapping rho scale & \(\E[\rho]\sim\sqrt{\pi N/2}\). \\
\bottomrule
\end{tabularx}
\end{center}

\section{Practical Study Plan}

\begin{enumerate}
  \item \textbf{Start with cost models.} Practice identifying which operation is counted and why.
  \item \textbf{Solve core recurrences.} Master mergesort, quicksort, binary search, BST construction, and basic divide-and-conquer forms.
  \item \textbf{Build generating-function fluency.} Know when to use OGF, EGF, PGF, and BGF.
  \item \textbf{Learn asymptotic extraction.} Practice converting exact formulas into leading terms and correction terms.
  \item \textbf{Map algorithms to structures.} Relate quicksort to BSTs, sorting to permutations, hashing to occupancy, and digital search to tries.
  \item \textbf{Validate empirically.} For each model, write code that counts the primitive operation and compares empirical means to the formula.
\end{enumerate}

\begin{keybox}{Engineering Interpretation}
The useful output of analysis is not a theorem alone. It is a decision aid: which algorithm to choose, how to tune it, how much memory to provision, how large a data set can be handled, and whether measured behavior is consistent with the intended implementation.
\end{keybox}

\newpage
\begin{thebibliography}{9}
\bibitem{SF2013}
Robert Sedgewick and Philippe Flajolet.
\newblock \emph{An Introduction to the Analysis of Algorithms}.
\newblock Second Edition. Pearson Education, 2013.
\end{thebibliography}

\end{document}
